---
title: "04_causal_forest"
author: Tyler L. McIntosh
date: last-modified
format:
  html:
    embed-resources: true
    html-math-method: katex
    css: styles.css
    toc: true
    toc-depth: 1
    toc-title: Contents
    toc-location: right
    code-fold: true
    code-summary: "Show the code"
---

# Setup for analysis

``` {r setupchunk}
#| echo: true
#| message: false
#| warning: false
#| results: 'hide'

rm(list = ls())

if(!requireNamespace("here", quietly = TRUE)) {
  install.packages("here")
}
library(here)

source(here::here("utils", "functions.R"))

install_and_load_packages(c(
  "tidyverse",
  "tictoc",
  "grf",
  "kernelshap",
  "hstats",
  "glue",
  "shapviz",
  "patchwork",
  "cobalt",
  "purrr",
  "corrplot",
  "gridExtra",
  "tictoc"))

dir_figs <- here::here('figs')
dir_derived <- here::here('data', 'derived')
dir_raw <- here::here('data', 'raw')
dir_manual <- here::here('data', 'manual')
dir_ensure(c(dir_figs,
             dir_derived,
             dir_manual,
             dir_raw))

seed = 1234
set.seed(seed)


```
# Load dats

```{r}

dats_long <- read_csv(here(dir_derived, "test_dats_long.csv"))


#Short dats
dats <- read_csv(here(dir_derived, "test_dats.csv"))
ecoregions <- unique(dats$ecoregion)

```

# Load and explore long data

We see that tree cover shifts lower more dramatically under high severity fire, which is what we would expect. However, these trends vary by ecoregion. Tree cover in the Southern Rockies shifts much more dramatically towards zero--since this is long data, we can theorize that this is potentially due to lack of regeneration (let's test this). We also see that the shift towards lower tree cover is more severe even under low/moderate severity fire in the Southern Rockies than it is in other ecoregions.

When we check the before/after densities of tree cover for our mock fires (i.e. the rest of the forested area used as control), we see that the after_burn values are higher than pre-burn, with the greatest change in the North Cascades. This is what we would expect as a result of growth rates in areas that have not been impacted by fires.

From the trajectories (both ribbon and boxplot) we see that there are varied recovery trajectories by ecoregion. Most ecoregions and fire severity groups have some element of delayed mortality. Most of these appear to last two years, although there are longer delayed mortality patterns in the North Cascades under low/moderate severity fire (~5 years) and under low/moderate severity fire in the Southern Rockies (~10 years!).

Interestingly, across all ecoregions, even in the 'unburned' areas (i.e. CBI < 0.1) we see a trajectory of long-term forest cover loss following fire occurrence, suggesting that the forest pattern changes from fire may have resulted in lowered resilience to other disturbances, or other fallout effects from nearby higher severity patches.

We find that the parallel trajectories hypothesis more or less holds for all ecoregions and fire severities with the exception of unburned (actual) ares in the Northern Cascades.

In the North Cascades and Sierra Nevada we see what looks like a signature of forest decline prior to low, moderate, and high severity burns.

*FINDINGS OF INTEREST:*
- Lack of recovery under high severity fire in S.Rockies, also under ANY fire in S.Nevada!
- Parallel assumption (more or less) holds
- Delayed mortality primarily in low/moderate severity areas (which is what we would expect)
  - QUESTION: Is this delayed mortality linked to post-fire drought?
- There is a notable signature of forest loss even in unburned areas within fire perimeters
  - QUESTION: Is this linked to other disturbances? Or to fire heterogeneity patterning?
- We see forest decline in the years just prior to burn in the North Cascades and Sierra Nevada
  - QUESTION: Is this linked to other disturbances? Does probability of burn / burn severity change depending on prior drought/biotic agents?
  

```{r}


# Plot density plots grouped by fire_group
ggplot(dats_long |>
         filter(fire_timeframe != "year_of_burn"),
       aes(x = rap_TRE, fill = fire_timeframe)) +
  geom_density(alpha = 0.5) +
  labs(
    title = "Density Plot of rap_TRE by Fire severity and pre/post burn",
    x = "Tree cover % (RAP)",
    fill = "Years Since Fire",
    caption = "NOTE: 'unburned_mock' represents the control sample \n'burned' in random years and given CBI values of 0"
  ) +
  theme_minimal() +
  facet_grid(ecoregion ~ burn_sev_tagged)

ggsave(filename = here(dir_figs, "density_p_pre_post_fire_by_sev.png"),
       plot = last_plot(),
       bg = "white")


# Examine trajectories by ecoregion

# BOXPLOTS

ggplot(data = dats_long) +
  geom_boxplot(
    aes(
      x = as.factor(year_from_fire_index),
      y = rap_TRE,
      fill = year_from_fire_index == 0  # TRUE for 0, FALSE otherwise
    )
  ) +
  scale_fill_manual(
    values = c("TRUE" = "red", "FALSE" = "gray70"),
    guide = "none"  # hide legend
  ) +
  labs(
    title = "Tree cover by years from fire",
    y = "Tree cover %",
    x = "Years from fire",
    caption = "NOTE: 'unburned_mock' represents the control sample \n'burned' in random years and given CBI values of 0"
  ) +
  theme_minimal() +
  facet_grid(ecoregion ~ burn_sev_tagged)
ggsave(filename = here(dir_figs, "boxplot_time_series.png"),
       plot = last_plot(),
       bg = "white",
       units = "px",
       width = 5000,
       height = 3000)



# LINE TRAJECTORIES
# Step 1: Summarize the data
summary_data <- dats_long %>%
  group_by(ecoregion, burn_sev_tagged, year_from_fire_index) %>%
  summarise(
    q25 = quantile(rap_TRE, 0.25, na.rm = TRUE),
    median = median(rap_TRE, na.rm = TRUE),
    q75 = quantile(rap_TRE, 0.75, na.rm = TRUE),
    .groups = "drop"
  )

# Step 2: Extract 'unburned-mock' data
unburned_mock <- summary_data %>%
  filter(burn_sev_tagged == "unburned-mock") %>%
  select(ecoregion, year_from_fire_index, median)

# Step 3: Plot
ggplot(summary_data, aes(x = year_from_fire_index, y = median)) +
  # IQR ribbon
  geom_ribbon(aes(ymin = q25, ymax = q75), fill = "gray80", alpha = 0.5) +
  
  # Main median line for each facet
  geom_line(color = "black", size = 1) +

  # Dotted blue overlay from unburned-mock
  geom_line(
    data = unburned_mock,
    aes(y = median),
    color = "blue",
    linetype = "dotted",
    inherit.aes = TRUE
  ) +

  # Solid blue line only in unburned-mock facet
  geom_line(
    data = summary_data %>% filter(burn_sev_tagged == "unburned-mock"),
    aes(y = median),
    color = "blue",
    size = 1,
    linetype = "solid"
  ) +

  geom_vline(xintercept = 0, linetype = "dotted", color = "red", size = 1) +
  
  facet_grid(ecoregion ~ burn_sev_tagged) +

  labs(
    title = "Tree Cover by Years from Fire (IQR shaded)",
    x = "Years from fire",
    y = "Tree cover %",
    caption = "NOTE: 'unburned_mock' represents the control sample \n'burned' in random years and given CBI values of 0"
  ) +
  theme_minimal()
ggsave(filename = here(dir_figs, "iqr_line_time_series.png"),
       plot = last_plot(),
       bg = "white",
       units = "px",
       width = 5000,
       height = 3000)

```


# Load and explore data
- Correlation matrix
- Covariate balance: love & density plots

*FINDINGS:*
- High correlation among post-event tree cover (expected)
- Burn severity and heterogeneity are most highly correlated with post-fire cover in the S.Rockies. Minimal correlation in S.Nevada & ID.Batholith, only with immediate post-fire in N.Cascades
- There's some correlation between long-term average AET/DEF and the HD/PDSI, but not as much as there might be
- The S.Rockies and S.Nevada have the best covariate balance out of the box, while the N.Cascades has extremely poor covariate balance, and IDBatholith isn't great. Both IDB & NC have imbalances in the biotic disturbance before and after fire, which is interesting
- We probably want to conduct pre-DiD matching



```{r}




# Function to assess correlation with corrplot and covariate balance with cobalt, output results
assess_covariate_balance <- function(d, nm) {
  
  # Correlation matrix
  correlations <- d |>
    select(where(is.numeric)) |>
    cor(use = "pairwise.complete.obs")
  png(here(dir_figs, glue('corr_plot_{nm}.png')), width = 2000, height = 2000)
  corrplot(correlations, method = 'circle', type = "lower", diag = FALSE, use = 'pairwise.complete.obs') 
  dev.off()
  
      
    
  cobalt::bal.tab(x = fire ~ aet + def + frg_reclass + srtm + tpi + chili + biotic_5_yrs_prior_sum + biotic_5_yrs_after_sum + pdsi_annual_5_yrs_prior_sum + pdsi_annual_5_yrs_after_sum + avg_forest_2_yr_pre_fire, 
                  data = d)
  
  png(here(dir_figs, glue("love_plot_fire_{nm}.png")))
  cobalt::love.plot(x = fire ~ aet + def + frg_reclass + srtm + tpi + chili + biotic_5_yrs_prior_sum + biotic_5_yrs_after_sum + pdsi_annual_5_yrs_prior_sum + pdsi_annual_5_yrs_after_sum + avg_forest_2_yr_pre_fire, 
                  data = d,
                  stats = c("mean.diffs", "ks.statistics"),
                  stars = "raw",
                  position = "bottom",
                  threshold = c(m = .1, ks = .05),
                  limits = list(c(-1, 1), c(0, 0.6)),
                  title = glue("Covariate balance love plot for FIRE treatment: {nm}"))
  dev.off()

  
    
  fire_covars <- c("aet", "def", "frg_reclass", "srtm", "tpi", "chili",
              "biotic_5_yrs_prior_sum", "biotic_5_yrs_after_sum",
              "pdsi_annual_5_yrs_prior_sum", "pdsi_annual_5_yrs_after_sum", "avg_forest_2_yr_pre_fire")
  
  # Make a working copy of your data
  dats_trans <- d
  
  # Transform biotic variables using log1p (handles 0s safely)
  for (v in fire_covars) {
    if (startsWith(v, "biotic")) {
      new_name <- paste0(v, "_log")
      dats_trans[[new_name]] <- log1p(dats_trans[[v]])
    }
  }
  
  # Plot all variables, using log-transformed versions when appropriate
  plot_list <- lapply(fire_covars, function(v) {
    # If biotic, use transformed version
    if (startsWith(v, "biotic")) {
      v_use <- paste0(v, "_log")
      label <- paste0(v, " (log)")
    } else {
      v_use <- v
      label <- v
    }
  
    # Construct formula and plot
    f <- reformulate(v_use, response = "fire")
    cobalt::bal.plot(
      x = f,
      data = dats_trans,
      var.name = v_use,
      which = "both"
    ) + ggtitle(label)
  })
  
  # Arrange in grid and save
  # Create a title grob
  title_grob <- textGrob(glue("Covariate Balance by FIRE Treatment Group: {nm}"),
                         gp = gpar(fontsize = 20, fontface = "bold"))
  
  # Add title and plots to a single grid
  grid_plot <- grid.arrange(
    title_grob,
    do.call(arrangeGrob, c(plot_list, ncol = 2)),
    nrow = 2,
    heights = c(0.05, 0.95)
  )
  # Save
  ggsave(here(dir_figs, glue("covariate_balance_grid_fire_{nm}.png")), grid_plot, units = "px", width = 3000, height = 3000, dpi = 300)
  
  
  
  
  
  
  # FOR CBI TREATMENT
  
  cobalt::bal.tab(x = burn_cbi ~ aet + def + frg_reclass + srtm + tpi + chili + biotic_5_yrs_prior_sum + biotic_5_yrs_after_sum + pdsi_annual_5_yrs_prior_sum + pdsi_annual_5_yrs_after_sum + avg_forest_2_yr_pre_fire + burn_cbi_1kmmwmedian + burn_cbi_1kmmwstd + mock_burn_year, 
                  data = d)
  
  
  burn_cbi_covars <- c("aet", "def", "frg_reclass", "srtm", "tpi", "chili",
              "biotic_5_yrs_prior_sum", "biotic_5_yrs_after_sum",
              "pdsi_annual_5_yrs_prior_sum", "pdsi_annual_5_yrs_after_sum", "avg_forest_2_yr_pre_fire", "burn_cbi_1kmmwmedian", "burn_cbi_1kmmwstd", "mock_burn_year")
  
  # Make a working copy of data
  dats_trans <- d
  
  # Transform biotic variables using log1p (handles 0s safely)
  for (v in burn_cbi_covars) {
    if (startsWith(v, "biotic")) {
      new_name <- paste0(v, "_log")
      dats_trans[[new_name]] <- log1p(dats_trans[[v]])
    }
  }
  
  # Plot all variables, using log-transformed versions when appropriate
  plot_list <- lapply(burn_cbi_covars, function(v) {
    # If biotic, use transformed version
    if (startsWith(v, "biotic")) {
      v_use <- paste0(v, "_log")
      label <- paste0(v, " (log)")
    } else {
      v_use <- v
      label <- v
    }
  
    # Construct formula and plot
    f <- reformulate(v_use, response = "burn_cbi")
    cobalt::bal.plot(
      x = f,
      data = dats_trans,
      var.name = v_use,
      which = "both"
    ) + ggtitle(label)
  })
  
  # Arrange in grid and save
  # Create a title grob
  title_grob <- textGrob("Covariate Balance by CBI Treatment Group",
                         gp = gpar(fontsize = 20, fontface = "bold"))
  
  # Add title and plots to a single grid
  grid_plot <- grid.arrange(
    title_grob,
    do.call(arrangeGrob, c(plot_list, ncol = 2)),
    nrow = 2,
    heights = c(0.05, 0.95)
  )
  # Save
  ggsave(here(dir_figs, glue("covariate_balance_grid_cbi{nm}.png")), grid_plot, units = "px", width = 3000, height = 6000, dpi = 300)
  
}


# USE FUNCTION FOR ECOREGIONS AND FULL DATSET
for(eco in ecoregions) {
  d <- dats |> filter(ecoregion == eco)
  assess_covariate_balance(d = d,
                           nm = eco)
}
assess_covariate_balance(d = dats,
                         nm = "ALL")







```


```{r}


cf_analysis <- function(dats, yvar, xvars, trt, data_nm, seed = 1234) {
  # Filter out rows with NA in the outcome variable
  dats_filt <- dats %>%
    filter(!is.na(.data[[yvar]]))
  
  n_total <- nrow(dats_filt)
  n_burned <- nrow(dats_filt |>
                      filter(fire == 1))
  
  # Select covariates and outcome
  covariates <- dats_filt %>%
    select(all_of(xvars))
  outcome <- dats_filt[[yvar]]
  treatment <- dats_filt[[trt]]

  # Fit causal forest
  fit <- grf::causal_forest(
    X = covariates,
    Y = outcome,
    W = treatment,
    seed = seed
  )

  # Variable importance - ensure normalized
  imp_raw <- sort(setNames(variable_importance(fit), xvars))
  imp_normalized <- imp_raw / sum(imp_raw)
  
  imp_df <- tibble(
    variable = names(imp_normalized),
    importance = as.numeric(imp_normalized)
  ) %>%
    arrange(importance)  # So the smallest is at the bottom of the horizontal bar plot

  imp_p <- ggplot(imp_df, aes(x = importance, y = reorder(variable, importance))) +
    geom_col(fill = "orange") +
    labs(
      x = "Variable Importance (Normalized)",
      y = NULL,
      title = "Causal Forest Variable Importance",
      subtitle = data_nm,
      caption = glue("Total sample size: {n_total}\nBurned sample size: {n_burned}")
    ) +
    theme_minimal(base_size = 13) +
    theme(
      axis.text.y = element_text(hjust = 1),
      plot.margin = margin(10, 20, 5, 5)
    )
  ggsave(plot = imp_p,
         filename = here(dir_figs, paste0("var_importance_", data_nm, ".png")),
         bg = "white")

  pred_fun <- function(object, newdata, ...) {
    predict(object, newdata, ...)$predictions
  }

  xvars_sorted <- rev(names(imp_normalized))

  pdps <- lapply(xvars_sorted, function(v) {
    pd <- partial_dep(fit, v, X = covariates, pred_fun = pred_fun)
    plot(pd) +
      theme(axis.title.y = element_blank())
  })

  #For adding y axis label to overall plot
  ylab_plot <- ggplot() +
    theme_void() +
    annotate("text", x = 0, y = 0.5, label = "Treatment effect", angle = 90, hjust = 0.5, size = 4)

  pdps_p <- wrap_plots(pdps, guides = "collect", ncol = 3) #&
    #ylab("Treatment effect")

  # Add format & add titles
  pdps_p <- ylab_plot + pdps_p +
    plot_layout(widths = c(0.05, 1)) +
    plot_annotation(
      title = "Partial Dependence Plots",
      subtitle = data_nm,
      caption = glue("NOTES:\nNegative values mean a stronger (positive) treatment effect (i.e. fire is causing a greater difference in forest cover change)\nPDPs are in order of variable importance\nTotal sample size: {n_total}\nBurned sample size: {n_burned}"))

  ggsave(filename = here(dir_figs, paste0("partial_dependence_", data_nm, ".png")),
         plot = pdps_p,
         units = 'px',
         width = 2000,
         height = 2000)

  return(list(fit = fit,
              imp_p = imp_p,
              pdps = pdps))
}
# 
# tic() #takes about 1.5 minutes to run
# test <- cf_analysis(dats = dats |> filter(ecoregion = "IDBatholith"),
#             yvar = "post_fire_forest_difference_10_yr",
#             xvars = c("aet", "def", "frg_reclass", "srtm", "tpi", "chili",
#            "burn_cbi", "burn_cbi_1kmmwmedian", "burn_cbi_1kmmwstd", "mock_burn_year",
#            "biotic_5_yrs_prior_sum", "biotic_5_yrs_after_sum",
#            "pdsi_annual_5_yrs_prior_sum", "pdsi_annual_5_yrs_after_sum", "avg_forest_2_yr_pre_fire"),
#             trt = "fire",
#             data_nm = "idBatholith_post_fire_forest_difference_10_yr",
#             seed = seed)
# toc()
# 
# 
# 
# tic() #takes about 1.5 minutes to run
# test <- cf_analysis(dats = dats |> filter(ecoregion = "IDBatholith"),
#             yvar = "post_fire_forest_slope_5_10_yrs",
#             xvars = c("aet", "def", "frg_reclass", "srtm", "tpi", "chili",
#            "burn_cbi", "burn_cbi_1kmmwmedian", "burn_cbi_1kmmwstd", "mock_burn_year",
#            "biotic_5_yrs_prior_sum", "biotic_5_yrs_after_sum",
#            "pdsi_annual_5_yrs_prior_sum", "pdsi_annual_5_yrs_after_sum",
#            "avg_forest_2_yr_pre_fire"),
#             trt = "fire",
#             data_nm = "idBatholith_post_fire_forest_slope_5_10_yrs",
#             seed = seed)
# toc()
# 
# tic() 
# test <- cf_analysis(dats = dats |> filter(ecoregion = "IDBatholith"),
#             yvar = "post_fire_forest_difference_10_yr",
#             xvars = c("aet", "def", "frg_reclass", "srtm", "tpi", "chili",
#            "burn_cbi_1kmmwmedian", "burn_cbi_1kmmwstd", "mock_burn_year",
#            "biotic_5_yrs_prior_sum", "biotic_5_yrs_after_sum",
#            "pdsi_annual_5_yrs_prior_sum", "pdsi_annual_5_yrs_after_sum",
#            "avg_forest_2_yr_pre_fire"),
#             trt = "burn_cbi",
#             data_nm = "CBITRT_idBatholith_post_fire_forest_difference_10_yr",
#             seed = seed)
# toc()




# OUTCOMES OF INTEREST
outcome_vars <- c("perc_recovery_10yr", "post_fire_forest_slope_2_10_yrs", "post_fire_forest_slope_5_15_yrs", "post_fire_forest_slope_10_15_yrs", "perc_recovery_15yr", "post_fire_forest_slope_5_10_yrs")

# Switch to use pwalk from purrr for ease of switch to parallelized furrr in future
# Create combinations of ecoregions and outcome variables
param_grid <- tidyr::expand_grid(
  eco = ecoregions,
  var = outcome_vars
)

# Walk through each combination and call cf_analysis() for side effects
pwalk(
  .l = list(eco = param_grid$eco, var = param_grid$var),
  .f = function(eco, var) {
    d <- dats |> filter(ecoregion == eco)

    # Binary fire treatment
    cf_analysis(
      dats = d,
      yvar = var,
      xvars = c("aet", "def", "frg_reclass", "srtm", "tpi", "chili",
                "burn_cbi", "burn_cbi_1kmmwmedian", "burn_cbi_1kmmwstd", "mock_burn_year",
                "biotic_5_yrs_prior_sum", "biotic_5_yrs_after_sum",
                "pdsi_annual_5_yrs_prior_sum", "pdsi_annual_5_yrs_after_sum",
                #"pdsi_annual_5_yrs_prior_mean", "pdsi_annual_5_yrs_after_mean",
                "hd_fingerprint_5_yrs_prior_sum", "hd_fingerprint_5_yrs_after_sum",
                #"hd_fingerprint_5_yrs_prior_mean", "hd_fingerprint_5_yrs_after_mean",
                #"avg_forest_2_yr_pre_fire"
                ),
      trt = "fire",
      data_nm = glue("{eco}_{var}_firebinary"),
      seed = seed
    )
    
    # Continuous CBI treatment
    cf_analysis(
      dats = d,
      yvar = var,
      xvars = c("aet", "def", "frg_reclass", "srtm", "tpi", "chili",
                "burn_cbi", "burn_cbi_1kmmwmedian", "burn_cbi_1kmmwstd", "mock_burn_year",
                "biotic_5_yrs_prior_sum", "biotic_5_yrs_after_sum",
                "pdsi_annual_5_yrs_prior_sum", "pdsi_annual_5_yrs_after_sum",
                #"pdsi_annual_5_yrs_prior_mean", "pdsi_annual_5_yrs_after_mean",
                "hd_fingerprint_5_yrs_prior_sum", "hd_fingerprint_5_yrs_after_sum",
                #"hd_fingerprint_5_yrs_prior_mean", "hd_fingerprint_5_yrs_after_mean",
                "avg_forest_2_yr_pre_fire"),
      trt = "burn_cbi",
      data_nm = glue("{eco}_{var}_cbi"),
      seed = seed
    )
  }
)

# RUN FOR FULL DATASET
walk(
  .x = outcome_vars,
  .f = function(var) {
    cf_analysis(
      dats = dats,
      yvar = var,
      xvars = c("aet", "def", "frg_reclass", "srtm", "tpi", "chili",
                "burn_cbi", "burn_cbi_1kmmwmedian", "burn_cbi_1kmmwstd", "mock_burn_year",
                "biotic_5_yrs_prior_sum", "biotic_5_yrs_after_sum",
                "pdsi_annual_5_yrs_prior_sum", "pdsi_annual_5_yrs_after_sum",
                # "pdsi_annual_5_yrs_prior_mean", "pdsi_annual_5_yrs_after_mean",
                "hd_fingerprint_5_yrs_prior_sum", "hd_fingerprint_5_yrs_after_sum",
                # "hd_fingerprint_5_yrs_prior_mean", "hd_fingerprint_5_yrs_after_mean",
                "avg_forest_2_yr_pre_fire"),
      trt = "fire",
      data_nm = glue("ALL_{var}_firebinary"),
      seed = seed
    )
  }
)

walk(
  .x = outcome_vars,
  .f = function(var) {
    cf_analysis(
      dats = dats,
      yvar = var,
      xvars = c("aet", "def", "frg_reclass", "srtm", "tpi", "chili",
                "burn_cbi", "burn_cbi_1kmmwmedian", "burn_cbi_1kmmwstd", "mock_burn_year",
                "biotic_5_yrs_prior_sum", "biotic_5_yrs_after_sum",
                "pdsi_annual_5_yrs_prior_sum", "pdsi_annual_5_yrs_after_sum",
                # "pdsi_annual_5_yrs_prior_mean", "pdsi_annual_5_yrs_after_mean",
                "hd_fingerprint_5_yrs_prior_sum", "hd_fingerprint_5_yrs_after_sum",
                # "hd_fingerprint_5_yrs_prior_mean", "hd_fingerprint_5_yrs_after_mean",
                "avg_forest_2_yr_pre_fire"),
      trt = "burn_cbi",
      data_nm = glue("ALL_{var}_cbi"),
      seed = seed
    )
  }
)




# RUN SIMPLE ANALYSIS
cf_analysis(
  dats = dats,
  yvar = "post_fire_forest_slope_2_10_yrs",
  xvars = c("aet", "def", "frg_reclass", "srtm", "tpi", "chili",
            "biotic_5_yrs_prior_sum", "biotic_5_yrs_after_sum",
            "pdsi_annual_5_yrs_prior_sum", "pdsi_annual_5_yrs_after_sum",
            "hd_fingerprint_5_yrs_prior_sum", "hd_fingerprint_5_yrs_after_sum",
            "avg_forest_2_yr_pre_fire"),
  trt = "fire",
  data_nm = glue("ALL_simple_post_fire_forest_slope_2_10_yrs_fire"),
  seed = seed
)


for(burn_l in c("low-moderate-sev", "unburned", "high-sev")) {
  df <- dats |>
    filter(burn_sev_tagged == burn_l | burn_sev_tagged == "unburned-mock")
  
  cf_analysis(
    dats = df,
    yvar = "post_fire_forest_slope_2_10_yrs",
    xvars = c("aet", "def", "frg_reclass", "srtm", "tpi", "chili",
              "biotic_5_yrs_prior_sum", "biotic_5_yrs_after_sum",
              "pdsi_annual_5_yrs_prior_sum", "pdsi_annual_5_yrs_after_sum",
              "hd_fingerprint_5_yrs_prior_sum", "hd_fingerprint_5_yrs_after_sum",
              "avg_forest_2_yr_pre_fire"),
    trt = "fire",
    data_nm = glue("ALL_simple_{burn_l}_post_fire_forest_slope_2_10_yrs_fire"),
    seed = seed
  )
  
}






# yvar = "post_fire_forest_difference_10_yr"
# xvars = c("aet", "def", "frg_reclass", "srtm", "tpi", "chili",
#            "burn_cbi", "burn_cbi_1kmmwmedian", "burn_cbi_1kmmwstd", "mock_burn_year",
#            "biotic_5_yrs_prior_sum", "biotic_5_yrs_after_sum",
#            "pdsi_annual_5_yrs_prior_sum", "pdsi_annual_5_yrs_after_sum")
# trt = "fire"
#   # Filter out rows with NA in the outcome variable
#   dats_filt <- dats %>%
#     filter(!is.na(.data[[yvar]]))
# 
#   # Select covariates and outcome
#   covariates <- dats_filt %>%
#     select(all_of(xvars))
#   outcome <- dats_filt[[yvar]]
#   treatment <- dats_filt[[trt]]
# 
#   # Fit causal forest
#   fit <- grf::causal_forest(
#     X = covariates,
#     Y = outcome,
#     W = treatment,
#     seed = seed
#   )
# 
#   # Variable importance
#   imp_raw <- sort(setNames(variable_importance(fit), xvars))
#   imp_normalized <- imp / sum(imp)
# 
#   imp_df <- tibble(
#     variable = names(imp_normalized),
#     importance = as.numeric(imp_normalized)
#   ) %>%
#     arrange(importance)  # So the smallest is at the bottom of the horizontal bar plot
# 
#   imp_p <- ggplot(imp_df, aes(x = importance, y = reorder(variable, importance))) +
#     geom_col(fill = "orange") +
#     labs(
#       x = "Variable Importance",
#       y = NULL,
#       title = "Causal Forest Variable Importance",
#       subtitle = "test"
#     ) +
#     theme_minimal(base_size = 13) +
#     theme(
#       axis.text.y = element_text(hjust = 1),
#       plot.margin = margin(10, 20, 5, 5)
#     )
#   imp_p
#   ggsave(plot = imp_p,
#          filename = here(dir_figs, paste0("var_importance_", data_nm, ".png")),
#          bg = "white")
# 
#   pred_fun <- function(object, newdata, ...) {
#     predict(object, newdata, ...)$predictions
#   }
# 
#   xvars_sorted <- rev(names(imp))
# 
#   pdps <- lapply(xvars_sorted, function(v) {
#     pd <- partial_dep(fit, v, X = covariates, pred_fun = pred_fun)
#     plot(pd) +
#       theme(axis.title.y = element_blank())
#   })
# 
#   #For adding y axis label to overall plot
#   ylab_plot <- ggplot() +
#     theme_void() +
#     annotate("text", x = 0, y = 0.5, label = "Treatment effect", angle = 90, hjust = 0.5, size = 4)
# 
#   pdps_p <- wrap_plots(pdps, guides = "collect", ncol = 3) #&
#     #ylab("Treatment effect")
# 
#   # Add format & add titles
#   pdps_p <- ylab_plot + pdps_p +
#     plot_layout(widths = c(0.05, 1)) +
#     plot_annotation(
#       title = "Partial Dependence Plots",
#       subtitle = data_nm,
#       caption = "NOTES:\nNegative values mean a stronger (positive) treatment effect (i.e. fire is causing a greater difference in forest cover change)\nPDPs are in order of variable importance")
# 
#   ggsave(filename = here(dir_figs, paste0("partial_dependence_", data_nm, ".png")),
#          plot = pdps_p,
#          units = 'px',
#          width = 2000,
#          height = 2000)
# 

```


These initial results mean that: (YES = makes sense, NO = doesn't make sense)
- As AET increases (there is more AET), there is greater difference between pre-and post-fire cover
- As DEF increases (sites get drier), there is decreased difference between pre-and post-fire cover
- YES - When there is biotic disturbance prior to burn there is a greater difference between pre and post-fire cover
- ??? - When there is more drought after fire there is greater difference between pre and post-fire cover; but then also a much larger difference between pre and post-fire cover with wetter conditions after fire?..
- YES - The difference between pre and post-fire is least when elevation is ~2000, and greater at low elevatoins or high elevations (at the edges of forest range forest causes more change in forest cover?)
- YES - The difference between pre and post fire forest cover is greatest when CHILI is high (very warm)
- The treatment effect is greatest for FRG classes 3,4,5,6
- YES? - The treatment effect is greatest for negative values of TPI (i.e. in valleys); there are denser forests pre-burn in valleys
- NO - The treatment effect is greatest earlier in the time frame


```{r}
hist(dats$pdsi_annual_5_yrs_after_sum)

```



<button class="accordion-button" type="button" data-bs-toggle="collapse" data-bs-target="#collapseOne" >Click to see Session Information</button><div id="collapseOne" class="accordion-collapse collapse"><div>

```{r sessionInfo}
sessionInfo()
```