---
title: "04_causal_forest"
author: Tyler L. McIntosh
date: last-modified
format:
  html:
    embed-resources: true
    html-math-method: katex
    css: styles.css
    toc: true
    toc-depth: 1
    toc-title: Contents
    toc-location: right
    code-fold: true
    code-summary: "Show the code"
---


``` {r setupchunk}
#| echo: true
#| message: false
#| warning: false
#| results: 'hide'

rm(list = ls())

if(!requireNamespace("here", quietly = TRUE)) {
  install.packages("here")
}
library(here)

source(here::here("utils", "functions.R"))

install_and_load_packages(c(
  "tidyverse",
  "tictoc",
  "grf",
  "kernelshap",
  "hstats",
  "shapviz",
  "patchwork",
  "cobalt",
  "corrplot",
  "gridExtra",
  "tictoc"))

dir_figs <- here::here('figs')
dir_derived <- here::here('data', 'derived')
dir_raw <- here::here('data', 'raw')
dir_manual <- here::here('data', 'manual')
dir_ensure(c(dir_figs,
             dir_derived,
             dir_manual,
             dir_raw))

seed = 1234
set.seed(seed)

```

# Load and explore data
- Correlation matrix
- Covariate balance: love & density plots

```{r}

dats <- read_csv(here(dir_derived, "test_dats.csv"))

hist(dats$post_fire_forest_difference_10_yr)

# Correlation matrix
correlations <- dats |>
  select(where(is.numeric)) |>
  cor(use = "pairwise.complete.obs")
png(here(dir_figs, 'corr_plot.png'), width = 2000, height = 2000)
corrplot(correlations, method = 'circle', type = "lower", diag = FALSE, use = 'pairwise.complete.obs') 
dev.off()



# Use cobalt to asses covariate balance for FIRE treatment

cobalt::bal.tab(x = fire ~ aet + def + frg_reclass + srtm + tpi + chili + biotic_5_yrs_prior_sum + biotic_5_yrs_after_sum + pdsi_annual_5_yrs_prior_sum + pdsi_annual_5_yrs_after_sum + avg_forest_2_yr_pre_fire, 
                data = dats)

png(here(dir_figs, "love_plot_fire.png"))
cobalt::love.plot(x = fire ~ aet + def + frg_reclass + srtm + tpi + chili + biotic_5_yrs_prior_sum + biotic_5_yrs_after_sum + pdsi_annual_5_yrs_prior_sum + pdsi_annual_5_yrs_after_sum + avg_forest_2_yr_pre_fire, 
                data = dats,
                stats = c("mean.diffs", "ks.statistics"),
                stars = "raw",
                position = "bottom",
                threshold = c(m = .1, ks = .05),
                title = "Covariate balance love plot for FIRE treatment")
dev.off()



fire_covars <- c("aet", "def", "frg_reclass", "srtm", "tpi", "chili",
            "biotic_5_yrs_prior_sum", "biotic_5_yrs_after_sum",
            "pdsi_annual_5_yrs_prior_sum", "pdsi_annual_5_yrs_after_sum", "avg_forest_2_yr_pre_fire")

# Make a working copy of your data
dats_trans <- dats

# Transform biotic variables using log1p (handles 0s safely)
for (v in fire_covars) {
  if (startsWith(v, "biotic")) {
    new_name <- paste0(v, "_log")
    dats_trans[[new_name]] <- log1p(dats_trans[[v]])
  }
}

# Plot all variables, using log-transformed versions when appropriate
plot_list <- lapply(fire_covars, function(v) {
  # If biotic, use transformed version
  if (startsWith(v, "biotic")) {
    v_use <- paste0(v, "_log")
    label <- paste0(v, " (log)")
  } else {
    v_use <- v
    label <- v
  }

  # Construct formula and plot
  f <- reformulate(v_use, response = "fire")
  cobalt::bal.plot(
    x = f,
    data = dats_trans,
    var.name = v_use,
    which = "both"
  ) + ggtitle(label)
})

# Arrange in grid and save
# Create a title grob
title_grob <- textGrob("Covariate Balance by FIRE Treatment Group",
                       gp = gpar(fontsize = 20, fontface = "bold"))

# Add title and plots to a single grid
grid_plot <- grid.arrange(
  title_grob,
  do.call(arrangeGrob, c(plot_list, ncol = 2)),
  nrow = 2,
  heights = c(0.05, 0.95)
)
# Save
ggsave(here(dir_figs, "covariate_balance_grid_fire.png"), grid_plot, units = "px", width = 3000, height = 3000, dpi = 300)






# FOR CBI TREATMENT

cobalt::bal.tab(x = burn_cbi ~ aet + def + frg_reclass + srtm + tpi + chili + biotic_5_yrs_prior_sum + biotic_5_yrs_after_sum + pdsi_annual_5_yrs_prior_sum + pdsi_annual_5_yrs_after_sum + avg_forest_2_yr_pre_fire + burn_cbi_1kmmwmedian + burn_cbi_1kmmwstd + mock_burn_year, 
                data = dats)


burn_cbi_covars <- c("aet", "def", "frg_reclass", "srtm", "tpi", "chili",
            "biotic_5_yrs_prior_sum", "biotic_5_yrs_after_sum",
            "pdsi_annual_5_yrs_prior_sum", "pdsi_annual_5_yrs_after_sum", "avg_forest_2_yr_pre_fire", "burn_cbi_1kmmwmedian", "burn_cbi_1kmmwstd", "mock_burn_year")

# Make a working copy of your data
dats_trans <- dats

# Transform biotic variables using log1p (handles 0s safely)
for (v in burn_cbi_covars) {
  if (startsWith(v, "biotic")) {
    new_name <- paste0(v, "_log")
    dats_trans[[new_name]] <- log1p(dats_trans[[v]])
  }
}

# Plot all variables, using log-transformed versions when appropriate
plot_list <- lapply(burn_cbi_covars, function(v) {
  # If biotic, use transformed version
  if (startsWith(v, "biotic")) {
    v_use <- paste0(v, "_log")
    label <- paste0(v, " (log)")
  } else {
    v_use <- v
    label <- v
  }

  # Construct formula and plot
  f <- reformulate(v_use, response = "burn_cbi")
  cobalt::bal.plot(
    x = f,
    data = dats_trans,
    var.name = v_use,
    which = "both"
  ) + ggtitle(label)
})

# Arrange in grid and save
# Create a title grob
title_grob <- textGrob("Covariate Balance by CBI Treatment Group",
                       gp = gpar(fontsize = 20, fontface = "bold"))

# Add title and plots to a single grid
grid_plot <- grid.arrange(
  title_grob,
  do.call(arrangeGrob, c(plot_list, ncol = 2)),
  nrow = 2,
  heights = c(0.05, 0.95)
)
# Save
ggsave(here(dir_figs, "covariate_balance_grid_cbi.png"), grid_plot, units = "px", width = 3000, height = 6000, dpi = 300)





```


```{r}


cf_analysis <- function(dats, yvar, xvars, trt, data_nm, seed = 1234) {
  # Filter out rows with NA in the outcome variable
  dats_filt <- dats %>%
    filter(!is.na(.data[[yvar]]))

  # Select covariates and outcome
  covariates <- dats_filt %>%
    select(all_of(xvars))
  outcome <- dats_filt[[yvar]]
  treatment <- dats_filt[[trt]]

  # Fit causal forest
  fit <- grf::causal_forest(
    X = covariates,
    Y = outcome,
    W = treatment,
    seed = seed
  )

  # Variable importance - ensure normalized
  imp_raw <- sort(setNames(variable_importance(fit), xvars))
  imp_normalized <- imp_raw / sum(imp_raw)
  
  imp_df <- tibble(
    variable = names(imp_normalized),
    importance = as.numeric(imp_normalized)
  ) %>%
    arrange(importance)  # So the smallest is at the bottom of the horizontal bar plot

  imp_p <- ggplot(imp_df, aes(x = importance, y = reorder(variable, importance))) +
    geom_col(fill = "orange") +
    labs(
      x = "Variable Importance (Normalized)",
      y = NULL,
      title = "Causal Forest Variable Importance",
      subtitle = data_nm
    ) +
    theme_minimal(base_size = 13) +
    theme(
      axis.text.y = element_text(hjust = 1),
      plot.margin = margin(10, 20, 5, 5)
    )
  ggsave(plot = imp_p,
         filename = here(dir_figs, paste0("var_importance_", data_nm, ".png")),
         bg = "white")

  pred_fun <- function(object, newdata, ...) {
    predict(object, newdata, ...)$predictions
  }

  xvars_sorted <- rev(names(imp_normalized))

  pdps <- lapply(xvars_sorted, function(v) {
    pd <- partial_dep(fit, v, X = covariates, pred_fun = pred_fun)
    plot(pd) +
      theme(axis.title.y = element_blank())
  })

  #For adding y axis label to overall plot
  ylab_plot <- ggplot() +
    theme_void() +
    annotate("text", x = 0, y = 0.5, label = "Treatment effect", angle = 90, hjust = 0.5, size = 4)

  pdps_p <- wrap_plots(pdps, guides = "collect", ncol = 3) #&
    #ylab("Treatment effect")

  # Add format & add titles
  pdps_p <- ylab_plot + pdps_p +
    plot_layout(widths = c(0.05, 1)) +
    plot_annotation(
      title = "Partial Dependence Plots",
      subtitle = data_nm,
      caption = "NOTES:\nNegative values mean a stronger (positive) treatment effect (i.e. fire is causing a greater difference in forest cover change)\nPDPs are in order of variable importance")

  ggsave(filename = here(dir_figs, paste0("partial_dependence_", data_nm, ".png")),
         plot = pdps_p,
         units = 'px',
         width = 2000,
         height = 2000)

  return(list(fit = fit,
              imp_p = imp_p,
              pdps = pdps))
}

tic() #takes about 1.5 minutes to run
test <- cf_analysis(dats = dats,
            yvar = "post_fire_forest_difference_10_yr",
            xvars = c("aet", "def", "frg_reclass", "srtm", "tpi", "chili",
           "burn_cbi", "burn_cbi_1kmmwmedian", "burn_cbi_1kmmwstd", "mock_burn_year",
           "biotic_5_yrs_prior_sum", "biotic_5_yrs_after_sum",
           "pdsi_annual_5_yrs_prior_sum", "pdsi_annual_5_yrs_after_sum", "avg_forest_2_yr_pre_fire"),
            trt = "fire",
            data_nm = "idBatholith_post_fire_forest_difference_10_yr",
            seed = seed)
toc()



tic() #takes about 1.5 minutes to run
test <- cf_analysis(dats = dats,
            yvar = "post_fire_forest_slope_5_10_yrs",
            xvars = c("aet", "def", "frg_reclass", "srtm", "tpi", "chili",
           "burn_cbi", "burn_cbi_1kmmwmedian", "burn_cbi_1kmmwstd", "mock_burn_year",
           "biotic_5_yrs_prior_sum", "biotic_5_yrs_after_sum",
           "pdsi_annual_5_yrs_prior_sum", "pdsi_annual_5_yrs_after_sum",
           "avg_forest_2_yr_pre_fire"),
            trt = "fire",
            data_nm = "idBatholith_post_fire_forest_slope_5_10_yrs",
            seed = seed)
toc()

tic() 
test <- cf_analysis(dats = dats,
            yvar = "post_fire_forest_difference_10_yr",
            xvars = c("aet", "def", "frg_reclass", "srtm", "tpi", "chili",
           "burn_cbi_1kmmwmedian", "burn_cbi_1kmmwstd", "mock_burn_year",
           "biotic_5_yrs_prior_sum", "biotic_5_yrs_after_sum",
           "pdsi_annual_5_yrs_prior_sum", "pdsi_annual_5_yrs_after_sum",
           "avg_forest_2_yr_pre_fire"),
            trt = "burn_cbi",
            data_nm = "CBITRT_idBatholith_post_fire_forest_difference_10_yr",
            seed = seed)
toc()




# yvar = "post_fire_forest_difference_10_yr"
# xvars = c("aet", "def", "frg_reclass", "srtm", "tpi", "chili",
#            "burn_cbi", "burn_cbi_1kmmwmedian", "burn_cbi_1kmmwstd", "mock_burn_year",
#            "biotic_5_yrs_prior_sum", "biotic_5_yrs_after_sum",
#            "pdsi_annual_5_yrs_prior_sum", "pdsi_annual_5_yrs_after_sum")
# trt = "fire"
#   # Filter out rows with NA in the outcome variable
#   dats_filt <- dats %>%
#     filter(!is.na(.data[[yvar]]))
# 
#   # Select covariates and outcome
#   covariates <- dats_filt %>%
#     select(all_of(xvars))
#   outcome <- dats_filt[[yvar]]
#   treatment <- dats_filt[[trt]]
# 
#   # Fit causal forest
#   fit <- grf::causal_forest(
#     X = covariates,
#     Y = outcome,
#     W = treatment,
#     seed = seed
#   )
# 
#   # Variable importance
#   imp_raw <- sort(setNames(variable_importance(fit), xvars))
#   imp_normalized <- imp / sum(imp)
# 
#   imp_df <- tibble(
#     variable = names(imp_normalized),
#     importance = as.numeric(imp_normalized)
#   ) %>%
#     arrange(importance)  # So the smallest is at the bottom of the horizontal bar plot
# 
#   imp_p <- ggplot(imp_df, aes(x = importance, y = reorder(variable, importance))) +
#     geom_col(fill = "orange") +
#     labs(
#       x = "Variable Importance",
#       y = NULL,
#       title = "Causal Forest Variable Importance",
#       subtitle = "test"
#     ) +
#     theme_minimal(base_size = 13) +
#     theme(
#       axis.text.y = element_text(hjust = 1),
#       plot.margin = margin(10, 20, 5, 5)
#     )
#   imp_p
#   ggsave(plot = imp_p,
#          filename = here(dir_figs, paste0("var_importance_", data_nm, ".png")),
#          bg = "white")
# 
#   pred_fun <- function(object, newdata, ...) {
#     predict(object, newdata, ...)$predictions
#   }
# 
#   xvars_sorted <- rev(names(imp))
# 
#   pdps <- lapply(xvars_sorted, function(v) {
#     pd <- partial_dep(fit, v, X = covariates, pred_fun = pred_fun)
#     plot(pd) +
#       theme(axis.title.y = element_blank())
#   })
# 
#   #For adding y axis label to overall plot
#   ylab_plot <- ggplot() +
#     theme_void() +
#     annotate("text", x = 0, y = 0.5, label = "Treatment effect", angle = 90, hjust = 0.5, size = 4)
# 
#   pdps_p <- wrap_plots(pdps, guides = "collect", ncol = 3) #&
#     #ylab("Treatment effect")
# 
#   # Add format & add titles
#   pdps_p <- ylab_plot + pdps_p +
#     plot_layout(widths = c(0.05, 1)) +
#     plot_annotation(
#       title = "Partial Dependence Plots",
#       subtitle = data_nm,
#       caption = "NOTES:\nNegative values mean a stronger (positive) treatment effect (i.e. fire is causing a greater difference in forest cover change)\nPDPs are in order of variable importance")
# 
#   ggsave(filename = here(dir_figs, paste0("partial_dependence_", data_nm, ".png")),
#          plot = pdps_p,
#          units = 'px',
#          width = 2000,
#          height = 2000)
# 

```


These initial results mean that: (YES = makes sense, NO = doesn't make sense)
- As AET increases (there is more AET), there is greater difference between pre-and post-fire cover
- As DEF increases (sites get drier), there is decreased difference between pre-and post-fire cover
- YES - When there is biotic disturbance prior to burn there is a greater difference between pre and post-fire cover
- ??? - When there is more drought after fire there is greater difference between pre and post-fire cover; but then also a much larger difference between pre and post-fire cover with wetter conditions after fire?..
- YES - The difference between pre and post-fire is least when elevation is ~2000, and greater at low elevatoins or high elevations (at the edges of forest range forest causes more change in forest cover?)
- YES - The difference between pre and post fire forest cover is greatest when CHILI is high (very warm)
- The treatment effect is greatest for FRG classes 3,4,5,6
- YES? - The treatment effect is greatest for negative values of TPI (i.e. in valleys); there are denser forests pre-burn in valleys
- NO - The treatment effect is greatest earlier in the time frame


```{r}
hist(dats$pdsi_annual_5_yrs_after_sum)

```



<button class="accordion-button" type="button" data-bs-toggle="collapse" data-bs-target="#collapseOne" >Click to see Session Information</button><div id="collapseOne" class="accordion-collapse collapse"><div>

```{r sessionInfo}
sessionInfo()
```