---
title: "04_causal_forest"
author: Tyler L. McIntosh
date: last-modified
format:
  html:
    embed-resources: true
    html-math-method: katex
    css: styles.css
    toc: true
    toc-depth: 1
    toc-title: Contents
    toc-location: right
    code-fold: true
    code-summary: "Show the code"
---

# Setup for analysis

``` {r setupchunk}
#| echo: true
#| message: false
#| warning: false
#| results: 'hide'

rm(list = ls())

if(!requireNamespace("here", quietly = TRUE)) {
  install.packages("here")
}
library(here)

source(here::here("utils", "functions.R"))

install_load_packages(c(
  "tidyverse",
  "tictoc",
  "grf",
  "kernelshap",
  "hstats",
  "glue",
  "shapviz",
  "patchwork",
  "cobalt",
  "purrr",
  "corrplot",
  "gridExtra",
  "rlang",
  "arrow",
  "tictoc"))

dir_figs <- here::here('figs', 'exploratory')
dir_derived <- here::here('data', 'derived')
dir_raw <- here::here('data', 'raw')
dir_manual <- here::here('data', 'manual')
dir_ensure(c(dir_figs,
             dir_derived,
             dir_manual,
             dir_raw))

seed = 1234
set.seed(seed)


```


# Load dats

```{r}

#Short dats
dats <- arrow::read_parquet(here(dir_derived, "test_dats_short.parquet"))
ecoregions <- unique(dats$ecoregion)

dats_long <- arrow::read_parquet(here(dir_derived, "test_dats_long.parquet"))


x <- dats_long |>
  select(starts_with("rap_tree"), burn_year, year, starts_with("avg_rap_tree"), starts_with("post_fire_rap_tree_perc"),  
         uniqueID)

xx <- filter(x, uniqueID == 25220) |> arrange(year)

```

# Examine MODIS vs RAP outcomes


```{r}
dats_f <- dats |>
  filter(fire == 1)

dats_no_f <- dats |>
  filter(fire != 1)

# 6-year pre-fire comparison
ggplot(
  dats_f |>
    filter(
      !is.na(avg_rap_tree_6_yr_pre_fire),
      !is.na(avg_vcf_tree_6_yr_pre_fire)
    )
) +
  geom_density(aes(x = avg_rap_tree_6_yr_pre_fire), color = "red") +
  geom_density(aes(x = avg_vcf_tree_6_yr_pre_fire), color = "blue")


# 3-year post-fire comparison
ggplot(
  dats_f |>
    filter(
      !is.na(avg_rap_tree_3_yr_post_fire),
      !is.na(avg_vcf_tree_3_yr_post_fire)
    )
) +
  geom_density(aes(x = avg_rap_tree_3_yr_post_fire), color = "red") +
  geom_density(aes(x = avg_vcf_tree_3_yr_post_fire), color = "blue")


# 10-year post-fire comparison
ggplot(
  dats_f |>
    filter(
      !is.na(avg_rap_tree_10_yr_post_fire),
      !is.na(avg_vcf_tree_10_yr_post_fire)
    )
) +
  geom_density(aes(x = avg_rap_tree_10_yr_post_fire), color = "red") +
  geom_density(aes(x = avg_vcf_tree_10_yr_post_fire), color = "blue")


# Percent recovery, 3 to 15 years post-fire
ggplot(
  dats_f |>
    filter(
      !is.na(post_fire_rap_tree_perc_recovery_n6_3_15_yr),
      !is.na(post_fire_vcf_tree_perc_recovery_n6_3_15_yr)
    )
) +
  geom_density(aes(x = post_fire_rap_tree_perc_recovery_n6_3_15_yr), color = "red") +
  geom_density(aes(x = post_fire_vcf_tree_perc_recovery_n6_3_15_yr), color = "blue") +
  xlim(c(-50, 50))


# Percent recovery, 3 to 10 years post-fire
ggplot(
  dats_f |>
    filter(
      !is.na(post_fire_rap_tree_perc_recovery_n6_3_10_yr),
      !is.na(post_fire_vcf_tree_perc_recovery_n6_3_10_yr)
    )
) +
  geom_density(aes(x = post_fire_rap_tree_perc_recovery_n6_3_10_yr), color = "red") +
  geom_density(aes(x = post_fire_vcf_tree_perc_recovery_n6_3_10_yr), color = "blue") +
  xlim(c(-50, 50))


ggplot(
  dats_f |>
    filter(
      !is.na(post_fire_rap_tree_slope_3_15_yrs),
      !is.na(post_fire_vcf_tree_slope_3_15_yrs)
    )
) +
  geom_density(aes(x = post_fire_rap_tree_slope_3_15_yrs), color = "red") +
  geom_density(aes(x = post_fire_vcf_tree_slope_3_15_yrs), color = "blue") +
  xlim(c(-2, 2))


ggplot(
  dats_f |>
    filter(
      !is.na(post_fire_rap_tree_slope_3_10_yrs),
      !is.na(post_fire_vcf_tree_slope_3_10_yrs)
    )
) +
  geom_density(aes(x = post_fire_rap_tree_slope_3_10_yrs), color = "red") +
  geom_density(aes(x = post_fire_vcf_tree_slope_3_10_yrs), color = "blue") +
  xlim(c(-2, 2))

# Check correlation between vcf and rap
# ... not very correlated...
dats_f_long <- dats_long |>
  filter(fire == 1)

ggplot(dats_f_long |>
         filter(fire_timeframe == "before_burn")) +
  geom_point(aes(x = rap_tree, y = vcf_tree)) +
  labs(title = "VCF vs RAP mean 250m, pre-burn")

ggplot(dats_f_long |>
         filter(fire_timeframe == "after_burn")) +
  geom_point(aes(x = rap_tree, y = vcf_tree)) +
  geom_smooth(aes(x = rap_tree, y = vcf_tree)) +
  labs(title = "VCF vs RAP mean 250m, post-burn")



ggplot(dats |> filter(post_fire_rap_tree_perc_recovery_n6_3_15_yr > -100 & post_fire_rap_tree_perc_recovery_n6_3_15_yr < 100 & post_fire_vcf_tree_perc_recovery_n6_3_15_yr > -100 & post_fire_vcf_tree_perc_recovery_n6_3_15_yr < 100 )) +
  geom_point(aes(x = post_fire_rap_tree_perc_recovery_n6_3_15_yr, y = post_fire_vcf_tree_perc_recovery_n6_3_15_yr), alpha = 0.1) +
  geom_smooth(aes(x = post_fire_rap_tree_perc_recovery_n6_3_15_yr, y = post_fire_vcf_tree_perc_recovery_n6_3_15_yr), alpha = 0.1) +
  labs(title = "VCF vs RAP recovery 250m")


ggplot(dats) +
  geom_point(aes(x = post_fire_rap_tree_slope_3_15_yrs, y = post_fire_vcf_tree_slope_3_15_yrs), alpha = 0.1) +
  geom_smooth(aes(x = post_fire_rap_tree_slope_3_15_yrs, y = post_fire_vcf_tree_slope_3_15_yrs), alpha = 0.1) +
  labs(title = "VCF vs RAP slope 250m")

```


# Load and explore long data

We see that tree cover shifts lower more dramatically under high severity fire, which is what we would expect. However, these trends vary by ecoregion. Tree cover in the Southern Rockies shifts much more dramatically towards zero--since this is long data, we can theorize that this is potentially due to lack of regeneration (let's test this). We also see that the shift towards lower tree cover is more severe even under low/moderate severity fire in the Southern Rockies than it is in other ecoregions.

When we check the before/after densities of tree cover for our mock fires (i.e. the rest of the forested area used as control), we see that the after_burn values are higher than pre-burn, with the greatest change in the North Cascades. This is what we would expect as a result of growth rates in areas that have not been impacted by fires.

From the trajectories (both ribbon and boxplot) we see that there are varied recovery trajectories by ecoregion. Most ecoregions and fire severity groups have some element of delayed mortality. Most of these appear to last two years, although there are longer delayed mortality patterns in the North Cascades under low/moderate severity fire (~5 years) and under low/moderate severity fire in the Southern Rockies (~10 years!).

Interestingly, across all ecoregions, even in the 'unburned' areas (i.e. CBI < 0.1) we see a trajectory of long-term forest cover loss following fire occurrence, suggesting that the forest pattern changes from fire may have resulted in lowered resilience to other disturbances, or other fallout effects from nearby higher severity patches.

We find that the parallel trajectories hypothesis more or less holds for all ecoregions and fire severities with the exception of unburned (actual) ares in the Northern Cascades.

In the North Cascades and Sierra Nevada we see what looks like a signature of forest decline prior to low, moderate, and high severity burns.

*FINDINGS OF INTEREST:*
- Lack of recovery under high severity fire in S.Rockies, also under ANY fire in S.Nevada!
- Parallel assumption (more or less) holds
- Delayed mortality primarily in low/moderate severity areas (which is what we would expect)
  - QUESTION: Is this delayed mortality linked to post-fire drought?
- There is a notable signature of forest loss even in unburned areas within fire perimeters
  - QUESTION: Is this linked to other disturbances? Or to fire heterogeneity patterning?
- We see forest decline in the years just prior to burn in the North Cascades and Sierra Nevada
  - QUESTION: Is this linked to other disturbances? Does probability of burn / burn severity change depending on prior drought/biotic agents?
  
  COMPARISON
  Observations comparing the two response variables
  - In general we see that VCF tends to document a lower tree cover percentage for the same pixel. this is consistently true across ecoregions, and relatively stable in never-burned units. We see a similar pattern in the negligibly burned areas.
  - The trends are closer together for the low/moderate and highly burned areas, both pre- and post-burn (WHY?), while VCF tends to document reduced recovery in later years 
  
  - In BOTH trajectories we see patterns that make sense in terms of forest density -> severity, but they vary a bit by ecoregion.
    - IDB: low/moderate fires are happening in lower-than-average density forests, while high severity fires are happening in ~ average density forests
    - SR: low/moderate fires are happening in average density forests, while high severity fires are happening in higher-than-average density forests
  - QUESTION: is this a difference in something ecological, or sampling due to the geography of the region (e.g. more high elevation forests in SR?)
  
  - Both show a pattern of delayed mortality in the southern rockies under low/moderate burn conditions

```{r}


make_tree_iqr_combined <- function(dats_long, dir_figs, suffix) {
  summary_data <- dats_long |>
    group_by(ecoregion, burn_sev, year_from_fire_index) |>
    summarise(
      q25_rap    = quantile(rap_tree, 0.25, na.rm = TRUE),
      median_rap = median(rap_tree, na.rm = TRUE),
      q75_rap    = quantile(rap_tree, 0.75, na.rm = TRUE),
      q25_vcf    = quantile(vcf_tree, 0.25, na.rm = TRUE),
      median_vcf = median(vcf_tree, na.rm = TRUE),
      q75_vcf    = quantile(vcf_tree, 0.75, na.rm = TRUE),
      .groups = "drop"
    )

  never_burned <- summary_data |>
    filter(burn_sev == "Never") |>
    select(ecoregion, year_from_fire_index, median_rap, median_vcf)

  p_iqr_combined <- ggplot(summary_data, aes(x = year_from_fire_index)) +
    # RAP IQR ribbon
    geom_ribbon(
      aes(ymin = q25_rap, ymax = q75_rap),
      fill = "red",
      alpha = 0.15
    ) +
    # MODIS IQR ribbon
    geom_ribbon(
      aes(ymin = q25_vcf, ymax = q75_vcf),
      fill = "blue",
      alpha = 0.15
    ) +
    # RAP median
    geom_line(aes(y = median_rap), color = "red", linewidth = 1) +
    # MODIS median
    geom_line(aes(y = median_vcf), color = "blue", linewidth = 1) +
    # Never baseline for RAP (lighter dashed red)
    geom_line(
      data = never_burned,
      aes(y = median_rap),
      color = "coral1",
      linetype = "dashed",
      linewidth = 0.9,
      inherit.aes = TRUE
    ) +
    # Never baseline for MODIS (lighter dashed blue)
    geom_line(
      data = never_burned,
      aes(y = median_vcf),
      color = "lightskyblue",
      linetype = "dashed",
      linewidth = 0.9,
      inherit.aes = TRUE
    ) +
    # Fire time = 0 line
    geom_vline(
      xintercept = 0,
      linetype = "dotted",
      color = "black",
      linewidth = 0.8
    ) +
    facet_grid(ecoregion ~ burn_sev) +
    labs(
      title = "Tree cover by years from fire (RAP vs MODIS, IQR shaded)",
      x = "Years from fire",
      y = "Tree cover %",
      caption = "RAP: red; MODIS VCF: blue. 'Never' represents control samples burned in random years with CBI = 0."
    ) +
    theme_minimal()

  ggsave(
    filename = here(dir_figs, glue("iqr_line_time_series_combined_rap_vcf_{suffix}.png")),
    plot = p_iqr_combined,
    bg = "white",
    units = "px",
    width = 5000,
    height = 3000
  )

  invisible(p_iqr_combined)
}

make_tree_iqr_combined(dats_long |> filter(!is.na(vcf_tree)), dir_figs = dir_figs, suffix = "matchedOnly")


```



# Assess recovery metrics that were calculated
Observations:
- Some individual trajectories aren't very clean, but others are
- in general the RAP data is important for getting a long enough timeframe to assess percent recovery 
tested:
x <- dats |> filter(!is.na(post_fire_vcf_tree_perc_recovery_n6_3_10_yr))
y <- dats |> filter(!is.na(post_fire_rap_tree_perc_recovery_n6_3_10_yr))
RAP provides a sample size of more than 2X the number of points -> USE RAP FOR THE RANDOM FOREST ANALYSIS

```{r}

dats_fh_long <- dats_long |>
  filter(fire == 1 & burn_sev == "High")

dats_fh <- dats |>
  filter(fire == 1 & burn_sev == "High")




plot_dual_tree_trajectories <- function(dats,
                                        dats_long,
                                        n = 10,
                                        dir_figs = NULL) {
  # high-severity fire pixels only
  dats_fh_long <- dats_long |>
    filter(fire == 1, burn_sev == "High", !is.na(post_fire_vcf_tree_perc_recovery_n6_3_15_yr))

  ids <- unique(dats_fh_long$uniqueID)
  n <- min(n, length(ids))

  plots <- vector("list", n)

  for (i in seq_len(n)) {
    this_id <- ids[i]

    d_s <- dats |>
      filter(uniqueID == this_id)

    # Averages for RAP
    rap_6pre   <- d_s$avg_rap_tree_6_yr_pre_fire[1]
    rap_3post  <- d_s$avg_rap_tree_3_yr_post_fire[1]
    rap_10post <- d_s$avg_rap_tree_10_yr_post_fire[1]
    rap_15post <- d_s$avg_rap_tree_15_yr_post_fire[1]

    # Averages for VCF
    vcf_6pre   <- d_s$avg_vcf_tree_6_yr_pre_fire[1]
    vcf_3post  <- d_s$avg_vcf_tree_3_yr_post_fire[1]
    vcf_10post <- d_s$avg_vcf_tree_10_yr_post_fire[1]
    vcf_15post <- d_s$avg_vcf_tree_15_yr_post_fire[1]

    # Slopes for caption (RAP)
    rap_s_3_10 <- d_s$post_fire_rap_tree_slope_3_10_yrs[1]
    rap_s_3_15 <- d_s$post_fire_rap_tree_slope_3_15_yrs[1]

    # Slopes for caption (VCF)
    vcf_s_3_10 <- d_s$post_fire_vcf_tree_slope_3_10_yrs[1]
    vcf_s_3_15 <- d_s$post_fire_vcf_tree_slope_3_15_yrs[1]
    
    # Slopes for caption (RAP)
    rap_p_3_10 <- d_s$post_fire_rap_tree_perc_recovery_n6_3_10_yr[1]
    rap_p_3_15 <- d_s$post_fire_rap_tree_perc_recovery_n6_3_15_yr[1]

    # Slopes for caption (VCF)
    vcf_p_3_10 <- d_s$post_fire_vcf_tree_perc_recovery_n6_3_10_yr[1]
    vcf_p_3_15 <- d_s$post_fire_vcf_tree_perc_recovery_n6_3_15_yr[1]

    d <- dats_fh_long |>
      filter(uniqueID == this_id)

    x <- d |>
      select(year, year_from_fire_index, rap_tree, vcf_tree, biotic) |>
      arrange(year)

    caption_text <- glue(
      "Dotted = RAP, DotDash = VCF\n",
      "RAP slopes: 3–10 = {round(rap_s_3_10, 2)}, 3–15 = {round(rap_s_3_15, 2)}\n",
      "VCF slopes: 3–10 = {round(vcf_s_3_10, 2)}, 3–15 = {round(vcf_s_3_15, 2)}\n",
      "RAP perc recovery: 3–10 = {round(rap_p_3_10, 2)}, 3–15 = {round(rap_p_3_15, 2)}\n",
      "VCF perc recovery: 3–10 = {round(vcf_p_3_10, 2)}, 3–15 = {round(vcf_p_3_15, 2)}"
    )

    p <- ggplot(x, aes(x = year_from_fire_index)) +
      # Trajectories
      geom_line(aes(y = rap_tree), color = "grey50", linetype = "dotted") +
      geom_line(aes(y = vcf_tree), color = "black", linetype = "dotdash") +

      # RAP context lines (grey, dotted)
      geom_vline(xintercept = -6,  color = "yellow", linetype = "dotted") +
      geom_vline(xintercept = 3,   color = "green", linetype = "dotted") +
      geom_vline(xintercept = 10,  color = "purple", linetype = "dotted") +
      geom_vline(xintercept = 15,  color = "blue", linetype = "dotted") +
      geom_hline(yintercept = rap_6pre,   color = "yellow", linetype = "dotted") +
      geom_hline(yintercept = rap_3post,  color = "green", linetype = "dotted") +
      geom_hline(yintercept = rap_10post, color = "purple", linetype = "dotted") +
      geom_hline(yintercept = rap_15post, color = "blue", linetype = "dotted") +

      # VCF context lines (black, dotdash)
      geom_hline(yintercept = vcf_6pre,   color = "yellow", linetype = "dotdash") +
      geom_hline(yintercept = vcf_3post,  color = "green", linetype = "dotdash") +
      geom_hline(yintercept = vcf_10post, color = "purple", linetype = "dotdash") +
      geom_hline(yintercept = vcf_15post, color = "blue", linetype = "dotdash") +
      
      geom_vline(xintercept = 0, color = "red") +

      labs(
        title = glue("RAP vs MODIS VCF tree cover trajectories (ID {this_id})"),
        x = "Years from fire",
        y = "Tree cover %",
        caption = caption_text
      ) +
      theme_minimal()

    print(p)
    plots[[i]] <- p

    if (!is.null(dir_figs)) {
      ggsave(
        filename = here(dir_figs, glue("trajectory_rap_vcf_id_{this_id}.png")),
        plot = p,
        bg = "white",
        width = 6,
        height = 4,
        units = "in",
        dpi = 300
      )
    }
  }

  invisible(plots)
}


plot_dual_tree_trajectories(
  dats = dats,
  dats_long = dats_long,
  n = 25,
  dir_figs = dir_figs
)



```






######FLAG: NOT UPDATED FOR NEW DATA YET
# CORRELATION & BALANCE
- Correlation matrix
- Covariate balance: love & density plots

*FINDINGS:*
- High correlation among post-event tree cover (expected)
- Burn severity and heterogeneity are most highly correlated with post-fire cover in the S.Rockies. Minimal correlation in S.Nevada & ID.Batholith, only with immediate post-fire in N.Cascades
- There's some correlation between long-term average AET/DEF and the HD/PDSI, but not as much as there might be
- The S.Rockies and S.Nevada have the best covariate balance out of the box, while the N.Cascades has extremely poor covariate balance, and IDBatholith isn't great. Both IDB & NC have imbalances in the biotic disturbance before and after fire, which is interesting
- We probably want to conduct pre-DiD matching



```{r}




# Function to assess correlation with corrplot and covariate balance with cobalt, output results
assess_covariate_balance <- function(d, nm) {
  
  # Correlation matrix
  correlations <- d |>
    select(where(is.numeric)) |>
    cor(use = "pairwise.complete.obs")
  png(here(dir_figs, glue('corr_plot_{nm}.png')), width = 2000, height = 2000)
  corrplot(correlations, method = 'circle', type = "lower", diag = FALSE, use = 'pairwise.complete.obs') 
  dev.off()
  
      
    
  cobalt::bal.tab(x = fire ~ aet + def + frg_reclass + srtm + tpi + chili + biotic_5_yrs_prior_sum + biotic_5_yrs_after_sum + pdsi_annual_5_yrs_prior_sum + pdsi_annual_5_yrs_after_sum + avg_forest_2_yr_pre_fire, 
                  data = d)
  
  png(here(dir_figs, glue("love_plot_fire_{nm}.png")))
  cobalt::love.plot(x = fire ~ aet + def + frg_reclass + srtm + tpi + chili + biotic_5_yrs_prior_sum + biotic_5_yrs_after_sum + pdsi_annual_5_yrs_prior_sum + pdsi_annual_5_yrs_after_sum + avg_forest_2_yr_pre_fire, 
                  data = d,
                  stats = c("mean.diffs", "ks.statistics"),
                  stars = "raw",
                  position = "bottom",
                  threshold = c(m = .1, ks = .05),
                  limits = list(c(-1, 1), c(0, 0.6)),
                  title = glue("Covariate balance love plot for FIRE treatment: {nm}"))
  dev.off()

  
    
  fire_covars <- c("aet", "def", "frg_reclass", "srtm", "tpi", "chili",
              "biotic_5_yrs_prior_sum", "biotic_5_yrs_after_sum",
              "pdsi_annual_5_yrs_prior_sum", "pdsi_annual_5_yrs_after_sum", "avg_forest_2_yr_pre_fire")
  
  # Make a working copy of data
  dats_trans <- d
  
  # Transform biotic variables using log1p (handles 0s safely)
  for (v in fire_covars) {
    if (startsWith(v, "biotic")) {
      new_name <- paste0(v, "_log")
      dats_trans[[new_name]] <- log1p(dats_trans[[v]])
    }
  }
  
  # Plot all variables, using log-transformed versions when appropriate
  plot_list <- lapply(fire_covars, function(v) {
    # If biotic, use transformed version
    if (startsWith(v, "biotic")) {
      v_use <- paste0(v, "_log")
      label <- paste0(v, " (log)")
    } else {
      v_use <- v
      label <- v
    }
  
    # Construct formula and plot
    f <- reformulate(v_use, response = "fire")
    cobalt::bal.plot(
      x = f,
      data = dats_trans,
      var.name = v_use,
      which = "both"
    ) + ggtitle(label)
  })
  
  # Arrange in grid and save
  # Create a title grob
  title_grob <- textGrob(glue("Covariate Balance by FIRE Treatment Group: {nm}"),
                         gp = gpar(fontsize = 20, fontface = "bold"))
  
  # Add title and plots to a single grid
  grid_plot <- grid.arrange(
    title_grob,
    do.call(arrangeGrob, c(plot_list, ncol = 2)),
    nrow = 2,
    heights = c(0.05, 0.95)
  )
  # Save
  ggsave(here(dir_figs, glue("covariate_balance_grid_fire_{nm}.png")), grid_plot, units = "px", width = 3000, height = 3000, dpi = 300)
  
  
  
  
  
  
  # FOR CBI TREATMENT
  
  cobalt::bal.tab(x = burn_cbi ~ aet + def + frg_reclass + srtm + tpi + chili + biotic_5_yrs_prior_sum + biotic_5_yrs_after_sum + pdsi_annual_5_yrs_prior_sum + pdsi_annual_5_yrs_after_sum + avg_forest_2_yr_pre_fire + burn_cbi_1kmmwmedian + burn_cbi_1kmmwstd + mock_burn_year, 
                  data = d)
  
  
  burn_cbi_covars <- c("aet", "def", "frg_reclass", "srtm", "tpi", "chili",
              "biotic_5_yrs_prior_sum", "biotic_5_yrs_after_sum",
              "pdsi_annual_5_yrs_prior_sum", "pdsi_annual_5_yrs_after_sum", "avg_forest_6_yr_pre_fire", "burn_cbi_1kmmwmedian", "burn_cbi_1kmmwstd", "mock_burn_year")
  
  # Make a working copy of data
  dats_trans <- d
  
  # Transform biotic variables using log1p (handles 0s safely)
  for (v in burn_cbi_covars) {
    if (startsWith(v, "biotic")) {
      new_name <- paste0(v, "_log")
      dats_trans[[new_name]] <- log1p(dats_trans[[v]])
    }
  }
  
  # Plot all variables, using log-transformed versions when appropriate
  plot_list <- lapply(burn_cbi_covars, function(v) {
    # If biotic, use transformed version
    if (startsWith(v, "biotic")) {
      v_use <- paste0(v, "_log")
      label <- paste0(v, " (log)")
    } else {
      v_use <- v
      label <- v
    }
  
    # Construct formula and plot
    f <- reformulate(v_use, response = "burn_cbi")
    cobalt::bal.plot(
      x = f,
      data = dats_trans,
      var.name = v_use,
      which = "both"
    ) + ggtitle(label)
  })
  
  # Arrange in grid and save
  # Create a title grob
  title_grob <- textGrob("Covariate Balance by CBI Treatment Group",
                         gp = gpar(fontsize = 20, fontface = "bold"))
  
  # Add title and plots to a single grid
  grid_plot <- grid.arrange(
    title_grob,
    do.call(arrangeGrob, c(plot_list, ncol = 2)),
    nrow = 2,
    heights = c(0.05, 0.95)
  )
  # Save
  ggsave(here(dir_figs, glue("covariate_balance_grid_cbi{nm}.png")), grid_plot, units = "px", width = 3000, height = 6000, dpi = 300)
  
}


# USE FUNCTION FOR ECOREGIONS AND FULL DATSET
for(eco in ecoregions) {
  d <- dats |> filter(ecoregion == eco)
  assess_covariate_balance(d = d,
                           nm = eco)
}
assess_covariate_balance(d = dats,
                         nm = "ALL")







```



<button class="accordion-button" type="button" data-bs-toggle="collapse" data-bs-target="#collapseOne" >Click to see Session Information</button><div id="collapseOne" class="accordion-collapse collapse"><div>

```{r sessionInfo}
sessionInfo()
```

