---
title: "04_data_exploration_2"
format: html
editor: visual
---

# Setup for analysis

```{r setupchunk}
#| echo: true
#| message: false
#| warning: false
#| results: 'hide'

rm(list = ls())

if(!requireNamespace("here", quietly = TRUE)) {
  install.packages("here")
}
library(here)

source(here::here("utils", "functions.R"))

install_load_packages(c(
  "tidyverse",
  "arrow",
  "sf",
  "scales",
  "rlang",
  "grid",
  "gridExtra",
  "ggpmisc",
  "h3jsr",
  "farver",
  "glue",
  "corrplot"))



gdrive_folder <- "GEE_resilience_v6_operational_ss500_ts50000"
version <- "v7"

dir_figs <- here::here('figs', gdrive_folder, version, "data_explore")
dir_derived <- here::here('data', 'derived', gdrive_folder)
dir_raw <- here::here('data', 'raw')
dir_manual <- here::here('data', 'manual')
dir_ensure(c(dir_figs,
             dir_derived,
             dir_manual,
             dir_raw))

seed = 1234
set.seed(seed)


```

# Data

```{r}

d_s <- open_dataset(here(dir_derived, "parquet_short"), format = "parquet") |>
  collect()


# d_l <- open_dataset(here(dir_derived, "parquet_long"), format = "parquet") |>
#   collect()
# 
# d_l_fire <- d_l |> filter(fire == 1)


l3pf <- perc_forest <- read_csv(here('data/derived', "l3_perc_usfs_forest.csv"))
l3_sf <- access_data_epa_l3_ecoregions_vsi()

d_s_fire <- d_s |> filter(fire == 1)
d_s_nfire <- d_s |> filter(fire == 0)

d_s_fire |>
  group_by(nfg_factor_clean) |>
  count() |>
  left_join(d_s_nfire |>
    group_by(nfg_factor_clean) |>
    count() |> rename(n2 = n)) |>
  mutate(total = n + n2) |>
  arrange(desc(total))

```

```{r}

# Check subset NFG split
x <- d_s |>
  dplyr::count(analysis_subset, nfg_factor_clean, name = "n") |>
  tidyr::complete(
    analysis_subset,
    nfg_factor_clean,
    fill = list(n = 0)
  ) |>
  dplyr::group_by(analysis_subset) |>
  dplyr::mutate(
    subset_total = sum(n),
    prop = n / subset_total
  ) |>
  dplyr::ungroup()


ggplot(x, aes(x = factor(analysis_subset), y = prop, fill = nfg_factor_clean)) +
  geom_col(position = "stack") +
  labs(
    x = "Analysis subset",
    y = "Proportion within subset",
    fill = "NFG",
    title = "Distribution of NFG groups across analysis subsets"
  ) +
  theme_minimal()

ggsave(here(dir_figs, "nfg_analysis_subset_split.png"))

```

# Examine density - DEPRECATED: see next chunk for functioned

```{r}
# 
# d_s <- d_s |>
#   mutate(biotic_relaxedforestnorm_sum_n5_to_p0_densityplot = ifelse(biotic_relaxedforestnorm_sum_n5_to_p0 < 0.1, NA, biotic_relaxedforestnorm_sum_n5_to_p0)) |>
#   mutate("bnonzero" = ifelse(biotic_relaxedforestnorm_sum_n5_to_p0 < 0.1, "< 0.1", ">= 0.1"),
#          "dbelowzero" = ifelse(pdsi_annual_sum_n5_to_p0 < 0, "< 0", ">= 0"))
# 
# 

# 
# 
# calc_deciles <- function(df, var) {
#   x <- dplyr::pull(df, var)
# 
#   x <- x[!is.na(x)]
# 
#   probs <- seq(0.1, 0.9, by = 0.1)
#   dec <- stats::quantile(x, probs = probs, na.rm = TRUE, names = TRUE)
# 
#   return(dec)
# }
# 
# calc_twentciles <- function(df, var) {
#   x <- dplyr::pull(df, var)
# 
#   x <- x[!is.na(x)]
# 
#   probs <- seq(0.1, 0.9, by = 0.05)
#   dec <- stats::quantile(x, probs = probs, na.rm = TRUE, names = TRUE)
# 
#   return(dec)
# }
# 
# 
# # deciles
# raw_f_deciles_bio <- calc_deciles(d_s_fire, "biotic_relaxedforestnorm_sum_n5_to_p0")
# raw_nf_deciles_bio <- calc_deciles(d_s_nfire, "biotic_relaxedforestnorm_sum_n5_to_p0")
# 
# raw_f_deciles_dr <- calc_deciles(d_s_fire, "pdsi_annual_sum_n5_to_p0")
# raw_nf_deciles_dr <- calc_deciles(d_s_nfire, "pdsi_annual_sum_n5_to_p0")
# 
# 
# raw_f_twentciles_bio <- calc_twentciles(d_s_fire, "biotic_relaxedforestnorm_sum_n5_to_p0")
# raw_nf_twentciles_bio <- calc_twentciles(d_s_nfire, "biotic_relaxedforestnorm_sum_n5_to_p0")
# 
# raw_f_twentciles_dr <- calc_twentciles(d_s_fire, "pdsi_annual_sum_n5_to_p0")
# raw_nf_twentciles_dr <- calc_twentciles(d_s_nfire, "pdsi_annual_sum_n5_to_p0")
# 
# 
# 
# 
# # figs
# bio_frac <- d_s_fire |>
#    group_by(bnonzero) |>
#    count() |>
#   ungroup()
# 
# dr_frac <- d_s_fire |>
#    group_by(dbelowzero) |>
#    count() |>
#   ungroup()
# 
# 
# low_pie_color <- "slategray3"
# high_pie_color <- "slategray4"
# 
# # pie?
# 
# bio_frac_plot <- bio_frac |>
#   dplyr::mutate(
#     bnonzero = factor(
#       bnonzero,
#       levels = c(">= 0.1", "< 0.1")
#     ),
#     prop = n / sum(n),
#     label = scales::percent(prop, accuracy = 1)
#   )
# 
# ggplot(bio_frac_plot, aes(x = "", y = prop, fill = bnonzero)) +
#   geom_col(width = 1, color = "white") +
#   geom_text(
#     aes(label = label),
#     position = position_stack(vjust = 0.5),
#     color = "white",
#     size = 3
#   ) +
#   coord_polar(theta = "y") +
#   scale_fill_manual(
#     values = c(
#       "< 0.1" = low_pie_color,
#       ">= 0.1" = high_pie_color
#     )
#   ) +
#   labs(
#     x = NULL,
#     y = NULL,
#     fill = "bnonzero"
#   ) +
#   theme_void() +
#   theme(
#     legend.title = element_blank()
#   )
# 
# ggsave(here(dir_figs, "biotic_frac_pie.png"),
#        units = "px",
#        width = 500,
#        height = 500)
# 
# 
# 
# 
# dr_frac_plot <- dr_frac |>
#   dplyr::mutate(
#     dbelowzero = factor(
#       dbelowzero,
#       levels = c(">= 0", "< 0")
#     ),
#     prop = n / sum(n),
#     label = scales::percent(prop, accuracy = 1)
#   )
# 
# 
# ggplot(dr_frac_plot, aes(x = "", y = prop, fill = dbelowzero)) +
#   geom_col(width = 1, color = "white") +
#   geom_text(
#     aes(label = label),
#     position = position_stack(vjust = 0.5),
#     color = "white",
#     size = 3
#   ) +
#   coord_polar(theta = "y") +
#   scale_fill_manual(
#     values = c(
#       "< 0" = low_pie_color,
#       ">= 0" = high_pie_color
#     )
#   ) +
#   labs(
#     x = NULL,
#     y = NULL,
#     fill = "dbelowzero"
#   ) +
#   theme_void() +
#   theme(
#     legend.title = element_blank()
#   )
# 
# ggsave(here(dir_figs, "drought_frac_pie.png"),
#        units = "px",
#        width = 500,
#        height = 500)
# 
# 
# 
# 
# 
# major_breaks <- 10^(-1:2)
# 
# minor_breaks <- unlist(lapply(-1:2, function(i) (2:9) * 10^i))
# 
# 
# ggplot() +
#   geom_density(
#     data = d_s_nfire,
#     aes(x = biotic_relaxedforestnorm_sum_n5_to_p0_densityplot),
#     color = "grey70",
#     lwd = 1
#   ) +
#   # geom_vline(xintercept = raw_nf_deciles_bio[9],
#   #            color = "grey70",
#   #            lty = 2,
#   #            lwd = 1) +
#   geom_density(
#     data = d_s_fire,
#     aes(x = biotic_relaxedforestnorm_sum_n5_to_p0_densityplot),
#     color = "maroon",
#     lwd = 1
#   ) +
#   geom_vline(xintercept = raw_f_deciles_bio[9],
#              color = "maroon",
#              lty = 2,
#              lwd = 1) +
#   scale_x_log10(
#     breaks = major_breaks,
#     minor_breaks = minor_breaks,
#     labels = scales::label_number()
#   ) +
#   theme_minimal() +
#   theme(
#     panel.grid.major.x = element_line(color = "grey85"),
#     panel.grid.minor.x = element_line(color = "grey95", linewidth = 0.3),
#     axis.text.y = element_blank(),
#     axis.title.y = element_blank()
#   ) +
#   labs(x = "6-year cumulative biotic disturbance")
# 
# ggsave(here(dir_figs, "biotic_density.png"),
#        units = "px",
#        width = 1000,
#        height = 700)
# 
# 
# 
# 
# ggplot() +
#   geom_density(
#     data = d_s_nfire,
#     aes(x = pdsi_annual_sum_n5_to_p0),
#     color = "grey70",
#     lwd = 1
#   ) +
#   # geom_vline(xintercept = raw_nf_deciles_dr[1],
#   #            color = "grey70",
#   #            lty = 2,
#   #            lwd = 1) +
#   geom_density(
#     data = d_s_fire,
#     aes(x = pdsi_annual_sum_n5_to_p0),
#     color = "maroon",
#     lwd = 1
#   ) +
#   geom_vline(xintercept = raw_f_deciles_dr[1],
#              color = "maroon",
#              lty = 2,
#              lwd = 1) +
#   theme_minimal() +
#   theme(
#     panel.grid.major.x = element_line(color = "grey85"),
#     panel.grid.minor.x = element_line(color = "grey95", linewidth = 0.3),
#     axis.text.y = element_blank(),
#     axis.title.y = element_blank()
#   ) +
#   labs(x = "6-year cumulative drought") +
#   scale_x_reverse()
# 
# ggsave(here(dir_figs, "drought_density.png"),
#        units = "px",
#        width = 1000,
#        height = 700)
# 



```

# examine densities functioned

```{r}

calc_deciles <- function(df, var, reverse = FALSE) {
  x <- dplyr::pull(df, {{ var }})
  x <- x[!is.na(x)]

  probs <- seq(0.1, 0.9, by = 0.1)

  dec <- stats::quantile(
    x,
    probs = probs,
    na.rm = TRUE,
    names = FALSE
  )

  if (reverse) {
    dec <- rev(dec)
  }

  names(dec) <- paste0("perc_", probs * 100)

  dec
}



disturbance_density_description <- function(
    df_background,
    df_interest,
    col_name,
    type,
    dir_figs,
    x_label,
    caption = FALSE,
    vline_decile_idx = 9,
    grouping = ""
) {
  type <- match.arg(type, c("biotic", "drought"))


  # create output subdirectory
  out_dir <- here::here(dir_figs, paste0("density_description_", col_name, "_", grouping))
  if (!dir.exists(out_dir)) dir.create(out_dir, recursive = TRUE)

  # ---- quantiles ----
  if(type == "biotic") {
    interest_deciles      <- calc_deciles(df_interest,    col_name)
    background_deciles    <- calc_deciles(df_background,  col_name)
  } else {
    interest_deciles      <- calc_deciles(df_interest,    col_name, reverse = TRUE)
    background_deciles    <- calc_deciles(df_background,  col_name, reverse = TRUE)
  }

  write.csv(data.frame(quantile = names(interest_deciles),      value = interest_deciles),
            file.path(out_dir, "interest_deciles.csv"),      row.names = FALSE)
  write.csv(data.frame(quantile = names(background_deciles),    value = background_deciles),
            file.path(out_dir, "background_deciles.csv"),    row.names = FALSE)

  low_pie_color  <- "slategray3"
  high_pie_color <- "slategray4"
  vline_x        <- interest_deciles[9]
  
  caption_text <- paste(
    paste(names(interest_deciles), round(interest_deciles, 3), sep = ": "),
    collapse = ", "
  )
  caption_text <- glue::glue("Index 9 used:\n{caption_text}")

  if (type == "biotic") {

    # ---- fraction pie ----
    frac_tbl <- df_interest |>
      dplyr::mutate(frac_group = ifelse(.data[[col_name]] < 0.1, "< 0.1", ">= 0.1")) |>
      dplyr::group_by(frac_group) |>
      dplyr::count() |>
      dplyr::ungroup() |>
      dplyr::mutate(
        frac_group = factor(frac_group, levels = c(">= 0.1", "< 0.1")),
        prop       = n / sum(n),
        label      = scales::percent(prop, accuracy = 1)
      )

    ggplot2::ggplot(frac_tbl, ggplot2::aes(x = "", y = prop, fill = frac_group)) +
      ggplot2::geom_col(width = 1, color = "white") +
      ggplot2::geom_text(ggplot2::aes(label = label),
                         position = ggplot2::position_stack(vjust = 0.5),
                         color = "white", size = 3) +
      ggplot2::coord_polar(theta = "y") +
      ggplot2::scale_fill_manual(values = c("< 0.1" = low_pie_color, ">= 0.1" = high_pie_color)) +
      ggplot2::labs(x = NULL, y = NULL) +
      ggplot2::theme_void() +
      ggplot2::theme(legend.title = ggplot2::element_blank())

    ggplot2::ggsave(file.path(out_dir, glue("frac_pie_{grouping}.png")), units = "px", width = 500, height = 500)

    # ---- density plot (log scale, values < 0.1 -> NA) ----
    density_col <- paste0(col_name, "_densityplot")
    mask_na     <- function(df) dplyr::mutate(df, !!density_col := ifelse(.data[[col_name]] < 0.1, NA, .data[[col_name]]))

    major_breaks <- 10^(-1:2)
    minor_breaks <- unlist(lapply(-1:2, function(i) (2:9) * 10^i))

    d <- ggplot2::ggplot() +
      ggplot2::geom_density(data = mask_na(df_background),
                            ggplot2::aes(x = .data[[density_col]]),
                            color = "grey70", lwd = 1) +
      ggplot2::geom_density(data = mask_na(df_interest),
                            ggplot2::aes(x = .data[[density_col]]),
                            color = "maroon", lwd = 1) +
      ggplot2::geom_vline(xintercept = vline_x, color = "maroon", lty = 2, lwd = 1) +
      ggplot2::scale_x_log10(breaks = major_breaks, minor_breaks = minor_breaks,
                             labels = scales::label_number()) +
      ggplot2::theme_minimal() +
      ggplot2::theme(panel.grid.major.x = ggplot2::element_line(color = "grey85"),
                     panel.grid.minor.x = ggplot2::element_line(color = "grey95", linewidth = 0.3),
                     axis.text.y        = ggplot2::element_blank(),
                     axis.title.y       = ggplot2::element_blank()) +
      ggplot2::labs(x = x_label)
    
    if(caption) {
      d <- d + 
        ggplot2::labs(caption = caption_text) +
        ggplot2::theme(plot.caption = ggplot2::element_text(size = 4))
    }

    ggplot2::ggsave(plot = d, filename = file.path(out_dir, glue("density_{grouping}.png")), units = "px", width = 1000, height = 700)

  } else {

    # ---- fraction pie ----
    frac_tbl <- df_interest |>
      dplyr::mutate(frac_group = ifelse(.data[[col_name]] < 0, "< 0", ">= 0")) |>
      dplyr::group_by(frac_group) |>
      dplyr::count() |>
      dplyr::ungroup() |>
      dplyr::mutate(
        frac_group = factor(frac_group, levels = c(">= 0", "< 0")),
        prop       = n / sum(n),
        label      = scales::percent(prop, accuracy = 1)
      )

    ggplot2::ggplot(frac_tbl, ggplot2::aes(x = "", y = prop, fill = frac_group)) +
      ggplot2::geom_col(width = 1, color = "white") +
      ggplot2::geom_text(ggplot2::aes(label = label),
                         position = ggplot2::position_stack(vjust = 0.5),
                         color = "white", size = 3) +
      ggplot2::coord_polar(theta = "y") +
      ggplot2::scale_fill_manual(values = c("< 0" = low_pie_color, ">= 0" = high_pie_color)) +
      ggplot2::labs(x = NULL, y = NULL) +
      ggplot2::theme_void() +
      ggplot2::theme(legend.title = ggplot2::element_blank())

    ggplot2::ggsave(file.path(out_dir, glue("frac_pie_{grouping}.png")), units = "px", width = 500, height = 500)

    # ---- density plot (reversed x axis) ----
    d <- ggplot2::ggplot() +
      ggplot2::geom_density(data = df_background,
                            ggplot2::aes(x = .data[[col_name]]),
                            color = "grey70", lwd = 1) +
      ggplot2::geom_density(data = df_interest,
                            ggplot2::aes(x = .data[[col_name]]),
                            color = "maroon", lwd = 1) +
      ggplot2::geom_vline(xintercept = vline_x, color = "maroon", lty = 2, lwd = 1) +
      ggplot2::scale_x_reverse() +
      ggplot2::theme_minimal() +
      ggplot2::theme(panel.grid.major.x = ggplot2::element_line(color = "grey85"),
                     panel.grid.minor.x = ggplot2::element_line(color = "grey95", linewidth = 0.3),
                     axis.text.y        = ggplot2::element_blank(),
                     axis.title.y       = ggplot2::element_blank()) +
      ggplot2::labs(x = x_label)
    
    if(caption) {
      d <- d + 
        ggplot2::labs(caption = caption_text) +
        ggplot2::theme(plot.caption = ggplot2::element_text(size = 4))
    }

    ggplot2::ggsave(plot = d, filename = file.path(out_dir, glue("density_{grouping}.png")), units = "px", width = 1000, height = 700)
  }
  
  frac_results <- frac_tbl |>
    dplyr::select(frac_group, label) |>
    tibble::deframe()
  
  results <- c(interest_deciles,
                        disturbance = col_name,
                        type = type,
                        grouping = grouping, frac_results)

  
  results <- tibble::as_tibble_row(
    c(
      as.list(interest_deciles),
      list(
        disturbance = col_name,
        type = type,
        grouping = grouping
      ),
      as.list(frac_results)
    ),
    .name_repair = "minimal"
  )
  
  return(results)
}



all_deciles <- tibble()

# Operational calls
raw_f_deciles_bio <- disturbance_density_description(
  df_background    = d_s_nfire,
  df_interest      = d_s_fire,
  col_name         = "biotic_relaxedforestnorm_sum_n5_to_p0",
  type             = "biotic",
  dir_figs         = dir_figs,
  x_label          = "6-year prior cumulative biotic disturbance"
)

raw_f_deciles_dr <- disturbance_density_description(
  df_background    = d_s_nfire,
  df_interest      = d_s_fire,
  col_name         = "pdsi_annual_sum_n5_to_p0",
  type             = "drought",
  dir_figs         = dir_figs,
  x_label          = "6-year prior cumulative drought"
)

all_deciles <- bind_rows(
  raw_f_deciles_bio,
  raw_f_deciles_dr
)

# check a few subsets - drought 90% threshold seems to be fairly consistent
# but biotic 90% is variable... 
nfgs <- unique(d_s_nfire |> filter(nfg_10k) |> pull(nfg_factor_clean))

dir_figs_density_nfg <- here(dir_figs, "nfg_densities")
dir_ensure(dir_figs_density_nfg)


for(i in nfgs) {

  bio_t <- disturbance_density_description(
    df_background    = d_s_nfire |> filter(nfg_factor_clean == i),
    df_interest      = d_s_fire |> filter(nfg_factor_clean == i),
    col_name         = "biotic_relaxedforestnorm_sum_n5_to_p0",
    type             = "biotic",
    dir_figs         = dir_figs_density_nfg,
    x_label          = "6-year prior cumulative biotic disturbance",
    grouping = i,
    caption = TRUE
  )
  
  dr_t <- disturbance_density_description(
    df_background    = d_s_nfire |> filter(nfg_factor_clean == i),
    df_interest      = d_s_fire |> filter(nfg_factor_clean == i),
    col_name         = "pdsi_annual_sum_n5_to_p0",
    type             = "drought",
    dir_figs         = dir_figs_density_nfg,
    x_label          = "6-year prior cumulative drought",
    grouping = i,
    caption = TRUE
  )
  
  all_deciles <- bind_rows(
    all_deciles,
    bio_t,
    dr_t
  )
    
}




ggplot(d_s_nfire |> filter(nfg_10k)) +
  geom_density(aes(x = gam_rap_tree_pre6_fit, color = nfg_factor_clean))

ggsave(here(dir_figs, "density_rap_pre6fit.png"),
       units = "px",
       width = 2000,
       height = 1000)


ggplot(d_s_nfire |> filter(nfg_10k)) +
  geom_density(aes(x = pdsi_annual_sum_n5_to_p0, color = nfg_factor_clean))

ggsave(here(dir_figs, "density_pdsi_nfg_6yr.png"),
       units = "px",
       width = 2000,
       height = 1000)


ggplot(d_s_nfire |> filter(nfg_10k)) +
  geom_density(aes(x = biotic_relaxedforestnorm_sum_n5_to_p0, color = nfg_factor_clean)) +
  scale_x_log10()

ggsave(here(dir_figs, "density_biotic_nfg_6yr.png"),
       units = "px",
       width = 2000,
       height = 1000)


# do a temporal split check
d <- disturbance_density_description(
  df_background    = d_s_nfire |> filter(mock_burn_year < 2012),
  df_interest      = d_s_fire |> filter(mock_burn_year < 2012),
  col_name         = "biotic_relaxedforestnorm_sum_n5_to_p0",
  type             = "biotic",
  dir_figs         = dir_figs,
  x_label          = "6-year prior cumulative biotic disturbance",
  grouping = "early",
  caption = TRUE
)
all_deciles <- bind_rows(all_deciles, d)

d <- disturbance_density_description(
  df_background    = d_s_nfire |> filter(mock_burn_year >= 2012),
  df_interest      = d_s_fire |> filter(mock_burn_year >= 2012),
  col_name         = "biotic_relaxedforestnorm_sum_n5_to_p0",
  type             = "biotic",
  dir_figs         = dir_figs,
  x_label          = "6-year prior cumulative biotic disturbance",
  grouping = "late",
  caption = TRUE
)
all_deciles <- bind_rows(all_deciles, d)

d <- disturbance_density_description(
  df_background    = d_s_nfire |> filter(mock_burn_year < 2012),
  df_interest      = d_s_fire |> filter(mock_burn_year < 2012),
  col_name         = "pdsi_annual_sum_n5_to_p0",
  type             = "drought",
  dir_figs         = dir_figs,
  x_label          = "6-year prior cumulative drought",
  grouping = "early",
  caption = TRUE
)
all_deciles <- bind_rows(all_deciles, d)


d <- disturbance_density_description(
  df_background    = d_s_nfire |> filter(mock_burn_year >= 2012),
  df_interest      = d_s_fire |> filter(mock_burn_year >= 2012),
  col_name         = "pdsi_annual_sum_n5_to_p0",
  type             = "drought",
  dir_figs         = dir_figs,
  x_label          = "6-year prior cumulative drought",
  grouping = "late",
  caption = TRUE
)
all_deciles <- bind_rows(all_deciles, d)



# pre-fire exploratory calls

d <- disturbance_density_description(
  df_background    = d_s_nfire,
  df_interest      = d_s_fire,
  col_name         = "biotic_relaxedforestnorm_sum_n2_to_p0",
  type             = "biotic",
  dir_figs         = dir_figs,
  x_label          = "3-year 2-0 prior cumulative biotic disturbance",
  caption = TRUE
)
all_deciles <- bind_rows(all_deciles, d)

d <- disturbance_density_description(
  df_background    = d_s_nfire,
  df_interest      = d_s_fire,
  col_name         = "biotic_relaxedforestnorm_sum_n5_to_n3",
  type             = "biotic",
  dir_figs         = dir_figs,
  x_label          = "3-year 5-3 prior cumulative biotic disturbance",
  caption = TRUE
)
all_deciles <- bind_rows(all_deciles, d)


d <- disturbance_density_description(
  df_background    = d_s_nfire,
  df_interest      = d_s_fire,
  col_name         = "pdsi_annual_sum_n2_to_p0",
  type             = "drought",
  dir_figs         = dir_figs,
  x_label          = "3-year 2-0 prior cumulative drought",
  caption = TRUE
)
all_deciles <- bind_rows(all_deciles, d)

d <- disturbance_density_description(
  df_background    = d_s_nfire,
  df_interest      = d_s_fire,
  col_name         = "pdsi_annual_sum_n5_to_n3",
  type             = "drought",
  dir_figs         = dir_figs,
  x_label          = "3-year 5-3 prior cumulative drought",
  caption = TRUE
)
all_deciles <- bind_rows(all_deciles, d)


# post-fire exploratory calls

d <- disturbance_density_description(
  df_background    = d_s_nfire,
  df_interest      = d_s_fire,
  col_name         = "biotic_relaxedforestnorm_sum_p1_to_p3",
  type             = "biotic",
  dir_figs         = dir_figs,
  x_label          = "3-year 1-3 post cumulative biotic disturbance",
  caption = TRUE
)
all_deciles <- bind_rows(all_deciles, d)

d <- disturbance_density_description(
  df_background    = d_s_nfire,
  df_interest      = d_s_fire,
  col_name         = "biotic_relaxedforestnorm_sum_p4_to_p6",
  type             = "biotic",
  dir_figs         = dir_figs,
  x_label          = "3-year 4-6 post cumulative biotic disturbance",
  caption = TRUE
)
all_deciles <- bind_rows(all_deciles, d)

d <- disturbance_density_description(
  df_background    = d_s_nfire,
  df_interest      = d_s_fire,
  col_name         = "biotic_relaxedforestnorm_sum_p1_to_p6",
  type             = "biotic",
  dir_figs         = dir_figs,
  x_label          = "6-year 1-6 post cumulative biotic disturbance",
  caption = TRUE
)
all_deciles <- bind_rows(all_deciles, d)


d <- disturbance_density_description(
  df_background    = d_s_nfire,
  df_interest      = d_s_fire,
  col_name         = "pdsi_annual_sum_p1_to_p3",
  type             = "drought",
  dir_figs         = dir_figs,
  x_label          = "3-year 1-3 post cumulative drought",
  caption = TRUE
)
all_deciles <- bind_rows(all_deciles, d)

d <- disturbance_density_description(
  df_background    = d_s_nfire,
  df_interest      = d_s_fire,
  col_name         = "pdsi_annual_sum_p4_to_p6",
  type             = "drought",
  dir_figs         = dir_figs,
  x_label          = "3-year 4-6 post cumulative drought",
  caption = TRUE
)
all_deciles <- bind_rows(all_deciles, d)

d <- disturbance_density_description(
  df_background    = d_s_nfire,
  df_interest      = d_s_fire,
  col_name         = "pdsi_annual_sum_p1_to_p6",
  type             = "drought",
  dir_figs         = dir_figs,
  x_label          = "6-year 1-6 post cumulative drought",
  caption = TRUE
)
all_deciles <- bind_rows(all_deciles, d) |>
  arrange(desc(disturbance))

write_csv(all_deciles, file = here(dir_figs, "deciles_all.csv"))


```

# Set CD groups

```{r}

set_cd_groups <- function(df,
                          b_nm,
                          d_nm,
                          cd_nm,
                          b_threshold,
                          d_threshold) {
  
  if(d_threshold < 0) {
    df_new <- df |>
    dplyr::mutate(
      "{cd_nm}" := dplyr::case_when(
        .data[[b_nm]] <  b_threshold & fire == 1 & .data[[d_nm]] >  d_threshold ~ "f",
        .data[[b_nm]] >= b_threshold & fire == 1 & .data[[d_nm]] >  d_threshold ~ "bf",
        .data[[b_nm]] <  b_threshold & fire == 1 & .data[[d_nm]] <= d_threshold ~ "df",
        .data[[b_nm]] >= b_threshold & fire == 1 & .data[[d_nm]] <= d_threshold ~ "bdf",
        TRUE ~ NA_character_
      )
    )
  } else {
    df_new <- df |>
    dplyr::mutate(
      "{cd_nm}" := dplyr::case_when(
        .data[[b_nm]] <  b_threshold & fire == 1 & .data[[d_nm]] <  d_threshold ~ "f",
        .data[[b_nm]] >= b_threshold & fire == 1 & .data[[d_nm]] <  d_threshold ~ "bf",
        .data[[b_nm]] <  b_threshold & fire == 1 & .data[[d_nm]] >= d_threshold ~ "df",
        .data[[b_nm]] >= b_threshold & fire == 1 & .data[[d_nm]] >= d_threshold ~ "bdf",
        TRUE ~ NA_character_
      )
    )
    
  }
  
  df_new <- df_new |>
    dplyr::mutate(
      "{cd_nm}" := forcats::fct_relevel(.data[[cd_nm]], "f")
    )

  return(df_new)
}




# deciles
raw_f_deciles_dr
raw_f_deciles_bio


d_s <- set_cd_groups(df = d_s,
                   b_nm = "biotic_relaxedforestnorm_sum_n5_to_p0",
                   d_nm = "pdsi_annual_sum_n5_to_p0",
                   cd_nm = "cd_group_b90_pdsin90",
                   b_threshold = raw_f_deciles_bio[[9]],
                   d_threshold = raw_f_deciles_dr[[9]]) |>
  set_cd_groups(b_nm = "biotic_relaxedforestnorm_sum_n5_to_p0",
                   d_nm = "pdsi_annual_sum_n5_to_p0",
                   cd_nm = "cd_group_b80_pdsin80",
                   b_threshold = raw_f_deciles_bio[[8]],
                   d_threshold = raw_f_deciles_dr[[8]])




```

# test some new grouping functions

```{r}


set_cd_groups_extended <- function(df,
                          b_nm,
                          d_nm,
                          cd_nm,
                          b_threshold,
                          d_threshold) {
  
  if(d_threshold < 0) {
    df_new <- df |>
    dplyr::mutate(
      "{cd_nm}" := dplyr::case_when(
        .data[[b_nm]] <  b_threshold &
          fire == 1 & 
          .data[[d_nm]] > d_threshold ~ "f",
        .data[[b_nm]] >= b_threshold & 
          fire == 1 & 
          .data[[d_nm]] > d_threshold ~ "bf",
        .data[[b_nm]] < b_threshold & 
          fire == 1 & 
          .data[[d_nm]] <= d_threshold ~ "df",
        .data[[b_nm]] >= b_threshold & 
          fire == 1 & 
          .data[[d_nm]] <= d_threshold ~ "bdf",
        .data[[b_nm]] >= b_threshold & 
          fire == 0 & 
          .data[[d_nm]] <= d_threshold ~ "bd",
        .data[[b_nm]] >= b_threshold & 
          fire == 0 & 
          .data[[d_nm]] > d_threshold ~ "b",
        .data[[b_nm]] < b_threshold & 
          fire == 0 & 
          .data[[d_nm]] <= d_threshold ~ "d",
        .data[[b_nm]] < b_threshold & 
          fire == 0 & 
          .data[[d_nm]] > d_threshold ~ NA_character_,
        TRUE ~ "DROP"
      )
    )
  } else {
    stop("please use drought threshold < 0")
    
  }
  
  df_new <- df_new |>
    dplyr::mutate(
      "{cd_nm}" := forcats::fct_relevel(.data[[cd_nm]], "f")
    )

  return(df_new)
}



set_cd_groups_extended_strict <- function(df,
                          b_nm,
                          d_nm,
                          b_rollmax_nm,
                          d_rollmax_nm,
                          cd_nm,
                          b_threshold,
                          d_threshold) {
  
  if(d_threshold < 0) {
    df_new <- df |>
    dplyr::mutate(
      "{cd_nm}" := dplyr::case_when(
        .data[[b_rollmax_nm]] <  b_threshold &
           .data[[b_nm]] <  b_threshold &
          fire == 1 &
          .data[[d_nm]] > d_threshold &
          .data[[d_rollmax_nm]] > d_threshold ~ "f",
        .data[[b_nm]] >  b_threshold &
          fire == 1 &
          .data[[d_nm]] > d_threshold &
          .data[[d_rollmax_nm]] > d_threshold ~ "bf",
        .data[[b_nm]] <  b_threshold &
          .data[[b_rollmax_nm]] < b_threshold &
          fire == 1 &
          .data[[d_nm]] <= d_threshold ~ "df",
        .data[[b_nm]] >= b_threshold & 
          fire == 1 & 
          .data[[d_nm]] <= d_threshold ~ "bdf",
        .data[[b_nm]] >= b_threshold & 
          fire == 0 & 
          .data[[d_nm]] <= d_threshold ~ "bd",
        .data[[b_nm]] >= b_threshold & 
          fire == 0 & 
          .data[[d_rollmax_nm]] > d_threshold &
          .data[[d_nm]] > d_threshold ~ "b",
        .data[[b_rollmax_nm]] < b_threshold & 
          .data[[b_nm]] < b_threshold &
          fire == 0 & 
          .data[[d_nm]] <= d_threshold ~ "d",
        .data[[b_rollmax_nm]] < b_threshold & 
          .data[[b_nm]] < b_threshold &
          fire == 0 & 
          .data[[d_nm]] > d_threshold &
          .data[[d_rollmax_nm]] > d_threshold ~ NA_character_,
        TRUE ~ "DROP"
      )
    )
  } else {
    stop("please use drought threshold < 0")
    
  }
  
  df_new <- df_new |>
    dplyr::filter(.data[[cd_nm]] != "DROP") |>
    dplyr::mutate(
      "{cd_nm}" := forcats::fct_relevel(.data[[cd_nm]], "f")
    )

  return(df_new)
}




set_cd_groups_extended_strict_v2 <- function(df,
                          b_nm,
                          d_nm,
                          b_rollmax_nm,
                          d_rollmax_nm,
                          cd_nm,
                          b_threshold,
                          d_threshold) {
  
  if(d_threshold < 0) {
    df_new <- df |>
    dplyr::mutate(
      "{cd_nm}" := dplyr::case_when(
        .data[[b_rollmax_nm]] <  b_threshold &
           .data[[b_nm]] <  b_threshold &
          fire == 1 &
          .data[[d_nm]] > d_threshold &
          .data[[d_rollmax_nm]] > d_threshold ~ "f",
        .data[[b_nm]] >  b_threshold &
          .data[[b_nm]] >= .data[[b_rollmax_nm]] &
          fire == 1 &
          .data[[d_nm]] > d_threshold &
          .data[[d_rollmax_nm]] > d_threshold ~ "bf",
        .data[[b_nm]] <  b_threshold &
          .data[[b_rollmax_nm]] < b_threshold &
          fire == 1 &
          .data[[d_nm]] <= .data[[d_rollmax_nm]] &
          .data[[d_nm]] <= d_threshold ~ "df",
        .data[[b_nm]] >= b_threshold & 
          .data[[b_nm]] >= .data[[b_rollmax_nm]] &
          fire == 1 & 
          .data[[d_nm]] <= .data[[d_rollmax_nm]] &
          .data[[d_nm]] <= d_threshold ~ "bdf",
        .data[[b_nm]] >= b_threshold & 
          .data[[b_nm]] >= .data[[b_rollmax_nm]] &
          fire == 0 & 
          .data[[d_nm]] <= .data[[d_rollmax_nm]] &
          .data[[d_nm]] <= d_threshold ~ "bd",
        .data[[b_nm]] >= b_threshold & 
          .data[[b_nm]] >= .data[[b_rollmax_nm]] &
          fire == 0 & 
          .data[[d_rollmax_nm]] > d_threshold &
          .data[[d_nm]] > d_threshold ~ "b",
        .data[[b_rollmax_nm]] < b_threshold & 
          .data[[b_nm]] < b_threshold &
          fire == 0 & 
          .data[[d_nm]] <= .data[[d_rollmax_nm]] &
          .data[[d_nm]] <= d_threshold ~ "d",
        .data[[b_rollmax_nm]] < b_threshold & 
          .data[[b_nm]] < b_threshold &
          fire == 0 & 
          .data[[d_nm]] > d_threshold &
          .data[[d_rollmax_nm]] > d_threshold ~ NA_character_,
        TRUE ~ "DROP"
      )
    )
  } else {
    stop("please use drought threshold < 0")
    
  }
  
  df_new <- df_new |>
    dplyr::filter(.data[[cd_nm]] != "DROP") |>
    dplyr::mutate(
      "{cd_nm}" := forcats::fct_relevel(.data[[cd_nm]], "f")
    )

  return(df_new)
}


# test 6 year
d_s <- set_cd_groups_extended(df = d_s,
                   b_nm = "biotic_relaxedforestnorm_sum_n5_to_p0",
                   d_nm = "pdsi_annual_sum_n5_to_p0",
                   cd_nm = "cd_90_extended",
                   b_threshold = raw_f_deciles_bio[[9]],
                   d_threshold = raw_f_deciles_dr[[9]])

d_s <- set_cd_groups_extended_strict(df = d_s,
                   b_nm = "biotic_relaxedforestnorm_sum_n5_to_p0",
                   d_nm = "pdsi_annual_sum_n5_to_p0",
                   b_rollmax_nm = "biotic_relaxedforestnorm_rollsum6max",
                   d_rollmax_nm = "pdsi_annual_rollsum6max",
                   cd_nm = "cd_90_extended_strict",
                   b_threshold = raw_f_deciles_bio[[9]],
                   d_threshold = raw_f_deciles_dr[[9]])

d_s <- set_cd_groups_extended_strict_v2(df = d_s,
                   b_nm = "biotic_relaxedforestnorm_sum_n5_to_p0",
                   d_nm = "pdsi_annual_sum_n5_to_p0",
                   b_rollmax_nm = "biotic_relaxedforestnorm_rollsum6max",
                   d_rollmax_nm = "pdsi_annual_rollsum6max",
                   cd_nm = "cd_90_extended_strict_v2",
                   b_threshold = raw_f_deciles_bio[[9]],
                   d_threshold = raw_f_deciles_dr[[9]])


# test 3 year
d_s <- set_cd_groups_extended(df = d_s,
                   b_nm = "biotic_relaxedforestnorm_sum_n2_to_p0",
                   d_nm = "pdsi_annual_sum_n2_to_p0",
                   cd_nm = "cd_90_extended_3yr",
                   b_threshold = raw_f_deciles_bio[[9]],
                   d_threshold = raw_f_deciles_dr[[9]])

d_s <- set_cd_groups_extended_strict(df = d_s,
                   b_nm = "biotic_relaxedforestnorm_sum_n2_to_p0",
                   d_nm = "pdsi_annual_sum_n2_to_p0",
                   b_rollmax_nm = "biotic_relaxedforestnorm_rollsum3max",
                   d_rollmax_nm = "pdsi_annual_rollsum3max",
                   cd_nm = "cd_90_extended_strict_3yr",
                   b_threshold = raw_f_deciles_bio[[9]],
                   d_threshold = raw_f_deciles_dr[[9]])

d_s <- set_cd_groups_extended_strict_v2(df = d_s,
                   b_nm = "biotic_relaxedforestnorm_sum_n2_to_p0",
                   d_nm = "pdsi_annual_sum_n2_to_p0",
                   b_rollmax_nm = "biotic_relaxedforestnorm_rollsum3max",
                   d_rollmax_nm = "pdsi_annual_rollsum3max",
                   cd_nm = "cd_90_extended_strict_v2_3yr",
                   b_threshold = raw_f_deciles_bio[[9]],
                   d_threshold = raw_f_deciles_dr[[9]])

```

# assess sample size and NON-fire impacted sample size

```{r}

# check 6-year
cd_counts_90 <- d_s |>
  group_by(cd_group_b90_pdsin90) |>
  count() |>
  ungroup()

cd_counts_90_extended <- d_s |>
  group_by(cd_90_extended) |>
  count() |>
  ungroup()

cd_counts_90_extended_strict <- d_s |>
  group_by(cd_90_extended_strict) |>
  count() |>
  ungroup()

cd_counts_90_extended_strict_v2 <- d_s |>
  group_by(cd_90_extended_strict_v2) |>
  count() |>
  ungroup()


analyze_strict_req <- cd_counts_90_extended |>
  rename(cd_group = cd_90_extended,
         cd_counts_90_extended = n) |>
  left_join(cd_counts_90_extended_strict |>
              rename(cd_group = cd_90_extended_strict,
                     cd_counts_90_extended_strict = n)) |>
  left_join(cd_counts_90_extended_strict_v2 |>
              rename(cd_group = cd_90_extended_strict_v2,
                     cd_counts_90_extended_strict_v2 = n)) |>
  mutate(strict_dropped = cd_counts_90_extended - cd_counts_90_extended_strict,
         perc_strict_dropped = round(strict_dropped / cd_counts_90_extended * 100),
         strict_v2_dropped = cd_counts_90_extended - cd_counts_90_extended_strict_v2,
         perc_strict_v2_dropped = round(strict_v2_dropped / cd_counts_90_extended * 100))


write_csv(analyze_strict_req, file = here(dir_figs, "analyze_strict_req.csv"))


# check 3-year

cd_counts_90_extended_3yr <- d_s |>
  group_by(cd_90_extended_3yr) |>
  count() |>
  ungroup()

cd_counts_90_extended_strict_3yr <- d_s |>
  group_by(cd_90_extended_strict_3yr) |>
  count() |>
  ungroup()

cd_counts_90_extended_strict_v2_3yr <- d_s |>
  group_by(cd_90_extended_strict_v2_3yr) |>
  count() |>
  ungroup()



analyze_strict_req_3yr <- cd_counts_90_extended_3yr |>
  rename(cd_group_3yr = cd_90_extended_3yr,
         cd_counts_90_extended_3yr = n) |>
  left_join(cd_counts_90_extended_strict_3yr |>
              rename(cd_group_3yr = cd_90_extended_strict_3yr,
                     cd_counts_90_extended_strict_3yr = n)) |>
  left_join(cd_counts_90_extended_strict_v2_3yr |>
              rename(cd_group_3yr = cd_90_extended_strict_v2_3yr,
                     cd_counts_90_extended_strict_v2_3yr = n)) |>
  mutate(strict_dropped_3yr = cd_counts_90_extended_3yr - cd_counts_90_extended_strict_3yr,
         perc_strict_dropped_3yr = round(strict_dropped_3yr / cd_counts_90_extended_3yr * 100),
         strict_v2_dropped_3yr = cd_counts_90_extended_3yr - cd_counts_90_extended_strict_v2_3yr,
         perc_strict_v2_dropped_3yr = round(strict_v2_dropped_3yr / cd_counts_90_extended_3yr * 100))


write_csv(analyze_strict_req_3yr, file = here(dir_figs, "analyze_strict_req_3yr.csv"))





cd_counts_80 <- d_s_fire |>
  group_by(cd_group_b80_pdsin80) |>
  count() |>
  ungroup()



d_s_fire |>
  group_by(nfg_factor_clean) |>
  count() |>
  left_join(d_s_nfire |>
    group_by(nfg_factor_clean) |>
    count() |> rename(n2 = n)) |>
  mutate(total = n + n2) |>
  arrange(desc(total))


d_s_fire |>
  group_by(nfg_broad_type) |>
  count() |>
  left_join(d_s_nfire |>
    group_by(nfg_broad_type) |>
    count() |> rename(n2 = n)) |>
  mutate(total = n + n2) |>
  arrange(desc(total))


```

# assess rolling maximums

How often are there multiple rollmaxes that pass the threshold?

How often is there a rollmax higher than the pre-fire rollmax for fire units?

Are fire versus non-fire biotic classes

```{r}

ggplot(d_s) +
  geom_density(aes(x = biotic_relaxedforestnorm_rollsum6max, color = factor(fire))) +
  scale_x_log10()

ggplot(d_s) +
  geom_density(aes(x = biotic_relaxedforestnorm_rollsum3max, color = factor(fire))) +
  scale_x_log10()






d_l <- open_dataset(here(dir_derived, "parquet_long"), format = "parquet") |>
  collect()



# d_l_fire <- d_l |> filter(fire == 1)


```

# temporal fig

```{r}

counts_by_year <- d_s_fire |>
  group_by(cd_group_b90_pdsin90, burn_year) |>
  count() |>
  ungroup()

ggplot(counts_by_year) +
  geom_col(aes(x = burn_year, y = n)) +
  facet_wrap( ~ cd_group_b90_pdsin90, scales = "free") +
  labs(title = "BDF counts over time: 90th percentile all time/all groups")

ggsave(filename = here(dir_figs, "bdf_counts_over_time_90.png"))





counts_by_year_extended <- d_s |>
  group_by(cd_90_extended, mock_burn_year) |>
  count() |>
  ungroup()

ggplot(counts_by_year_extended) +
  geom_col(aes(x = mock_burn_year, y = n)) +
  facet_wrap( ~ cd_90_extended, scales = "free") +
  labs(title = "BDF counts over time: 90th percentile all time/all groups")

ggsave(filename = here(dir_figs, "bdf_counts_over_time_90_extended.png"))




counts_by_year_extended_strict <- d_s |>
  group_by(cd_90_extended_strict, mock_burn_year) |>
  count() |>
  ungroup()

ggplot(counts_by_year_extended_strict) +
  geom_col(aes(x = mock_burn_year, y = n)) +
  facet_wrap( ~ cd_90_extended_strict, scales = "free") +
  labs(title = "BDF counts over time: 90th percentile all time/all groups")

ggsave(filename = here(dir_figs, "bdf_counts_over_time_90_extended_strict.png"))

```

# ecoregions map

```{r}

ecors <- l3pf |>
  filter(perc_both >= 30) |>
  pull(na_l3name)

ecors_sf <- l3_sf |>
  filter(NA_L3NAME %in% ecors) |>
  sf::st_transform('EPSG:4326')

states <- tigris::states() |>
  sf::st_transform('EPSG:4326') |>
  sf::st_filter(ecors_sf) |>
  filter(STUSPS != "TX" & STUSPS != "SD")

#mapview::mapview(states)


h3_counts <- d_s_fire |>
  count(h3jsr_5, name = "n_units") |>
  filter(!is.na(h3jsr_5))

h3_geom_sf <- h3jsr::cell_to_polygon(
  h3_counts$h3jsr_5,
  simple = FALSE
)

h3_counts_sf <- h3_geom_sf |>
  mutate(
    h3jsr_5 = h3_counts$h3jsr_5,
    n_units = h3_counts$n_units
  ) |>
  st_transform(st_crs(ecors_sf))



ggplot() +
  geom_sf(
    data = states,
    fill = "gray97",
    color = "gray85",
    linewidth = 0.2
  ) +
  geom_sf(
    data = ecors_sf,
    fill = NA,
    color = "gray20",
    linewidth = 0.2
  ) +
  geom_sf(
    data = h3_counts_sf,
    aes(fill = n_units),
    color = NA,
    alpha = 0.9
  ) +
  scale_fill_viridis_c(
    option = "viridis",
    trans = "sqrt",
    labels = comma,
    name = "Units per hex"
  ) +
  theme_minimal() +
  labs(
    x = NULL,
    y = NULL
  )

ggsave(here(dir_figs, "spatial_domain_units_per_hex.png"),
       units = "px",
       width = 2000,
       height = 2000)


```

# cd percentage separate maps

```{r}



h3_medians <- d_s_fire |>
  dplyr::filter(!is.na(h3jsr_5)) |>
  dplyr::group_by(h3jsr_5) |>
  dplyr::summarize(
    n_units = dplyr::n(),
    median_bio = median(biotic_relaxedforestnorm_sum_n5_to_p0, na.rm = TRUE),
    median_dr  = median(pdsi_annual_sum_n5_to_p0, na.rm = TRUE),
    .groups = "drop"
  )


h3_medians_sf <- h3_geom_sf |>
  dplyr::mutate(
    h3jsr_5 = h3_medians$h3jsr_5,
    n_units = h3_medians$n_units,
    median_bio = h3_medians$median_bio,
    median_dr = h3_medians$median_dr
  ) |>
  sf::st_transform(sf::st_crs(ecors_sf))



ggplot() +
  geom_sf(
    data = states,
    fill = "gray97",
    color = "gray85",
    linewidth = 0.2
  ) +
  geom_sf(
    data = ecors_sf,
    fill = NA,
    color = "gray20",
    linewidth = 0.2
  ) +
  geom_sf(
    data = h3_medians_sf,
    aes(fill = median_bio),
    color = NA,
    alpha = 0.9
  ) +
  scale_fill_viridis_c(
    option = "viridis",
    trans = "sqrt",
    labels = comma,
    name = "Median\n6-yr\npre-fire\nbiotic"
  ) +
  theme_minimal() +
  labs(
    x = NULL,
    y = NULL
  )

ggsave(here(dir_figs, "spatial_domain_median_biotic.png"),
       units = "px",
       width = 2000,
       height = 2000)


ggplot() +
  geom_sf(
    data = states,
    fill = "gray97",
    color = "gray85",
    linewidth = 0.2
  ) +
  geom_sf(
    data = ecors_sf,
    fill = NA,
    color = "gray20",
    linewidth = 0.2
  ) +
  geom_sf(
    data = h3_medians_sf,
    aes(fill = median_dr),
    color = NA,
    alpha = 0.9
  ) +
  scale_fill_viridis_c(
    option = "viridis",
    direction = -1,
    labels = comma,
    name = "Median\n6-yr\npre-fire\ndrought"
  ) +
  theme_minimal() +
  labs(
    x = NULL,
    y = NULL
  )

ggsave(
  here(dir_figs, "spatial_domain_median_drought.png"),
  units = "px",
  width = 2000,
  height = 2000
)


```

# v2 - cd proportions map

## assessing colors

```{r}
library(colorspace)

# ---- Option 1: Okabe-Ito inspired (blue/orange is a classic CVD-safe pair) ----
group_cols1 <- c(
  f   = hex(polarLUV(L = 82, C = 8,  H = 95)),   # near-neutral warm gray, barely tinted
  bf  = hex(polarLUV(L = 62, C = 45, H = 230)),  # sky blue
  df  = hex(polarLUV(L = 62, C = 45, H = 65)),   # orange
  bdf = hex(polarLUV(L = 25, C = 40, H = 300))   # dark purple
)

# ---- Option 2: earthy/aquatic (teal + gold) ----
group_cols2 <- c(
  f   = hex(polarLUV(L = 80, C = 10, H = 70)),   # warm sand/beige
  bf  = hex(polarLUV(L = 58, C = 45, H = 195)),  # teal
  df  = hex(polarLUV(L = 58, C = 45, H = 85)),   # goldenrod
  bdf = hex(polarLUV(L = 22, C = 35, H = 270))   # deep indigo
)

# ---- Option 3: cool violet/yellow ----
group_cols3 <- c(
  f   = hex(polarLUV(L = 80, C = 8,  H = 250)),  # pale cool gray-blue
  bf  = hex(polarLUV(L = 62, C = 55, H = 300)),  # violet
  df  = hex(polarLUV(L = 62, C = 55, H = 95)),   # yellow
  bdf = hex(polarLUV(L = 25, C = 28, H = 200))   # near-black teal
)

# ---- Option 4: monochrome-blue family with a magenta "pop" for bdf ----
group_cols4 <- c(
  f   = hex(polarLUV(L = 85, C = 6,  H = 250)),  # very pale slate blue
  bf  = hex(polarLUV(L = 55, C = 45, H = 260)),  # blue-violet
  df  = hex(polarLUV(L = 55, C = 43, H = 190)),  # cyan-teal
  bdf = hex(polarLUV(L = 22, C = 45, H = 330))   # deep magenta, off the blue family entirely
)

group_cols <- c(
  f   = "#C9BFB4",
  bf  = "#0072B2",  
  df  = "#E69F00",
  bdf = "#7A0177"   
)

check_valid_cols <- function(cols) {
  bad <- names(cols)[is.na(cols)]
  if (length(bad) > 0) {
    stop(
      "these group_cols entries are NA (likely out-of-gamut polarLUV values, ",
      "try lowering C or raising L): ", paste(bad, collapse = ", ")
    )
  }
}


# ---------------------------------------------------------------------------
# 2. color mixing: blend the 3 "edge" groups (f, bf, df) in Lab space
# (Lab avoids the muddy grays you get averaging raw sRGB), then pull toward
# a pure bdf color in proportion to p_bdf itself -- p_bdf = 0 gives a pure
# 3-way blend, p_bdf = 1 gives pure bdf, everything between fades smoothly
# ---------------------------------------------------------------------------
mix_ternary_cols <- function(p_f, p_bf, p_df, p_bdf, cols = group_cols) {

  edge_sum <- p_f + p_bf + p_df
  # guard divide-by-zero for hexes that are 100% bdf -- these edge weights
  # get wiped out by the p_bdf = 1 term below anyway, so the value here
  # doesn't matter
  edge_sum[edge_sum == 0] <- 1

  w_f  <- p_f  / edge_sum
  w_bf <- p_bf / edge_sum
  w_df <- p_df / edge_sum

  lab_mat <- farver::decode_colour(cols[c("f", "bf", "df")], to = "lab")
  bdf_lab <- farver::decode_colour(cols["bdf"], to = "lab")

  edge_lab <- outer(w_f, lab_mat[1, ]) +
    outer(w_bf, lab_mat[2, ]) +
    outer(w_df, lab_mat[3, ])

  mixed_lab <- (1 - p_bdf) * edge_lab + p_bdf * matrix(
    bdf_lab, nrow = length(p_bdf), ncol = 3, byrow = TRUE
  )

  farver::encode_colour(mixed_lab, from = "lab")
}

make_ternary_legend <- function(cols = group_cols, resolution = 300) {

  tri_height <- sqrt(3) / 2

  grid <- expand.grid(
    x = seq(0, 1, length.out = resolution),
    y = seq(0, tri_height, length.out = resolution)
  )

  # inverse barycentric transform for corners at (0,0), (1,0), (0.5, tri_height)
  grid <- grid |>
    mutate(
      w_df = y / tri_height,
      w_bf = x - w_df * 0.5,
      w_f  = 1 - w_bf - w_df
    ) |>
    filter(w_f >= -1e-6, w_bf >= -1e-6, w_df >= -1e-6) |>
    mutate(across(c(w_f, w_bf, w_df), \(w) pmax(w, 0)))

  # min(w_f, w_bf, w_df) peaks at exactly 1/3 at the centroid and hits 0 on
  # any edge or corner, so scaling by 3 gives a clean 0-to-1 fade-in for bdf - lower numbers will be less
  grid$w_bdf <- pmin(1, 2 * pmin(grid$w_f, grid$w_bf, grid$w_df))

  grid$fill_col <- mix_ternary_cols(
    p_f = grid$w_f, p_bf = grid$w_bf, p_df = grid$w_df, p_bdf = grid$w_bdf,
    cols = cols
  )

  # pick a readable label color for "bdf" against whatever bdf color is set
  bdf_lab_l <- farver::decode_colour(cols["bdf"], to = "lab")[1, 1]
  bdf_label_col <- if (bdf_lab_l > 60) "black" else "white"

  ggplot(grid, aes(x, y, fill = fill_col)) +
    geom_raster() +
    scale_fill_identity() +
    annotate("text", x = 0,   y = -0.05,             label = "F",   fontface = "bold") +
    annotate("text", x = 1,   y = -0.05,             label = "BF",  fontface = "bold") +
    annotate("text", x = 0.5, y = tri_height + 0.05,  label = "DF",  fontface = "bold") +
    annotate("text", x = 0.5, y = tri_height / 3,     label = "BDF", fontface = "bold", color = bdf_label_col) +
    coord_fixed(clip = "off") +
    theme_void()
}


check_valid_cols(group_cols1)
check_valid_cols(group_cols2)
check_valid_cols(group_cols3)
check_valid_cols(group_cols4)


make_ternary_legend(group_cols1)
make_ternary_legend(group_cols2)
#make_ternary_legend(group_cols3)
make_ternary_legend(group_cols4)
make_ternary_legend(group_cols)

```

```{r}



# ---------------------------------------------------------------------------
# f / bf / df sit at the triangle's three corners; bdf is its own pure
# color that the fill fades toward at the triangle's center
# ---------------------------------------------------------------------------

group_cols <- c(
  f   = "#C9BFB4",
  bf  = "#0072B2",  
  df  = "#E69F00",
  bdf = "#7A0177"   
)

# ---------------------------------------------------------------------------
# 1. proportion of each group within each h3 hex
# ---------------------------------------------------------------------------
h3_props <- d_s_fire |>
  st_drop_geometry() |>
  filter(
    !is.na(h3jsr_5),
    !is.na(cd_group_b90_pdsin90)
  ) |>
  count(h3jsr_5, cd_group_b90_pdsin90, name = "n") |>
  group_by(h3jsr_5) |>
  mutate(
    n_units = sum(n),
    prop = n / n_units
  ) |>
  ungroup() |>
  select(h3jsr_5, n_units, cd_group_b90_pdsin90, prop) |>
  pivot_wider(
    names_from = cd_group_b90_pdsin90,
    values_from = prop,
    values_fill = 0,
    names_prefix = "p_"
  )

for (nm in c("p_f", "p_bf", "p_df", "p_bdf")) {
  if (!nm %in% names(h3_props)) h3_props[[nm]] <- 0
}


h3_props <- h3_props |>
  mutate(
    fill_col = mix_ternary_cols(p_f, p_bf, p_df, p_bdf)
  )

# ---------------------------------------------------------------------------
# 3. convert h3 cells to polygons and join on the index
# ---------------------------------------------------------------------------
h3_geom_sf <- h3jsr::cell_to_polygon(h3_props$h3jsr_5, simple = FALSE)

idx_col <- setdiff(names(h3_geom_sf), "geometry")[1]

h3_props_sf <- h3_geom_sf |>
  rename(h3jsr_5 = all_of(idx_col)) |>
  left_join(h3_props, by = "h3jsr_5") |>
  st_transform(st_crs(ecors_sf))

# ---------------------------------------------------------------------------
# 4. the map itself
# ---------------------------------------------------------------------------
map_plot <- ggplot() +
  geom_sf(
    data = states,
    fill = "gray97",
    color = "gray85",
    linewidth = 0.2
  ) +
  geom_sf(
    data = ecors_sf,
    fill = NA,
    color = "gray20",
    linewidth = 0.2
  ) +
  geom_sf(
    data = h3_props_sf,
    aes(fill = fill_col),
    color = NA,
    alpha = 0.95
  ) +
  scale_fill_identity() +
  theme_minimal() +
  labs(x = NULL, y = NULL)


legend_plot <- make_ternary_legend()

# ---------------------------------------------------------------------------
# 6. combine map + legend and save
# ---------------------------------------------------------------------------
combined_plot <- map_plot +
  inset_element(
    legend_plot,
    left = 0.02, bottom = 0.02, right = 0.22, top = 0.22
  )

ggsave(
  here(dir_figs, "spatial_domain_group_composition_per_hex_v2.png"),
  plot = combined_plot,
  units = "px",
  width = 2000,
  height = 2000
)


```

# Outcomes

```{r}
ggplot(d_s |> filter(fire == 1 & !is.na(cd_group_b10_pdsisumn10))) +
  geom_density(aes(x = raw_post_fire_rap_tree_difference_n6_3_yr, color = cd_group_b10_pdsisumn10))



ggplot(d_s |> filter(fire == 1)) +
  geom_point(aes(x = burn_cbi,
                 y = raw_post_fire_rap_tree_difference_n6_3_yr),
             alpha = 0.2) +
  geom_smooth(aes(x = burn_cbi,
                 y = raw_post_fire_rap_tree_difference_n6_3_yr))

ggsave(here(dir_figs, "loss_by_cbi.png"))

```

# correlations

```{r}

safe_path_component <- function(x) gsub("[^[:alnum:]_\\-]+", "_", as.character(x))

dir_correlations <- here::here(dir_figs, "correlations")
dir_ensure(dir_correlations)

# Correlation matrix
correlations <- d_s |>
  select(aet, def, gam_rap_tree_pre6_fit, srtm, tpi, chili,
         pdsi_annual_5_yrs_prior_sum_yot, biotic_relaxedforestnorm_5_yrs_prior_sum_yot) |>
  cor(use = "pairwise.complete.obs")
png(here(dir_correlations, glue('corr_plot_1_all.png')), width = 600, height = 600)
corrplot(correlations, method = 'number', type = "lower", diag = FALSE, use = 'pairwise.complete.obs') 
dev.off()





drought_correlations <- d_s |>
  select(hd_fingerprint_year_of, pdsi_annual_year_of, pdsi_annual_5_yrs_prior_sum_yot, pdsi_annual_5_yrs_prior_threshold_n4_yot, pdsi_annual_5_yrs_prior_threshold_n3_yot,  hd_fingerprint_5_yrs_prior_threshold_6_yot, pdsi_annual_5_yrs_after_sum, pdsi_annual_5_yrs_prior_mean_yot, pdsi_annual_5_yrs_after_mean) |>
  cor(use = "pairwise.complete.obs")
png(here(dir_correlations, glue('corr_plot_2_all.png')), width = 800, height = 800)
corrplot(drought_correlations, method = 'number', type = "lower", diag = FALSE, use = 'pairwise.complete.obs') 
dev.off()


for(ecor in unique(d_s$nfg_factor)) {
  
  correlations <- d_s |>
    filter(nfg_factor == ecor) |>
    select(aet, def, gam_rap_tree_pre6_fit, srtm, tpi, chili,
         pdsi_annual_5_yrs_prior_sum_yot, biotic_relaxedforestnorm_5_yrs_prior_sum_yot) |>
    cor(use = "pairwise.complete.obs")
  png(here(dir_correlations, glue('corr_plot_1_{safe_path_component(ecor)}.png')), width = 600, height = 600)
  corrplot(correlations, method = 'number', type = "lower", diag = FALSE, use = 'pairwise.complete.obs') 
  dev.off()
  
  
  
  drought_correlations <- d_s |>
    filter(nfg_factor == ecor) |>
    select(hd_fingerprint_year_of, pdsi_annual_year_of, pdsi_annual_5_yrs_prior_sum_yot, pdsi_annual_5_yrs_prior_threshold_n4_yot, pdsi_annual_5_yrs_prior_threshold_n3_yot, hd_fingerprint_5_yrs_prior_threshold_6_yot, pdsi_annual_5_yrs_after_sum, pdsi_annual_5_yrs_prior_mean_yot, pdsi_annual_5_yrs_after_mean) |>
    cor(use = "pairwise.complete.obs")
  png(here(dir_correlations, glue('corr_plot_2_{safe_path_component(ecor)}.png')), width = 800, height = 800)
  corrplot(drought_correlations, method = 'number', type = "lower", diag = FALSE, use = 'pairwise.complete.obs') 
  dev.off()

}



ggplot(d_s_fire |> filter(cd_group_b10_pdsisumn10 == "bdf" | cd_group_b10_pdsisumn10 == "bf")) +
  geom_density(aes(x = biotic_relaxedforestnorm_5_yrs_prior_sum_yot, color = cd_group_b10_pdsisumn10)) +
  theme_minimal()
ggsave(here(dir_figs, "biotic_check.png"))

ggplot(d_s_fire |> filter(cd_group_b10_pdsisumn10 == "bdf" | cd_group_b10_pdsisumn10 == "bf")) +
  geom_density(aes(x = biotic_relaxedforestnorm_5_yrs_prior_sum_yot, color = cd_group_b10_pdsisumn10)) +
  theme_minimal() +
  facet_wrap(~nfg_factor)
ggsave(here(dir_figs, "biotic_check_ft.png"))


ggplot(d_s_fire |> filter(cd_group_b10_pdsisumn10 == "bdf" | cd_group_b10_pdsisumn10 == "df")) +
  geom_density(aes(x = pdsi_annual_5_yrs_prior_sum_yot, color = cd_group_b10_pdsisumn10)) +
  theme_minimal() 
ggsave(here(dir_figs, "drought_check.png"))


ggplot(d_s_fire) +
  geom_density(aes(x = pdsi_annual_5_yrs_prior_sum_yot, color = cd_group_b10_pdsisumn10)) +
  theme_minimal() 
ggsave(here(dir_figs, "drought_check_all_groups.png"))



ggplot(d_s_fire |> filter(cd_group_b10_pdsisumn10 == "bdf" | cd_group_b10_pdsisumn10 == "df" )) +
  geom_density(aes(x = pdsi_annual_5_yrs_prior_sum_yot, color = cd_group_b10_pdsisumn10)) +
  theme_minimal() +
  facet_wrap(~nfg_factor)
ggsave(here(dir_figs, "drought_check_ft.png"))



ggplot(d_s_fire) +
  geom_point(aes(x = pdsi_annual_5_yrs_prior_sum_yot, y = biotic_relaxedforestnorm_5_yrs_prior_sum_yot), alpha = 0.2) +
  geom_smooth(aes(x = pdsi_annual_5_yrs_prior_sum_yot, y = biotic_relaxedforestnorm_5_yrs_prior_sum_yot), method = "lm") +
  theme_minimal()
ggsave(here(dir_figs, "biotic_vs_drought.png"))


ggplot(d_s_fire |> filter(biotic_relaxedforestnorm_5_yrs_prior_sum_yot >= 5)) +
  geom_point(aes(x = pdsi_annual_5_yrs_prior_sum_yot, y = biotic_relaxedforestnorm_5_yrs_prior_sum_yot), alpha = 0.2) +
  geom_smooth(aes(x = pdsi_annual_5_yrs_prior_sum_yot, y = biotic_relaxedforestnorm_5_yrs_prior_sum_yot), method = "lm") +
  theme_minimal()
ggsave(here(dir_figs, "biotic_vs_drought_greater_than_5b.png"))



```

# Extend covariate distributions to add quant

```{r}

library(ggplot2)
library(dplyr)
library(purrr)
library(patchwork)
library(rlang)
library(forcats)
library(readr)
library(stringr)
library(tidyr)

dir_density_assessment <- here(dir_figs, "density_assessment")
dir_ensure(dir_density_assessment)

safe_filename <- function(x) {
  x |>
    stringr::str_replace_all("[^[:alnum:]_ -]", "_") |>
    stringr::str_replace_all("[ /]+", "_") |>
    stringr::str_replace_all("_+", "_") |>
    stringr::str_trim(side = "both")
}
compute_density_overlap_metrics <- function(df,
                                            x_var,
                                            group_var,
                                            n_grid = 512,
                                            min_n_group = 20) {
  
  d <- tibble::tibble(
    x = df[[x_var]],
    group = df[[group_var]]
  ) |>
    dplyr::filter(
      !is.na(x),
      !is.na(group)
    ) |>
    dplyr::mutate(group = as.character(group))
  
  if (nrow(d) == 0) {
    return(tibble::tibble(
      n = 0L,
      n_groups = 0L,
      groups_used = NA_character_,
      multigroup_overlap = NA_real_,
      min_pairwise_overlap = NA_real_,
      mean_pairwise_overlap = NA_real_
    ))
  }
  
  group_counts <- d |>
    dplyr::count(group, name = "n_group") |>
    dplyr::filter(n_group >= min_n_group)
  
  d <- d |>
    dplyr::semi_join(group_counts, by = "group")
  
  if (nrow(d) == 0) {
    return(tibble::tibble(
      n = 0L,
      n_groups = 0L,
      groups_used = NA_character_,
      multigroup_overlap = NA_real_,
      min_pairwise_overlap = NA_real_,
      mean_pairwise_overlap = NA_real_
    ))
  }
  
  groups <- sort(unique(d$group))
  
  if (length(groups) < 2) {
    return(tibble::tibble(
      n = nrow(d),
      n_groups = length(groups),
      groups_used = paste(groups, collapse = ", "),
      multigroup_overlap = NA_real_,
      min_pairwise_overlap = NA_real_,
      mean_pairwise_overlap = NA_real_
    ))
  }
  
  x_range <- range(d$x, na.rm = TRUE)
  
  if (!is.finite(x_range[1]) || !is.finite(x_range[2]) || x_range[1] == x_range[2]) {
    return(tibble::tibble(
      n = nrow(d),
      n_groups = length(groups),
      groups_used = paste(groups, collapse = ", "),
      multigroup_overlap = NA_real_,
      min_pairwise_overlap = NA_real_,
      mean_pairwise_overlap = NA_real_
    ))
  }
  
  dens_list <- purrr::map(
    groups,
    function(gr) {
      xg <- d |>
        dplyr::filter(group == gr) |>
        dplyr::pull(x)
      
      stats::density(
        x = xg,
        from = x_range[1],
        to = x_range[2],
        n = n_grid,
        na.rm = TRUE
      )
    }
  )
  
  x_grid <- dens_list[[1]]$x
  dx <- x_grid[2] - x_grid[1]
  
  y_mat <- do.call(
    cbind,
    purrr::map(dens_list, \(dd) dd$y)
  )
  
  multigroup_overlap <- sum(apply(y_mat, 1, min)) * dx
  
  pair_idx <- utils::combn(seq_along(groups), 2, simplify = FALSE)
  
  pairwise_tbl <- purrr::map_dfr(
    pair_idx,
    function(idx) {
      ovl <- sum(pmin(y_mat[, idx[1]], y_mat[, idx[2]])) * dx
      
      tibble::tibble(
        group_1 = groups[idx[1]],
        group_2 = groups[idx[2]],
        pairwise_overlap = ovl
      )
    }
  )
  
  tibble::tibble(
    n = nrow(d),
    n_groups = length(groups),
    groups_used = paste(groups, collapse = ", "),
    multigroup_overlap = multigroup_overlap,
    min_pairwise_overlap = min(pairwise_tbl$pairwise_overlap, na.rm = TRUE),
    mean_pairwise_overlap = mean(pairwise_tbl$pairwise_overlap, na.rm = TRUE)
  )
}

plot_density_triplets <- function(df,
                                  x_vars = c(
                                    "aet",
                                    "def",
                                    "gam_rap_tree_pre6_fit",
                                    "srtm",
                                    "chili",
                                    "tpi"
                                  ),
                                  group_vars = c(
                                    "cd_group_b10_pdsisumn10",
                                    "cd_group_b10_pdsin3t1",
                                    "cd_group_b10_pdsin4t1"
                                  ),
                                  n_grid = 512,
                                  min_n_group = 20,
                                  plot_title_prefix = NULL) {
  
  metrics_tbl <- tidyr::expand_grid(
    x_var = x_vars,
    group_var = group_vars
  ) |>
    dplyr::mutate(
      metrics = purrr::map2(
        x_var,
        group_var,
        ~ compute_density_overlap_metrics(
          df = df,
          x_var = .x,
          group_var = .y,
          n_grid = n_grid,
          min_n_group = min_n_group
        )
      )
    ) |>
    tidyr::unnest(metrics)
  
  make_one_plot <- function(x_var, group_var) {
    
    metrics_row <- metrics_tbl |>
      dplyr::filter(
        x_var == !!x_var,
        group_var == !!group_var
      )
    
    label_txt <- paste0(
      "OVL(all): ",
      ifelse(is.na(metrics_row$multigroup_overlap), "NA",
             sprintf("%.3f", metrics_row$multigroup_overlap)),
      "\nMin pair: ",
      ifelse(is.na(metrics_row$min_pairwise_overlap), "NA",
             sprintf("%.3f", metrics_row$min_pairwise_overlap)),
      "\nMean pair: ",
      ifelse(is.na(metrics_row$mean_pairwise_overlap), "NA",
             sprintf("%.3f", metrics_row$mean_pairwise_overlap)),
      "\nN: ",
      metrics_row$n,
      "\nGroups: ",
      metrics_row$n_groups
    )
    
    d_plot <- tibble::tibble(
      x = df[[x_var]],
      group = df[[group_var]]
    ) |>
      dplyr::filter(
        !is.na(x),
        !is.na(group)
      ) |>
      dplyr::mutate(group = as.character(group)) |>
      dplyr::add_count(group, name = "n_group") |>
      dplyr::filter(n_group >= min_n_group)
    
    ggplot(
      data = d_plot,
      aes(
        x = x,
        color = group
      )
    ) +
      geom_density(linewidth = 0.7) +
      annotate(
        geom = "text",
        x = Inf,
        y = Inf,
        label = label_txt,
        hjust = 1.05,
        vjust = 1.1,
        size = 3.2
      ) +
      theme_minimal() +
      labs(
        x = x_var,
        y = "Density",
        color = group_var,
        title = group_var
      )
  }
  
  plot_list <- purrr::map(
    x_vars,
    function(x_var) {
      purrr::map(
        group_vars,
        ~ make_one_plot(x_var = x_var, group_var = .x)
      ) |>
        patchwork::wrap_plots(nrow = 1) +
        patchwork::plot_annotation(
          title = if (is.null(plot_title_prefix)) {
            x_var
          } else {
            paste0(plot_title_prefix, " - ", x_var)
          }
        )
    }
  ) |>
    rlang::set_names(x_vars)
  
  list(
    plots = plot_list,
    metrics = metrics_tbl
  )
}

save_density_plot_set <- function(plot_list, out_dir, prefix) {
  dir_ensure(out_dir)
  
  purrr::iwalk(
    plot_list,
    ~ ggplot2::ggsave(
      filename = file.path(out_dir, paste0(prefix, .y, "_density_triplet.png")),
      plot = .x,
      width = 12,
      height = 4,
      dpi = 300
    )
  )
}

save_density_metrics_table <- function(metrics_tbl, out_dir, filename) {
  dir_ensure(out_dir)
  
  readr::write_csv(
    metrics_tbl,
    file.path(out_dir, filename)
  )
}


# all-data plots + summary table
density_all <- plot_density_triplets(
  df = d_s,
  plot_title_prefix = "All",
  min_n_group = 20
)

save_density_plot_set(
  plot_list = density_all$plots,
  out_dir = dir_density_assessment,
  prefix = "all_"
)

save_density_metrics_table(
  metrics_tbl = density_all$metrics |>
    dplyr::mutate(ecoregion_code_name = "All") |>
    dplyr::relocate(ecoregion_code_name),
  out_dir = dir_density_assessment,
  filename = "all_density_overlap_metrics.csv"
)

ecoregions <- d_s |>
  dplyr::distinct(ecoregion_code_name) |>
  dplyr::pull(ecoregion_code_name) |>
  stats::na.omit() |>
  as.character() |>
  sort()

purrr::walk(
  ecoregions,
  function(ecoregion_i) {
    
    ecoregion_dir <- file.path(
      dir_density_assessment,
      safe_filename(ecoregion_i)
    )
    dir_ensure(ecoregion_dir)
    
    d_eco <- d_s |>
      dplyr::filter(ecoregion_code_name == ecoregion_i)
    
    density_eco <- plot_density_triplets(
      df = d_eco,
      plot_title_prefix = ecoregion_i,
      min_n_group = 20
    )
    
    save_density_plot_set(
      plot_list = density_eco$plots,
      out_dir = ecoregion_dir,
      prefix = paste0(safe_filename(ecoregion_i), "_")
    )
    
    save_density_metrics_table(
      metrics_tbl = density_eco$metrics |>
        dplyr::mutate(ecoregion_code_name = ecoregion_i) |>
        dplyr::relocate(ecoregion_code_name),
      out_dir = ecoregion_dir,
      filename = paste0(safe_filename(ecoregion_i), "_density_overlap_metrics.csv")
    )
  }
)

density_metrics_by_ecoregion <- purrr::map_dfr(
  ecoregions,
  function(ecoregion_i) {
    d_s |>
      dplyr::filter(ecoregion_code_name == ecoregion_i) |>
      plot_density_triplets(
        plot_title_prefix = ecoregion_i,
        min_n_group = 20
      ) |>
      purrr::pluck("metrics") |>
      dplyr::mutate(ecoregion_code_name = ecoregion_i) |>
      dplyr::relocate(ecoregion_code_name)
  }
) |>
  dplyr::mutate(
    x_var = factor(x_var, levels = c("aet", "def", "chili", "tpi", "srtm", "gam_rap_tree_pre6_fit"))
  )



readr::write_csv(
  density_metrics_by_ecoregion,
  file.path(dir_density_assessment, "density_overlap_metrics_by_ecoregion.csv")
)










forest_types <- d_s |>
  dplyr::distinct() |>
  dplyr::pull(nfg_factor) |>
  stats::na.omit() |>
  as.character() |>
  sort()

purrr::walk(
  forest_types,
  function(ft_i) {
    
    ft_dir <- file.path(
      dir_density_assessment,
      safe_filename(ft_i)
    )
    dir_ensure(ft_dir)
    
    d_ft <- d_s |>
      dplyr::filter(nfg_factor == ft_i)
    
    density_ft <- plot_density_triplets(
      df = d_ft,
      plot_title_prefix = ft_i,
      min_n_group = 20
    )
    
    save_density_plot_set(
      plot_list = density_ft$plots,
      out_dir = ft_dir,
      prefix = paste0(safe_filename(ft_i), "_")
    )
    
    save_density_metrics_table(
      metrics_tbl = density_ft$metrics |>
        dplyr::mutate(nfg_factor = ft_i) |>
        dplyr::relocate(nfg_factor),
      out_dir = ft_dir,
      filename = paste0(safe_filename(ft_i), "_density_overlap_metrics.csv")
    )
  }
)

density_metrics_by_ft <- purrr::map_dfr(
  forest_types,
  function(ft_i) {
    d_s |>
      dplyr::filter(nfg_factor == ft_i) |>
      plot_density_triplets(
        plot_title_prefix = ft_i,
        min_n_group = 20
      ) |>
      purrr::pluck("metrics") |>
      dplyr::mutate(ft_code_name = ft_i) |>
      dplyr::relocate(ft_code_name)
  }
) |>
  dplyr::mutate(
    x_var = factor(x_var, levels = c("aet", "def", "chili", "tpi", "srtm", "gam_rap_tree_pre6_fit"))
  )



readr::write_csv(
  density_metrics_by_ft,
  file.path(dir_density_assessment, "density_overlap_metrics_by_forest_types.csv")
)

```

# assess density overlap metrics

```{r}

ggplot(data = density_metrics_by_ecoregion) +
  geom_density(aes(x = multigroup_overlap, color = x_var)) +
  facet_wrap(~group_var)

ggplot(data = density_metrics_by_ecoregion) +
  geom_density(aes(x = multigroup_overlap, color = ecoregion_code_name)) +
  facet_wrap(~group_var)


ggplot(data = density_metrics_by_ecoregion |>
  dplyr::filter(!stringr::str_starts(x_var, "tt_")) |>
    dplyr::filter(group_var == "cd_group_b10_pdsin3t1")) +
  geom_col(aes(x = x_var, y = multigroup_overlap)) +
  facet_wrap(~ecoregion_code_name) +
  labs(title = "Multigroup overlap for all ecoregions for key covariates: cd_group_b10_pdsin3t1") +
  theme_minimal() +
  theme(
    axis.text.x = element_text(
      angle = 45,
      hjust = 1,
      vjust = 1
    )
  ) +
  geom_hline(aes(yintercept = 0.3), color = "red") +
  geom_hline(aes(yintercept = 0.6))

ggsave(filename = here(dir_density_assessment, "multigroup_overlap.png"))

```

# Assess standardized mean difference (SMD) using cobalt

```{r}


library(cobalt)
   
cobalt::bal.tab(x = cd_group_b10_pdsisumn10 ~ aet + def + srtm + tpi + chili + gam_rap_tree_pre6_fit + tt_change_aet + tt_change_def, 
                data = d_s |> filter(fire == 1 & !is.na(cd_group_b10_pdsisumn10)))

png(here(dir_figs, glue("love_plot_fire_cd_group_b10_pdsisumn10.png")))
cobalt::love.plot(x = cd_group_b10_pdsisumn10 ~ aet + def + srtm + tpi + chili + gam_rap_tree_pre6_fit + tt_change_aet + tt_change_def, 
                data = d_s |> filter(fire == 1 & !is.na(cd_group_b10_pdsisumn10)),
                stats = c("mean.diffs", "ks.statistics"),
                stars = "raw",
                position = "bottom",
                threshold = c(m = .1, ks = .05),
                limits = list(c(-1, 1), c(0, 0.6)),
                title = glue("Covariate balance love plot for CD treatment: cd_group_b10_pdsisumn10"))
dev.off()



library(cobalt)
library(dplyr)
library(glue)

d_plot <- d_s |>
  filter(fire == 1, !is.na(cd_group_b10_pdsisumn10)) |>
  mutate(
    cd_group_b10_pdsisumn10 = factor(cd_group_b10_pdsisumn10)
  )

bal_out <- cobalt::bal.tab(
  cd_group_b10_pdsisumn10 ~ aet + def + srtm + tpi + chili +
    gam_rap_tree_pre6_fit + tt_change_aet + tt_change_def,
  data = d_plot,
  pairwise = TRUE,
  stats = c("mean.diffs", "ks.statistics")
)

png(
  here::here(dir_figs, "love_plot_fire_cd_group_b10_pdsisumn10_smd.png"),
  width = 1400, height = 1000, res = 150
)

cobalt::love.plot(
  bal_out,
  stat = "mean.diffs",
  abs = TRUE,
  agg.fun = NULL,
  facet = "contrast",   # <-- separates pairs
  thresholds = c(m = .1)
)

dev.off()

```

```{r}


# GAM METRIC DISTRIBUTIONS
# 
# plot_density_by_group_with_n <- function(df, x_var, group_var, x_label = NULL) {
# 
#   x_quo <- enquo(x_var)
#   g_quo <- enquo(group_var)
# 
#   x_name <- as_name(x_quo)
#   g_name <- as_name(g_quo)
# 
#   # Filter missing groups
#   d_plot <- df |>
#     filter(!is.na(!!g_quo))
# 
#   # Compute N per group
#   group_n <- d_plot |>
#     count(!!g_quo, name = "N") |>
#     mutate(label = paste0(!!g_quo, " (N=", N, ")"))
# 
#   lab_vec <- setNames(group_n$label, group_n[[g_name]])
# 
#   ggplot(d_plot, aes(x = !!x_quo, color = !!g_quo)) +
#     geom_density() +
#     scale_color_discrete(labels = lab_vec) +
#     labs(
#       x = ifelse(is.null(x_label), x_name, x_label),
#       color = g_name
#     ) +
#     theme_minimal()
# }
# 
# plot_density_by_group_with_n(
#   d_s,
#   gam_rap_tree_diff_pre6_to_post_min,
#   cd_group_b10_pdsin4t1,
#   x_label = "RAP total loss"
# )
# 
# plot_density_by_group_with_n(
#   d_s,
#   gam_rap_tree_diff_pre6_to_post_min,
#   cd_group_b10_pdsin3t1,
#   x_label = "RAP total loss"
# )
# 
# plot_density_by_group_with_n(
#   d_s,
#   gam_rap_tree_diff_pre6_to_post_min,
#   cd_group_b10_pdsin4t3,
#   x_label = "RAP total loss"
# )
# 
# 
# plot_density_by_group_with_n(
#   d_s,
#   gam_rap_tree_diff_pre6_to_post_min,
#   cd_group_b10_pdsisumn10,
#   x_label = "RAP total loss"
# )


# Boxplots
ggplot(
  d_s |>
    filter(!is.na(cd_group_b10_pdsin4t1)) |>
    add_count(cd_group_b10_pdsin4t1, name = "n_group"),
  aes(
    x = paste0(cd_group_b10_pdsin4t1, " (N=", n_group, ")"),
    y = gam_rap_tree_diff_pre6_to_post_min
  )
) +
  geom_boxplot() +
  labs(y = "RAP total loss") +
  theme_minimal()
ggsave(here(dir_figs, "boxplot_raploss_cd_group_b10_pdsin4t1.png"))


ggplot(
  d_s |>
    filter(!is.na(cd_group_b10_pdsin4t3)) |>
    add_count(cd_group_b10_pdsin4t3, name = "n_group"),
  aes(
    x = paste0(cd_group_b10_pdsin4t3, " (N=", n_group, ")"),
    y = gam_rap_tree_diff_pre6_to_post_min
  )
) +
  geom_boxplot() +
  labs(y = "RAP total loss") +
  theme_minimal()
ggsave(here(dir_figs, "boxplot_raploss_cd_group_b10_pdsin4t3.png"))


ggplot(
  d_s |>
    filter(!is.na(cd_group_b10_pdsisumn10)) |>
    add_count(cd_group_b10_pdsisumn10, name = "n_group"),
  aes(
    x = paste0(cd_group_b10_pdsisumn10, " (N=", n_group, ")"),
    y = gam_rap_tree_diff_pre6_to_post_min
  )
) +
  geom_boxplot() +
  labs(y = "RAP total loss") +
  theme_minimal()
ggsave(here(dir_figs, "boxplot_raploss_cd_group_b10_pdsisumn10.png"))




ggplot(
  d_s |>
    filter(!is.na(cd_group_b10_pdsin4t3)) |>
    add_count(cd_group_b10_pdsin4t3, name = "n_group"),
  aes(
    x = paste0(cd_group_b10_pdsin4t3, " (N=", n_group, ")"),
    y = gam_rap_tree_slope_post_min_to_p15
  )
) +
  geom_boxplot() +
  labs(y = "RAP min to 15 yr recovery slope") +
  theme_minimal()
ggsave(here(dir_figs, "boxplot_rapgamslope_cd_group_b10_pdsin4t3.png"))




ggplot(
  d_s |>
    filter(!is.na(cd_group_b10_pdsin4t3)) |>
    add_count(cd_group_b10_pdsin4t3, name = "n_group"),
  aes(
    x = paste0(cd_group_b10_pdsin4t3, " (N=", n_group, ")"),
    y = raw_post_fire_rap_tree_slope_3_15_yrs
  )
) +
  geom_boxplot() +
  labs(y = "RAP 3-15 yr recovery slope") +
  theme_minimal()
ggsave(here(dir_figs, "boxplot_raprawslope_cd_group_b10_pdsin4t3.png"))




# 
# 
# 
# bin_cd_and_boxplot <- function(df, cd_nm = cd_nm, outcome = outcome, fire = fire, n_bins = 10) {
# 
#   cd_quo <- rlang::enquo(cd_nm)
#   out_quo <- rlang::enquo(outcome)
#   fire_quo <- rlang::enquo(fire)
# 
#   cd_name <- rlang::as_name(cd_quo)
# 
#   x <- df[[cd_name]]
# 
#   if (inherits(x, "integer")) {
# 
#     df2 <- df |>
#       dplyr::mutate(
#         bin_cd = ifelse(!!fire_quo == 0, NA, as.character(!!cd_quo))
#       )
# 
#   } else {
# 
#     rng <- range(x, na.rm = TRUE, finite = TRUE)
# 
#     # if there isn't a usable range, bins can't be formed
#     if (!is.finite(rng[1]) || !is.finite(rng[2]) || rng[1] == rng[2]) {
#       df2 <- df |>
#         dplyr::mutate(bin_cd = NA_character_)
#     } else {
# 
#       breaks <- seq(from = rng[1], to = rng[2], length.out = n_bins + 1)
# 
#       labels <- paste0(
#         round(head(breaks, -1), 1),
#         "–",
#         round(tail(breaks, -1), 1)
#       )
# 
#       df2 <- df |>
#         dplyr::mutate(
#           bin_cd = cut(
#             !!cd_quo,
#             breaks = breaks,
#             include.lowest = TRUE,
#             labels = labels
#           ),
#           bin_cd = ifelse(!!fire_quo == 0, NA, as.character(bin_cd))
#         )
#     }
#   }
# 
#   df2 <- df2 |> dplyr::mutate(bin_cd = factor(bin_cd))
# 
#   p <- ggplot2::ggplot(df2 |> dplyr::filter(!is.na(bin_cd)),
#               ggplot2::aes(x = bin_cd, y = !!out_quo)) +
#     ggplot2::geom_boxplot(outlier.alpha = 0.2) +
#     ggplot2::labs(
#       x = paste0(cd_name, " bins"),
#       y = rlang::as_name(out_quo)
#     ) +
#     ggplot2::theme_minimal() +
#     ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 30, hjust = 1))
# 
#   list(data = df2, plot = p)
# }
# 
# 
# res <- bin_cd_and_boxplot(d_s, biotic_relaxedforestnorm_5_yrs_prior_sum_yot, gam_rap_tree_diff_pre6_to_post_min)
# res$plot
# 
# res <- bin_cd_and_boxplot(d_s, pdsi_annual_5_yrs_prior_threshold_n3_yot, gam_rap_tree_diff_pre6_to_post_min)
# res$plot
# 
# 
# res <- bin_cd_and_boxplot(d_s, pdsi_annual_5_yrs_prior_threshold_n4_yot, gam_rap_tree_diff_pre6_to_post_min)
# res$plot
# 
# res <- bin_cd_and_boxplot(d_s, pdsi_annual_5_yrs_prior_sum_yot, gam_rap_tree_diff_pre6_to_post_min)
# res$plot
# 
# res <- bin_cd_and_boxplot(d_s,  hd_fingerprint_5_yrs_prior_threshold_6_yot, gam_rap_tree_diff_pre6_to_post_min)
# res$plot
# 



```

# Examine multi-covariate space

```{r}

plot_order <- c("f", "df", "bf", "bdf")

d_plot <- d_s_fire |>
  dplyr::mutate(
    cd_group_b10_pdsisumn10 = factor(
      cd_group_b10_pdsisumn10,
      levels = plot_order
    )
  ) |>
  dplyr::arrange(cd_group_b10_pdsisumn10)


covars <- c(
  "gam_rap_tree_pre6_fit",
  "aet",
  "srtm",
  "def"
)

density_line_diag <- function(data, mapping, ...) {
  ggplot(data = data, mapping = mapping) +
    geom_density(
      aes(color = cd_group_b10_pdsisumn10),
      fill = NA,
      linewidth = 0.8,
      alpha = 1
    )
}

d_plot |>
  dplyr::select(
    cd_group_b10_pdsisumn10,
    dplyr::all_of(covars)
  ) |>
  GGally::ggpairs(
    aes(color = cd_group_b10_pdsisumn10, alpha = 0.25),
    columns = 2:(length(covars) + 1),
    upper = list(continuous = GGally::wrap("cor", size = 3)),
    lower = list(continuous = GGally::wrap("points", alpha = 0.15, size = 0.4)),
    diag  = list(continuous = density_line_diag)
  ) + theme_minimal()

ggsave(here(dir_figs, "ggpairs.png"),
            units = "px",
            width = 3000,
            height = 3000)







pca_dat <- d_plot |>
  dplyr::select(
    cd_group_b10_pdsisumn10,
    dplyr::all_of(covars)
  ) |>
  tidyr::drop_na()

pca_fit <- prcomp(
  pca_dat |> dplyr::select(dplyr::all_of(covars)),
  center = TRUE,
  scale. = TRUE
)

pca_scores <- pca_dat |>
  dplyr::bind_cols(
    as.data.frame(pca_fit$x[, 1:4])
  )

var_explained <- summary(pca_fit)$importance["Proportion of Variance", ]
var_explained


ggplot(
  pca_scores |>
    dplyr::arrange(cd_group_b10_pdsisumn10),
  aes(
    x = PC1,
    y = PC2,
    color = cd_group_b10_pdsisumn10
  )
) +
  geom_point(alpha = 0.2, size = 0.8) +
  stat_ellipse(linewidth = 0.8, alpha = 0.8) +
  labs(
    x = paste0("PC1 (", round(100 * var_explained[1], 1), "%)"),
    y = paste0("PC2 (", round(100 * var_explained[2], 1), "%)"),
    color = "Group"
  ) +
  theme_bw()

ggsave(here(dir_figs, "pca.png"))



# Functionalized

make_covariate_space_figures <- function(
    data,
    group_col,
    covars,
    plot_order = NULL,
    fig_dir = NULL,
    file_prefix = "covariate_space",
    ggpairs_width = 3000,
    ggpairs_height = 3000,
    ggpairs_units = "px",
    pca_width = 7,
    pca_height = 7,
    pca_units = "in",
    point_alpha = 0.15,
    point_size = 0.4,
    density_linewidth = 0.8,
    ellipse = TRUE
) {

  stopifnot(group_col %in% names(data))
  stopifnot(all(covars %in% names(data)))

  if (is.null(plot_order)) {
    plot_order <- data |>
      dplyr::pull(dplyr::all_of(group_col)) |>
      unique() |>
      sort()
  }

  d_plot <- data |>
    dplyr::mutate(
      .plot_group = factor(.data[[group_col]], levels = plot_order)
    ) |>
    dplyr::arrange(.plot_group)

  density_line_diag <- function(data, mapping, ...) {
    ggplot2::ggplot(data = data, mapping = mapping) +
      ggplot2::geom_density(
        fill = NA,
        linewidth = density_linewidth,
        alpha = 1
      )
  }

  ggpairs_plot <- d_plot |>
    dplyr::select(
      .plot_group,
      dplyr::all_of(covars)
    ) |>
    GGally::ggpairs(
      mapping = ggplot2::aes(color = .plot_group),
      columns = 2:(length(covars) + 1),
      upper = list(
        continuous = GGally::wrap("cor", size = 3)
      ),
      lower = list(
        continuous = GGally::wrap(
          "points",
          alpha = point_alpha,
          size = point_size
        )
      ),
      diag = list(
        continuous = density_line_diag
      )
    ) +
    ggplot2::labs(color = group_col) +
    ggplot2::theme_minimal()

  pca_dat <- d_plot |>
    dplyr::select(
      .plot_group,
      dplyr::all_of(covars)
    ) |>
    tidyr::drop_na()

  pca_covars <- covars[
    vapply(
      pca_dat[covars],
      function(x) stats::sd(x, na.rm = TRUE) > 0,
      logical(1)
    )
  ]

  if (length(pca_covars) < 2) {
    pca_fit <- NULL
    pca_scores <- NULL
    var_explained <- NULL
    pca_plot <- NULL

    warning(
      "PCA was not run because fewer than two covariates had non-zero variance."
    )

  } else {

    pca_fit <- stats::prcomp(
      pca_dat |> dplyr::select(dplyr::all_of(pca_covars)),
      center = TRUE,
      scale. = TRUE
    )

    n_pcs <- min(4, ncol(pca_fit$x))

    pca_scores <- pca_dat |>
      dplyr::bind_cols(
        as.data.frame(pca_fit$x[, 1:n_pcs, drop = FALSE])
      )

    var_explained <- summary(pca_fit)$importance["Proportion of Variance", ]

    pca_plot <- ggplot2::ggplot(
      pca_scores |>
        dplyr::arrange(.plot_group),
      ggplot2::aes(
        x = PC1,
        y = PC2,
        color = .plot_group
      )
    ) +
      ggplot2::geom_point(alpha = 0.2, size = 0.8) +
      ggplot2::labs(
        x = paste0("PC1 (", round(100 * var_explained[1], 1), "%)"),
        y = paste0("PC2 (", round(100 * var_explained[2], 1), "%)"),
        color = group_col
      ) +
      ggplot2::theme_bw()

    if (ellipse) {
      pca_plot <- pca_plot +
        ggplot2::stat_ellipse(linewidth = 0.8, alpha = 0.8)
    }
  }

  if (!is.null(fig_dir)) {
    if (!dir.exists(fig_dir)) {
      dir.create(fig_dir, recursive = TRUE)
    }

    ggplot2::ggsave(
      filename = file.path(fig_dir, paste0(file_prefix, "_ggpairs.png")),
      plot = ggpairs_plot,
      units = ggpairs_units,
      width = ggpairs_width,
      height = ggpairs_height
    )

    if (!is.null(pca_plot)) {
      ggplot2::ggsave(
        filename = file.path(fig_dir, paste0(file_prefix, "_pca.png")),
        plot = pca_plot,
        units = pca_units,
        width = pca_width,
        height = pca_height
      )
    }
  }

  list(
    ggpairs = ggpairs_plot,
    pca = pca_plot,
    pca_fit = pca_fit,
    pca_scores = pca_scores,
    var_explained = var_explained,
    data = d_plot,
    covars_used_for_pca = pca_covars
  )
}

make_covariate_space_figures_by_group <- function(
    data,
    split_col,
    group_col,
    covars,
    plot_order = NULL,
    fig_dir = NULL,
    file_prefix = "covariate_space",
    ...
) {

  stopifnot(split_col %in% names(data))

  split_vals <- data |>
    dplyr::pull(dplyr::all_of(split_col)) |>
    unique() |>
    sort()

  out <- purrr::map(
    split_vals,
    function(split_val) {

      d_sub <- data |>
        dplyr::filter(.data[[split_col]] == split_val)

      safe_split_val <- split_val |>
        as.character() |>
        stringr::str_replace_all("[^A-Za-z0-9_\\-]+", "_")

      make_covariate_space_figures(
        data = d_sub,
        group_col = group_col,
        covars = covars,
        plot_order = plot_order,
        fig_dir = fig_dir,
        file_prefix = paste0(file_prefix, "_", split_col, "_", safe_split_val),
        ...
      )
    }
  )

  names(out) <- split_vals

  out
}


# Use
plot_order <- c("f", "df", "bf", "bdf")

covars <- c(
  "gam_rap_tree_pre6_fit",
  "aet",
  "srtm",
  "def"
)

figs <- make_covariate_space_figures(
  data = d_s_fire,
  group_col = "cd_group_b10_pdsisumn10",
  covars = covars,
  plot_order = plot_order,
  fig_dir = here::here(dir_figs),
  file_prefix = "cd_group_b10_pdsisumn10"
)



figs_by_ecoregion <- make_covariate_space_figures_by_group(
  data = d_s_fire,
  split_col = "ecoregion",
  group_col = "cd_group_b10_pdsisumn10",
  covars = covars,
  plot_order = plot_order,
  fig_dir = here::here(dir_figs),
  file_prefix = "cd_group_b10_pdsisumn10"
)

```

# VCF vs RAP plots

Both metrics & raw data

Takeaways: we find that RAP aggregated to 250m and VCF have RMSE of around 12.4. This remains similar across subgroups of data, although there is a slightly lower error in prefire and no-fire environments than postfire (\~12 vs \~13). The postfire error climbs slightly with vcf above 50, to \~14.

In terms of estimating resilience metrics, there is higher agreement between VCF and RAP when estimating pre-fire cover using an average window or GAM (as opposed to using single raw values). In estimating total loss, recovery, slope, etc, averaged windows provide lower RMSE between the two datasets, likely due to the fact that the GAMs are identifying different minimum values post fire between the two datasets.

```{r}

plot_lm_r2 <- function(df, x, y, title, xlim = NULL, ylim = NULL) {
  p <- ggplot(df, aes(x = {{ x }}, y = {{ y }})) +
    geom_point(alpha = 0.1) +
    geom_smooth(method = "lm", alpha = 0.1) +
    ggpmisc::stat_poly_eq(
      formula = y ~ x,
      aes(label = after_stat(rr.label)),
      parse = TRUE,
      label.x = "right",
      label.y = "bottom"
    ) +
    labs(title = title)

  if (!is.null(xlim)) p <- p + xlim(xlim)
  if (!is.null(ylim)) p <- p + ylim(ylim)

  p
}

extract_r2_rmse_table <- function(models) {
  tibble(
    model_name = names(models),
    model = models
  ) |>
    mutate(
      stats = purrr::map(model, broom::glance),
      rmse = purrr::map_dbl(model, \(m) {
        sqrt(mean(residuals(m)^2, na.rm = TRUE))
      })
    ) |>
    tidyr::unnest(stats) |>
    select(model_name, r.squared, rmse)
}

lm_safe <- function(formula, data) {
  vars <- all.vars(formula)
  d <- data |>
    dplyr::select(dplyr::all_of(vars)) |>
    dplyr::filter(dplyr::if_all(dplyr::everything(), is.finite))
  stats::lm(formula, data = d)
}

# 
# p_loss <- plot_lm_r2(
#   d_s,
#   gam_rap_tree_diff_pre6_to_post_min,
#   gam_vcf_tree_diff_pre6_to_post_min,
#   "VCF vs RAP total loss 250m"
# )
# 
# p_loss_raw <- plot_lm_r2(
#   d_s,
#   raw_post_fire_rap_tree_difference_n6_3_yr,
#   raw_post_fire_vcf_tree_difference_n6_3_yr,
#   "VCF vs RAP total loss 250m -6 to 3 yr raw windows"
# )
# 
# 
# 
# p_recov15 <- plot_lm_r2(
#   d_s,
#   gam_rap_tree_perc_recov_p15,
#   gam_vcf_tree_perc_recov_p15,
#   "VCF vs RAP recovery percent min-15yr 250m",
#   xlim = c(0, 100),
#   ylim = c(0, 100)
# )
# 
# p_raw_recov15 <- plot_lm_r2(
#   d_s,
#   raw_post_fire_rap_tree_perc_recovery_n6_3_15_yr,
#   raw_post_fire_vcf_tree_perc_recovery_n6_3_15_yr,
#   "VCF vs RAP raw recovery percent 3-15yr 250m",
#   xlim = c(0, 100),
#   ylim = c(0, 100)
# )
# 
# 
# p_slope15 <- plot_lm_r2(
#   d_s,
#   gam_rap_tree_slope_post_min_to_p15,
#   gam_vcf_tree_slope_post_min_to_p15,
#   "VCF vs RAP slope gam min to 15yr slope 250m"
# )
# 
# p_rawslope15 <- plot_lm_r2(
#   d_s,
#   raw_post_fire_rap_tree_slope_3_15_yrs,
#   raw_post_fire_vcf_tree_slope_3_15_yrs,
#   "VCF vs RAP slope raw 3 to 15yr slope 250m"
# )
# 
# 
# 
# p_pre6 <- plot_lm_r2(
#   d_s,
#   gam_rap_tree_pre6_fit,
#   gam_vcf_tree_pre6_fit,
#   "VCF vs RAP 6-yr prior fit 250m"
# )
# 
# p_pre6raw <- plot_lm_r2(
#   d_l |> filter(Time_to_Treatment == -6),
#   rap_tree,
#   vcf_tree,
#   "VCF vs RAP 6-yr-prior raw 250m"
# )
# 
# p_pre6raw_win <- plot_lm_r2(
#   d_s,
#   raw_avg_rap_tree_6_yr_pre_fire,
#   raw_avg_vcf_tree_6_yr_pre_fire,
#   "VCF vs RAP 6-yr-prior-window 250m"
# )
# 
# 
# 
# p_min <- plot_lm_r2(
#   d_s,
#   gam_rap_tree_post_min,
#   gam_vcf_tree_post_min,
#   "VCF vs RAP minimum fit 250m"
# )
# 
# p_3yr_raw <- plot_lm_r2(
#   d_s,
#   raw_avg_rap_tree_3_yr_post_fire,
#   raw_avg_vcf_tree_3_yr_post_fire,
#   "VCF vs RAP post-fire 3 yr raw"
# )
# 
# 
# 
# p_recov_tot <- plot_lm_r2(
#   d_s,
#   gam_rap_tree_diff_post_min_to_p15,
#   gam_vcf_tree_diff_post_min_to_p15,
#   "VCF vs RAP post-fire min to 15 gam"
# )
# 
# p_recov_tot_raw <- plot_lm_r2(
#   d_s,
#   raw_post_fire_rap_tree_increase_3_15_yr,
#   raw_post_fire_vcf_tree_increase_3_15_yr,
#   "VCF vs RAP post-fire 3 to 15 raw"
# )
# 
# 
# p_timetomin <- plot_lm_r2(
#   d_s,
#   gam_rap_tree_yrs_ref_to_post_min,
#   gam_vcf_tree_yrs_ref_to_post_min,
#   "VCF vs RAP time to minimum 250m"
# )
# 
# 
# 
# # Print
# p_loss
# p_loss_raw
# 
# p_recov15
# p_raw_recov15
# 
# p_slope15
# p_rawslope15
# 
# p_pre6
# p_pre6raw
# p_pre6raw_win
# 
# p_min
# p_3yr_raw
# 
# p_recov_tot
# p_recov_tot_raw
# 
# p_timetomin
# 
# ggplot(d_s) +
#   geom_bin2d(aes(x = gam_rap_tree_yrs_ref_to_post_min, y = gam_vcf_tree_yrs_ref_to_post_min))
# 




set.seed(1234)
d_s_2006 <- d_s |> dplyr::filter(burn_year > 2005)
d_l_2006 <- d_l |> dplyr::filter(burn_year > 2005)

models_xy <- list(
  loss_gam = lm_safe(gam_vcf_tree_diff_pre6_to_post_min ~ gam_rap_tree_diff_pre6_to_post_min, d_s_2006),
  loss_raw = lm_safe(raw_post_fire_vcf_tree_difference_n6_3_yr ~ raw_post_fire_rap_tree_difference_n6_3_yr, d_s_2006),

  recov15_gam = lm_safe(gam_vcf_tree_perc_recov_p15 ~ gam_rap_tree_perc_recov_p15, d_s_2006),
  recov15_raw = lm_safe(raw_post_fire_vcf_tree_perc_recovery_n6_3_15_yr ~ raw_post_fire_rap_tree_perc_recovery_n6_3_15_yr, d_s_2006),

  slope15_gam = lm_safe(gam_vcf_tree_slope_post_min_to_p15 ~ gam_rap_tree_slope_post_min_to_p15, d_s_2006),
  slope15_raw = lm_safe(raw_post_fire_vcf_tree_slope_3_15_yrs ~ raw_post_fire_rap_tree_slope_3_15_yrs, d_s_2006),

  pre6_fit_gam = lm_safe(gam_vcf_tree_pre6_fit ~ gam_rap_tree_pre6_fit, d_s_2006),
  pre6_raw_tminus6 = lm_safe(vcf_tree ~ rap_tree, d_l_2006 |> dplyr::filter(Time_to_Treatment == -6)),
  pre6_raw_window = lm_safe(raw_avg_vcf_tree_6_yr_pre_fire ~ raw_avg_rap_tree_6_yr_pre_fire, d_s_2006),

  postmin_gam = lm_safe(gam_vcf_tree_post_min ~ gam_rap_tree_post_min, d_s_2006),
  post3yr_raw = lm_safe(raw_avg_vcf_tree_3_yr_post_fire ~ raw_avg_rap_tree_3_yr_post_fire, d_s_2006),

  recov_tot_gam = lm_safe(gam_vcf_tree_diff_post_min_to_p15 ~ gam_rap_tree_diff_post_min_to_p15, d_s_2006),
  recov_tot_raw = lm_safe(raw_post_fire_vcf_tree_increase_3_15_yr ~ raw_post_fire_rap_tree_increase_3_15_yr, d_s_2006),

  time_to_min = lm_safe(gam_vcf_tree_yrs_ref_to_post_min ~ gam_rap_tree_yrs_ref_to_post_min, d_s_2006)
)

r2_rmse_table <- extract_r2_rmse_table(models_xy)

r2_rmse_table_g <- tableGrob(
  r2_rmse_table,
  theme = ttheme_minimal(
    core = list(bg_params = list(fill = "white", col = NA)),
    colhead = list(bg_params = list(fill = "white", col = NA))
  )
)

title1 <- textGrob(
  "VCF-RAP resilience metric \nR² and RMSE Comparisons \n(burn year >2005)",
  gp = gpar(fontsize = 16, fontface = "bold")
)

table_with_title1 <- arrangeGrob(
  r2_rmse_table_g,
  top = title1
)

ggsave(
  here(dir_figs, "rap_vs_vfc_metrics_table.png"),
  table_with_title1,
  width = 6,
  height = 7,
  bg = "white"
)




# Comparing raw data



ggplot(d_l) +
  geom_density(aes(x = rap_tree), color = "red") +
  geom_density(aes(x = vcf_tree), color = "blue") +
  labs(xlab = "Percent tree cover", caption = "RAP = red, VCF = blue")
ggsave(here(dir_figs, "raw_rap_vcf_density.png"))

set.seed(1234)
models <- list(
  all = lm(rap_tree ~ vcf_tree, d_l |> slice_sample(n = 1000000)),
  nofire = lm(rap_tree ~ vcf_tree, d_l |> filter(fire == 0) |> slice_sample(n = 1000000)),
  prefire = lm(rap_tree ~ vcf_tree, d_l |> filter(fire == 1 & treated == 0) |> slice_sample(n = 1000000)),
  postfire = lm(rap_tree ~ vcf_tree, d_l |> filter(fire == 1 & treated == 1) |> slice_sample(n = 1000000)),
  biotic_above10 = lm(rap_tree ~ vcf_tree, d_l |> filter(fire == 0 & biotic_6roll_max > 10) |> slice_sample(n = 1000000)),
  all_vcf_above50 = lm(rap_tree ~ vcf_tree, d_l |> filter(vcf_tree > 50) |> slice_sample(n = 1000000)),
  all_vcf_below50 = lm(rap_tree ~ vcf_tree, d_l |> filter(vcf_tree <= 50) |> slice_sample(n = 1000000)),
  postfire_vcf_above50 = lm(rap_tree ~ vcf_tree, d_l |> filter(vcf_tree > 50 & fire == 1 & treated == 1) |> slice_sample(n = 1000000)),
  postfire_vcf_below50 = lm(rap_tree ~ vcf_tree, d_l |> filter(vcf_tree <= 50 & fire == 1 & treated == 1) |> slice_sample(n = 1000000))
)


tbl <- tableGrob(
  extract_r2_rmse_table(models),
  theme = ttheme_minimal(
    core = list(bg_params = list(fill = "white", col = NA)),
    colhead = list(bg_params = list(fill = "white", col = NA))
  )
)

title <- textGrob(
  "Raw VCF-RAP data R² and RMSE Comparisons\n1 million samples per model",
  gp = gpar(fontsize = 16, fontface = "bold")
)

table_with_title <- arrangeGrob(
  tbl,
  top = title
)

ggsave(
  here(dir_figs, "rap_vs_vfc_raw_table.png"),
  table_with_title,
  width = 6,
  height = 4,
  bg = "white"
)





```

```{r}

dats_quick <- d_l |>
  filter(!is.na(cd_group)) |>
  slice_head(n = 100000) |>
  mutate(fire_frg = interaction(burn_sev, frg_reclass_desc),
         fire_pre_drought = interaction(burn_sev, pdsi_annual_5_yrs_prior_threshold),
         fire_pre_drought = interaction(burn_sev, pdsi_annual_5_yrs_after_threshold),
         biotic_cut = cut(biotic,
             breaks = c(0, 20, 40, 60, 80, 100),
             labels = c()
             )
         )

ggplot(dats_quick) +
  geom_smooth(method = 'gam', aes(x = Time_to_Treatment, y = rap_tree, color = cd_group)) +
  xlim(c(-10, 20)) +
  labs(title = "GAM fits sample")

ggplot(dats_quick) +
  geom_smooth(method = 'gam', aes(x = Time_to_Treatment, y = vcf_tree, color = cd_group)) +
  xlim(c(-10, 20)) +
  labs(title = "GAM fits sample")

ggplot(dats_quick |> filter(burn_sev != "Negligible")) +
  geom_smooth(method = 'gam', aes(x = Time_to_Treatment, y = rap_tree, color = frg_reclass_desc, lty = burn_sev)) +
  xlim(c(-10, 20)) +
  labs(title = "GAM fits sample")


```

# Disturbance over time

```{r}
d_s_f <- d_s |> filter(fire == 1)
d_s_nfire <- d_s |> filter(fire == 0)

d_s_f_summary <- d_s_f |>
  group_by(cd_group_b10_pdsin3t1, cd_group_b10_pdsisumn10, cd_group_b10_pdsin4t1, burn_year) |>
  summary(count = n())

ggplot(d_s_f, aes(x = burn_year, y = biotic_relaxedforestnorm_5_yrs_prior_sum_yot)) +
  geom_point(alpha = 0.1) +
  geom_smooth(method = "lm") +
  xlim(2000, 2020)

ggplot(d_s_f, aes(x = burn_year, y = pdsi_annual_5_yrs_prior_sum_yot)) +
  geom_point(alpha = 0.1) +
  geom_smooth(method = "lm") +
  xlim(2000, 2020)



ggplot(d_s_f |> filter(biotic_relaxedforestnorm_5_yrs_prior_sum_yot >= 10), aes(x = burn_year, y = biotic_relaxedforestnorm_5_yrs_prior_sum_yot)) +
  geom_point(alpha = 0.1) +
  geom_smooth(method = "lm") +
  xlim(2000, 2020) +
  labs(title = "Limited to >= 10")

ggplot(d_s_f |> filter(pdsi_annual_5_yrs_prior_sum_yot <= -10), aes(x = burn_year, y = pdsi_annual_5_yrs_prior_sum_yot)) +
  geom_point(alpha = 0.1) +
  geom_smooth(method = "lm") +
  xlim(2000, 2020) +
  labs(title = "Limited to <= -10")



ggplot(d_s_f |> filter(!is.na(cd_group_b10_pdsin3t1))) +
  geom_bar(aes(x = burn_year), position = "dodge") +
  facet_wrap(~ cd_group_b10_pdsin3t1, scales = "free_y"
             ) +
  #geom_smooth(method = "lm") +
  xlim(2000, 2020) +
  labs(title = "cd_group_b10_pdsin3t1")






ggplot(d_s_f |> filter(!is.na(cd_group_b10_pdsisumn10))) +
  geom_bar(aes(x = burn_year), position = "dodge") +
  facet_wrap(~ cd_group_b10_pdsisumn10, scales = "free_y"
             ) +
  #geom_smooth(method = "lm") +
  xlim(2000, 2020) +
  labs(title = "cd_group_b10_pdsisumn10")



ggplot(d_s_f |> filter(!is.na(cd_group_b10_pdsin4t1))) +
  geom_bar(aes(x = burn_year), position = "dodge") +
  facet_wrap(~ cd_group_b10_pdsin4t1, scales = "free_y"
             ) +
  #geom_smooth(method = "lm") +
  xlim(2000, 2020) +
  labs(title = "cd_group_b10_pdsin4t1")


```

<button class="accordion-button" type="button" data-bs-toggle="collapse" data-bs-target="#collapseOne">

Click to see Session Information

</button>

:::: {#collapseOne .accordion-collapse .collapse}
::: {}
```{r sessionInfo}
sessionInfo()
```
:::
::::
