---
title: "04_data_exploration_2"
format: html
editor: visual
---

# Setup for analysis

```{r setupchunk}
#| echo: true
#| message: false
#| warning: false
#| results: 'hide'

rm(list = ls())

if(!requireNamespace("here", quietly = TRUE)) {
  install.packages("here")
}
library(here)

source(here::here("utils", "functions.R"))

install_load_packages(c(
  "tidyverse",
  "arrow",
  "sf",
  "scales",
  "rlang",
  "grid",
  "gridExtra",
  "ggpmisc",
  "corrplot"))

gdrive_folder <- "GEE_resilience_v6_operational_ss500_ts50000"

dir_figs <- here::here('figs', gdrive_folder)
dir_derived <- here::here('data', 'derived', gdrive_folder)
dir_raw <- here::here('data', 'raw')
dir_manual <- here::here('data', 'manual')
dir_ensure(c(dir_figs,
             dir_derived,
             dir_manual,
             dir_raw))

seed = 1234
set.seed(seed)


```

# Data

```{r}

d_s_f <- list.files(
  here::here(dir_derived, "parquet"),
  pattern = "dats_short.*\\.parquet$",
  full.names = TRUE
)

d_s <- open_dataset(d_s_f, format = "parquet") |>
  collect()

d_s_fire <- d_s |> filter(fire == 1)
d_s_nfire <- d_s |> filter(fire == 0)

d_l_f <- list.files(
  here::here(dir_derived, "parquet"),
  pattern = "dats_long.*\\.parquet$",
  full.names = TRUE
)

d_l <- open_dataset(d_l_f, format = "parquet") |>
  collect()

d_l_fire <- d_l |> filter(fire == 1)

```


# Set CD groups

```{r}

set_cd_groups <- function(df,
                          b_nm,
                          d_nm,
                          cd_nm,
                          b_threshold,
                          d_threshold) {
  
  if(d_threshold < 0) {
    df_new <- df |>
    dplyr::mutate(
      "{cd_nm}" := dplyr::case_when(
        .data[[b_nm]] <  b_threshold & fire == 1 & .data[[d_nm]] >  d_threshold ~ "f",
        .data[[b_nm]] >= b_threshold & fire == 1 & .data[[d_nm]] >  d_threshold ~ "bf",
        .data[[b_nm]] <  b_threshold & fire == 1 & .data[[d_nm]] <= d_threshold ~ "df",
        .data[[b_nm]] >= b_threshold & fire == 1 & .data[[d_nm]] <= d_threshold ~ "bdf",
        TRUE ~ NA_character_
      )
    )
  } else {
    df_new <- df |>
    dplyr::mutate(
      "{cd_nm}" := dplyr::case_when(
        .data[[b_nm]] <  b_threshold & fire == 1 & .data[[d_nm]] <  d_threshold ~ "f",
        .data[[b_nm]] >= b_threshold & fire == 1 & .data[[d_nm]] <  d_threshold ~ "bf",
        .data[[b_nm]] <  b_threshold & fire == 1 & .data[[d_nm]] >= d_threshold ~ "df",
        .data[[b_nm]] >= b_threshold & fire == 1 & .data[[d_nm]] >= d_threshold ~ "bdf",
        TRUE ~ NA_character_
      )
    )
    
  }
  
  df_new <- df_new |>
    dplyr::mutate(
      "{cd_nm}" := forcats::fct_relevel(.data[[cd_nm]], "f")
    )

  return(df_new)
}



cd_group_table <- function(df, specs) {

  purrr::map_dfr(specs, function(s) {

    df2 <- set_cd_groups(
      df = df,
      b_nm = s$b_nm,
      d_nm = s$d_nm,
      cd_nm = s$cd_nm,
      b_threshold = s$b_threshold,
      d_threshold = s$d_threshold
    )

    # row counts
    counts_rows <- df2 |>
      count(.data[[s$cd_nm]], name = "n") |>
      mutate(group = .data[[s$cd_nm]]) |>
      select(group, n) |>
      tidyr::complete(group = c("f", "df", "bf", "bdf"), fill = list(n = 0)) |>
      tidyr::pivot_wider(names_from = group, values_from = n)

    # unique fire counts
    counts_fire <- df2 |>
      filter(!is.na(.data[[s$cd_nm]])) |>
      group_by(group = .data[[s$cd_nm]]) |>
      summarise(nf = n_distinct(fireid), .groups = "drop") |>
      tidyr::complete(group = c("f", "df", "bf", "bdf"), fill = list(nf = 0)) |>
      mutate(group = paste0(group, "_nf")) |>
      tidyr::pivot_wider(names_from = group, values_from = nf)

    out <- dplyr::bind_cols(counts_rows, counts_fire) |>
      mutate(
        cd_nm = s$cd_nm,
        b_nm = s$b_nm,
        d_nm = s$d_nm,
        b_threshold = s$b_threshold,
        d_threshold = s$d_threshold
      ) |>
      select(
        cd_nm, b_nm, b_threshold, d_nm, d_threshold,
        f, df, bf, bdf,
        f_nf, df_nf, bf_nf, bdf_nf
      ) |>
      mutate(
        total = f + df + bf + bdf,
        f_p   = round(100 * f   / total),
        df_p  = round(100 * df  / total),
        bf_p  = round(100 * bf  / total),
        bdf_p = round(100 * bdf / total)
      ) |>
      select(-total)

    out
  })
}

expand_cd_specs <- function(specs) {
  # specs: list of lists; each list can include:
  #   b_nm, d_nm, cd_prefix (optional), b_thresholds, d_thresholds
  #
  # returns a list of fully-expanded spec lists:
  #   b_nm, d_nm, cd_nm, b_threshold, d_threshold
  
  expanded <- purrr::map_dfr(specs, function(s) {
    cd_prefix <- s$cd_prefix %||% paste0("cd_", s$b_nm, "__", s$d_nm)

    tidyr::crossing(
      b_threshold = s$b_thresholds,
      d_threshold = s$d_thresholds
    ) |>
      mutate(
        b_nm = s$b_nm,
        d_nm = s$d_nm,
        cd_nm = paste0(
          cd_prefix,
          "_b", b_threshold,
          "_d", d_threshold
        )
      ) |>
      select(cd_nm, b_nm, d_nm, b_threshold, d_threshold)
  })

  # convert rows to list-of-lists (what cd_group_table expects)
  purrr::pmap(
    expanded,
    \(cd_nm, b_nm, d_nm, b_threshold, d_threshold)
      list(cd_nm = cd_nm, b_nm = b_nm, d_nm = d_nm,
           b_threshold = b_threshold, d_threshold = d_threshold)
  )
}

`%||%` <- function(x, y) if (!is.null(x)) x else y



specs_grid1 <- list(
  list(
    b_nm = "biotic_relaxedforestnorm_5_yrs_prior_sum_yot",
    d_nm = "pdsi_annual_5_yrs_prior_sum_yot",
    b_thresholds = c(10, 20, 50),
    d_thresholds = c(-10)
  )
)

specs_grid2 <- list(
  list(
    b_nm = "biotic_relaxedforestnorm_5_yrs_prior_sum_yot",
    d_nm = "pdsi_annual_5_yrs_prior_threshold_n4_yot",
    b_thresholds = c(10, 20, 50),
    d_thresholds = c(1, 2)
  )
)

specs_grid3 <- list(
  list(
    b_nm = "biotic_relaxedforestnorm_5_yrs_prior_sum_yot",
    d_nm = "pdsi_annual_5_yrs_prior_threshold_n3_yot",
    b_thresholds = c(10, 20, 50),
    d_thresholds = c(1, 2, 3)
  )
)

specs_grid4 <- list(
  list(
    b_nm = "biotic_relaxedforestnorm_5_yrs_prior_sum_yot",
    d_nm = "hd_fingerprint_5_yrs_prior_threshold_6_yot",
    b_thresholds = c(10, 20, 50),
    d_thresholds = c(1, 2)
  )
)

expanded_specs <- c(
  expand_cd_specs(specs_grid1),
  expand_cd_specs(specs_grid2),
  expand_cd_specs(specs_grid3),
  expand_cd_specs(specs_grid4)
)

cd_summary_grid <- cd_group_table(d_s, expanded_specs) |>
  arrange(b_threshold, d_nm) |>
  filter(!if_any(where(is.numeric), ~ . == 0)) |>
  select(-cd_nm)


cd_tbl <- tableGrob(
  cd_summary_grid,
  theme = ttheme_minimal(
    core = list(bg_params = list(fill = "white", col = NA)),
    colhead = list(bg_params = list(fill = "white", col = NA))
  )
)

cd_title <- textGrob(
  "Counts for possible CD groups",
  gp = gpar(fontsize = 16, fontface = "bold")
)

cd_table_with_title <- arrangeGrob(
  cd_tbl,
  top = cd_title
)

ggsave(
  here(dir_figs, "cd_group_counts.png"),
  cd_table_with_title,
  width = 20,
  height = 6,
  bg = "white"
)














d_s <- set_cd_groups(df = d_s,
                   b_nm = "biotic_relaxedforestnorm_5_yrs_prior_sum_yot",
                   d_nm = "pdsi_annual_5_yrs_prior_threshold_n4_yot",
                   cd_nm = "cd_group_b10_pdsin4t1",
                   b_threshold = 10,
                   d_threshold = 1)

d_s <- set_cd_groups(df = d_s,
                   b_nm = "biotic_relaxedforestnorm_5_yrs_prior_sum_yot",
                   d_nm = "pdsi_annual_5_yrs_prior_threshold_n3_yot",
                   cd_nm = "cd_group_b10_pdsin3t1",
                   b_threshold = 10,
                   d_threshold = 1)

d_s <- set_cd_groups(df = d_s,
                   b_nm = "biotic_relaxedforestnorm_5_yrs_prior_sum_yot",
                   d_nm = "pdsi_annual_5_yrs_prior_sum_yot",
                   cd_nm = "cd_group_b10_pdsisumn10",
                   b_threshold = 10,
                   d_threshold = -10)


```

# Hist & N for each (burn_year 2005-2020)

```{r}

plot_density_with_deciles <- function(df, var, title, x_label = NULL, log_scale = FALSE) {

  var_quo <- enquo(var)
  var_name <- as_name(var_quo)

  # Handle log scale
  if (log_scale) {
    df_plot <- df |> filter(!!var_quo > 0)
  } else {
    df_plot <- df
  }

  # Compute deciles
  probs <- seq(0.1, 0.9, by = 0.1)
  deciles <- quantile(
    df_plot[[var_name]],
    probs = probs,
    na.rm = TRUE
  )

  caption_text <- paste0(
    "Deciles (10–90%): ",
    paste0(
      names(deciles),
      " = ",
      round(deciles, 3),
      collapse = " | "
    )
  )

  p <- ggplot(df_plot, aes(x = !!var_quo)) +
    geom_density() +
    geom_vline(
      xintercept = deciles,
      linetype = "dotted",
      color = "red"
    ) +
    labs(
      x = ifelse(is.null(x_label), var_name, x_label),
      y = "Density",
      caption = caption_text,
      title = title
    ) +
    theme_minimal()

  if (log_scale) {
    p <- p + scale_x_log10(labels = label_number())
  }

  p
}





compare_densities_with_deciles <- function(
    df1,
    df2,
    var,
    title,
    x_label = NULL,
    log_scale = FALSE,
    df1_name = deparse(substitute(df1)),
    df2_name = deparse(substitute(df2))
) {

  var_quo <- rlang::enquo(var)
  var_name <- rlang::as_name(var_quo)

  prep_one <- function(df, name) {
    x <- dplyr::pull(df, !!var_quo)

    if (log_scale) {
      keep <- !is.na(x) & x > 0
      x <- x[keep]
      df <- df[keep, , drop = FALSE]
    } else {
      x <- x[!is.na(x)]
    }

    probs <- seq(0.1, 0.9, by = 0.1)
    dec <- stats::quantile(x, probs = probs, na.rm = TRUE, names = TRUE)

    list(
      df = df,
      name = name,
      deciles = dec
    )
  }

  a <- prep_one(df1, df1_name)
  b <- prep_one(df2, df2_name)

  df_plot <- dplyr::bind_rows(
    dplyr::mutate(a$df, .which_df = a$name),
    dplyr::mutate(b$df, .which_df = b$name)
  )

  caption_text <- paste0(
    a$name, " deciles: ",
    paste0(names(a$deciles), "=", round(a$deciles, 3), collapse = " | "),
    "\n",
    b$name, " deciles: ",
    paste0(names(b$deciles), "=", round(b$deciles, 3), collapse = " | ")
  )

  p <- ggplot2::ggplot(
    df_plot,
    ggplot2::aes(x = !!var_quo, color = .data[[".which_df"]])
  ) +
    ggplot2::geom_density() +
    ggplot2::geom_vline(
      xintercept = unname(a$deciles),
      linetype = "dotted",
      color = "red"
    ) +
    ggplot2::geom_vline(
      xintercept = unname(b$deciles),
      linetype = "dotted",
      color = "blue"
    ) +
    ggplot2::annotate(
      "text",
      x = unname(a$deciles),
      y = 0,
      label = names(a$deciles),
      color = "red",
      vjust = 0.3,
      size = 3
    ) +
    ggplot2::annotate(
      "text",
      x = unname(b$deciles),
      y = 0,
      label = names(b$deciles),
      color = "blue",
      vjust = 1.2,
      size = 3
    ) +
    ggplot2::scale_color_manual(
      values = stats::setNames(c("red", "blue"), c(a$name, b$name)),
      breaks = c(a$name, b$name)
    ) +
    ggplot2::labs(
      x = ifelse(is.null(x_label), var_name, x_label),
      y = "Density",
      caption = caption_text,
      title = title,
      color = NULL
    ) +
    ggplot2::theme_minimal() +
    theme(
      plot.caption = element_text(size = 8)
  )

  if (log_scale) {
    p <- p + ggplot2::scale_x_log10(labels = scales::label_number())
  }

  p
}

compare_densities_with_deciles(
  d_s_fire,
  d_s_nfire,
  biotic_relaxedforestnorm_5_yrs_prior_sum_yot,
  title = "Biotic prior sum: fire-only vs no fire",
  x_label = "Biotic 5-yr prior sum",
  log_scale = TRUE
)
ggsave(here(dir_figs, "density_biotic_prior.png"))

compare_densities_with_deciles(
  d_s_fire,
  d_s_nfire,
  biotic_6roll_max,
  title = "Biotic roll max : fire-only vs no fire",
  x_label = "Biotic 5-yr roll max",
  log_scale = TRUE
)
ggsave(here(dir_figs, "density_biotic_rollmax.png"))


compare_densities_with_deciles(
  d_s_fire,
  d_s_nfire,
  pdsi_annual_5_yrs_prior_sum_yot,
  title = "PDSI sum prior : fire-only vs no fire",
  x_label = "PDSI 5-yr sum",
  log_scale = FALSE
)
ggsave(here(dir_figs, "density_pdsi_prior.png"))

compare_densities_with_deciles(
  d_s_fire,
  d_s_nfire,
  pdsi_annual_year_of,
  title = "PDSI year of : fire-only vs no fire",
  x_label = "PDSI",
  log_scale = FALSE
)
ggsave(here(dir_figs, "density_pdsi_yo.png"))





plot_group_prop <- function(df, x_var) {
  d_plot <- df |>
    dplyr::filter(!is.na(.data[[x_var]]), !is.na(fire)) |>
    dplyr::count(fire, value = .data[[x_var]], name = "n") |>
    dplyr::group_by(fire) |>
    dplyr::mutate(prop = n / sum(n)) |>
    dplyr::ungroup()
  
  ggplot(
    d_plot,
    aes(
      x = factor(value),
      y = prop,
      fill = factor(fire)
    )
  ) +
    geom_col(position = "dodge") +
    scale_y_continuous(labels = scales::percent) +
    labs(
      x = x_var,
      y = "Proportion within fire group",
      fill = "fire"
    ) +
    theme_minimal()
}

p1 <- plot_group_prop(d_s, "pdsi_annual_5_yrs_prior_threshold_n3_yot")
p2 <- plot_group_prop(d_s, "pdsi_annual_5_yrs_prior_threshold_n4_yot")
p3 <- plot_group_prop(d_s, "hd_fingerprint_5_yrs_prior_threshold_6_yot")
p4 <- plot_group_prop(d_s, "hd_fingerprint_year_of")

ggsave(here(dir_figs, "pdsin3t_prior.png"), plot = p1)
ggsave(here(dir_figs, "pdsin4t_prior.png"), plot = p2)
ggsave(here(dir_figs, "hd6t_prior.png"), plot = p3)
ggsave(here(dir_figs, "hd_yo.png"), plot = p4)





# Burn severity

ggplot(d_s |> filter(fire == 1)) +
  geom_density(aes(x = burn_cbi, color = cd_group_b10_pdsisumn10))
ggsave(here(dir_figs, "cbi_by_cd_group_b10_pdsisumn10.png"))

ggplot(d_s |> filter(fire == 1)) +
  geom_density(aes(x = burn_cbi, color = cd_group_b10_pdsin3t1))
ggsave(here(dir_figs, "cbi_by_cd_group_b10_pdsin3t1.png"))

ggplot(d_s |> filter(fire == 1)) +
  geom_density(aes(x = burn_cbi, color = cd_group_b10_pdsin4t1))
ggsave(here(dir_figs, "cbi_by_cd_group_b10_pdsin4t1.png"))




d_plot <- d_s_fire |>
  dplyr::filter(burn_cbi >= 0 & biotic_relaxedforestnorm_5_yrs_prior_sum_yot <= 100 & !is.na(biotic_relaxedforestnorm_5_yrs_prior_sum_yot)) |>
  dplyr::mutate(
    biotic_bin = cut(
      biotic_relaxedforestnorm_5_yrs_prior_sum_yot,
      breaks = 10
    )
  )

ggplot(d_plot, aes(x = biotic_bin, y = burn_cbi)) +
  geom_boxplot() +
  theme_minimal()
ggsave(here(dir_figs, "cbi_by_biotic_bin.png"))




d_plot <- d_s_fire |>
  dplyr::filter(burn_cbi >= 0 & !is.na(pdsi_annual_5_yrs_prior_sum_yot)) |>
  dplyr::mutate(
    pdsi_bin = cut(
      pdsi_annual_5_yrs_prior_sum_yot,
      breaks = 10
    )
  )

ggplot(d_plot, aes(x = pdsi_bin, y = burn_cbi)) +
  geom_boxplot() +
  theme_minimal()
ggsave(here(dir_figs, "cbi_by_pdsisumprior_bin.png"))


d_plot <- d_s_fire |>
  dplyr::filter(burn_cbi >= 0 & !is.na(pdsi_annual_year_of)) |>
  dplyr::mutate(
    pdsiyo_bin = cut(
      pdsi_annual_year_of,
      breaks = 10
    )
  )

ggplot(d_plot, aes(x = pdsiyo_bin, y = burn_cbi)) +
  geom_boxplot() +
  theme_minimal()
ggsave(here(dir_figs, "cbi_by_pdsiyo_bin.png"))



ggplot(d_s_fire, aes(x = as.character(hd_fingerprint_year_of), y = burn_cbi)) +
  geom_boxplot() +
  theme_minimal()
ggsave(here(dir_figs, "cbi_by_hdyo_bin.png"))


```




# Outcomes

```{r}
ggplot(d_s |> filter(fire == 1 & !is.na(cd_group_b10_pdsisumn10))) +
  geom_density(aes(x = raw_post_fire_rap_tree_difference_n6_3_yr, color = cd_group_b10_pdsisumn10))



ggplot(d_s |> filter(fire == 1)) +
  geom_point(aes(x = burn_cbi,
                 y = raw_post_fire_rap_tree_difference_n6_3_yr),
             alpha = 0.2) +
  geom_smooth(aes(x = burn_cbi,
                 y = raw_post_fire_rap_tree_difference_n6_3_yr))


```


# correlations
```{r}

dir_correlations <- here::here(dir_figs, "correlations")
dir_ensure(dir_correlations)

# Correlation matrix
correlations <- d_s |>
  select(aet, def, gam_rap_tree_pre6_fit, srtm, tpi, chili) |>
  cor(use = "pairwise.complete.obs")
png(here(dir_correlations, glue('corr_plot_1_all.png')), width = 400, height = 400)
corrplot(correlations, method = 'number', type = "lower", diag = FALSE, use = 'pairwise.complete.obs') 
dev.off()



drought_correlations <- d_s |>
  select(hd_fingerprint_year_of, pdsi_annual_year_of, pdsi_annual_5_yrs_prior_sum_yot, pdsi_annual_5_yrs_prior_threshold_n4_yot, pdsi_annual_5_yrs_prior_threshold_n3_yot, hd_fingerprint_5_yrs_prior_threshold_6_yot) |>
  cor(use = "pairwise.complete.obs")
png(here(dir_correlations, glue('corr_plot_2_all.png')), width = 800, height = 800)
corrplot(drought_correlations, method = 'number', type = "lower", diag = FALSE, use = 'pairwise.complete.obs') 
dev.off()


for(ecor in unique(d_s$ecoregion_code_name)) {
  
  correlations <- d_s |>
    filter(ecoregion_code_name == ecor) |>
    select(aet, def, gam_rap_tree_pre6_fit, srtm, tpi, chili) |>
    cor(use = "pairwise.complete.obs")
  png(here(dir_correlations, glue('corr_plot_1_{ecor}.png')), width = 400, height = 400)
  corrplot(correlations, method = 'number', type = "lower", diag = FALSE, use = 'pairwise.complete.obs') 
  dev.off()
  
  
  
  drought_correlations <- d_s |>
    filter(ecoregion_code_name == ecor) |>
    select(hd_fingerprint_year_of, pdsi_annual_year_of, pdsi_annual_5_yrs_prior_sum_yot, pdsi_annual_5_yrs_prior_threshold_n4_yot, pdsi_annual_5_yrs_prior_threshold_n3_yot, hd_fingerprint_5_yrs_prior_threshold_6_yot) |>
    cor(use = "pairwise.complete.obs")
  png(here(dir_correlations, glue('corr_plot_2_{ecor}.png')), width = 800, height = 800)
  corrplot(drought_correlations, method = 'number', type = "lower", diag = FALSE, use = 'pairwise.complete.obs') 
  dev.off()

}




```
      


# Extend covariate distributions to add quant

```{r}

library(ggplot2)
library(dplyr)
library(purrr)
library(patchwork)
library(rlang)
library(forcats)
library(readr)
library(stringr)
library(tidyr)

dir_density_assessment <- here(dir_figs, "density_assessment")
dir_ensure(dir_density_assessment)

safe_filename <- function(x) {
  x |>
    stringr::str_replace_all("[^[:alnum:]_ -]", "_") |>
    stringr::str_replace_all("[ /]+", "_") |>
    stringr::str_replace_all("_+", "_") |>
    stringr::str_trim(side = "both")
}
compute_density_overlap_metrics <- function(df,
                                            x_var,
                                            group_var,
                                            n_grid = 512,
                                            min_n_group = 20) {
  
  d <- tibble::tibble(
    x = df[[x_var]],
    group = df[[group_var]]
  ) |>
    dplyr::filter(
      !is.na(x),
      !is.na(group)
    ) |>
    dplyr::mutate(group = as.character(group))
  
  if (nrow(d) == 0) {
    return(tibble::tibble(
      n = 0L,
      n_groups = 0L,
      groups_used = NA_character_,
      multigroup_overlap = NA_real_,
      min_pairwise_overlap = NA_real_,
      mean_pairwise_overlap = NA_real_
    ))
  }
  
  group_counts <- d |>
    dplyr::count(group, name = "n_group") |>
    dplyr::filter(n_group >= min_n_group)
  
  d <- d |>
    dplyr::semi_join(group_counts, by = "group")
  
  if (nrow(d) == 0) {
    return(tibble::tibble(
      n = 0L,
      n_groups = 0L,
      groups_used = NA_character_,
      multigroup_overlap = NA_real_,
      min_pairwise_overlap = NA_real_,
      mean_pairwise_overlap = NA_real_
    ))
  }
  
  groups <- sort(unique(d$group))
  
  if (length(groups) < 2) {
    return(tibble::tibble(
      n = nrow(d),
      n_groups = length(groups),
      groups_used = paste(groups, collapse = ", "),
      multigroup_overlap = NA_real_,
      min_pairwise_overlap = NA_real_,
      mean_pairwise_overlap = NA_real_
    ))
  }
  
  x_range <- range(d$x, na.rm = TRUE)
  
  if (!is.finite(x_range[1]) || !is.finite(x_range[2]) || x_range[1] == x_range[2]) {
    return(tibble::tibble(
      n = nrow(d),
      n_groups = length(groups),
      groups_used = paste(groups, collapse = ", "),
      multigroup_overlap = NA_real_,
      min_pairwise_overlap = NA_real_,
      mean_pairwise_overlap = NA_real_
    ))
  }
  
  dens_list <- purrr::map(
    groups,
    function(gr) {
      xg <- d |>
        dplyr::filter(group == gr) |>
        dplyr::pull(x)
      
      stats::density(
        x = xg,
        from = x_range[1],
        to = x_range[2],
        n = n_grid,
        na.rm = TRUE
      )
    }
  )
  
  x_grid <- dens_list[[1]]$x
  dx <- x_grid[2] - x_grid[1]
  
  y_mat <- do.call(
    cbind,
    purrr::map(dens_list, \(dd) dd$y)
  )
  
  multigroup_overlap <- sum(apply(y_mat, 1, min)) * dx
  
  pair_idx <- utils::combn(seq_along(groups), 2, simplify = FALSE)
  
  pairwise_tbl <- purrr::map_dfr(
    pair_idx,
    function(idx) {
      ovl <- sum(pmin(y_mat[, idx[1]], y_mat[, idx[2]])) * dx
      
      tibble::tibble(
        group_1 = groups[idx[1]],
        group_2 = groups[idx[2]],
        pairwise_overlap = ovl
      )
    }
  )
  
  tibble::tibble(
    n = nrow(d),
    n_groups = length(groups),
    groups_used = paste(groups, collapse = ", "),
    multigroup_overlap = multigroup_overlap,
    min_pairwise_overlap = min(pairwise_tbl$pairwise_overlap, na.rm = TRUE),
    mean_pairwise_overlap = mean(pairwise_tbl$pairwise_overlap, na.rm = TRUE)
  )
}

plot_density_triplets <- function(df,
                                  x_vars = c(
                                    "aet",
                                    "def",
                                    "gam_rap_tree_pre6_fit",
                                    "srtm",
                                    "chili",
                                    "tpi"
                                  ),
                                  group_vars = c(
                                    "cd_group_b10_pdsisumn10",
                                    "cd_group_b10_pdsin3t1",
                                    "cd_group_b10_pdsin4t1"
                                  ),
                                  n_grid = 512,
                                  min_n_group = 20,
                                  plot_title_prefix = NULL) {
  
  metrics_tbl <- tidyr::expand_grid(
    x_var = x_vars,
    group_var = group_vars
  ) |>
    dplyr::mutate(
      metrics = purrr::map2(
        x_var,
        group_var,
        ~ compute_density_overlap_metrics(
          df = df,
          x_var = .x,
          group_var = .y,
          n_grid = n_grid,
          min_n_group = min_n_group
        )
      )
    ) |>
    tidyr::unnest(metrics)
  
  make_one_plot <- function(x_var, group_var) {
    
    metrics_row <- metrics_tbl |>
      dplyr::filter(
        x_var == !!x_var,
        group_var == !!group_var
      )
    
    label_txt <- paste0(
      "OVL(all): ",
      ifelse(is.na(metrics_row$multigroup_overlap), "NA",
             sprintf("%.3f", metrics_row$multigroup_overlap)),
      "\nMin pair: ",
      ifelse(is.na(metrics_row$min_pairwise_overlap), "NA",
             sprintf("%.3f", metrics_row$min_pairwise_overlap)),
      "\nMean pair: ",
      ifelse(is.na(metrics_row$mean_pairwise_overlap), "NA",
             sprintf("%.3f", metrics_row$mean_pairwise_overlap)),
      "\nN: ",
      metrics_row$n,
      "\nGroups: ",
      metrics_row$n_groups
    )
    
    d_plot <- tibble::tibble(
      x = df[[x_var]],
      group = df[[group_var]]
    ) |>
      dplyr::filter(
        !is.na(x),
        !is.na(group)
      ) |>
      dplyr::mutate(group = as.character(group)) |>
      dplyr::add_count(group, name = "n_group") |>
      dplyr::filter(n_group >= min_n_group)
    
    ggplot(
      data = d_plot,
      aes(
        x = x,
        color = group
      )
    ) +
      geom_density(linewidth = 0.7) +
      annotate(
        geom = "text",
        x = Inf,
        y = Inf,
        label = label_txt,
        hjust = 1.05,
        vjust = 1.1,
        size = 3.2
      ) +
      theme_minimal() +
      labs(
        x = x_var,
        y = "Density",
        color = group_var,
        title = group_var
      )
  }
  
  plot_list <- purrr::map(
    x_vars,
    function(x_var) {
      purrr::map(
        group_vars,
        ~ make_one_plot(x_var = x_var, group_var = .x)
      ) |>
        patchwork::wrap_plots(nrow = 1) +
        patchwork::plot_annotation(
          title = if (is.null(plot_title_prefix)) {
            x_var
          } else {
            paste0(plot_title_prefix, " - ", x_var)
          }
        )
    }
  ) |>
    rlang::set_names(x_vars)
  
  list(
    plots = plot_list,
    metrics = metrics_tbl
  )
}

save_density_plot_set <- function(plot_list, out_dir, prefix) {
  dir_ensure(out_dir)
  
  purrr::iwalk(
    plot_list,
    ~ ggplot2::ggsave(
      filename = file.path(out_dir, paste0(prefix, .y, "_density_triplet.png")),
      plot = .x,
      width = 12,
      height = 4,
      dpi = 300
    )
  )
}

save_density_metrics_table <- function(metrics_tbl, out_dir, filename) {
  dir_ensure(out_dir)
  
  readr::write_csv(
    metrics_tbl,
    file.path(out_dir, filename)
  )
}


# all-data plots + summary table
density_all <- plot_density_triplets(
  df = d_s,
  plot_title_prefix = "All",
  min_n_group = 20
)

save_density_plot_set(
  plot_list = density_all$plots,
  out_dir = dir_density_assessment,
  prefix = "all_"
)

save_density_metrics_table(
  metrics_tbl = density_all$metrics |>
    dplyr::mutate(ecoregion_code_name = "All") |>
    dplyr::relocate(ecoregion_code_name),
  out_dir = dir_density_assessment,
  filename = "all_density_overlap_metrics.csv"
)

ecoregions <- d_s |>
  dplyr::distinct(ecoregion_code_name) |>
  dplyr::pull(ecoregion_code_name) |>
  stats::na.omit() |>
  as.character() |>
  sort()

purrr::walk(
  ecoregions,
  function(ecoregion_i) {
    
    ecoregion_dir <- file.path(
      dir_density_assessment,
      safe_filename(ecoregion_i)
    )
    dir_ensure(ecoregion_dir)
    
    d_eco <- d_s |>
      dplyr::filter(ecoregion_code_name == ecoregion_i)
    
    density_eco <- plot_density_triplets(
      df = d_eco,
      plot_title_prefix = ecoregion_i,
      min_n_group = 20
    )
    
    save_density_plot_set(
      plot_list = density_eco$plots,
      out_dir = ecoregion_dir,
      prefix = paste0(safe_filename(ecoregion_i), "_")
    )
    
    save_density_metrics_table(
      metrics_tbl = density_eco$metrics |>
        dplyr::mutate(ecoregion_code_name = ecoregion_i) |>
        dplyr::relocate(ecoregion_code_name),
      out_dir = ecoregion_dir,
      filename = paste0(safe_filename(ecoregion_i), "_density_overlap_metrics.csv")
    )
  }
)

density_metrics_by_ecoregion <- purrr::map_dfr(
  ecoregions,
  function(ecoregion_i) {
    d_s |>
      dplyr::filter(ecoregion_code_name == ecoregion_i) |>
      plot_density_triplets(
        plot_title_prefix = ecoregion_i,
        min_n_group = 20
      ) |>
      purrr::pluck("metrics") |>
      dplyr::mutate(ecoregion_code_name = ecoregion_i) |>
      dplyr::relocate(ecoregion_code_name)
  }
) |>
  dplyr::mutate(
    x_var = factor(x_var, levels = c("aet", "def", "chili", "tpi", "srtm", "gam_rap_tree_pre6_fit"))
  )



readr::write_csv(
  density_metrics_by_ecoregion,
  file.path(dir_density_assessment, "density_overlap_metrics_by_ecoregion.csv")
)

```

# assess density overlap metrics
```{r}

ggplot(data = density_metrics_by_ecoregion) +
  geom_density(aes(x = multigroup_overlap, color = x_var)) +
  facet_wrap(~group_var)

ggplot(data = density_metrics_by_ecoregion) +
  geom_density(aes(x = multigroup_overlap, color = ecoregion_code_name)) +
  facet_wrap(~group_var)


ggplot(data = density_metrics_by_ecoregion |>
  dplyr::filter(!stringr::str_starts(x_var, "tt_")) |>
    dplyr::filter(group_var == "cd_group_b10_pdsin3t1")) +
  geom_col(aes(x = x_var, y = multigroup_overlap)) +
  facet_wrap(~ecoregion_code_name) +
  labs(title = "Multigroup overlap for all ecoregions for key covariates: cd_group_b10_pdsin3t1") +
  theme_minimal() +
  theme(
    axis.text.x = element_text(
      angle = 45,
      hjust = 1,
      vjust = 1
    )
  ) +
  geom_hline(aes(yintercept = 0.3), color = "red") +
  geom_hline(aes(yintercept = 0.6))

ggsave(filename = here(dir_density_assessment, "multigroup_overlap.png"))

```

# Assess standardized mean difference (SMD) using cobalt

```{r}


library(cobalt)
   
cobalt::bal.tab(x = cd_group_b10_pdsisumn10 ~ aet + def + srtm + tpi + chili + gam_rap_tree_pre6_fit + tt_change_aet + tt_change_def, 
                data = d_s |> filter(fire == 1 & !is.na(cd_group_b10_pdsisumn10)))

png(here(dir_figs, glue("love_plot_fire_cd_group_b10_pdsisumn10.png")))
cobalt::love.plot(x = cd_group_b10_pdsisumn10 ~ aet + def + srtm + tpi + chili + gam_rap_tree_pre6_fit + tt_change_aet + tt_change_def, 
                data = d_s |> filter(fire == 1 & !is.na(cd_group_b10_pdsisumn10)),
                stats = c("mean.diffs", "ks.statistics"),
                stars = "raw",
                position = "bottom",
                threshold = c(m = .1, ks = .05),
                limits = list(c(-1, 1), c(0, 0.6)),
                title = glue("Covariate balance love plot for CD treatment: cd_group_b10_pdsisumn10"))
dev.off()



library(cobalt)
library(dplyr)
library(glue)

d_plot <- d_s |>
  filter(fire == 1, !is.na(cd_group_b10_pdsisumn10)) |>
  mutate(
    cd_group_b10_pdsisumn10 = factor(cd_group_b10_pdsisumn10)
  )

bal_out <- cobalt::bal.tab(
  cd_group_b10_pdsisumn10 ~ aet + def + srtm + tpi + chili +
    gam_rap_tree_pre6_fit + tt_change_aet + tt_change_def,
  data = d_plot,
  pairwise = TRUE,
  stats = c("mean.diffs", "ks.statistics")
)

png(
  here::here(dir_figs, "love_plot_fire_cd_group_b10_pdsisumn10_smd.png"),
  width = 1400, height = 1000, res = 150
)

cobalt::love.plot(
  bal_out,
  stat = "mean.diffs",
  abs = TRUE,
  agg.fun = NULL,
  facet = "contrast",   # <-- separates pairs
  thresholds = c(m = .1)
)

dev.off()

```


```{r}


# GAM METRIC DISTRIBUTIONS
# 
# plot_density_by_group_with_n <- function(df, x_var, group_var, x_label = NULL) {
# 
#   x_quo <- enquo(x_var)
#   g_quo <- enquo(group_var)
# 
#   x_name <- as_name(x_quo)
#   g_name <- as_name(g_quo)
# 
#   # Filter missing groups
#   d_plot <- df |>
#     filter(!is.na(!!g_quo))
# 
#   # Compute N per group
#   group_n <- d_plot |>
#     count(!!g_quo, name = "N") |>
#     mutate(label = paste0(!!g_quo, " (N=", N, ")"))
# 
#   lab_vec <- setNames(group_n$label, group_n[[g_name]])
# 
#   ggplot(d_plot, aes(x = !!x_quo, color = !!g_quo)) +
#     geom_density() +
#     scale_color_discrete(labels = lab_vec) +
#     labs(
#       x = ifelse(is.null(x_label), x_name, x_label),
#       color = g_name
#     ) +
#     theme_minimal()
# }
# 
# plot_density_by_group_with_n(
#   d_s,
#   gam_rap_tree_diff_pre6_to_post_min,
#   cd_group_b10_pdsin4t1,
#   x_label = "RAP total loss"
# )
# 
# plot_density_by_group_with_n(
#   d_s,
#   gam_rap_tree_diff_pre6_to_post_min,
#   cd_group_b10_pdsin3t1,
#   x_label = "RAP total loss"
# )
# 
# plot_density_by_group_with_n(
#   d_s,
#   gam_rap_tree_diff_pre6_to_post_min,
#   cd_group_b10_pdsin4t3,
#   x_label = "RAP total loss"
# )
# 
# 
# plot_density_by_group_with_n(
#   d_s,
#   gam_rap_tree_diff_pre6_to_post_min,
#   cd_group_b10_pdsisumn10,
#   x_label = "RAP total loss"
# )


# Boxplots
ggplot(
  d_s |>
    filter(!is.na(cd_group_b10_pdsin4t1)) |>
    add_count(cd_group_b10_pdsin4t1, name = "n_group"),
  aes(
    x = paste0(cd_group_b10_pdsin4t1, " (N=", n_group, ")"),
    y = gam_rap_tree_diff_pre6_to_post_min
  )
) +
  geom_boxplot() +
  labs(y = "RAP total loss") +
  theme_minimal()
ggsave(here(dir_figs, "boxplot_raploss_cd_group_b10_pdsin4t1.png"))


ggplot(
  d_s |>
    filter(!is.na(cd_group_b10_pdsin4t3)) |>
    add_count(cd_group_b10_pdsin4t3, name = "n_group"),
  aes(
    x = paste0(cd_group_b10_pdsin4t3, " (N=", n_group, ")"),
    y = gam_rap_tree_diff_pre6_to_post_min
  )
) +
  geom_boxplot() +
  labs(y = "RAP total loss") +
  theme_minimal()
ggsave(here(dir_figs, "boxplot_raploss_cd_group_b10_pdsin4t3.png"))


ggplot(
  d_s |>
    filter(!is.na(cd_group_b10_pdsisumn10)) |>
    add_count(cd_group_b10_pdsisumn10, name = "n_group"),
  aes(
    x = paste0(cd_group_b10_pdsisumn10, " (N=", n_group, ")"),
    y = gam_rap_tree_diff_pre6_to_post_min
  )
) +
  geom_boxplot() +
  labs(y = "RAP total loss") +
  theme_minimal()
ggsave(here(dir_figs, "boxplot_raploss_cd_group_b10_pdsisumn10.png"))




ggplot(
  d_s |>
    filter(!is.na(cd_group_b10_pdsin4t3)) |>
    add_count(cd_group_b10_pdsin4t3, name = "n_group"),
  aes(
    x = paste0(cd_group_b10_pdsin4t3, " (N=", n_group, ")"),
    y = gam_rap_tree_slope_post_min_to_p15
  )
) +
  geom_boxplot() +
  labs(y = "RAP min to 15 yr recovery slope") +
  theme_minimal()
ggsave(here(dir_figs, "boxplot_rapgamslope_cd_group_b10_pdsin4t3.png"))




ggplot(
  d_s |>
    filter(!is.na(cd_group_b10_pdsin4t3)) |>
    add_count(cd_group_b10_pdsin4t3, name = "n_group"),
  aes(
    x = paste0(cd_group_b10_pdsin4t3, " (N=", n_group, ")"),
    y = raw_post_fire_rap_tree_slope_3_15_yrs
  )
) +
  geom_boxplot() +
  labs(y = "RAP 3-15 yr recovery slope") +
  theme_minimal()
ggsave(here(dir_figs, "boxplot_raprawslope_cd_group_b10_pdsin4t3.png"))




# 
# 
# 
# bin_cd_and_boxplot <- function(df, cd_nm = cd_nm, outcome = outcome, fire = fire, n_bins = 10) {
# 
#   cd_quo <- rlang::enquo(cd_nm)
#   out_quo <- rlang::enquo(outcome)
#   fire_quo <- rlang::enquo(fire)
# 
#   cd_name <- rlang::as_name(cd_quo)
# 
#   x <- df[[cd_name]]
# 
#   if (inherits(x, "integer")) {
# 
#     df2 <- df |>
#       dplyr::mutate(
#         bin_cd = ifelse(!!fire_quo == 0, NA, as.character(!!cd_quo))
#       )
# 
#   } else {
# 
#     rng <- range(x, na.rm = TRUE, finite = TRUE)
# 
#     # if there isn't a usable range, bins can't be formed
#     if (!is.finite(rng[1]) || !is.finite(rng[2]) || rng[1] == rng[2]) {
#       df2 <- df |>
#         dplyr::mutate(bin_cd = NA_character_)
#     } else {
# 
#       breaks <- seq(from = rng[1], to = rng[2], length.out = n_bins + 1)
# 
#       labels <- paste0(
#         round(head(breaks, -1), 1),
#         "–",
#         round(tail(breaks, -1), 1)
#       )
# 
#       df2 <- df |>
#         dplyr::mutate(
#           bin_cd = cut(
#             !!cd_quo,
#             breaks = breaks,
#             include.lowest = TRUE,
#             labels = labels
#           ),
#           bin_cd = ifelse(!!fire_quo == 0, NA, as.character(bin_cd))
#         )
#     }
#   }
# 
#   df2 <- df2 |> dplyr::mutate(bin_cd = factor(bin_cd))
# 
#   p <- ggplot2::ggplot(df2 |> dplyr::filter(!is.na(bin_cd)),
#               ggplot2::aes(x = bin_cd, y = !!out_quo)) +
#     ggplot2::geom_boxplot(outlier.alpha = 0.2) +
#     ggplot2::labs(
#       x = paste0(cd_name, " bins"),
#       y = rlang::as_name(out_quo)
#     ) +
#     ggplot2::theme_minimal() +
#     ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 30, hjust = 1))
# 
#   list(data = df2, plot = p)
# }
# 
# 
# res <- bin_cd_and_boxplot(d_s, biotic_relaxedforestnorm_5_yrs_prior_sum_yot, gam_rap_tree_diff_pre6_to_post_min)
# res$plot
# 
# res <- bin_cd_and_boxplot(d_s, pdsi_annual_5_yrs_prior_threshold_n3_yot, gam_rap_tree_diff_pre6_to_post_min)
# res$plot
# 
# 
# res <- bin_cd_and_boxplot(d_s, pdsi_annual_5_yrs_prior_threshold_n4_yot, gam_rap_tree_diff_pre6_to_post_min)
# res$plot
# 
# res <- bin_cd_and_boxplot(d_s, pdsi_annual_5_yrs_prior_sum_yot, gam_rap_tree_diff_pre6_to_post_min)
# res$plot
# 
# res <- bin_cd_and_boxplot(d_s,  hd_fingerprint_5_yrs_prior_threshold_6_yot, gam_rap_tree_diff_pre6_to_post_min)
# res$plot
# 



```

# VCF vs RAP plots

Both metrics & raw data

Takeaways: we find that RAP aggregated to 250m and VCF have RMSE of around 12.4. This remains similar across subgroups of data, although there is a slightly lower error in prefire and no-fire environments than postfire (\~12 vs \~13). The postfire error climbs slightly with vcf above 50, to \~14.

In terms of estimating resilience metrics, there is higher agreement between VCF and RAP when estimating pre-fire cover using an average window or GAM (as opposed to using single raw values). In estimating total loss, recovery, slope, etc, averaged windows provide lower RMSE between the two datasets, likely due to the fact that the GAMs are identifying different minimum values post fire between the two datasets.

```{r}

plot_lm_r2 <- function(df, x, y, title, xlim = NULL, ylim = NULL) {
  p <- ggplot(df, aes(x = {{ x }}, y = {{ y }})) +
    geom_point(alpha = 0.1) +
    geom_smooth(method = "lm", alpha = 0.1) +
    ggpmisc::stat_poly_eq(
      formula = y ~ x,
      aes(label = after_stat(rr.label)),
      parse = TRUE,
      label.x = "right",
      label.y = "bottom"
    ) +
    labs(title = title)

  if (!is.null(xlim)) p <- p + xlim(xlim)
  if (!is.null(ylim)) p <- p + ylim(ylim)

  p
}

extract_r2_rmse_table <- function(models) {
  tibble(
    model_name = names(models),
    model = models
  ) |>
    mutate(
      stats = purrr::map(model, broom::glance),
      rmse = purrr::map_dbl(model, \(m) {
        sqrt(mean(residuals(m)^2, na.rm = TRUE))
      })
    ) |>
    tidyr::unnest(stats) |>
    select(model_name, r.squared, rmse)
}

lm_safe <- function(formula, data) {
  vars <- all.vars(formula)
  d <- data |>
    dplyr::select(dplyr::all_of(vars)) |>
    dplyr::filter(dplyr::if_all(dplyr::everything(), is.finite))
  stats::lm(formula, data = d)
}

# 
# p_loss <- plot_lm_r2(
#   d_s,
#   gam_rap_tree_diff_pre6_to_post_min,
#   gam_vcf_tree_diff_pre6_to_post_min,
#   "VCF vs RAP total loss 250m"
# )
# 
# p_loss_raw <- plot_lm_r2(
#   d_s,
#   raw_post_fire_rap_tree_difference_n6_3_yr,
#   raw_post_fire_vcf_tree_difference_n6_3_yr,
#   "VCF vs RAP total loss 250m -6 to 3 yr raw windows"
# )
# 
# 
# 
# p_recov15 <- plot_lm_r2(
#   d_s,
#   gam_rap_tree_perc_recov_p15,
#   gam_vcf_tree_perc_recov_p15,
#   "VCF vs RAP recovery percent min-15yr 250m",
#   xlim = c(0, 100),
#   ylim = c(0, 100)
# )
# 
# p_raw_recov15 <- plot_lm_r2(
#   d_s,
#   raw_post_fire_rap_tree_perc_recovery_n6_3_15_yr,
#   raw_post_fire_vcf_tree_perc_recovery_n6_3_15_yr,
#   "VCF vs RAP raw recovery percent 3-15yr 250m",
#   xlim = c(0, 100),
#   ylim = c(0, 100)
# )
# 
# 
# p_slope15 <- plot_lm_r2(
#   d_s,
#   gam_rap_tree_slope_post_min_to_p15,
#   gam_vcf_tree_slope_post_min_to_p15,
#   "VCF vs RAP slope gam min to 15yr slope 250m"
# )
# 
# p_rawslope15 <- plot_lm_r2(
#   d_s,
#   raw_post_fire_rap_tree_slope_3_15_yrs,
#   raw_post_fire_vcf_tree_slope_3_15_yrs,
#   "VCF vs RAP slope raw 3 to 15yr slope 250m"
# )
# 
# 
# 
# p_pre6 <- plot_lm_r2(
#   d_s,
#   gam_rap_tree_pre6_fit,
#   gam_vcf_tree_pre6_fit,
#   "VCF vs RAP 6-yr prior fit 250m"
# )
# 
# p_pre6raw <- plot_lm_r2(
#   d_l |> filter(Time_to_Treatment == -6),
#   rap_tree,
#   vcf_tree,
#   "VCF vs RAP 6-yr-prior raw 250m"
# )
# 
# p_pre6raw_win <- plot_lm_r2(
#   d_s,
#   raw_avg_rap_tree_6_yr_pre_fire,
#   raw_avg_vcf_tree_6_yr_pre_fire,
#   "VCF vs RAP 6-yr-prior-window 250m"
# )
# 
# 
# 
# p_min <- plot_lm_r2(
#   d_s,
#   gam_rap_tree_post_min,
#   gam_vcf_tree_post_min,
#   "VCF vs RAP minimum fit 250m"
# )
# 
# p_3yr_raw <- plot_lm_r2(
#   d_s,
#   raw_avg_rap_tree_3_yr_post_fire,
#   raw_avg_vcf_tree_3_yr_post_fire,
#   "VCF vs RAP post-fire 3 yr raw"
# )
# 
# 
# 
# p_recov_tot <- plot_lm_r2(
#   d_s,
#   gam_rap_tree_diff_post_min_to_p15,
#   gam_vcf_tree_diff_post_min_to_p15,
#   "VCF vs RAP post-fire min to 15 gam"
# )
# 
# p_recov_tot_raw <- plot_lm_r2(
#   d_s,
#   raw_post_fire_rap_tree_increase_3_15_yr,
#   raw_post_fire_vcf_tree_increase_3_15_yr,
#   "VCF vs RAP post-fire 3 to 15 raw"
# )
# 
# 
# p_timetomin <- plot_lm_r2(
#   d_s,
#   gam_rap_tree_yrs_ref_to_post_min,
#   gam_vcf_tree_yrs_ref_to_post_min,
#   "VCF vs RAP time to minimum 250m"
# )
# 
# 
# 
# # Print
# p_loss
# p_loss_raw
# 
# p_recov15
# p_raw_recov15
# 
# p_slope15
# p_rawslope15
# 
# p_pre6
# p_pre6raw
# p_pre6raw_win
# 
# p_min
# p_3yr_raw
# 
# p_recov_tot
# p_recov_tot_raw
# 
# p_timetomin
# 
# ggplot(d_s) +
#   geom_bin2d(aes(x = gam_rap_tree_yrs_ref_to_post_min, y = gam_vcf_tree_yrs_ref_to_post_min))
# 




set.seed(1234)
d_s_2006 <- d_s |> dplyr::filter(burn_year > 2005)
d_l_2006 <- d_l |> dplyr::filter(burn_year > 2005)

models_xy <- list(
  loss_gam = lm_safe(gam_vcf_tree_diff_pre6_to_post_min ~ gam_rap_tree_diff_pre6_to_post_min, d_s_2006),
  loss_raw = lm_safe(raw_post_fire_vcf_tree_difference_n6_3_yr ~ raw_post_fire_rap_tree_difference_n6_3_yr, d_s_2006),

  recov15_gam = lm_safe(gam_vcf_tree_perc_recov_p15 ~ gam_rap_tree_perc_recov_p15, d_s_2006),
  recov15_raw = lm_safe(raw_post_fire_vcf_tree_perc_recovery_n6_3_15_yr ~ raw_post_fire_rap_tree_perc_recovery_n6_3_15_yr, d_s_2006),

  slope15_gam = lm_safe(gam_vcf_tree_slope_post_min_to_p15 ~ gam_rap_tree_slope_post_min_to_p15, d_s_2006),
  slope15_raw = lm_safe(raw_post_fire_vcf_tree_slope_3_15_yrs ~ raw_post_fire_rap_tree_slope_3_15_yrs, d_s_2006),

  pre6_fit_gam = lm_safe(gam_vcf_tree_pre6_fit ~ gam_rap_tree_pre6_fit, d_s_2006),
  pre6_raw_tminus6 = lm_safe(vcf_tree ~ rap_tree, d_l_2006 |> dplyr::filter(Time_to_Treatment == -6)),
  pre6_raw_window = lm_safe(raw_avg_vcf_tree_6_yr_pre_fire ~ raw_avg_rap_tree_6_yr_pre_fire, d_s_2006),

  postmin_gam = lm_safe(gam_vcf_tree_post_min ~ gam_rap_tree_post_min, d_s_2006),
  post3yr_raw = lm_safe(raw_avg_vcf_tree_3_yr_post_fire ~ raw_avg_rap_tree_3_yr_post_fire, d_s_2006),

  recov_tot_gam = lm_safe(gam_vcf_tree_diff_post_min_to_p15 ~ gam_rap_tree_diff_post_min_to_p15, d_s_2006),
  recov_tot_raw = lm_safe(raw_post_fire_vcf_tree_increase_3_15_yr ~ raw_post_fire_rap_tree_increase_3_15_yr, d_s_2006),

  time_to_min = lm_safe(gam_vcf_tree_yrs_ref_to_post_min ~ gam_rap_tree_yrs_ref_to_post_min, d_s_2006)
)

r2_rmse_table <- extract_r2_rmse_table(models_xy)

r2_rmse_table_g <- tableGrob(
  r2_rmse_table,
  theme = ttheme_minimal(
    core = list(bg_params = list(fill = "white", col = NA)),
    colhead = list(bg_params = list(fill = "white", col = NA))
  )
)

title1 <- textGrob(
  "VCF-RAP resilience metric \nR² and RMSE Comparisons \n(burn year >2005)",
  gp = gpar(fontsize = 16, fontface = "bold")
)

table_with_title1 <- arrangeGrob(
  r2_rmse_table_g,
  top = title1
)

ggsave(
  here(dir_figs, "rap_vs_vfc_metrics_table.png"),
  table_with_title1,
  width = 6,
  height = 7,
  bg = "white"
)




# Comparing raw data



ggplot(d_l) +
  geom_density(aes(x = rap_tree), color = "red") +
  geom_density(aes(x = vcf_tree), color = "blue") +
  labs(xlab = "Percent tree cover", caption = "RAP = red, VCF = blue")
ggsave(here(dir_figs, "raw_rap_vcf_density.png"))

set.seed(1234)
models <- list(
  all = lm(rap_tree ~ vcf_tree, d_l |> slice_sample(n = 1000000)),
  nofire = lm(rap_tree ~ vcf_tree, d_l |> filter(fire == 0) |> slice_sample(n = 1000000)),
  prefire = lm(rap_tree ~ vcf_tree, d_l |> filter(fire == 1 & treated == 0) |> slice_sample(n = 1000000)),
  postfire = lm(rap_tree ~ vcf_tree, d_l |> filter(fire == 1 & treated == 1) |> slice_sample(n = 1000000)),
  biotic_above10 = lm(rap_tree ~ vcf_tree, d_l |> filter(fire == 0 & biotic_6roll_max > 10) |> slice_sample(n = 1000000)),
  all_vcf_above50 = lm(rap_tree ~ vcf_tree, d_l |> filter(vcf_tree > 50) |> slice_sample(n = 1000000)),
  all_vcf_below50 = lm(rap_tree ~ vcf_tree, d_l |> filter(vcf_tree <= 50) |> slice_sample(n = 1000000)),
  postfire_vcf_above50 = lm(rap_tree ~ vcf_tree, d_l |> filter(vcf_tree > 50 & fire == 1 & treated == 1) |> slice_sample(n = 1000000)),
  postfire_vcf_below50 = lm(rap_tree ~ vcf_tree, d_l |> filter(vcf_tree <= 50 & fire == 1 & treated == 1) |> slice_sample(n = 1000000))
)


tbl <- tableGrob(
  extract_r2_rmse_table(models),
  theme = ttheme_minimal(
    core = list(bg_params = list(fill = "white", col = NA)),
    colhead = list(bg_params = list(fill = "white", col = NA))
  )
)

title <- textGrob(
  "Raw VCF-RAP data R² and RMSE Comparisons\n1 million samples per model",
  gp = gpar(fontsize = 16, fontface = "bold")
)

table_with_title <- arrangeGrob(
  tbl,
  top = title
)

ggsave(
  here(dir_figs, "rap_vs_vfc_raw_table.png"),
  table_with_title,
  width = 6,
  height = 4,
  bg = "white"
)





```

```{r}

dats_quick <- d_l |>
  filter(!is.na(cd_group)) |>
  slice_head(n = 100000) |>
  mutate(fire_frg = interaction(burn_sev, frg_reclass_desc),
         fire_pre_drought = interaction(burn_sev, pdsi_annual_5_yrs_prior_threshold),
         fire_pre_drought = interaction(burn_sev, pdsi_annual_5_yrs_after_threshold),
         biotic_cut = cut(biotic,
             breaks = c(0, 20, 40, 60, 80, 100),
             labels = c()
             )
         )

ggplot(dats_quick) +
  geom_smooth(method = 'gam', aes(x = Time_to_Treatment, y = rap_tree, color = cd_group)) +
  xlim(c(-10, 20)) +
  labs(title = "GAM fits sample")

ggplot(dats_quick) +
  geom_smooth(method = 'gam', aes(x = Time_to_Treatment, y = vcf_tree, color = cd_group)) +
  xlim(c(-10, 20)) +
  labs(title = "GAM fits sample")

ggplot(dats_quick |> filter(burn_sev != "Negligible")) +
  geom_smooth(method = 'gam', aes(x = Time_to_Treatment, y = rap_tree, color = frg_reclass_desc, lty = burn_sev)) +
  xlim(c(-10, 20)) +
  labs(title = "GAM fits sample")


```

<button class="accordion-button" type="button" data-bs-toggle="collapse" data-bs-target="#collapseOne">

Click to see Session Information

</button>

:::: {#collapseOne .accordion-collapse .collapse}
<div>

```{r sessionInfo}
sessionInfo()
```

</div>
::::
