---
title: "04_data_exploration_2"
format: html
editor: visual
---

# Setup for analysis

```{r setupchunk}
#| echo: true
#| message: false
#| warning: false
#| results: 'hide'

rm(list = ls())

if(!requireNamespace("here", quietly = TRUE)) {
  install.packages("here")
}
library(here)

source(here::here("utils", "functions.R"))

install_load_packages(c(
  "tidyverse",
  "arrow",
  "sf",
  "scales",
  "rlang",
  "grid",
  "gridExtra",
  "ggpmisc"))

gdrive_folder <- "GEE_resilience_v6_operational_ss500_ts50000"

dir_figs <- here::here('figs', gdrive_folder)
dir_derived <- here::here('data', 'derived', gdrive_folder)
dir_raw <- here::here('data', 'raw')
dir_manual <- here::here('data', 'manual')
dir_ensure(c(dir_figs,
             dir_derived,
             dir_manual,
             dir_raw))

seed = 1234
set.seed(seed)


```

# Data

```{r}

d_s_f <- list.files(
  here::here(dir_derived, "parquet"),
  pattern = "dats_short.*\\.parquet$",
  full.names = TRUE
)

d_s <- open_dataset(d_s_f, format = "parquet") |>
  collect()

d_s_fire <- d_s |> filter(fire == 1)
d_s_nfire <- d_s |> filter(fire == 0)

d_l_f <- list.files(
  here::here(dir_derived, "parquet"),
  pattern = "dats_long.*\\.parquet$",
  full.names = TRUE
)

d_l <- open_dataset(d_l_f, format = "parquet") |>
  collect()

d_l_fire <- d_l |> filter(fire == 1)

```

# Hist & N for each (burn_year 2005-2020)

```{r}

plot_density_with_deciles <- function(df, var, title, x_label = NULL, log_scale = FALSE) {

  var_quo <- enquo(var)
  var_name <- as_name(var_quo)

  # Handle log scale
  if (log_scale) {
    df_plot <- df |> filter(!!var_quo > 0)
  } else {
    df_plot <- df
  }

  # Compute deciles
  probs <- seq(0.1, 0.9, by = 0.1)
  deciles <- quantile(
    df_plot[[var_name]],
    probs = probs,
    na.rm = TRUE
  )

  caption_text <- paste0(
    "Deciles (10–90%): ",
    paste0(
      names(deciles),
      " = ",
      round(deciles, 3),
      collapse = " | "
    )
  )

  p <- ggplot(df_plot, aes(x = !!var_quo)) +
    geom_density() +
    geom_vline(
      xintercept = deciles,
      linetype = "dotted",
      color = "red"
    ) +
    labs(
      x = ifelse(is.null(x_label), var_name, x_label),
      y = "Density",
      caption = caption_text,
      title = title
    ) +
    theme_minimal()

  if (log_scale) {
    p <- p + scale_x_log10(labels = label_number())
  }

  p
}





compare_densities_with_deciles <- function(
    df1,
    df2,
    var,
    title,
    x_label = NULL,
    log_scale = FALSE,
    df1_name = deparse(substitute(df1)),
    df2_name = deparse(substitute(df2))
) {

  var_quo <- rlang::enquo(var)
  var_name <- rlang::as_name(var_quo)

  prep_one <- function(df, name) {
    x <- dplyr::pull(df, !!var_quo)

    if (log_scale) {
      keep <- !is.na(x) & x > 0
      x <- x[keep]
      df <- df[keep, , drop = FALSE]
    } else {
      x <- x[!is.na(x)]
    }

    probs <- seq(0.1, 0.9, by = 0.1)
    dec <- stats::quantile(x, probs = probs, na.rm = TRUE, names = TRUE)

    list(
      df = df,
      name = name,
      deciles = dec
    )
  }

  a <- prep_one(df1, df1_name)
  b <- prep_one(df2, df2_name)

  df_plot <- dplyr::bind_rows(
    dplyr::mutate(a$df, .which_df = a$name),
    dplyr::mutate(b$df, .which_df = b$name)
  )

  caption_text <- paste0(
    a$name, " deciles: ",
    paste0(names(a$deciles), "=", round(a$deciles, 3), collapse = " | "),
    "\n",
    b$name, " deciles: ",
    paste0(names(b$deciles), "=", round(b$deciles, 3), collapse = " | ")
  )

  p <- ggplot2::ggplot(
    df_plot,
    ggplot2::aes(x = !!var_quo, color = .data[[".which_df"]])
  ) +
    ggplot2::geom_density() +
    ggplot2::geom_vline(
      xintercept = unname(a$deciles),
      linetype = "dotted",
      color = "red"
    ) +
    ggplot2::geom_vline(
      xintercept = unname(b$deciles),
      linetype = "dotted",
      color = "blue"
    ) +
    ggplot2::annotate(
      "text",
      x = unname(a$deciles),
      y = 0,
      label = names(a$deciles),
      color = "red",
      vjust = 0.3,
      size = 3
    ) +
    ggplot2::annotate(
      "text",
      x = unname(b$deciles),
      y = 0,
      label = names(b$deciles),
      color = "blue",
      vjust = 1.2,
      size = 3
    ) +
    ggplot2::scale_color_manual(
      values = stats::setNames(c("red", "blue"), c(a$name, b$name)),
      breaks = c(a$name, b$name)
    ) +
    ggplot2::labs(
      x = ifelse(is.null(x_label), var_name, x_label),
      y = "Density",
      caption = caption_text,
      title = title,
      color = NULL
    ) +
    ggplot2::theme_minimal() +
    theme(
      plot.caption = element_text(size = 8)
  )

  if (log_scale) {
    p <- p + ggplot2::scale_x_log10(labels = scales::label_number())
  }

  p
}

compare_densities_with_deciles(
  d_s_fire,
  d_s_nfire,
  biotic_relaxedforestnorm_5_yrs_prior_sum_yot,
  title = "Biotic prior sum: fire-only vs no fire",
  x_label = "Biotic 5-yr prior sum",
  log_scale = TRUE
)
ggsave(here(dir_figs, "density_biotic_prior.png"))

compare_densities_with_deciles(
  d_s_fire,
  d_s_nfire,
  biotic_6roll_max,
  title = "Biotic roll max : fire-only vs no fire",
  x_label = "Biotic 5-yr roll max",
  log_scale = TRUE
)
ggsave(here(dir_figs, "density_biotic_rollmax.png"))


compare_densities_with_deciles(
  d_s_fire,
  d_s_nfire,
  pdsi_annual_5_yrs_prior_sum_yot,
  title = "PDSI sum prior : fire-only vs no fire",
  x_label = "PDSI 5-yr sum",
  log_scale = FALSE
)
ggsave(here(dir_figs, "density_pdsi_prior.png"))


 # histograms for thresholded data?


```

```{r}

set_cd_groups <- function(df,
                          b_nm,
                          d_nm,
                          cd_nm,
                          b_threshold,
                          d_threshold) {
  
  if(d_threshold < 0) {
    df_new <- df |>
    dplyr::mutate(
      "{cd_nm}" := dplyr::case_when(
        .data[[b_nm]] <  b_threshold & fire == 1 & .data[[d_nm]] >  d_threshold ~ "f",
        .data[[b_nm]] >= b_threshold & fire == 1 & .data[[d_nm]] >  d_threshold ~ "bf",
        .data[[b_nm]] <  b_threshold & fire == 1 & .data[[d_nm]] <= d_threshold ~ "df",
        .data[[b_nm]] >= b_threshold & fire == 1 & .data[[d_nm]] <= d_threshold ~ "bdf",
        TRUE ~ NA_character_
      )
    )
  } else {
    df_new <- df |>
    dplyr::mutate(
      "{cd_nm}" := dplyr::case_when(
        .data[[b_nm]] <  b_threshold & fire == 1 & .data[[d_nm]] <  d_threshold ~ "f",
        .data[[b_nm]] >= b_threshold & fire == 1 & .data[[d_nm]] <  d_threshold ~ "bf",
        .data[[b_nm]] <  b_threshold & fire == 1 & .data[[d_nm]] >= d_threshold ~ "df",
        .data[[b_nm]] >= b_threshold & fire == 1 & .data[[d_nm]] >= d_threshold ~ "bdf",
        TRUE ~ NA_character_
      )
    )
    
  }

  
  return(df_new)
}



cd_group_table <- function(df, specs) {

  purrr::map_dfr(specs, function(s) {

    df2 <- set_cd_groups(
      df = df,
      b_nm = s$b_nm,
      d_nm = s$d_nm,
      cd_nm = s$cd_nm,
      b_threshold = s$b_threshold,
      d_threshold = s$d_threshold
    )

    # row counts
    counts_rows <- df2 |>
      count(.data[[s$cd_nm]], name = "n") |>
      mutate(group = .data[[s$cd_nm]]) |>
      select(group, n) |>
      tidyr::complete(group = c("f", "df", "bf", "bdf"), fill = list(n = 0)) |>
      tidyr::pivot_wider(names_from = group, values_from = n)

    # unique fire counts
    counts_fire <- df2 |>
      filter(!is.na(.data[[s$cd_nm]])) |>
      group_by(group = .data[[s$cd_nm]]) |>
      summarise(nf = n_distinct(fireid), .groups = "drop") |>
      tidyr::complete(group = c("f", "df", "bf", "bdf"), fill = list(nf = 0)) |>
      mutate(group = paste0(group, "_nf")) |>
      tidyr::pivot_wider(names_from = group, values_from = nf)

    out <- dplyr::bind_cols(counts_rows, counts_fire) |>
      mutate(
        cd_nm = s$cd_nm,
        b_nm = s$b_nm,
        d_nm = s$d_nm,
        b_threshold = s$b_threshold,
        d_threshold = s$d_threshold
      ) |>
      select(
        cd_nm, b_nm, b_threshold, d_nm, d_threshold,
        f, df, bf, bdf,
        f_nf, df_nf, bf_nf, bdf_nf
      ) |>
      mutate(
        total = f + df + bf + bdf,
        f_p   = round(100 * f   / total),
        df_p  = round(100 * df  / total),
        bf_p  = round(100 * bf  / total),
        bdf_p = round(100 * bdf / total)
      ) |>
      select(-total)

    out
  })
}

expand_cd_specs <- function(specs) {
  # specs: list of lists; each list can include:
  #   b_nm, d_nm, cd_prefix (optional), b_thresholds, d_thresholds
  #
  # returns a list of fully-expanded spec lists:
  #   b_nm, d_nm, cd_nm, b_threshold, d_threshold
  
  expanded <- purrr::map_dfr(specs, function(s) {
    cd_prefix <- s$cd_prefix %||% paste0("cd_", s$b_nm, "__", s$d_nm)

    tidyr::crossing(
      b_threshold = s$b_thresholds,
      d_threshold = s$d_thresholds
    ) |>
      mutate(
        b_nm = s$b_nm,
        d_nm = s$d_nm,
        cd_nm = paste0(
          cd_prefix,
          "_b", b_threshold,
          "_d", d_threshold
        )
      ) |>
      select(cd_nm, b_nm, d_nm, b_threshold, d_threshold)
  })

  # convert rows to list-of-lists (what cd_group_table expects)
  purrr::pmap(
    expanded,
    \(cd_nm, b_nm, d_nm, b_threshold, d_threshold)
      list(cd_nm = cd_nm, b_nm = b_nm, d_nm = d_nm,
           b_threshold = b_threshold, d_threshold = d_threshold)
  )
}

`%||%` <- function(x, y) if (!is.null(x)) x else y



specs_grid1 <- list(
  list(
    b_nm = "biotic_relaxedforestnorm_5_yrs_prior_sum_yot",
    d_nm = "pdsi_annual_5_yrs_prior_sum_yot",
    b_thresholds = c(10, 20, 50),
    d_thresholds = c(-10)
  )
)

specs_grid2 <- list(
  list(
    b_nm = "biotic_relaxedforestnorm_5_yrs_prior_sum_yot",
    d_nm = "pdsi_annual_5_yrs_prior_threshold_n4_yot",
    b_thresholds = c(10, 20, 50),
    d_thresholds = c(1, 2)
  )
)

specs_grid3 <- list(
  list(
    b_nm = "biotic_relaxedforestnorm_5_yrs_prior_sum_yot",
    d_nm = "pdsi_annual_5_yrs_prior_threshold_n3_yot",
    b_thresholds = c(10, 20, 50),
    d_thresholds = c(1, 2, 3)
  )
)

specs_grid4 <- list(
  list(
    b_nm = "biotic_relaxedforestnorm_5_yrs_prior_sum_yot",
    d_nm = "hd_fingerprint_5_yrs_prior_threshold_6_yot",
    b_thresholds = c(10, 20, 50),
    d_thresholds = c(1, 2)
  )
)

expanded_specs <- c(
  expand_cd_specs(specs_grid1),
  expand_cd_specs(specs_grid2),
  expand_cd_specs(specs_grid3),
  expand_cd_specs(specs_grid4)
)

cd_summary_grid <- cd_group_table(d_s, expanded_specs) |>
  arrange(b_threshold, d_nm) |>
  filter(!if_any(where(is.numeric), ~ . == 0)) |>
  select(-cd_nm)


cd_tbl <- tableGrob(
  cd_summary_grid,
  theme = ttheme_minimal(
    core = list(bg_params = list(fill = "white", col = NA)),
    colhead = list(bg_params = list(fill = "white", col = NA))
  )
)

cd_title <- textGrob(
  "Counts for possible CD groups",
  gp = gpar(fontsize = 16, fontface = "bold")
)

cd_table_with_title <- arrangeGrob(
  cd_tbl,
  top = cd_title
)

ggsave(
  here(dir_figs, "cd_group_counts.png"),
  cd_table_with_title,
  width = 20,
  height = 6,
  bg = "white"
)














d_s <- set_cd_groups(df = d_s,
                   b_nm = "biotic_relaxedforestnorm_5_yrs_prior_sum_yot",
                   d_nm = "pdsi_annual_5_yrs_prior_threshold_n4_yot",
                   cd_nm = "cd_group_b10_pdsin4t1",
                   b_threshold = 10,
                   d_threshold = 1)

d_s <- set_cd_groups(df = d_s,
                   b_nm = "biotic_relaxedforestnorm_5_yrs_prior_sum_yot",
                   d_nm = "pdsi_annual_5_yrs_prior_threshold_n3_yot",
                   cd_nm = "cd_group_b10_pdsin3t1",
                   b_threshold = 10,
                   d_threshold = 1)

d_s <- set_cd_groups(df = d_s,
                   b_nm = "biotic_relaxedforestnorm_5_yrs_prior_sum_yot",
                   d_nm = "pdsi_annual_5_yrs_prior_threshold_n3_yot",
                   cd_nm = "cd_group_b10_pdsin4t3",
                   b_threshold = 10,
                   d_threshold = 3)

d_s <- set_cd_groups(df = d_s,
                   b_nm = "biotic_relaxedforestnorm_5_yrs_prior_sum_yot",
                   d_nm = "pdsi_annual_5_yrs_prior_sum_yot",
                   cd_nm = "cd_group_b10_pdsisumn10",
                   b_threshold = 10,
                   d_threshold = -10)





# GAM METRIC DISTRIBUTIONS
# 
# plot_density_by_group_with_n <- function(df, x_var, group_var, x_label = NULL) {
# 
#   x_quo <- enquo(x_var)
#   g_quo <- enquo(group_var)
# 
#   x_name <- as_name(x_quo)
#   g_name <- as_name(g_quo)
# 
#   # Filter missing groups
#   d_plot <- df |>
#     filter(!is.na(!!g_quo))
# 
#   # Compute N per group
#   group_n <- d_plot |>
#     count(!!g_quo, name = "N") |>
#     mutate(label = paste0(!!g_quo, " (N=", N, ")"))
# 
#   lab_vec <- setNames(group_n$label, group_n[[g_name]])
# 
#   ggplot(d_plot, aes(x = !!x_quo, color = !!g_quo)) +
#     geom_density() +
#     scale_color_discrete(labels = lab_vec) +
#     labs(
#       x = ifelse(is.null(x_label), x_name, x_label),
#       color = g_name
#     ) +
#     theme_minimal()
# }
# 
# plot_density_by_group_with_n(
#   d_s,
#   gam_rap_tree_diff_pre6_to_post_min,
#   cd_group_b10_pdsin4t1,
#   x_label = "RAP total loss"
# )
# 
# plot_density_by_group_with_n(
#   d_s,
#   gam_rap_tree_diff_pre6_to_post_min,
#   cd_group_b10_pdsin3t1,
#   x_label = "RAP total loss"
# )
# 
# plot_density_by_group_with_n(
#   d_s,
#   gam_rap_tree_diff_pre6_to_post_min,
#   cd_group_b10_pdsin4t3,
#   x_label = "RAP total loss"
# )
# 
# 
# plot_density_by_group_with_n(
#   d_s,
#   gam_rap_tree_diff_pre6_to_post_min,
#   cd_group_b10_pdsisumn10,
#   x_label = "RAP total loss"
# )


# Boxplots
ggplot(
  d_s |>
    filter(!is.na(cd_group_b10_pdsin4t1)) |>
    add_count(cd_group_b10_pdsin4t1, name = "n_group"),
  aes(
    x = paste0(cd_group_b10_pdsin4t1, " (N=", n_group, ")"),
    y = gam_rap_tree_diff_pre6_to_post_min
  )
) +
  geom_boxplot() +
  labs(y = "RAP total loss") +
  theme_minimal()
ggsave(here(dir_figs, "boxplot_raploss_cd_group_b10_pdsin4t1.png"))


ggplot(
  d_s |>
    filter(!is.na(cd_group_b10_pdsin4t3)) |>
    add_count(cd_group_b10_pdsin4t3, name = "n_group"),
  aes(
    x = paste0(cd_group_b10_pdsin4t3, " (N=", n_group, ")"),
    y = gam_rap_tree_diff_pre6_to_post_min
  )
) +
  geom_boxplot() +
  labs(y = "RAP total loss") +
  theme_minimal()
ggsave(here(dir_figs, "boxplot_raploss_cd_group_b10_pdsin4t3.png"))


ggplot(
  d_s |>
    filter(!is.na(cd_group_b10_pdsisumn10)) |>
    add_count(cd_group_b10_pdsisumn10, name = "n_group"),
  aes(
    x = paste0(cd_group_b10_pdsisumn10, " (N=", n_group, ")"),
    y = gam_rap_tree_diff_pre6_to_post_min
  )
) +
  geom_boxplot() +
  labs(y = "RAP total loss") +
  theme_minimal()
ggsave(here(dir_figs, "boxplot_raploss_cd_group_b10_pdsisumn10.png"))




ggplot(
  d_s |>
    filter(!is.na(cd_group_b10_pdsin4t3)) |>
    add_count(cd_group_b10_pdsin4t3, name = "n_group"),
  aes(
    x = paste0(cd_group_b10_pdsin4t3, " (N=", n_group, ")"),
    y = gam_rap_tree_slope_post_min_to_p15
  )
) +
  geom_boxplot() +
  labs(y = "RAP min to 15 yr recovery slope") +
  theme_minimal()
ggsave(here(dir_figs, "boxplot_rapgamslope_cd_group_b10_pdsin4t3.png"))




ggplot(
  d_s |>
    filter(!is.na(cd_group_b10_pdsin4t3)) |>
    add_count(cd_group_b10_pdsin4t3, name = "n_group"),
  aes(
    x = paste0(cd_group_b10_pdsin4t3, " (N=", n_group, ")"),
    y = raw_post_fire_rap_tree_slope_3_15_yrs
  )
) +
  geom_boxplot() +
  labs(y = "RAP 3-15 yr recovery slope") +
  theme_minimal()
ggsave(here(dir_figs, "boxplot_raprawslope_cd_group_b10_pdsin4t3.png"))




# 
# 
# 
# bin_cd_and_boxplot <- function(df, cd_nm = cd_nm, outcome = outcome, fire = fire, n_bins = 10) {
# 
#   cd_quo <- rlang::enquo(cd_nm)
#   out_quo <- rlang::enquo(outcome)
#   fire_quo <- rlang::enquo(fire)
# 
#   cd_name <- rlang::as_name(cd_quo)
# 
#   x <- df[[cd_name]]
# 
#   if (inherits(x, "integer")) {
# 
#     df2 <- df |>
#       dplyr::mutate(
#         bin_cd = ifelse(!!fire_quo == 0, NA, as.character(!!cd_quo))
#       )
# 
#   } else {
# 
#     rng <- range(x, na.rm = TRUE, finite = TRUE)
# 
#     # if there isn't a usable range, bins can't be formed
#     if (!is.finite(rng[1]) || !is.finite(rng[2]) || rng[1] == rng[2]) {
#       df2 <- df |>
#         dplyr::mutate(bin_cd = NA_character_)
#     } else {
# 
#       breaks <- seq(from = rng[1], to = rng[2], length.out = n_bins + 1)
# 
#       labels <- paste0(
#         round(head(breaks, -1), 1),
#         "–",
#         round(tail(breaks, -1), 1)
#       )
# 
#       df2 <- df |>
#         dplyr::mutate(
#           bin_cd = cut(
#             !!cd_quo,
#             breaks = breaks,
#             include.lowest = TRUE,
#             labels = labels
#           ),
#           bin_cd = ifelse(!!fire_quo == 0, NA, as.character(bin_cd))
#         )
#     }
#   }
# 
#   df2 <- df2 |> dplyr::mutate(bin_cd = factor(bin_cd))
# 
#   p <- ggplot2::ggplot(df2 |> dplyr::filter(!is.na(bin_cd)),
#               ggplot2::aes(x = bin_cd, y = !!out_quo)) +
#     ggplot2::geom_boxplot(outlier.alpha = 0.2) +
#     ggplot2::labs(
#       x = paste0(cd_name, " bins"),
#       y = rlang::as_name(out_quo)
#     ) +
#     ggplot2::theme_minimal() +
#     ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 30, hjust = 1))
# 
#   list(data = df2, plot = p)
# }
# 
# 
# res <- bin_cd_and_boxplot(d_s, biotic_relaxedforestnorm_5_yrs_prior_sum_yot, gam_rap_tree_diff_pre6_to_post_min)
# res$plot
# 
# res <- bin_cd_and_boxplot(d_s, pdsi_annual_5_yrs_prior_threshold_n3_yot, gam_rap_tree_diff_pre6_to_post_min)
# res$plot
# 
# 
# res <- bin_cd_and_boxplot(d_s, pdsi_annual_5_yrs_prior_threshold_n4_yot, gam_rap_tree_diff_pre6_to_post_min)
# res$plot
# 
# res <- bin_cd_and_boxplot(d_s, pdsi_annual_5_yrs_prior_sum_yot, gam_rap_tree_diff_pre6_to_post_min)
# res$plot
# 
# res <- bin_cd_and_boxplot(d_s,  hd_fingerprint_5_yrs_prior_threshold_6_yot, gam_rap_tree_diff_pre6_to_post_min)
# res$plot
# 



```

# VCF vs RAP plots

Both metrics & raw data

Takeaways: we find that RAP aggregated to 250m and VCF have RMSE of around 12.4. This remains similar across subgroups of data, although there is a slightly lower error in prefire and no-fire environments than postfire (\~12 vs \~13). The postfire error climbs slightly with vcf above 50, to \~14.

In terms of estimating resilience metrics, there is higher agreement between VCF and RAP when estimating pre-fire cover using an average window or GAM (as opposed to using single raw values). In estimating total loss, recovery, slope, etc, averaged windows provide lower RMSE between the two datasets, likely due to the fact that the GAMs are identifying different minimum values post fire between the two datasets.

```{r}

plot_lm_r2 <- function(df, x, y, title, xlim = NULL, ylim = NULL) {
  p <- ggplot(df, aes(x = {{ x }}, y = {{ y }})) +
    geom_point(alpha = 0.1) +
    geom_smooth(method = "lm", alpha = 0.1) +
    ggpmisc::stat_poly_eq(
      formula = y ~ x,
      aes(label = after_stat(rr.label)),
      parse = TRUE,
      label.x = "right",
      label.y = "bottom"
    ) +
    labs(title = title)

  if (!is.null(xlim)) p <- p + xlim(xlim)
  if (!is.null(ylim)) p <- p + ylim(ylim)

  p
}

extract_r2_rmse_table <- function(models) {
  tibble(
    model_name = names(models),
    model = models
  ) |>
    mutate(
      stats = purrr::map(model, broom::glance),
      rmse = purrr::map_dbl(model, \(m) {
        sqrt(mean(residuals(m)^2, na.rm = TRUE))
      })
    ) |>
    tidyr::unnest(stats) |>
    select(model_name, r.squared, rmse)
}

lm_safe <- function(formula, data) {
  vars <- all.vars(formula)
  d <- data |>
    dplyr::select(dplyr::all_of(vars)) |>
    dplyr::filter(dplyr::if_all(dplyr::everything(), is.finite))
  stats::lm(formula, data = d)
}

# 
# p_loss <- plot_lm_r2(
#   d_s,
#   gam_rap_tree_diff_pre6_to_post_min,
#   gam_vcf_tree_diff_pre6_to_post_min,
#   "VCF vs RAP total loss 250m"
# )
# 
# p_loss_raw <- plot_lm_r2(
#   d_s,
#   raw_post_fire_rap_tree_difference_n6_3_yr,
#   raw_post_fire_vcf_tree_difference_n6_3_yr,
#   "VCF vs RAP total loss 250m -6 to 3 yr raw windows"
# )
# 
# 
# 
# p_recov15 <- plot_lm_r2(
#   d_s,
#   gam_rap_tree_perc_recov_p15,
#   gam_vcf_tree_perc_recov_p15,
#   "VCF vs RAP recovery percent min-15yr 250m",
#   xlim = c(0, 100),
#   ylim = c(0, 100)
# )
# 
# p_raw_recov15 <- plot_lm_r2(
#   d_s,
#   raw_post_fire_rap_tree_perc_recovery_n6_3_15_yr,
#   raw_post_fire_vcf_tree_perc_recovery_n6_3_15_yr,
#   "VCF vs RAP raw recovery percent 3-15yr 250m",
#   xlim = c(0, 100),
#   ylim = c(0, 100)
# )
# 
# 
# p_slope15 <- plot_lm_r2(
#   d_s,
#   gam_rap_tree_slope_post_min_to_p15,
#   gam_vcf_tree_slope_post_min_to_p15,
#   "VCF vs RAP slope gam min to 15yr slope 250m"
# )
# 
# p_rawslope15 <- plot_lm_r2(
#   d_s,
#   raw_post_fire_rap_tree_slope_3_15_yrs,
#   raw_post_fire_vcf_tree_slope_3_15_yrs,
#   "VCF vs RAP slope raw 3 to 15yr slope 250m"
# )
# 
# 
# 
# p_pre6 <- plot_lm_r2(
#   d_s,
#   gam_rap_tree_pre6_fit,
#   gam_vcf_tree_pre6_fit,
#   "VCF vs RAP 6-yr prior fit 250m"
# )
# 
# p_pre6raw <- plot_lm_r2(
#   d_l |> filter(Time_to_Treatment == -6),
#   rap_tree,
#   vcf_tree,
#   "VCF vs RAP 6-yr-prior raw 250m"
# )
# 
# p_pre6raw_win <- plot_lm_r2(
#   d_s,
#   raw_avg_rap_tree_6_yr_pre_fire,
#   raw_avg_vcf_tree_6_yr_pre_fire,
#   "VCF vs RAP 6-yr-prior-window 250m"
# )
# 
# 
# 
# p_min <- plot_lm_r2(
#   d_s,
#   gam_rap_tree_post_min,
#   gam_vcf_tree_post_min,
#   "VCF vs RAP minimum fit 250m"
# )
# 
# p_3yr_raw <- plot_lm_r2(
#   d_s,
#   raw_avg_rap_tree_3_yr_post_fire,
#   raw_avg_vcf_tree_3_yr_post_fire,
#   "VCF vs RAP post-fire 3 yr raw"
# )
# 
# 
# 
# p_recov_tot <- plot_lm_r2(
#   d_s,
#   gam_rap_tree_diff_post_min_to_p15,
#   gam_vcf_tree_diff_post_min_to_p15,
#   "VCF vs RAP post-fire min to 15 gam"
# )
# 
# p_recov_tot_raw <- plot_lm_r2(
#   d_s,
#   raw_post_fire_rap_tree_increase_3_15_yr,
#   raw_post_fire_vcf_tree_increase_3_15_yr,
#   "VCF vs RAP post-fire 3 to 15 raw"
# )
# 
# 
# p_timetomin <- plot_lm_r2(
#   d_s,
#   gam_rap_tree_yrs_ref_to_post_min,
#   gam_vcf_tree_yrs_ref_to_post_min,
#   "VCF vs RAP time to minimum 250m"
# )
# 
# 
# 
# # Print
# p_loss
# p_loss_raw
# 
# p_recov15
# p_raw_recov15
# 
# p_slope15
# p_rawslope15
# 
# p_pre6
# p_pre6raw
# p_pre6raw_win
# 
# p_min
# p_3yr_raw
# 
# p_recov_tot
# p_recov_tot_raw
# 
# p_timetomin
# 
# ggplot(d_s) +
#   geom_bin2d(aes(x = gam_rap_tree_yrs_ref_to_post_min, y = gam_vcf_tree_yrs_ref_to_post_min))
# 




set.seed(1234)
d_s_2006 <- d_s |> dplyr::filter(burn_year > 2005)
d_l_2006 <- d_l |> dplyr::filter(burn_year > 2005)

models_xy <- list(
  loss_gam = lm_safe(gam_vcf_tree_diff_pre6_to_post_min ~ gam_rap_tree_diff_pre6_to_post_min, d_s_2006),
  loss_raw = lm_safe(raw_post_fire_vcf_tree_difference_n6_3_yr ~ raw_post_fire_rap_tree_difference_n6_3_yr, d_s_2006),

  recov15_gam = lm_safe(gam_vcf_tree_perc_recov_p15 ~ gam_rap_tree_perc_recov_p15, d_s_2006),
  recov15_raw = lm_safe(raw_post_fire_vcf_tree_perc_recovery_n6_3_15_yr ~ raw_post_fire_rap_tree_perc_recovery_n6_3_15_yr, d_s_2006),

  slope15_gam = lm_safe(gam_vcf_tree_slope_post_min_to_p15 ~ gam_rap_tree_slope_post_min_to_p15, d_s_2006),
  slope15_raw = lm_safe(raw_post_fire_vcf_tree_slope_3_15_yrs ~ raw_post_fire_rap_tree_slope_3_15_yrs, d_s_2006),

  pre6_fit_gam = lm_safe(gam_vcf_tree_pre6_fit ~ gam_rap_tree_pre6_fit, d_s_2006),
  pre6_raw_tminus6 = lm_safe(vcf_tree ~ rap_tree, d_l_2006 |> dplyr::filter(Time_to_Treatment == -6)),
  pre6_raw_window = lm_safe(raw_avg_vcf_tree_6_yr_pre_fire ~ raw_avg_rap_tree_6_yr_pre_fire, d_s_2006),

  postmin_gam = lm_safe(gam_vcf_tree_post_min ~ gam_rap_tree_post_min, d_s_2006),
  post3yr_raw = lm_safe(raw_avg_vcf_tree_3_yr_post_fire ~ raw_avg_rap_tree_3_yr_post_fire, d_s_2006),

  recov_tot_gam = lm_safe(gam_vcf_tree_diff_post_min_to_p15 ~ gam_rap_tree_diff_post_min_to_p15, d_s_2006),
  recov_tot_raw = lm_safe(raw_post_fire_vcf_tree_increase_3_15_yr ~ raw_post_fire_rap_tree_increase_3_15_yr, d_s_2006),

  time_to_min = lm_safe(gam_vcf_tree_yrs_ref_to_post_min ~ gam_rap_tree_yrs_ref_to_post_min, d_s_2006)
)

r2_rmse_table <- extract_r2_rmse_table(models_xy)

r2_rmse_table_g <- tableGrob(
  r2_rmse_table,
  theme = ttheme_minimal(
    core = list(bg_params = list(fill = "white", col = NA)),
    colhead = list(bg_params = list(fill = "white", col = NA))
  )
)

title1 <- textGrob(
  "VCF-RAP resilience metric \nR² and RMSE Comparisons \n(burn year >2005)",
  gp = gpar(fontsize = 16, fontface = "bold")
)

table_with_title1 <- arrangeGrob(
  r2_rmse_table_g,
  top = title1
)

ggsave(
  here(dir_figs, "rap_vs_vfc_metrics_table.png"),
  table_with_title1,
  width = 6,
  height = 7,
  bg = "white"
)




# Comparing raw data



ggplot(d_l) +
  geom_density(aes(x = rap_tree), color = "red") +
  geom_density(aes(x = vcf_tree), color = "blue") +
  labs(xlab = "Percent tree cover", caption = "RAP = red, VCF = blue")
ggsave(here(dir_figs, "raw_rap_vcf_density.png"))

set.seed(1234)
models <- list(
  all = lm(rap_tree ~ vcf_tree, d_l |> slice_sample(n = 1000000)),
  nofire = lm(rap_tree ~ vcf_tree, d_l |> filter(fire == 0) |> slice_sample(n = 1000000)),
  prefire = lm(rap_tree ~ vcf_tree, d_l |> filter(fire == 1 & treated == 0) |> slice_sample(n = 1000000)),
  postfire = lm(rap_tree ~ vcf_tree, d_l |> filter(fire == 1 & treated == 1) |> slice_sample(n = 1000000)),
  biotic_above10 = lm(rap_tree ~ vcf_tree, d_l |> filter(fire == 0 & biotic_6roll_max > 10) |> slice_sample(n = 1000000)),
  all_vcf_above50 = lm(rap_tree ~ vcf_tree, d_l |> filter(vcf_tree > 50) |> slice_sample(n = 1000000)),
  all_vcf_below50 = lm(rap_tree ~ vcf_tree, d_l |> filter(vcf_tree <= 50) |> slice_sample(n = 1000000)),
  postfire_vcf_above50 = lm(rap_tree ~ vcf_tree, d_l |> filter(vcf_tree > 50 & fire == 1 & treated == 1) |> slice_sample(n = 1000000)),
  postfire_vcf_below50 = lm(rap_tree ~ vcf_tree, d_l |> filter(vcf_tree <= 50 & fire == 1 & treated == 1) |> slice_sample(n = 1000000))
)


tbl <- tableGrob(
  extract_r2_rmse_table(models),
  theme = ttheme_minimal(
    core = list(bg_params = list(fill = "white", col = NA)),
    colhead = list(bg_params = list(fill = "white", col = NA))
  )
)

title <- textGrob(
  "Raw VCF-RAP data R² and RMSE Comparisons\n1 million samples per model",
  gp = gpar(fontsize = 16, fontface = "bold")
)

table_with_title <- arrangeGrob(
  tbl,
  top = title
)

ggsave(
  here(dir_figs, "rap_vs_vfc_raw_table.png"),
  table_with_title,
  width = 6,
  height = 4,
  bg = "white"
)





```

```{r}

dats_quick <- d_l |>
  filter(!is.na(cd_group)) |>
  slice_head(n = 100000) |>
  mutate(fire_frg = interaction(burn_sev, frg_reclass_desc),
         fire_pre_drought = interaction(burn_sev, pdsi_annual_5_yrs_prior_threshold),
         fire_pre_drought = interaction(burn_sev, pdsi_annual_5_yrs_after_threshold),
         biotic_cut = cut(biotic,
             breaks = c(0, 20, 40, 60, 80, 100),
             labels = c()
             )
         )

ggplot(dats_quick) +
  geom_smooth(method = 'gam', aes(x = Time_to_Treatment, y = rap_tree, color = cd_group)) +
  xlim(c(-10, 20)) +
  labs(title = "GAM fits sample")

ggplot(dats_quick) +
  geom_smooth(method = 'gam', aes(x = Time_to_Treatment, y = vcf_tree, color = cd_group)) +
  xlim(c(-10, 20)) +
  labs(title = "GAM fits sample")

ggplot(dats_quick |> filter(burn_sev != "Negligible")) +
  geom_smooth(method = 'gam', aes(x = Time_to_Treatment, y = rap_tree, color = frg_reclass_desc, lty = burn_sev)) +
  xlim(c(-10, 20)) +
  labs(title = "GAM fits sample")


```

<button class="accordion-button" type="button" data-bs-toggle="collapse" data-bs-target="#collapseOne">

Click to see Session Information

</button>

:::: {#collapseOne .accordion-collapse .collapse}
<div>

```{r sessionInfo}
sessionInfo()
```

</div>
::::
