---
title: "04_causal_forest"
author: Tyler L. McIntosh
date: last-modified
format:
  html:
    embed-resources: true
    html-math-method: katex
    css: styles.css
    toc: true
    toc-depth: 1
    toc-title: Contents
    toc-location: right
    code-fold: true
    code-summary: "Show the code"
---

# Setup for analysis

``` {r setupchunk}
#| echo: true
#| message: false
#| warning: false
#| results: 'hide'

rm(list = ls())

if(!requireNamespace("here", quietly = TRUE)) {
  install.packages("here")
}
library(here)

source(here::here("utils", "functions.R"))

install_load_packages(c(
  "tidyverse",
  "tictoc",
  "grf",
  "kernelshap",
  "hstats",
  "glue",
  "shapviz",
  "patchwork",
  "cobalt",
  "purrr",
  "corrplot",
  "gridExtra",
  "arrow",
  "tictoc"))

devtools::install_github("grf-labs/sufrep")
library(sufrep)

dir_figs <- here::here('figs')
dir_derived <- here::here('data', 'derived')
dir_raw <- here::here('data', 'raw')
dir_manual <- here::here('data', 'manual')
dir_ensure(c(dir_figs,
             dir_derived,
             dir_manual,
             dir_raw))

seed = 1234
set.seed(seed)


```
# Load dats

```{r}

#Short dats
dats <- read_parquet(here(dir_derived, "test_dats_short.parquet"))
ecoregions <- unique(dats$ecoregion)

```

# Sufrep function
```{r}

# A function to use sufrep::make_encoder such that the encoded variables are added to the dataframe
sufrep_encode_append <- function(dats,
                                 method = 'means',
                                 X_cols,
                                 G_col,
                                 remove_G_col = TRUE,
                                 ...) {
  
  #Select variables for encoding
  x_df <- dats |> 
    select(all_of(X_cols)) |>
    as.data.frame()
  g_df <- dats |>
    pull(G_col)
  
  #Encode
  encoder <- sufrep::make_encoder(method = method, X = x_df, G = g_df, ...)
  dats_encoded <- encoder(x_df, G = g_df)
  
  #Make interpretable encoding columns
  encoded_cols <- dats_encoded |> select(starts_with("ENC"))
  names(encoded_cols) <- paste0(names(encoded_cols), "_", G_col)
  
  dats_out <- dats |>
    cbind(encoded_cols)
  
  if(remove_G_col) {
    dats_out <- dats_out |>
      select(-all_of(G_col))
  }
  
  return(dats_out)
}


```

# Prep data for causal forest machine learning
- Normalize continuous covariates to control for covariates with different numerical scales (avoid large-scale variables dominating and stabilize splits)
- Encode categorical variables as a lower-dimensional real-valued representation for inclusion in causal forest (use sufficient representations instead of one-hot encoding since random forests "may largely ignore high-cardinality one-hot encoded factors" - see Johannemann et al 2021). Low-rank encoding appears to be the most applicable to real-world data.

```{r}

cont_covars <- c("aet", "def", "srtm", "tpi", "chili", "avg_rap_TRE_6_yr_pre_fire")
cat_covars <- c("frg_reclass", "mock_burn_year", "ecoregion")
outcomes <- dats |>
  select(starts_with("post_fire")) |>
  names()
trts <- c("fire", "burn_sev")



all_cols <- c(cont_covars, cat_covars, outcomes, trts)

# Scale all continuous covariates
# filter out anything with NAs
# ensure categorical covariates are factors
# encode categorical as one-hot dummies or sufficient representations
dats_ml <- dats |>
  select(all_of(all_cols)) |>
  filter(dplyr::if_all(everything(), ~ !is.na(.x))) |>
  mutate(
    across(
      all_of(cont_covar),
      \(x) as.numeric(scale(x))
    )
  ) |>
  mutate(
    across(
      all_of(cat_covar),
      \(x) as.factor(x)
    )
  ) |>
  # fastDummies::dummy_cols(select_columns = c("frg_reclass"),
  #                         remove_first_dummy = FALSE, #Keep full one-hot encoding for tree-based methods
  #                         remove_selected_columns = TRUE) |> 
  sufrep_encode_append(method = 'low_rank', # Use low-rank sufficient representation encoding to reduce to 2 dimensions
                       X_cols = cont_covars,
                       G_col = "mock_burn_year",
                       num_components = 2) |> 
  sufrep_encode_append(method = 'low_rank', # Use low-rank sufficient representation encoding to reduce to 2 dimensions
                       X_cols = cont_covars,
                       G_col = "frg_reclass",
                       num_components = 2)

enc_xvars <- dats_ml |>
  select(starts_with("ENC")) |>
  names()

```


# Test model tuning time bumps

Tuning all parameters as opposed to a select few increases training time by ~50%.

GRF package tuning parameters:

tune.parameters	
A vector of parameter names to tune. If "all": all tunable parameters are tuned by cross-validation. The following parameters are tunable: ("sample.fraction", "mtry", "min.node.size", "honesty.fraction", "honesty.prune.leaves", "alpha", "imbalance.penalty"). If honesty is FALSE the honesty.* parameters are not tuned. Default is "none" (no parameters are tuned).

tune.num.trees	
The number of trees in each 'mini forest' used to fit the tuning model. Default is 200.

tune.num.reps	
The number of forests used to fit the tuning model. Default is 50.

tune.num.draws	
The number of random parameter values considered when using the model to select the optimal parameters. Default is 1000.

```{r}

df_test <- dats_ml |> filter(ecoregion == "Idaho Batholith")


tic() # 253 seconds
cf <- causal_forest(
  X=df_test |> select(-fire, -post_fire_rap_TRE_difference_n2_3_yr, -ecoregion),
  Y=df_test$post_fire_rap_TRE_difference_n2_3_yr,
  W=df_test$fire,
  tune.parameters = c("min.node.size","sample.fraction", "mtry"),
  tune.num.trees = 1000,
  tune.num.reps = 50,
  tune.num.draws = 1000,
  num.trees=2000,
  seed=seed
)
toc()


tic() # 390 seconds
cf <-causal_forest(
  X=df_test |> select(-fire, -post_fire_rap_TRE_difference_n2_3_yr, -ecoregion),
  Y=df_test$post_fire_rap_TRE_difference_n2_3_yr,
  W=df_test$fire,
  tune.parameters = c("all"),
  tune.num.trees = 1000,
  tune.num.reps = 50,
  tune.num.draws = 1000,
  num.trees=2000,
  seed=seed
)
toc()


imp_raw <- sort(setNames(variable_importance(cf), names(df_test |> select(-fire, -post_fire_rap_TRE_difference_n2_3_yr, -ecoregion))))
imp_normalized <- imp_raw / sum(imp_raw)

imp_df <- tibble(
  variable = names(imp_normalized),
  importance = as.numeric(imp_normalized)
) %>%
  arrange(importance)  # So the smallest is at the bottom of the horizontal bar plot

imp_p <- ggplot(imp_df, aes(x = importance, y = reorder(variable, importance))) +
  geom_col(fill = "orange") +
  labs(
    x = "Variable Importance (Normalized)",
    y = NULL,
    title = "Causal Forest Variable Importance"#,
    #subtitle = "test",
    #caption = glue("Total sample size: {n_total}\nBurned sample size: {n_burned}")
  ) +
  theme_minimal(base_size = 13) +
  theme(
    axis.text.y = element_text(hjust = 1),
    plot.margin = margin(10, 20, 5, 5)
  )

imp_p
```



# Variable interpretation
- TPI: low = valleys, high = ridges
- CHILI: low = cold, high = hot
- SRTM: low = low elevation, high = high elevation
- AET: low = wet, high = dry
- DEF: low = wet, high = dry

```{r}


cf_analysis <- function(dats,
                        yvar,
                        xvars,
                        trt,
                        data_nm,
                        seed = 1234,
                        dir_out,
                        ...) {
  # Filter out rows with NA in the outcome variable
  dats_filt <- dats %>%
    filter(!is.na(.data[[yvar]]))
  
  n_total <- nrow(dats_filt)
  n_burned <- nrow(dats_filt |>
                      filter(fire == 1))
  
  # Select covariates and outcome
  covariates <- dats_filt %>%
    select(all_of(xvars))
  outcome <- dats_filt[[yvar]]
  treatment <- dats_filt[[trt]]

  # Fit causal forest
  fit <- grf::causal_forest(
    X = covariates,
    Y = outcome,
    W = treatment,
    seed = seed,
    ...
  )

  # Variable importance - ensure normalized
  imp_raw <- sort(setNames(variable_importance(fit), xvars))
  imp_normalized <- imp_raw / sum(imp_raw)
  
  imp_df <- tibble(
    variable = names(imp_normalized),
    importance = as.numeric(imp_normalized)
  ) %>%
    arrange(importance)  # So the smallest is at the bottom of the horizontal bar plot

  imp_p <- ggplot(imp_df, aes(x = importance, y = reorder(variable, importance))) +
    geom_col(fill = "orange") +
    labs(
      x = "Variable Importance (Normalized)",
      y = NULL,
      title = glue("Causal Forest Variable Importance, trt = {yvar}"),
      subtitle = data_nm,
      caption = glue("Total sample size: {n_total}\nBurned sample size: {n_burned}")
    ) +
    theme_minimal(base_size = 13) +
    theme(
      axis.text.y = element_text(hjust = 1),
      plot.margin = margin(10, 20, 5, 5)
    )
  ggsave(plot = imp_p,
         filename = here(dir_out, glue("var_importance_{data_nm}_{yvar}.png")),
         bg = "white")

  pred_fun <- function(object, newdata, ...) {
    predict(object, newdata, ...)$predictions
  }

  xvars_sorted <- rev(names(imp_normalized))

  pdps <- lapply(xvars_sorted, function(v) {
    pd <- partial_dep(fit, v, X = covariates, pred_fun = pred_fun)
    plot(pd) +
      theme(axis.title.y = element_blank())
  })

  #For adding y axis label to overall plot
  ylab_plot <- ggplot() +
    theme_void() +
    annotate("text", x = 0, y = 0.5, label = "Treatment effect", angle = 90, hjust = 0.5, size = 4)

  pdps_p <- wrap_plots(pdps, guides = "collect", ncol = 3) #&
    #ylab("Treatment effect")

  # Add format & add titles
  pdps_p <- ylab_plot + pdps_p +
    plot_layout(widths = c(0.05, 1)) +
    plot_annotation(
      title = glue("Partial Dependence Plots: trt = {yvar}"),
      subtitle = data_nm,
      caption = glue("NOTES:\nNegative values mean a stronger (positive) treatment effect (i.e. fire is causing a greater difference in forest cover change)\nPDPs are in order of variable importance\nTotal sample size: {n_total}\nBurned sample size: {n_burned}"))

  ggsave(filename = here(dir_out, glue("partial_dependence_{data_nm}_{yvar}.png")),
         plot = pdps_p,
         units = 'px',
         width = 2000,
         height = 2000)

  return(list(fit = fit,
              imp_p = imp_p,
              pdps = pdps))
}



dir_test <- here(dir_figs, "dir_test_causal_multi")
dir_ensure(dir_test)

# 
tic() #takes about 4.25 minutes to run for ID Batholith
test <- cf_analysis(dats = df_test,
            yvar = "post_fire_rap_TRE_difference_n2_3_yr",
            xvars = c("aet", "def", "srtm", "tpi", "chili", "avg_rap_TRE_6_yr_pre_fire",
                      enc_xvars),
            trt = "fire",
            data_nm = "idBatholith",
            seed = seed,
            dir_out = dir_test,
            tune.parameters = c("all"),
            tune.num.trees = 1000,
            tune.num.reps = 50,
            tune.num.draws = 1000,
            num.trees = 2000
           )
toc()


tic() #takes about 4.25 minutes to run for ID Batholith
test <- cf_analysis(dats = df_test,
            yvar = "post_fire_rap_TRE_difference_n2_3_yr",
            xvars = c("aet", "def", "srtm", "tpi", "chili", "avg_rap_TRE_6_yr_pre_fire",
                      enc_xvars),
            trt = "fire",
            data_nm = "idBatholith",
            seed = seed,
            dir_out = dir_test,
            tune.parameters = c("all"),
            tune.num.trees = 1000,
            tune.num.reps = 50,
            tune.num.draws = 1000,
            num.trees = 2000
           )
toc()

tic() #takes about 4.25 minutes to run for ID Batholith
test <- cf_analysis(dats = df_test,
            yvar = "post_fire_rap_TRE_difference_n2_5_yr",
            xvars = c("aet", "def", "srtm", "tpi", "chili", "avg_rap_TRE_6_yr_pre_fire",
                      enc_xvars),
            trt = "fire",
            data_nm = "idBatholith",
            seed = seed,
            dir_out = dir_test,
            tune.parameters = c("all"),
            tune.num.trees = 1000,
            tune.num.reps = 50,
            tune.num.draws = 1000,
            num.trees = 2000
           )
toc()

tic() #takes about 4.25 minutes to run for ID Batholith
test <- cf_analysis(dats = df_test,
            yvar = "post_fire_rap_TRE_difference_n2_10_yr",
            xvars = c("aet", "def", "srtm", "tpi", "chili", "avg_rap_TRE_6_yr_pre_fire",
                      enc_xvars),
            trt = "fire",
            data_nm = "idBatholith",
            seed = seed,
            dir_out = dir_test,
            tune.parameters = c("all"),
            tune.num.trees = 1000,
            tune.num.reps = 50,
            tune.num.draws = 1000,
            num.trees = 2000
           )
toc()

tic() #takes about 4.25 minutes to run for ID Batholith
test <- cf_analysis(dats = df_test,
            yvar = "post_fire_rap_TRE_difference_n2_15_yr",
            xvars = c("aet", "def", "srtm", "tpi", "chili", "avg_rap_TRE_6_yr_pre_fire",
                      enc_xvars),
            trt = "fire",
            data_nm = "idBatholith",
            seed = seed,
            dir_out = dir_test,
            tune.parameters = c("all"),
            tune.num.trees = 1000,
            tune.num.reps = 50,
            tune.num.draws = 1000,
            num.trees = 2000
           )
toc()


# 
# 
# 
# tic() #takes about 1.5 minutes to run
# test <- cf_analysis(dats = dats |> filter(ecoregion = "IDBatholith"),
#             yvar = "post_fire_forest_slope_5_10_yrs",
#             xvars = c("aet", "def", "frg_reclass", "srtm", "tpi", "chili",
#            "burn_cbi", "burn_cbi_1kmmwmedian", "burn_cbi_1kmmwstd", "mock_burn_year",
#            "biotic_5_yrs_prior_sum", "biotic_5_yrs_after_sum",
#            "pdsi_annual_5_yrs_prior_sum", "pdsi_annual_5_yrs_after_sum",
#            "avg_forest_2_yr_pre_fire"),
#             trt = "fire",
#             data_nm = "idBatholith_post_fire_forest_slope_5_10_yrs",
#             seed = seed)
# toc()
# 
# tic() 
# test <- cf_analysis(dats = dats |> filter(ecoregion = "IDBatholith"),
#             yvar = "post_fire_forest_difference_10_yr",
#             xvars = c("aet", "def", "frg_reclass", "srtm", "tpi", "chili",
#            "burn_cbi_1kmmwmedian", "burn_cbi_1kmmwstd", "mock_burn_year",
#            "biotic_5_yrs_prior_sum", "biotic_5_yrs_after_sum",
#            "pdsi_annual_5_yrs_prior_sum", "pdsi_annual_5_yrs_after_sum",
#            "avg_forest_2_yr_pre_fire"),
#             trt = "burn_cbi",
#             data_nm = "CBITRT_idBatholith_post_fire_forest_difference_10_yr",
#             seed = seed)
# toc()




# OUTCOMES OF INTEREST
outcome_vars <- c("perc_recovery_10yr", "post_fire_forest_slope_2_10_yrs", "post_fire_forest_slope_5_15_yrs", "post_fire_forest_slope_10_15_yrs", "perc_recovery_15yr", "post_fire_forest_slope_5_10_yrs")

# Switch to use pwalk from purrr for ease of switch to parallelized furrr in future
# Create combinations of ecoregions and outcome variables
param_grid <- tidyr::expand_grid(
  eco = ecoregions,
  var = outcome_vars
)

# Walk through each combination and call cf_analysis() for side effects
pwalk(
  .l = list(eco = param_grid$eco, var = param_grid$var),
  .f = function(eco, var) {
    d <- dats |> filter(ecoregion == eco)

    # Binary fire treatment
    cf_analysis(
      dats = d,
      yvar = var,
      xvars = c("aet", "def", "frg_reclass", "srtm", "tpi", "chili",
                "burn_cbi", "burn_cbi_1kmmwmedian", "burn_cbi_1kmmwstd", "mock_burn_year",
                "biotic_5_yrs_prior_sum", "biotic_5_yrs_after_sum",
                "pdsi_annual_5_yrs_prior_sum", "pdsi_annual_5_yrs_after_sum",
                #"pdsi_annual_5_yrs_prior_mean", "pdsi_annual_5_yrs_after_mean",
                "hd_fingerprint_5_yrs_prior_sum", "hd_fingerprint_5_yrs_after_sum",
                #"hd_fingerprint_5_yrs_prior_mean", "hd_fingerprint_5_yrs_after_mean",
                #"avg_forest_2_yr_pre_fire"
                ),
      trt = "fire",
      data_nm = glue("{eco}_{var}_firebinary"),
      seed = seed
    )
    
    # Continuous CBI treatment
    cf_analysis(
      dats = d,
      yvar = var,
      xvars = c("aet", "def", "frg_reclass", "srtm", "tpi", "chili",
                "burn_cbi", "burn_cbi_1kmmwmedian", "burn_cbi_1kmmwstd", "mock_burn_year",
                "biotic_5_yrs_prior_sum", "biotic_5_yrs_after_sum",
                "pdsi_annual_5_yrs_prior_sum", "pdsi_annual_5_yrs_after_sum",
                #"pdsi_annual_5_yrs_prior_mean", "pdsi_annual_5_yrs_after_mean",
                "hd_fingerprint_5_yrs_prior_sum", "hd_fingerprint_5_yrs_after_sum",
                #"hd_fingerprint_5_yrs_prior_mean", "hd_fingerprint_5_yrs_after_mean",
                "avg_forest_2_yr_pre_fire"),
      trt = "burn_cbi",
      data_nm = glue("{eco}_{var}_cbi"),
      seed = seed
    )
  }
)

# RUN FOR FULL DATASET
walk(
  .x = outcome_vars,
  .f = function(var) {
    cf_analysis(
      dats = dats,
      yvar = var,
      xvars = c("aet", "def", "frg_reclass", "srtm", "tpi", "chili",
                "burn_cbi", "burn_cbi_1kmmwmedian", "burn_cbi_1kmmwstd", "mock_burn_year",
                "biotic_5_yrs_prior_sum", "biotic_5_yrs_after_sum",
                "pdsi_annual_5_yrs_prior_sum", "pdsi_annual_5_yrs_after_sum",
                # "pdsi_annual_5_yrs_prior_mean", "pdsi_annual_5_yrs_after_mean",
                "hd_fingerprint_5_yrs_prior_sum", "hd_fingerprint_5_yrs_after_sum",
                # "hd_fingerprint_5_yrs_prior_mean", "hd_fingerprint_5_yrs_after_mean",
                "avg_forest_2_yr_pre_fire"),
      trt = "fire",
      data_nm = glue("ALL_{var}_firebinary"),
      seed = seed
    )
  }
)

walk(
  .x = outcome_vars,
  .f = function(var) {
    cf_analysis(
      dats = dats,
      yvar = var,
      xvars = c("aet", "def", "frg_reclass", "srtm", "tpi", "chili",
                "burn_cbi", "burn_cbi_1kmmwmedian", "burn_cbi_1kmmwstd", "mock_burn_year",
                "biotic_5_yrs_prior_sum", "biotic_5_yrs_after_sum",
                "pdsi_annual_5_yrs_prior_sum", "pdsi_annual_5_yrs_after_sum",
                # "pdsi_annual_5_yrs_prior_mean", "pdsi_annual_5_yrs_after_mean",
                "hd_fingerprint_5_yrs_prior_sum", "hd_fingerprint_5_yrs_after_sum",
                # "hd_fingerprint_5_yrs_prior_mean", "hd_fingerprint_5_yrs_after_mean",
                "avg_forest_2_yr_pre_fire"),
      trt = "burn_cbi",
      data_nm = glue("ALL_{var}_cbi"),
      seed = seed
    )
  }
)




# RUN SIMPLE ANALYSIS
cf_analysis(
  dats = dats,
  yvar = "post_fire_forest_slope_2_10_yrs",
  xvars = c("aet", "def", "frg_reclass", "srtm", "tpi", "chili",
            "biotic_5_yrs_prior_sum", "biotic_5_yrs_after_sum",
            "pdsi_annual_5_yrs_prior_sum", "pdsi_annual_5_yrs_after_sum",
            "hd_fingerprint_5_yrs_prior_sum", "hd_fingerprint_5_yrs_after_sum",
            "avg_forest_2_yr_pre_fire"),
  trt = "fire",
  data_nm = glue("ALL_simple_post_fire_forest_slope_2_10_yrs_fire"),
  seed = seed
)


for(burn_l in c("low-moderate-sev", "unburned", "high-sev")) {
  df <- dats |>
    filter(burn_sev_tagged == burn_l | burn_sev_tagged == "unburned-mock")
  
  cf_analysis(
    dats = df,
    yvar = "post_fire_forest_slope_2_10_yrs",
    xvars = c("aet", "def", "frg_reclass", "srtm", "tpi", "chili",
              "biotic_5_yrs_prior_sum", "biotic_5_yrs_after_sum",
              "pdsi_annual_5_yrs_prior_sum", "pdsi_annual_5_yrs_after_sum",
              "hd_fingerprint_5_yrs_prior_sum", "hd_fingerprint_5_yrs_after_sum",
              "avg_forest_2_yr_pre_fire"),
    trt = "fire",
    data_nm = glue("ALL_simple_{burn_l}_post_fire_forest_slope_2_10_yrs_fire"),
    seed = seed
  )
  
}






# yvar = "post_fire_forest_difference_10_yr"
# xvars = c("aet", "def", "frg_reclass", "srtm", "tpi", "chili",
#            "burn_cbi", "burn_cbi_1kmmwmedian", "burn_cbi_1kmmwstd", "mock_burn_year",
#            "biotic_5_yrs_prior_sum", "biotic_5_yrs_after_sum",
#            "pdsi_annual_5_yrs_prior_sum", "pdsi_annual_5_yrs_after_sum")
# trt = "fire"
#   # Filter out rows with NA in the outcome variable
#   dats_filt <- dats %>%
#     filter(!is.na(.data[[yvar]]))
# 
#   # Select covariates and outcome
#   covariates <- dats_filt %>%
#     select(all_of(xvars))
#   outcome <- dats_filt[[yvar]]
#   treatment <- dats_filt[[trt]]
# 
#   # Fit causal forest
#   fit <- grf::causal_forest(
#     X = covariates,
#     Y = outcome,
#     W = treatment,
#     seed = seed
#   )
# 
#   # Variable importance
#   imp_raw <- sort(setNames(variable_importance(fit), xvars))
#   imp_normalized <- imp / sum(imp)
# 
#   imp_df <- tibble(
#     variable = names(imp_normalized),
#     importance = as.numeric(imp_normalized)
#   ) %>%
#     arrange(importance)  # So the smallest is at the bottom of the horizontal bar plot
# 
#   imp_p <- ggplot(imp_df, aes(x = importance, y = reorder(variable, importance))) +
#     geom_col(fill = "orange") +
#     labs(
#       x = "Variable Importance",
#       y = NULL,
#       title = "Causal Forest Variable Importance",
#       subtitle = "test"
#     ) +
#     theme_minimal(base_size = 13) +
#     theme(
#       axis.text.y = element_text(hjust = 1),
#       plot.margin = margin(10, 20, 5, 5)
#     )
#   imp_p
#   ggsave(plot = imp_p,
#          filename = here(dir_figs, paste0("var_importance_", data_nm, ".png")),
#          bg = "white")
# 
#   pred_fun <- function(object, newdata, ...) {
#     predict(object, newdata, ...)$predictions
#   }
# 
#   xvars_sorted <- rev(names(imp))
# 
#   pdps <- lapply(xvars_sorted, function(v) {
#     pd <- partial_dep(fit, v, X = covariates, pred_fun = pred_fun)
#     plot(pd) +
#       theme(axis.title.y = element_blank())
#   })
# 
#   #For adding y axis label to overall plot
#   ylab_plot <- ggplot() +
#     theme_void() +
#     annotate("text", x = 0, y = 0.5, label = "Treatment effect", angle = 90, hjust = 0.5, size = 4)
# 
#   pdps_p <- wrap_plots(pdps, guides = "collect", ncol = 3) #&
#     #ylab("Treatment effect")
# 
#   # Add format & add titles
#   pdps_p <- ylab_plot + pdps_p +
#     plot_layout(widths = c(0.05, 1)) +
#     plot_annotation(
#       title = "Partial Dependence Plots",
#       subtitle = data_nm,
#       caption = "NOTES:\nNegative values mean a stronger (positive) treatment effect (i.e. fire is causing a greater difference in forest cover change)\nPDPs are in order of variable importance")
# 
#   ggsave(filename = here(dir_figs, paste0("partial_dependence_", data_nm, ".png")),
#          plot = pdps_p,
#          units = 'px',
#          width = 2000,
#          height = 2000)
# 

```


These initial results mean that: (YES = makes sense, NO = doesn't make sense)
- As AET increases (there is more AET), there is greater difference between pre-and post-fire cover
- As DEF increases (sites get drier), there is decreased difference between pre-and post-fire cover
- YES - When there is biotic disturbance prior to burn there is a greater difference between pre and post-fire cover
- ??? - When there is more drought after fire there is greater difference between pre and post-fire cover; but then also a much larger difference between pre and post-fire cover with wetter conditions after fire?..
- YES - The difference between pre and post-fire is least when elevation is ~2000, and greater at low elevatoins or high elevations (at the edges of forest range forest causes more change in forest cover?)
- YES - The difference between pre and post fire forest cover is greatest when CHILI is high (very warm)
- The treatment effect is greatest for FRG classes 3,4,5,6
- YES? - The treatment effect is greatest for negative values of TPI (i.e. in valleys); there are denser forests pre-burn in valleys
- NO - The treatment effect is greatest earlier in the time frame


```{r}
hist(dats$pdsi_annual_5_yrs_after_sum)

```



<button class="accordion-button" type="button" data-bs-toggle="collapse" data-bs-target="#collapseOne" >Click to see Session Information</button><div id="collapseOne" class="accordion-collapse collapse"><div>

```{r sessionInfo}
sessionInfo()
```