---
title: "04_causal_forest"
author: Tyler L. McIntosh
date: last-modified
format:
  html:
    embed-resources: true
    html-math-method: katex
    css: styles.css
    toc: true
    toc-depth: 1
    toc-title: Contents
    toc-location: right
    code-fold: true
    code-summary: "Show the code"
---

# Setup for analysis

``` {r setupchunk}
#| echo: true
#| message: false
#| warning: false
#| results: 'hide'

rm(list = ls())

if(!requireNamespace("here", quietly = TRUE)) {
  install.packages("here")
}
library(here)

source(here::here("utils", "functions.R"))

install_load_packages(c(
  "tidyverse",
  "tictoc",
  "hstats",
  "glue",
  "shapviz",
  "patchwork",
  "cobalt",
  "purrr",
  "corrplot",
  "gridExtra",
  "tictoc",
  "fect",
  "fixest",
  "did",
  "panelView",
  "PanelMatch",
  "bacondecomp",
  "didimputation",
  "doParallel",
  "HonestDid",
  "HonestDidFEct",
  "DIDmultiplegtDYN"))

# # install packages from CRAN
# packages <- c("fixest", "did", "didimputation", 
#          "panelView", "bacondecomp", "HonestDiD",
#               "DIDmultiplegtDYN", "PanelMatch")
# install.packages(setdiff(packages, rownames(installed.packages())))  
# 
# # install most up-to-date "fect" from Github
# if ("fect" %in% rownames(installed.packages()) == FALSE) {
#   devtools:: install_github("xuyiqing/fect")
# }
# 
# # install forked "HonestDiD" package compatible with "fect"
# if ("HonestDiDFEct" %in% rownames(installed.packages()) == FALSE) {
#   devtools:: install_github("lzy318/HonestDiDFEct")
# }


dir_figs <- here::here('figs')
dir_derived <- here::here('data', 'derived')
dir_raw <- here::here('data', 'raw')
dir_manual <- here::here('data', 'manual')
dir_ensure(c(dir_figs,
             dir_derived,
             dir_manual,
             dir_raw))

seed = 1234
set.seed(seed)


```

#

```{r}

run_nm <- paste0("did_conley_", timestamp('for_file'))
dir_figs <- here(dir_figs, run_nm)
dir_ensure(dir_figs)
  
dats_long <- read_csv(here(dir_derived, "test_dats_long.csv"))
dats_short <- read_csv(here(dir_derived, "test_dats.csv"))


```



# Test key assumptions

- Parallel trends
- No anticipatory behavior
- Treatment effect heterogeneity - “Since treatment effects heterogeneity violates Assumption 3 and can alter how we interpret μg , it is important to think through when different cohorts likely experience different paths of treatment effect. Such heterogeneity could arise for many reasons. For example, cohorts may differ in their covariates, which affect how they respond to treatment. We will explore a concrete example in our application: if treatment effects differ with age, and there is variation in age across units first treated at different times, we will have heterogeneous effects (see Section 5 for details)” ([Sun and Abraham, 2021, p. 182](zotero://select/library/items/86R62KQ4)) ([pdf](zotero://open-pdf/library/items/3WRC5CP7?page=8&annotation=QUPBLF69))

IT APPEARS THAT WE HAVE A PROBLEM OF HETEROGENEOUS TREATMENT EFFECTS - We need to use an interaction-weighted estimator (i.e. dynamic DiD from Sun & Abraham)



```{r}


# Parallel trends - see results from 04_causal_forest


# Heterogeneous treatment effects?

ggplot() +
  geom_boxplot(data = dats_short |> filter(fire == 1), aes(x = as.factor(burn_year), y = biotic_5_yrs_prior_sum))

ggplot() +
  geom_boxplot(data = dats_short |> filter(fire == 1), aes(x = as.factor(burn_year), y = biotic_5_yrs_after_sum))

ggplot() +
  geom_boxplot(data = dats_short |> filter(fire == 1), aes(x = as.factor(burn_year), y = pdsi_annual_5_yrs_prior_sum))

ggplot() +
  geom_boxplot(data = dats_short |> filter(fire == 1), aes(x = as.factor(burn_year), y = pdsi_annual_5_yrs_after_sum))

ggplot() +
  geom_boxplot(data = dats_short |> filter(fire == 1), aes(x = as.factor(burn_year), y = chili))

ggplot() +
  geom_boxplot(data = dats_short |> filter(fire == 1), aes(x = as.factor(burn_year), y = def))


# Number of points for disturbance interactions
n_bio_greater_10 <- dats_short |>
  filter(fire == 1 & biotic_5_yrs_prior_sum >= 10)
n_hd_greater_17 <- dats_short |>
  filter(fire == 1 & hd_fingerprint_5_yrs_after_sum >= 17)

n <- dats_short |>
  filter(fire == 1 & pdsi_annual_5_yrs_after_threshold >= 1)


only_burned <- dats_short |>
  filter(fire == 1)

ggplot(only_burned, aes(x = biotic_5_yrs_prior_sum, y = burn_cbi)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", se = TRUE, color = "blue") +
  labs(
    x = "Biotic Sum (5 Years Prior)",
    y = "Burn Severity",
    title = "Linear Relationship: Burn Severity vs Biotic Sum"
  ) +
  theme_minimal()



burn_sev_model <- lm(burn_cbi ~ biotic_5_yrs_prior_sum + def + aet + avg_forest_2_yr_pre_fire + pdsi_annual_5_yrs_prior_sum + tpi + ecoregion + chili + frg_reclass, data = only_burned)
summary(burn_sev_model)

```

```{r}



# #Visualize treatment status over time
# panelview(rap_TRE ~ treated, data = testset, index = c("uniqueID", "year"),
#   xlab = "Year", ylab = "Unit", display.all = T,
#   gridOff = TRUE, by.timing = TRUE)
# 
# 
# hist(testset |> filter(fire_timeframe == "year_of_burn") |> pull(year))

```

# Run Dynamic DiD with Sun-Abraham estimators and two-way fixed effects

PROBLEM: Our error bars pre-event are often NOT covering zero; this suggests that we are violating the parallel trends assumption. We are missing important covariates (?) and/or need to perform pre-fire matching/weighting.



```{r}

# model.twfe.0 <- feols(rap_TRE~treated|uniqueID+year,
#                       data=dats_long, cluster = "uniqueID") 
# print(model.twfe.0)


# # We don't have any always-treated units, so we don't need to worry about the Goodman-Bacon decomposition


dats_long_did <- get.cohort(dats_long, D = "treated",
                                 index = c("uniqueID", "year"),
                                 start0 = TRUE)
dats_long_did[which(is.na(dats_long_did$FirstTreat)),"FirstTreat"] <- 1000 
df.use <- dats_long_did |>
  mutate(biotic_before_10t = case_when(biotic_5_yrs_prior_sum >= 10 ~ "biotic_prior_10t",
                                  biotic_5_yrs_prior_sum < 10 ~ "no_biotic_prior_10t"),
         biotic_before_20t = case_when(biotic_5_yrs_prior_sum >= 20 ~ "biotic_prior_20t",
                                  biotic_5_yrs_prior_sum < 20 ~ "no_biotic_prior_20t"),
         hd_after_17t = case_when(hd_fingerprint_5_yrs_after_sum >= 17 ~ "hd_after_17t",
                              hd_fingerprint_5_yrs_after_sum < 17 ~ "no_hd_after_17t"),
         pdsi_before_n500t = case_when(pdsi_annual_5_yrs_prior_sum <= -500 ~ "pdsi_before_n500t",
                                 pdsi_annual_5_yrs_prior_sum > -500 ~ "no_pdsi_before_n500t"),
         pdsi_before_n1000t = case_when(pdsi_annual_5_yrs_prior_sum <= -1000 ~ "pdsi_before_n1000t",
                                 pdsi_annual_5_yrs_prior_sum > -1000 ~ "no_pdsi_before_n1000t"))



# 
# n_bio_greater_10 <- dats_short |>
#     filter(ecoregion == "IDBatholith") |>
#   filter(fire == 1 & biotic_5_yrs_prior_sum >= 10)
# n_bio_greater_10_control <- dats_short |>
#     filter(ecoregion == "IDBatholith") |>
#   filter(fire == 0 & biotic_5_yrs_prior_sum >= 10)

# # Dynamic TWFE
# df.twfe <- df.use
# 
# # drop always treated units
# #df.twfe$treat <- as.numeric(df.twfe$treatment_mean>0) 
# df.twfe[which(is.na(df.twfe$Time_to_Treatment)),'Time_to_Treatment'] <- -999 # can be an arbitrary value
# twfe.est <- feols(rap_TRE ~ i(Time_to_Treatment, fire, ref = c(-1, -999))| uniqueID + year, 
#                   data = df.twfe,
#                   cluster = "uniqueID")
# 
# twfe.output <- as.matrix(twfe.est$coeftable)
# print(round(twfe.output, 3))
# 
# twfe.est$collin.var
# 
# table(df.twfe$treated, useNA = "always")
# table(df.twfe$treated, is.na(df.twfe$Time_to_Treatment))
# 
# 
# twfe.output <- as.data.frame(twfe.output)
# twfe.output$Time <- c(c(-29:-2),c(0:21))+1 
# p.twfe <- esplot(twfe.output,Period = 'Time',Estimate = 'Estimate',
#                                SE = 'Std. Error', xlim = c(-29,21))
# p.twfe




# 
# model <- feols(
#   rap_TRE ~ sunab(FirstTreat, year) | year + uniqueID,
#   data = testset |>
#     filter(biotic_before == 0),
#   cluster = "uniqueID"
# )
# summary(model)
# plot_sunab_models(model_list = list(model),
#                   model_labels = "test2 - no biotic before")




# Multi-model functions ----

iw_estimator_subg <- function(subgrp_col, grps, df) {
  # Convert string to symbol
  col_sym <- sym(subgrp_col)
  
  if(subgrp_col == "burn_sev_tagged") {
    df_sub <- df %>%
      filter((!!col_sym) == grps | burn_sev_tagged == "unburned-mock")
  } else {
    df_sub <- df %>%
      filter((!!col_sym) == grps)    
  }

  model <- feols(
    rap_TRE ~ sunab(FirstTreat, year) | uniqueID + year,
    data = df_sub,
    mem.clean = TRUE,
    lean = TRUE
  )
  return(model)
}




plot_sunab_models <- function(model_list, model_labels = NULL, title = "Dynamic-DiD Event-Study Estimates from Multiple Models") {
  
  # If no labels provided, auto-label as Model 1, Model 2, etc.
  if (is.null(model_labels)) {
    model_labels <- paste0("Model ", seq_along(model_list))
  }
  
  # Extract coefficients for each model and attach model label
  plot_data <- map2_dfr(model_list, model_labels, function(model, label) {
    df <- as.data.frame(as.matrix(model$coeftable))
    df$term <- rownames(model$coeftable)
    
    df %>%
      filter(str_detect(term, "^year::")) %>%
      mutate(
        Time = as.numeric(str_extract(term, "-?\\d+")),
        CI_upper = Estimate + 1.96 * `Std. Error`,
        CI_lower = Estimate - 1.96 * `Std. Error`,
        Model = label
      )
  })

  # Plot
  ggplot(plot_data, aes(x = Time, y = Estimate, color = Model, fill = Model)) +
    geom_line(linewidth = 1) +
    geom_point(size = 2) +
    geom_ribbon(aes(ymin = CI_lower, ymax = CI_upper), alpha = 0.2, color = NA) +
    geom_vline(xintercept = -1, linetype = "dashed", color = "gray40") +
    geom_hline(yintercept = 0, color = "black") +
    labs(
      x = "Event Time",
      y = "Treatment Effect of fire on forest cover",
      title = title,
      caption = "95% confidence intervals shown; Interaction-weighted (Sun-Abraham) estimates used"
    ) +
    theme_minimal(base_size = 13) +
    theme(legend.position = "bottom")
}


plot_all_together_did <- function(subg_col, df, title = "Interaction-Weighted Estimates for all Ecoregions") {
  
  if(subg_col == "burn_sev_tagged") {
    subgs <- c("low-moderate-sev", "unburned", "high-sev")
  } else {
    subgs <- unique(df |> pull(subg_col))
  }  
  
    # Fit models for each subgroup level
  iw_results <- subgs %>%
    set_names() %>%
    map(~ iw_estimator_subg(grps = .x,
                            subgrp_col = subg_col,
                            df = df))
  
  # Generate the overlay plot
  p <- plot_sunab_models(
    model_list = iw_results,
    model_labels = subgs,
    title = title
  )
  return(p)
}


plot_ecoregion_did <- function(ecoregion_name, subg_col, df) {
  df_ecoregion <- df %>% filter(ecoregion == ecoregion_name)
  
  if(subg_col == "burn_sev_tagged") {
    subgs <- c("low-moderate-sev", "unburned", "high-sev")
  } else {
    subgs <- unique(df_ecoregion |> pull(subg_col))
  }
  
  # Fit models for each subgroup level
  iw_results <- subgs %>%
    set_names() %>%
    map(~ iw_estimator_subg(grps = .x,
                            subgrp_col = subg_col,
                            df = df_ecoregion))
  
  # Generate the overlay plot
  p <- plot_sunab_models(
    model_list = iw_results,
    model_labels = subgs,
    title = ecoregion_name
  )
  return(p)
}

plot_all_ecoregions_did <- function(subg_col, df, title = "Interaction-Weighted Estimates by Ecoregion") {
  # Get unique ecoregions
  ecoregions <- unique(df$ecoregion)
  
  # Generate a plot per ecoregion
  ecoregion_plots <- map(ecoregions, ~ plot_ecoregion_did(ecoregion_name = .x,
                                                          subg_col = subg_col,
                                                          df = df))
  
  # Combine into grid
  final_plot <- wrap_plots(ecoregion_plots, ncol = 2) +
    plot_annotation(title = title )
  
  return(final_plot)
  
}


# Operate ----

# USE FUNCTIONS TO VIEW SEVERITY BY ECOREGION

all_by_sev_p <- plot_all_ecoregions_did(subg_col = "burn_sev_tagged",
                                        df = df.use,
                                        title = "Interaction-Weighted Estimates \nby Ecoregion & Burn Severity, conley100km")


ggsave(plot = all_by_sev_p,
       filename = here(dir_figs, "all_did_by_eco_and_sev_plot.png"),
       bg = 'white',
       units = "px",
       width = 4000,
       height = 4000)




all_by_biotic_p <- plot_all_ecoregions_did(subg_col = "biotic_before_10t",
                                        df = df.use,
                                        title = "Interaction-Weighted Estimates by \nEcoregion & Pre-Fire Biotic (10%, 5-yr), conley100km")

ggsave(plot = all_by_biotic_p,
       filename = here(dir_figs, "all_did_by_eco_and_biotic10t_plot.png"),
       bg = 'white',
       units = "px",
       width = 4000,
       height = 4000)



all_by_hd_p <- plot_all_ecoregions_did(subg_col = "hd_after",
                                        df = df.use,
                                        title = "Interaction-Weighted Estimates by Ecoregion & Post-fire HD (17, sum, 5-yr)")

ggsave(plot = all_by_hd_p,
       filename = here(dir_figs, "all_did_by_eco_and_hd_plot.png"),
       bg = 'white',
       units = "px",
       width = 4000,
       height = 4000)



all_by_pdsi_p <- plot_all_ecoregions_did(subg_col = "pdsi_before",
                                        df = df.use,
                                        title = "Interaction-Weighted Estimates by Ecoregion & Pre-Fire PDSI (-500, sum, 5-yr)")

ggsave(plot = all_by_pdsi_p,
       filename = here(dir_figs, "all_did_by_eco_and_pdsi_plot.png"),
       bg = 'white',
       units = "px",
       width = 4000,
       height = 4000)




# RESULTS FOR ALL ECOREGIONS MERGED
together_by_sev_p <- plot_all_together_did(subg_col = "burn_sev_tagged",
                                           df = df.use,
                                           title = "Interaction-Weighted Estimates for all Ecoregions by severity")

ggsave(plot = together_by_sev_p,
       filename = here(dir_figs, "together_did_plot_sev.png"),
       bg = 'white')




together_by_biotic_p <- plot_all_together_did(subg_col = "biotic_before",
                                           df = df.use,
                                           title = "Interaction-Weighted Estimates for all Ecoregions by pre-fire biotic (10%, 5-yr)")

ggsave(plot = together_by_biotic_p,
       filename = here(dir_figs, "together_did_plot_biotic.png"),
       bg = 'white')




together_by_hd_p <- plot_all_together_did(subg_col = "hd_after",
                                           df = df.use,
                                           title = "Interaction-Weighted Estimates for all Ecoregions by post-fire hd (17, sum, 5yr)")

ggsave(plot = together_by_hd_p,
       filename = here(dir_figs, "together_did_plot_hd.png"),
       bg = 'white')





together_by_pdsi_p <- plot_all_together_did(subg_col = "pdsi_before",
                                           df = df.use,
                                           title = "Interaction-Weighted Estimates for all Ecoregions by pre-fire PDSI (-500, sum, 5yr)")

ggsave(plot = together_by_pdsi_p,
       filename = here(dir_figs, "together_did_plot_pdsi.png"),
       bg = 'white')






# MULTI-SPLIT

for(burn_l in c("low-moderate-sev", "unburned", "high-sev")) {
  df <- df.use |>
    filter(burn_sev_tagged == burn_l | burn_sev_tagged == "unburned-mock")
  
  by_biotic_sev_p <- plot_all_ecoregions_did(subg_col = "biotic_before_20t",
                                          df = df,
                                          title = paste0(burn_l, ": Interaction-Weighted Estimates by Ecoregion & Pre-Fire Biotic (10%, 5-yr)"))
  
  ggsave(plot = by_biotic_sev_p,
         filename = here(dir_figs, paste0("did_by_eco_and_biotic_20t_plot_", burn_l, ".png")),
         bg = 'white',
         units = "px",
         width = 4000,
         height = 4000)
  
  gc()

}


# 
# objs <- ls(envir = .GlobalEnv)
# sizes <- sapply(objs, function(x) object.size(get(x)))
# sizes_df <- data.frame(
#   Object = objs,
#   Size_Bytes = as.numeric(sizes),
#   Size_GB = round(as.numeric(sizes) / 1024^3, 3)
# )
# sizes_df[order(sizes_df$Size_Bytes, decreasing = TRUE), ]

```

```{r}


# SIMPLE
testset <- df.use |>
  filter(ecoregion == "IDBatholith")
model <- feols(
  rap_TRE ~ sunab(FirstTreat, year) | uniqueID + year,
  data = testset,
  cluster = "uniqueID")

# INTERACTED
model_interact <- feols(
  rap_TRE ~ sunab(FirstTreat, year):biotic_before_10t | uniqueID + year,
  data = testset,
  cluster = "uniqueID")

# SUBGROUPED
p_test <- plot_all_together_did(subg_col = "biotic_before_10t",
                      df = testset, title = "Test subg")


# RESULTS
summary(model)
plot_sunab_models(model_list = list(model),
                  model_labels = "test")

summary(model_interact)
interacted_coef <- summary(model_interact)$coefficients
interacted_coef


plot_sunab_models(model_list = list(model_interact),
                  model_labels = "test interact")

p_test


# TRYING TO PLOT THE INTERACTED RESULTS

etable(model_interact, keep = "%year::")


agg_main <- aggregate(model_interact, agg = "cohort", fun = "mean", keep = "^(?!.*biotic_before_10t)")
agg_inter <- aggregate(model_interact, agg = "cohort", fun = "mean", keep = "biotic_before_10t")


tidy_main <- broom::tidy(agg_main) %>%
  mutate(type = "Baseline (biotic_before_10t = 0)")

tidy_inter <- broom::tidy(agg_inter) %>%
  mutate(type = "Effect of biotic_before_10t (per unit)")



plot_df <- bind_rows(tidy_main, tidy_inter) %>%
  mutate(rel_year = as.numeric(gsub("year::(-?\\d+).*", "\\1", term)))

ggplot(plot_df, aes(x = rel_year, y = estimate, color = type)) +
  geom_line() +
  geom_point() +
  geom_errorbar(aes(ymin = estimate - 1.96 * std.error,
                    ymax = estimate + 1.96 * std.error),
                width = 0.2) +
  geom_vline(xintercept = -1, linetype = "dashed") +
  theme_minimal() +
  labs(x = "Event time (years)", y = "Estimated effect",
       color = "Series")

```




<button class="accordion-button" type="button" data-bs-toggle="collapse" data-bs-target="#collapseOne" >Click to see Session Information</button><div id="collapseOne" class="accordion-collapse collapse"><div>

```{r sessionInfo}
sessionInfo()
```