---
title: "random_forest"
author: Tyler L. McIntosh
date: last-modified
format:
  html:
    embed-resources: true
    html-math-method: katex
    css: styles.css
    toc: true
    toc-depth: 1
    toc-title: Contents
    toc-location: right
    code-fold: true
    code-summary: "Show the code"
---

# Setup for analysis

``` {r setupchunk}
#| echo: true
#| message: false
#| warning: false
#| results: 'hide'

rm(list = ls())

if(!requireNamespace("here", quietly = TRUE)) {
  install.packages("here")
}
library(here)

source(here::here("utils", "functions.R"))

install_load_packages(c(
  "tidyverse",
  "tictoc",
  "ranger",
  "spmodel",
  "iml",
  "fastshap",
  "patchwork",
  "purrr",
  "arrow",
  "vip",
  "pdp",
  "tictoc"))

# devtools::install_github("grf-labs/sufrep")
# library(sufrep)

dir_figs <- here::here('figs')
dir_derived <- here::here('data', 'derived')
dir_raw <- here::here('data', 'raw')
dir_manual <- here::here('data', 'manual')

dir_rf_figs <- here(dir_figs, "rf")

dir_ensure(c(dir_figs,
             dir_derived,
             dir_manual,
             dir_raw,
             dir_rf_figs))

seed = 1234
set.seed(seed)


```
# Load dats

```{r}

#Short dats
dats <- read_parquet(here(dir_derived, "test_dats_short.parquet"))
ecoregions <- unique(dats$ecoregion)

```




# Sufrep function
```{r}

# A function to use sufrep::make_encoder such that the encoded variables are added to the dataframe
sufrep_encode_append <- function(dats,
                                 method = 'means',
                                 X_cols,
                                 G_col,
                                 remove_G_col = TRUE,
                                 ...) {
  
  #Select variables for encoding
  x_df <- dats |> 
    select(all_of(X_cols)) |>
    as.data.frame()
  g_df <- dats |>
    pull(G_col)
  
  #Encode
  encoder <- sufrep::make_encoder(method = method, X = x_df, G = g_df, ...)
  dats_encoded <- encoder(x_df, G = g_df)
  
  #Make interpretable encoding columns
  encoded_cols <- dats_encoded |> select(starts_with("ENC"))
  names(encoded_cols) <- paste0(names(encoded_cols), "_", G_col)
  
  dats_out <- dats |>
    cbind(encoded_cols)
  
  if(remove_G_col) {
    dats_out <- dats_out |>
      select(-all_of(G_col))
  }
  
  return(dats_out)
}


```

# Prep data for causal forest machine learning
- Normalize continuous covariates to control for covariates with different numerical scales (avoid large-scale variables dominating and stabilize splits)
- Encode categorical variables as a lower-dimensional real-valued representation for inclusion in causal forest (use sufficient representations instead of one-hot encoding since random forests "may largely ignore high-cardinality one-hot encoded factors" - see Johannemann et al 2021). Low-rank encoding appears to be the most applicable to real-world data.

```{r}

disturbance_covars <- dats |>
  select(starts_with("burn_cbi") | starts_with("biotic") | starts_with("hd") | starts_with("pdsi")) |>
  names()

cont_covars <- c("aet", "def", "srtm", "tpi", "chili", "avg_rap_TRE_6_yr_pre_fire", "lat", "long", disturbance_covars)
cat_covars <- c("frg_reclass_desc", "mock_burn_year", "ecoregion")
outcomes <- dats |>
  select(starts_with("post_fire")) |>
  names()
trts <- c("fire", "burn_sev")


all_cols <- c(cont_covars, cat_covars, outcomes, trts)

# Scale all continuous covariates
# filter out anything with NAs
# ensure categorical covariates are factors
# encode categorical as one-hot dummies or sufficient representations
dats_ml <- dats |>
  select(all_of(all_cols)) |>
  filter(dplyr::if_all(everything(), ~ !is.na(.x))) |>
  mutate(
    across(
      all_of(cont_covars),
      \(x) as.numeric(scale(x))
    )
  ) |>
  mutate(
    across(
      all_of(cat_covars),
      \(x) as.factor(x)
    )
  ) |>
  # fastDummies::dummy_cols(select_columns = c("frg_reclass"),
  #                         remove_first_dummy = FALSE, #Keep full one-hot encoding for tree-based methods
  #                         remove_selected_columns = TRUE) |> 
  sufrep_encode_append(method = 'low_rank', # Use low-rank sufficient representation encoding to reduce to 2 dimensions
                       X_cols = cont_covars,
                       G_col = "mock_burn_year",
                       num_components = 2) |> 
  sufrep_encode_append(method = 'low_rank', # Use low-rank sufficient representation encoding to reduce to 2 dimensions
                       X_cols = cont_covars,
                       G_col = "frg_reclass_desc",
                       num_components = 2)

enc_xvars <- dats_ml |>
  select(starts_with("ENC")) |>
  names()

all_covars <- c(cont_covars, enc_xvars)


```


# Test model tuning time bumps

Tuning all parameters as opposed to a select few increases training time by ~50%.

GRF package tuning parameters:

tune.parameters	
A vector of parameter names to tune. If "all": all tunable parameters are tuned by cross-validation. The following parameters are tunable: ("sample.fraction", "mtry", "min.node.size", "honesty.fraction", "honesty.prune.leaves", "alpha", "imbalance.penalty"). If honesty is FALSE the honesty.* parameters are not tuned. Default is "none" (no parameters are tuned).

tune.num.trees	
The number of trees in each 'mini forest' used to fit the tuning model. Default is 200.

tune.num.reps	
The number of forests used to fit the tuning model. Default is 50.

tune.num.draws	
The number of random parameter values considered when using the model to select the optimal parameters. Default is 1000.

```{r}

df_test <- dats_ml |>
  filter(ecoregion == "Idaho Batholith" & fire == 1)



```



# Variable interpretation
- TPI: low = valleys, high = ridges
- CHILI: low = cold, high = hot
- SRTM: low = low elevation, high = high elevation
- AET: low = wet, high = dry
- DEF: low = wet, high = dry
- PDSI: low = dry, high = wet

```{r}
df_test <- df_test |>
  st_as_sf(coords = c("long", "lat"), remove = FALSE)

form_str <- paste(
  "post_fire_rap_TRE_slope_2_10_yrs ~ burn_sev + ",
  paste(all_covars, collapse = " + ")
)

form <- as.formula(form_str)

tic()
test_spmlmRF <- splmRF(form,
                  data = df_test,
                  spcov_type = "exponential",
                  importance = 'permutation',
                  scale.permutation.importance = TRUE,
                  num.threads = 0,
                  mtry = floor(sqrt(ncol(dats) - 1)),
                  num.trees = 1000,
                  seed = seed)
toc()
summary(test_spmlmRF)
print(test_spmlmRF$ranger)
#500 trees mtry 3, MSE 213, R2 0.49
#1000 trees mtry 8, basically same
#1000t, mtry8, include disturbance MSE 2.17, R^2 0.14


vip(test_spmlmRF$ranger,
    num_features = 40)
pdp(test_spmlmRF$ranger)




```

#################
# From run with all PDSI data, the cumulative drought sums appear to be more important than the means, which are more important than the thresholds

```{r}

narrower_covars <- all_covars[!str_ends(all_covars, "mean") &
                                !str_ends(all_covars, "threshold") &
                                !str_starts(all_covars, "pdsi_summer")]


form_str <- paste(
  "post_fire_rap_TRE_slope_2_10_yrs ~ burn_sev + ",
  paste(narrower_covars, collapse = " + ")
)

form <- as.formula(form_str)

tic()
test_spmlmRF <- splmRF(form,
                  data = df_test,
                  spcov_type = "exponential",
                  importance = 'permutation',
                  scale.permutation.importance = TRUE,
                  num.threads = 0,
                  mtry = floor(sqrt(ncol(dats) - 1)),
                  num.trees = 1000,
                  seed = seed)
toc()
summary(test_spmlmRF)
print(test_spmlmRF$ranger)
print(test_spmlmRF$call)

#500 trees mtry 3, MSE 213, R2 0.49
#1000 trees mtry 8, basically same
#1000t, mtry8, include disturbance MSE 2.17, R^2 0.14


vip <- vip(test_spmlmRF$ranger,
    num_features = 25)
vip_tbl <- vip(test_spmlmRF$ranger, num_features = 25)$data |>
  as_tibble() |>
  arrange(desc(Importance)) |>   # highest first
  mutate(variable = factor(Variable, levels = Variable))

ggsave(plot = vip, filename = here(dir_rf_figs, "vip_idbath_v1.png"), units = "px", width = 3000, height = 4000, )


pdp_cont <- c(narrower_covars) |>
  set_names() |>
  map_dfr(\(v) {
    # run PDP
    pd <- partial(
      object   = test_spmlmRF$ranger,
      pred.var = v,
      train    = df_test |> st_drop_geometry()
    )

    # rename predictor column to a common name 'x'
    pd |>
      rename(x = !!sym(v)) |>
      mutate(variable = v, .before = x)
  })

pdp_sorted <- pdp_cont |>
  inner_join(
    vip_tbl |> select(variable = Variable),
    by = "variable"
  ) |>
  mutate(variable = factor(variable, levels = vip_tbl$Variable))

plot <- pdp_sorted |>
  ggplot(aes(x = x, y = yhat)) +
  geom_line() +
  facet_wrap(~ variable, scales = "free_x") +
  labs(
    x = "Predictor value",
    y = "Partial dependence (yhat)",
    title = "Partial dependence for continuous covariates"
  )

ggsave(plot = plot, filename = here(dir_rf_figs, "pdp_idbath_v1.png"), units = "px", width = 3000, height = 4000, )


```




<button class="accordion-button" type="button" data-bs-toggle="collapse" data-bs-target="#collapseOne" >Click to see Session Information</button><div id="collapseOne" class="accordion-collapse collapse"><div>

```{r sessionInfo}
sessionInfo()
```