---
title: "random_forest"
author: Tyler L. McIntosh
date: last-modified
format:
  html:
    embed-resources: true
    html-math-method: katex
    css: styles.css
    toc: true
    toc-depth: 1
    toc-title: Contents
    toc-location: right
    code-fold: true
    code-summary: "Show the code"
---

# Setup for analysis

``` {r setupchunk}
#| echo: true
#| message: false
#| warning: false
#| results: 'hide'

rm(list = ls())

if(!requireNamespace("here", quietly = TRUE)) {
  install.packages("here")
}
library(here)

source(here::here("utils", "functions.R"))

install_load_packages(c(
  "tidyverse",
  "glue",
  "patchwork",
  "tictoc",
  "ranger",
  "spmodel",
  "iml",
  "patchwork",
  "tidymodels",
  "purrr",
  "arrow",
  "vip",
  "pdp",
  "fastshap",
  "shapviz",
  "dials",
  "sf",
  "tictoc"))

# devtools::install_github("grf-labs/sufrep")
# library(sufrep)

dir_figs <- here::here('figs')
dir_derived <- here::here('data', 'derived')
dir_raw <- here::here('data', 'raw')
dir_manual <- here::here('data', 'manual')
dir_models <- here('models/v2')

dir_rf_figs <- here(dir_figs, "rf")

dir_ensure(c(dir_figs,
             dir_derived,
             dir_manual,
             dir_raw,
             dir_rf_figs,
             dir_models))

seed = 1234
set.seed(seed)


```
# Load dats

```{r}

#Short dats
dats <- read_parquet(here(dir_derived, "test_dats_short.parquet"))
ecoregions <- unique(dats$ecoregion)

```





# Prep data for random forest machine learning
- We do not need to normalize data for RF
- We do want to encode categorical variables as a lower-dimensional real-valued representation for inclusion in random/causal forest (use sufficient representations instead of one-hot encoding since random forests "may largely ignore high-cardinality one-hot encoded factors" - see Johannemann et al 2021). Low-rank encoding appears to be the most applicable to real-world data.

```{r}

# A function to use sufrep::make_encoder such that the encoded variables are added to the dataframe
# used in prep_ml_dats
sufrep_encode_append <- function(dats,
                                 method = 'means',
                                 X_cols,
                                 G_col,
                                 remove_G_col = TRUE,
                                 ...) {
  
  #Select variables for encoding
  x_df <- dats |> 
    select(all_of(X_cols)) |>
    as.data.frame()
  
  # Grouping variable as factor with dropped unused levels
  # The function breaks if we don't drop unused levels because sufrep assumes that all levels are represented at least once in the data
  g_raw <- dats[[G_col]]

  if (!is.factor(g_raw)) {
    g_fac <- factor(g_raw)
  } else {
    g_fac <- droplevels(g_raw)
  }
  
  #Encode
  encoder <- sufrep::make_encoder(method = method, X = x_df, G = g_fac, ...)
  dats_encoded <- encoder(x_df, G = g_fac)
  
  #Make interpretable encoding columns
  encoded_cols <- dats_encoded |> select(starts_with("ENC"))
  names(encoded_cols) <- paste0(names(encoded_cols), "_", G_col)
  
  dats_out <- dats |>
    cbind(encoded_cols)
  
  if(remove_G_col) {
    dats_out <- dats_out |>
      select(-all_of(G_col))
  }
  
  return(dats_out)
}

# Select relevant columns, filter out NAs, ensure correct types, and sufrep encode
prep_ml_dats <- function(dats, target) {
  disturbance_covars <- dats |>
    select(starts_with("burn_cbi") | 
             starts_with("burn_distance") | 
             #starts_with("biotic") | 
             starts_with("hd") | 
             starts_with("pdsi") | 
             starts_with("pdsi")) |>
    names()
  
  cont_covars <- c("aet", "def", "srtm", "tpi", "chili", "avg_rap_tree_6_yr_pre_fire",
                   #"lat", "long",
                   "mock_burn_year",  disturbance_covars, grep("^tt", names(dats), value = TRUE))
  cat_covars <- c("frg_reclass_desc", "ecoregion", "nfg_factor")
  trts <- c("fire", "burn_sev")
  
  all_cols <- c(cont_covars, cat_covars, target, trts)

  
  # Scale all continuous covariates
  # filter out anything with NAs
  # ensure categorical covariates are factors
  # encode categorical as one-hot dummies or sufficient representations
  dats_ml <- dats |>
    select(all_of(all_cols)) |>
    filter(dplyr::if_all(everything(), ~ !is.na(.x))) |>
    filter(!is.infinite(!!sym(target))) |>
    mutate(
      across(
        all_of(cont_covars),
        \(x) as.numeric(x)
      )
    ) |>
    mutate(
      across(
        all_of(cat_covars),
        \(x) as.factor(x)
      )
    ) |>
    # fastDummies::dummy_cols(select_columns = c("frg_reclass"),
    #                         remove_first_dummy = FALSE, #Keep full one-hot encoding for tree-based methods
    #                         remove_selected_columns = TRUE) |> 
    sufrep_encode_append(method = 'low_rank', # Use low-rank sufficient representation encoding to reduce to 2 dimensions
                         X_cols = cont_covars,
                         G_col = "nfg_factor",
                         num_components = 2) |> 
    sufrep_encode_append(method = 'low_rank', # Use low-rank sufficient representation encoding to reduce to 2 dimensions
                         X_cols = cont_covars,
                         G_col = "frg_reclass_desc",
                         num_components = 2)
  
  enc_xvars <- dats_ml |>
    select(starts_with("ENC")) |>
    names()
  
  all_covars <- c(cont_covars, enc_xvars)
  
  return(list("dats_ml" = dats_ml,
              "target" = target,
              "covars" = all_covars))
}



#Create operational datasets - we need a separate one for each because the sufrep encodings will be different for each
perc10d <- dats |>
  prep_ml_dats(target = "post_fire_rap_tree_perc_recovery_n6_3_10_yr")

perc15d <- dats |>
  prep_ml_dats(target = "post_fire_rap_tree_perc_recovery_n6_3_15_yr")

perc20d <- dats |>
  prep_ml_dats(target = "post_fire_rap_tree_perc_recovery_n6_3_20_yr")

slope15d <- dats |>
  prep_ml_dats(target = "post_fire_rap_tree_slope_3_15_yrs")


```


# Test model tuning time bumps

Tuning all parameters as opposed to a select few increases training time by ~50%.

GRF package tuning parameters:

tune.parameters	
A vector of parameter names to tune. If "all": all tunable parameters are tuned by cross-validation. The following parameters are tunable: ("sample.fraction", "mtry", "min.node.size", "honesty.fraction", "honesty.prune.leaves", "alpha", "imbalance.penalty"). If honesty is FALSE the honesty.* parameters are not tuned. Default is "none" (no parameters are tuned).

tune.num.trees	
The number of trees in each 'mini forest' used to fit the tuning model. Default is 200.

tune.num.reps	
The number of forests used to fit the tuning model. Default is 50.

tune.num.draws	
The number of random parameter values considered when using the model to select the optimal parameters. Default is 1000.

```{r}

ml_ready_data <- slope15d

target <- ml_ready_data$target

covars <- ml_ready_data$covars

df_use <- ml_ready_data$dats_ml #|>
  # #filter(ecoregion == "Idaho Batholith") |>
  # filter(!!sym(target) > -100 & !!sym(target) < 100) |>
  # filter(burn_sev == "High")


ggplot(df_use) +
  geom_histogram(aes(x = !!sym(target))) +
  xlim(c(-100, 100))

nrow(df_use)


```



```{r}

form <- reformulate(termlabels = covars, response = target)
form

lm_fit <- lm(form,
             data = df_use)
               # data = df_use |> filter(post_fire_rap_tree_perc_recovery_n6_3_15_yr < 100 & post_fire_rap_tree_perc_recovery_n6_3_15_yr > -100))
summary(lm_fit)


```

# Build and tune a ranger model


```{r}

form <- reformulate(termlabels = covars, response = target)

tic("Random forest train test")
data_split <- rsample::initial_split(df_use, prop = 0.8, strata = ecoregion) #stratify by ecoregion
train_data <- rsample::training(data_split)
test_data  <- rsample::testing(data_split)

rf_recipe <- recipes::recipe(form, data = train_data) |>
  step_zv(all_predictors()) #remove any zero-variance predictors

#Build our random forest specification
rf_spec <- parsnip::rand_forest(
  mtry  = tune(),
  trees = 1000,
  min_n = tune()) |>
  parsnip::set_engine(
    "ranger",
    importance = "permutation") |>
  parsnip::set_mode("regression")

rf_wf <- workflows::workflow() |>
  workflows::add_model(rf_spec) |>
  workflows::add_recipe(rf_recipe)

folds <- rsample::vfold_cv(train_data, v = 5)

mtry_default <- floor(sqrt(length(covars)))
rf_grid <- dials::grid_regular(
  dials::mtry(c(mtry_default - 2, mtry_default + 2)),
  dials::min_n(),
  levels = 5)

rf_metrics <- yardstick::metric_set(yardstick::rmse, yardstick::rsq)

rf_tuned <- rf_wf |>
  tune::tune_grid(
    resamples = folds,
    grid = rf_grid,
    metrics = rf_metrics)

rf_tuned |>
  tune::collect_metrics()

autoplot(rf_tuned)

show_best(rf_tuned)

best_params <- rf_tuned |>
  select_best(metric = "rmse")

best_params


rf_final_wf <- rf_wf |>
  finalize_workflow(best_params)

rf_final_fit <- rf_final_wf |>
  last_fit(split = data_split)

rf_final_fit |>
  collect_metrics()

rf_test_preds <- rf_final_fit |>
  collect_predictions()

rf_fit <- rf_final_fit$.workflow[[1]] |>
  workflows::extract_fit_parsnip()

saveRDS(rf_fit, here(dir_models, "rf_small.rds"))

toc()

# show permutation-based variable importance (included in model specification)
vip_rf <- vip::vip(rf_fit,
         num_features = 40)

vip_rf


#?
X_train <- train_data

# PDPs
rf_aet <- pdp::partial(
  object   = rf_fit,
  pred.var = "aet",
  train    = X_train,
  grid.resolution = 30
)

autoplot(rf_aet)

rf_cbi <- pdp::partial(
  object   = rf_fit,
  pred.var = "burn_cbi",
  train    = X_train,
  grid.resolution = 30
)

autoplot(rf_cbi)

# ICE plots
p_aet_ice <- partial(
  object   = rf_fit$fit,
  pred.var = "aet",
  train    = X_train,
  ice      = TRUE
)

autoplot(p_aet_ice, alpha = 0.1) 


# Examine interaction importance using IML package



# Predictor object with correct prediction wrapper
pred_fun <- function(object, newdata) {
  preds <- predict(object, new_data = newdata)
  preds$.pred
}

predictor <- Predictor$new(
  model = rf_fit,          # parsnip model_fit
  data = X_train, 
  predict.fun = pred_fun
)

# # Measure interaction strength for all features
# interaction_strength <- Interaction$new(predictor)
# interaction_strength$results
# 
# plot(interaction_strength)






# 
# # NOT WORKING YET
# tic()
# # Compute SHAP values using fastshap
# pfun <- function(object, newdata) {
#   predict(object, data = newdata)$predictions
# }
# 
# shap_vals <- fastshap::explain(
#   object = rf_fit$fit,
#   X = X_train,
#   pred_wrapper = pfun,
#   nsim = 10
# )
# 
# # Wrap for visualization
# sv <- shapviz::shapviz(
#   S = shap_vals,
#   X = X_train
# )
# 
# 
# # Global: beeswarm summary (like SHAP summary plot)
# sv_importance <- shapviz::sv_importance(sv)   # optional
# plot(sv)                             # beeswarm by default
# 
# # Dependence plot for one variable
# shapviz::sv_dependence(sv, "aet")
# 
# # Grouped by feature groups (e.g. "climate", "topography", etc.)
# groups <- c(
#   aet = "climate",
#   def = "climate",
#   srtm = "topography",
#   tpi  = "topography",
#   chili = "topography",
#   avg_rap_TRE_6_yr_pre_fire = "veg"
# )
# sv_groups <- add_groups(sv, groups)
# shapviz::sv_importance(sv_groups, kind = "group")
# 
# toc()







```

# Build and train a boosted gradient model

```{r}


tic("Gradient boost train test")

# ------------------------------------------------------------
# 1. Formula and data split
# ------------------------------------------------------------

form <- reformulate(termlabels = covars, response = target)

data_split <- rsample::initial_split(df_use, prop = 0.8, strata = ecoregion)
train_data <- rsample::training(data_split)
test_data  <- rsample::testing(data_split)

# ------------------------------------------------------------
# 2. Recipe (xgboost wants numeric predictors)
# ------------------------------------------------------------

xgb_recipe <- recipes::recipe(form, data = train_data) |>
  recipes::step_zv(all_predictors()) |>
  recipes::step_dummy(all_nominal_predictors(), one_hot = TRUE)

# ------------------------------------------------------------
# 3. XGBoost model specification
# ------------------------------------------------------------

xgb_spec <- parsnip::boost_tree(
  trees          = 1000,
  tree_depth     = tune(),
  learn_rate     = tune(),
  loss_reduction = tune(),  # gamma
  mtry           = tune(),  # number of predictors sampled at each split
  min_n          = tune()   # min observations per node
) |>
  parsnip::set_engine("xgboost") |>
  parsnip::set_mode("regression")

xgb_wf <- workflows::workflow() |>
  workflows::add_model(xgb_spec) |>
  workflows::add_recipe(xgb_recipe)

# ------------------------------------------------------------
# 4. Resampling
# ------------------------------------------------------------

folds <- rsample::vfold_cv(train_data, v = 5)

# ------------------------------------------------------------
# 5. Hyperparameter ranges and grid
# ------------------------------------------------------------

# Prep recipe to see predictor space after dummy encoding
xgb_prep <- xgb_recipe |>
  recipes::prep(training = train_data)

xgb_train_baked <- recipes::bake(xgb_prep, new_data = NULL)

# Drop outcome column to count predictors
n_pred <- ncol(xgb_train_baked) - 1L
mtry_default <- floor(sqrt(n_pred))

xgb_params <- dials::parameters(
  dials::mtry(range = c(max(1L, mtry_default - 2L), mtry_default + 2L)),
  dials::tree_depth(),
  dials::learn_rate(),
  dials::loss_reduction(),
  dials::min_n()
)

set.seed(123)
xgb_grid <- dials::grid_latin_hypercube(
  xgb_params,
  size = 30
)

# ------------------------------------------------------------
# 6. Tuning
# ------------------------------------------------------------

xgb_metrics <- yardstick::metric_set(yardstick::rmse, yardstick::rsq)

xgb_tuned <- xgb_wf |>
  tune::tune_grid(
    resamples = folds,
    grid      = xgb_grid,
    metrics   = xgb_metrics
  )

xgb_tuned |>
  tune::collect_metrics()

autoplot(xgb_tuned)

tune::show_best(xgb_tuned, metric = "rmse")

best_params_xgb <- xgb_tuned |>
  tune::select_best(metric = "rmse")

best_params_xgb

# ------------------------------------------------------------
# 7. Finalize workflow and fit on full training data
# ------------------------------------------------------------

xgb_final_wf <- xgb_wf |>
  tune::finalize_workflow(best_params_xgb)

xgb_final_fit <- xgb_final_wf |>
  tune::last_fit(split = data_split)

xgb_final_fit |>
  tune::collect_metrics()

xgb_test_preds <- xgb_final_fit |>
  tune::collect_predictions()

xgb_test_preds

# ------------------------------------------------------------
# 8. Extract fitted xgboost model and run VIP
# ------------------------------------------------------------

xgb_fit <- xgb_final_fit$.workflow[[1]] |>
  workflows::extract_fit_parsnip()

saveRDS(xgb_fit, here(dir_models, "xgb_small.rds"))

toc()


vip_xgb <- vip::vip(
  xgb_fit,
  num_features = 40
)

vip_xgb


xgb_aet <- pdp::partial(
  object   = xgb_fit,
  pred.var = "aet",
  train    = X_train,
  grid.resolution = 30
)

autoplot(rf_aet)



```

# Set up for robust post-hoc analysis

```{r}

plot_pdp_ice_ale <- function(fit, train, pred_var, fig_dir,
                             grid_resolution = 30,
                             ice_alpha = 0.1,
                             width = 6, height = 9, dpi = 300) {
  # ---------------------------------------------------------------------------
  # 0. Basic checks & setup
  # ---------------------------------------------------------------------------
  if (!inherits(fit, "model_fit")) {
    stop("`fit` should be a parsnip model_fit (e.g., from extract_fit_parsnip()).")
  }
  
  if (!pred_var %in% names(train)) {
    stop("pred_var '", pred_var, "' not found in 'train'.")
  }
  
  if (!dir.exists(fig_dir)) {
    dir.create(fig_dir, recursive = TRUE)
  }
  
  train_df <- as.data.frame(train)
  
  # ---------------------------------------------------------------------------
  # 1. PDP and ICE via pdp using model_fit directly
  #    (this is exactly what you did manually)
  # ---------------------------------------------------------------------------
  pd_pdp <- pdp::partial(
    object   = fit,
    pred.var = pred_var,
    train    = train_df,
    grid.resolution = grid_resolution,
  )
  
  pd_ice <- pdp::partial(
    object   = fit,
    pred.var = pred_var,
    train    = train_df,
    grid.resolution = grid_resolution,
    ice      = TRUE,
  )
  
  # Figure out PDP column names
  # Typically: pred_var + "yhat"
  if (!all(c(pred_var, "yhat") %in% names(pd_pdp))) {
    stop("PDP object does not contain expected columns '", pred_var, "' and 'yhat'.")
  }
  
  # ICE: need pred_var, yhat, and an ID column
  if (!all(c(pred_var, "yhat") %in% names(pd_ice))) {
    stop("ICE object does not contain expected columns '", pred_var, "' and 'yhat'.")
  }
  
  id_col <- NULL
  if (".id" %in% names(pd_ice)) {
    id_col <- ".id"
  } else if ("yhat.id" %in% names(pd_ice)) {
    id_col <- "yhat.id"
  } else {
    pd_ice$.ice_id <- seq_len(nrow(pd_ice))
    id_col <- ".ice_id"
  }
  
  # ---------------------------------------------------------------------------
  # 2. ALE via iml with a parsnip-friendly predict.fun
  # ---------------------------------------------------------------------------
  pred_fun_iml <- function(object, newdata) {
    preds <- predict(object, new_data = newdata)
    if (".pred" %in% names(preds)) {
      preds$.pred
    } else {
      # fallback: first numeric column
      num_cols <- vapply(preds, is.numeric, logical(1))
      if (!any(num_cols)) stop("predict() output has no numeric columns.")
      preds[[which(num_cols)[1]]]
    }
  }
  
  predictor_iml <- iml::Predictor$new(
    model = fit,
    data  = train_df,
    predict.fun = pred_fun_iml
  )
  
  eff_ale <- iml::FeatureEffect$new(
    predictor = predictor_iml,
    feature   = pred_var,
    method    = "ale"
  )
  
  ale_df <- eff_ale$results
  
  if (!pred_var %in% names(ale_df)) {
    x_col <- names(ale_df)[1]
  } else {
    x_col <- pred_var
  }
  y_cols <- setdiff(names(ale_df), x_col)
  if (length(y_cols) < 1) {
    stop("ALE results do not contain an effect column.")
  }
  y_col <- y_cols[1]
  
  names(ale_df)[names(ale_df) == x_col] <- "x"
  names(ale_df)[names(ale_df) == y_col] <- "ale"
  
  # ---------------------------------------------------------------------------
  # 3. Build the three ggplots: PDP, ICE, ALE
  # ---------------------------------------------------------------------------
  library(ggplot2)
  library(patchwork)
  
  p_pdp <- ggplot(pd_pdp, aes_string(x = pred_var, y = "yhat")) +
    geom_line() +
    labs(
      title = paste0("Partial Dependence (PDP): ", pred_var),
      x = pred_var,
      y = "Predicted outcome (average)"
    ) +
    theme_minimal()
  
  p_ice <- ggplot(
    pd_ice,
    aes_string(x = pred_var, y = "yhat", group = id_col)
  ) +
    geom_line(alpha = ice_alpha) +
    labs(
      title = paste0("Individual Conditional Expectation (ICE): ", pred_var),
      x = pred_var,
      y = "Predicted outcome"
    ) +
    theme_minimal()
  
  vals <- round(range(ale_df$.value), 2)
  p_ale <- ggplot(ale_df, aes(x = x, y = .value)) +
    geom_line() +
    labs(
      title = paste0("Accumulated Local Effects (ALE): ", pred_var),
      x = pred_var,
      y = "ALE",
      caption = glue("ALE Range: {vals[1]}, {vals[2]}")
    ) +
    theme_minimal()
  
  combined_plot <- p_pdp / p_ice / p_ale
  
  # ---------------------------------------------------------------------------
  # 4. Save PNG
  # ---------------------------------------------------------------------------
  fit_label <- class(fit)[1]
  fit_label <- gsub("[^A-Za-z0-9_]+", "_", fit_label)
  pred_label <- gsub("[^A-Za-z0-9_]+", "_", pred_var)
  
  file_name <- paste0("pdp_ice_ale_", fit_label, "_", pred_label, ".png")
  out_path  <- file.path(fig_dir, file_name)
  
  ggsave(
    filename = out_path,
    plot     = combined_plot,
    width    = width,
    height   = height,
    dpi      = dpi
  )
  
  message("Saved: ", out_path)
  return(list("ale_df" = ale_df))
}

plot_pdp_ice_ale(
  fit      = rf_fit, 
  train    = X_train,
  pred_var = "srtm",
  fig_dir  = dir_figs
)

plot_pdp_ice_ale(
  fit      = rf_fit, 
  train    = X_train,
  pred_var = "pdsi_annual_5_yrs_after_mean",
  fig_dir  = dir_figs
)



```




# Variable interpretation
- TPI: low = valleys, high = ridges
- CHILI: low = cold, high = hot
- SRTM: low = low elevation, high = high elevation
- AET: low = wet, high = dry
- DEF: low = wet, high = dry
- PDSI: low = dry, high = wet

```{r}


rf_fit <- ranger(
  formula   = form,
  data      = df_use,
  num.trees = 1000,
  mtry      = 30,
  importance = "permutation",
  min.node.size = 5,
  num.threads   = parallel::detectCores(),
  seed = 42
)
print(rf_fit)
oob_mse  <- rf_fit$prediction.error
oob_rmse <- sqrt(oob_mse)
oob_rmse


# True and OOB-predicted values
y_true <- df_use[[target]]
y_pred <- predict(rf_fit, df_use)$predictions

# RMSE
rmse <- sqrt(mean((y_true - y_pred)^2))

# %RMSE (range-based)
perc_rmse <- (rmse / (max(y_true) - min(y_true))) * 100

# %RMSE (mean-based)
perc_rmse_mean <- (rmse / mean(y_true)) * 100

# Percent Variance Explained
sse <- sum((y_true - y_pred)^2)
sst <- sum((y_true - mean(y_true))^2)

perc_var_explained <- (1 - sse/sst) * 100

list(
  RMSE = rmse,
  percent_RMSE_range = perc_rmse,
  percent_RMSE_mean = perc_rmse_mean,
  percent_variance_explained = perc_var_explained
)



tic()
spmlmRF <- splmRF(form,
                  data = df_use ,
                  spcov_type = "exponential",
                  importance = 'permutation',
                  scale.permutation.importance = TRUE,
                  num.threads = 0,
                  mtry = floor(sqrt(ncol(dats) - 1)),
                  num.trees = 1000,
                  seed = seed)
toc()
summary(spmlmRF)
print(spmlmRF$ranger)
#500 trees mtry 3, MSE 213, R2 0.49
#1000 trees mtry 8, basically same
#1000t, mtry8, include disturbance MSE 2.17, R^2 0.14


vip(test_spmlmRF$ranger,
    num_features = 40)
pdp(test_spmlmRF$ranger)




```


# MIR
```{r}
compute_mir_groups <- function(
  model_spec,
  data_split,
  outcome,
  predictors,
  groups,
  metrics    = yardstick::metric_set(yardstick::rmse),
  recipe_fun = NULL
) {
  # Split data
  train_data <- rsample::training(data_split)
  test_data  <- rsample::testing(data_split)
  
  # Basic checks
  if (!outcome %in% names(train_data)) {
    stop("Outcome '", outcome, "' not found in training data.")
  }
  if (!all(predictors %in% names(train_data))) {
    missing <- setdiff(predictors, names(train_data))
    stop(
      "Some predictors not found in training data: ",
      paste(missing, collapse = ", ")
    )
  }
  
  # Helper: build a recipe for a given predictor set
  build_recipe <- function(preds, data) {
    form <- reformulate(termlabels = preds, response = outcome)
    
    if (is.null(recipe_fun)) {
      recipes::recipe(form, data = data) |>
        recipes::step_zv(recipes::all_predictors())
    } else {
      recipe_fun(form, data)
    }
  }
  
  # ------------------------
  # Baseline model (all predictors)
  # ------------------------
  base_recipe <- build_recipe(predictors, train_data)
  
  base_wf <- workflows::workflow() |>
    workflows::add_model(model_spec) |>
    workflows::add_recipe(base_recipe)
  
  base_fit <- base_wf |>
    fit(data = train_data)
  
  base_pred <- predict(base_fit, new_data = test_data) |>
    dplyr::bind_cols(
      truth = test_data[[outcome]]
    )
  
  base_metrics <- metrics(
    data    = base_pred,
    truth   = truth,
    estimate = .pred
  ) |>
    dplyr::mutate(
      group              = "baseline",
      removed_predictors = NA_character_
    )
  
  # ------------------------
  # Function to evaluate one group
  # ------------------------
  # IMPORTANT: first argument = group vars (.x), second = group name (.y)
  eval_group <- function(g_vars, g_name) {
    preds_reduced <- setdiff(predictors, g_vars)
    
    rec_reduced <- build_recipe(preds_reduced, train_data)
    
    wf_reduced <- workflows::workflow() |>
      workflows::add_model(model_spec) |>
      workflows::add_recipe(rec_reduced)
    
    fit_reduced <- wf_reduced |>
      fit(data = train_data)
    
    pred_reduced <- predict(fit_reduced, new_data = test_data) |>
      dplyr::bind_cols(
        truth = test_data[[outcome]]
      )
    
    met_reduced <- metrics(
      data    = pred_reduced,
      truth   = truth,
      estimate = .pred
    ) |>
      dplyr::mutate(
        group              = g_name,
        removed_predictors = paste(g_vars, collapse = ", ")
      )
    
    met_reduced
  }
  
  # Apply over groups (named list)
  group_results <- purrr::imap_dfr(groups, eval_group)
  
  # ------------------------
  # Combine + compute MIR
  # ------------------------
  combined <- dplyr::bind_rows(base_metrics, group_results)
  
  # Infer metric direction from name (no yardstick::direction)
  combined <- combined |>
    dplyr::mutate(
      direction = dplyr::case_when(
        .metric %in% c("rmse", "mae", "mape", "mse") ~ "minimize",
        .metric %in% c("rsq", "rsq_trad", "rsq_adj") ~ "maximize",
        TRUE ~ "minimize"   # default; adjust if you add others
      )
    )
  
  combined |>
    dplyr::left_join(
      combined |>
        dplyr::filter(group == "baseline") |>
        dplyr::select(.metric, base = .estimate),
      by = ".metric"
    ) |>
    dplyr::mutate(
      delta = dplyr::case_when(
        direction == "minimize" ~ .estimate - base,  # worse = larger value
        direction == "maximize" ~ base - .estimate,  # worse = smaller value
        TRUE ~ NA_real_
      ),
      mir = dplyr::case_when(
        base != 0 ~ delta / abs(base),
        TRUE ~ NA_real_
      )
    ) |>
    dplyr::select(
      group,
      removed_predictors,
      .metric,
      direction,
      base,
      dropped = .estimate,
      delta,
      mir
    ) |>
    dplyr::arrange(.metric, dplyr::desc(mir))
}




rf_spec_final <- rf_final_wf |>
  workflows::extract_spec_parsnip()

# Outcome and predictors based on your existing variables
outcome    <- target       # e.g. "post_fire_rap_TRE_difference_n2_3_yr"
predictors <- covars       # character vector of predictor column names

mir_groups <- list(
  climate_before = c(
    "pdsi_annual_5_yrs_prior_sum",
    "pdsi_annual_5_yrs_prior_mean",
    "pdsi_summer_5_yrs_prior_sum",
    "pdsi_summer_5_yrs_prior_mean",
    "pdsi_annual_5_yrs_prior_threshold",
    "pdsi_summer_5_yrs_prior_threshold"
  ),
  
  climate_after = c(
    "pdsi_annual_5_yrs_after_sum",
    "pdsi_annual_5_yrs_after_mean",
    "pdsi_summer_5_yrs_after_sum",
    "pdsi_summer_5_yrs_after_mean",
    "pdsi_annual_5_yrs_after_threshold",
    "pdsi_summer_5_yrs_after_threshold"
  ),
  
  topo = c(
    "srtm",
    "tpi",
    "chili"
  ),
  
  prefire_tree = c(
    "avg_rap_tree_6_yr_pre_fire"
  ),
  
  burn_severity = c(
    "burn_cbi",
    "burn_cbi_500mmwmean",
    "burn_cbi_500mmwstd",
    "burn_distance_forest_after"
  ),
  
  climate_normals = c(
    "tt_change_aet",
    "tt_change_def",
    "tt_early_aet",
    "tt_early_def",
    "tt_later_aet",
    "tt_later_def",
    "tt_normal_aet",
    "tt_normal_def",
    "tt_future_aet",
    "tt_future_def",
    "aet",
    "def"
  ),
  
  forest_group = c(
    "ENC1_nfg_factor",
    "ENC2_nfg_factor"
  ),
  
  regime_group = c(
    "ENC1_frg_reclass_desc",
    "ENC2_frg_reclass_desc"
  ),
  
  location = c(
    "lat",
    "long"
  ),
  
  year = c(
    "mock_burn_year"
  )
)

mir_results <- compute_mir_groups(
  model_spec  = rf_spec_final,
  data_split  = data_split,
  outcome     = outcome,
  predictors  = predictors,
  groups      = mir_groups,
  metrics     = yardstick::metric_set(yardstick::rmse, yardstick::rsq)
)

mir_results

ggplot(mir_results) +
  geom_col(aes(x = mir, y = group)) +
  facet_grid(~.metric)


```




#################
# From run with all PDSI data, the cumulative drought sums appear to be more important than the means, which are more important than the thresholds

```{r}

narrower_covars <- all_covars[!str_ends(all_covars, "mean") &
                                !str_ends(all_covars, "threshold") &
                                !str_starts(all_covars, "pdsi_summer")]


form_str <- paste(
  "post_fire_rap_TRE_slope_2_10_yrs ~ burn_sev + ",
  paste(narrower_covars, collapse = " + ")
)

form <- as.formula(form_str)

tic()
test_spmlmRF <- splmRF(form,
                  data = df_test,
                  spcov_type = "exponential",
                  importance = 'permutation',
                  scale.permutation.importance = TRUE,
                  num.threads = 0,
                  mtry = floor(sqrt(ncol(dats) - 1)),
                  num.trees = 1000,
                  seed = seed)
toc()
summary(test_spmlmRF)
print(test_spmlmRF$ranger)
print(test_spmlmRF$call)

#500 trees mtry 3, MSE 213, R2 0.49
#1000 trees mtry 8, basically same
#1000t, mtry8, include disturbance MSE 2.17, R^2 0.14


vip <- vip(test_spmlmRF$ranger,
    num_features = 25)
vip_tbl <- vip(test_spmlmRF$ranger, num_features = 25)$data |>
  as_tibble() |>
  arrange(desc(Importance)) |>   # highest first
  mutate(variable = factor(Variable, levels = Variable))

ggsave(plot = vip, filename = here(dir_rf_figs, "vip_idbath_v1.png"), units = "px", width = 3000, height = 4000, )


pdp_cont <- c(narrower_covars) |>
  set_names() |>
  map_dfr(\(v) {
    # run PDP
    pd <- partial(
      object   = test_spmlmRF$ranger,
      pred.var = v,
      train    = df_test |> st_drop_geometry()
    )

    # rename predictor column to a common name 'x'
    pd |>
      rename(x = !!sym(v)) |>
      mutate(variable = v, .before = x)
  })

pdp_sorted <- pdp_cont |>
  inner_join(
    vip_tbl |> select(variable = Variable),
    by = "variable"
  ) |>
  mutate(variable = factor(variable, levels = vip_tbl$Variable))

plot <- pdp_sorted |>
  ggplot(aes(x = x, y = yhat)) +
  geom_line() +
  facet_wrap(~ variable, scales = "free_x") +
  labs(
    x = "Predictor value",
    y = "Partial dependence (yhat)",
    title = "Partial dependence for continuous covariates"
  )

ggsave(plot = plot, filename = here(dir_rf_figs, "pdp_idbath_v1.png"), units = "px", width = 3000, height = 4000, )


```




<button class="accordion-button" type="button" data-bs-toggle="collapse" data-bs-target="#collapseOne" >Click to see Session Information</button><div id="collapseOne" class="accordion-collapse collapse"><div>

```{r sessionInfo}
sessionInfo()
```