---
title: "random_forest"
author: Tyler L. McIntosh
date: last-modified
format:
  html:
    embed-resources: true
    html-math-method: katex
    css: styles.css
    toc: true
    toc-depth: 1
    toc-title: Contents
    toc-location: right
    code-fold: true
    code-summary: "Show the code"
---

# Setup for analysis

``` {r setupchunk}
#| echo: true
#| message: false
#| warning: false
#| results: 'hide'

rm(list = ls())

if(!requireNamespace("here", quietly = TRUE)) {
  install.packages("here")
}
library(here)

source(here::here("utils", "functions.R"))

install_load_packages(c(
  "tidyverse",
  "glue",
  "patchwork",
  "tictoc",
  "ranger",
  "spmodel",
  "iml",
  "patchwork",
  "spatialsample",
  "tidymodels",
  "tidysdm",
  "purrr",
  "arrow",
  "vip",
  "pdp",
  "fastshap",
  "shapviz",
  "dials",
  "sf",
  "ggrepel",
  "corrplot",
  "tictoc"))

# devtools::install_github("grf-labs/sufrep")
# library(sufrep)

run_nm <- 'v3_spatialcv_20251209'

dir_figs <- here::here('figs', run_nm)
dir_derived <- here::here('data', 'derived')
dir_raw <- here::here('data', 'raw')
dir_manual <- here::here('data', 'manual')
dir_models <- here('models', run_nm)

dir_rf_figs <- here(dir_figs, "rf")

dir_ensure(c(dir_figs,
             dir_derived,
             dir_manual,
             dir_raw,
             dir_rf_figs,
             dir_models))

seed = 1234
set.seed(seed)


```
# Load dats

```{r}

#Short dats
dats <- read_parquet(here(dir_derived, "test_dats_short.parquet"))
ecoregions <- unique(dats$ecoregion)

geo_dats <- st_read(here(dir_derived, "test_dats_sf.gpkg")) |>
  st_transform('EPSG:5070') |>
  select(uniqueID)

```





# Prep data for random forest machine learning
- We do not need to normalize data for RF
- We do want to encode categorical variables as a lower-dimensional real-valued representation for inclusion in random/causal forest (use sufficient representations instead of one-hot encoding since random forests "may largely ignore high-cardinality one-hot encoded factors" - see Johannemann et al 2021). Low-rank encoding appears to be the most applicable to real-world data.

```{r}

# A function to use sufrep::make_encoder such that the encoded variables are added to the dataframe
# used in prep_ml_dats
sufrep_encode_append <- function(dats,
                                 method = 'means',
                                 X_cols,
                                 G_col,
                                 remove_G_col = TRUE,
                                 ...) {
  
  #Select variables for encoding
  x_df <- dats |> 
    select(all_of(X_cols)) |>
    as.data.frame()
  
  # Grouping variable as factor with dropped unused levels
  # The function breaks if we don't drop unused levels because sufrep assumes that all levels are represented at least once in the data
  g_raw <- dats[[G_col]]

  if (!is.factor(g_raw)) {
    g_fac <- factor(g_raw)
  } else {
    g_fac <- droplevels(g_raw)
  }
  
  #Encode
  encoder <- sufrep::make_encoder(method = method, X = x_df, G = g_fac, ...)
  dats_encoded <- encoder(x_df, G = g_fac)
  
  #Make interpretable encoding columns
  encoded_cols <- dats_encoded |> select(starts_with("ENC"))
  names(encoded_cols) <- paste0(names(encoded_cols), "_", G_col)
  
  dats_out <- dats |>
    cbind(encoded_cols)
  
  if(remove_G_col) {
    dats_out <- dats_out |>
      select(-all_of(G_col))
  }
  
  return(dats_out)
}

# Select relevant columns, filter out NAs, ensure correct types, and sufrep encode
prep_ml_dats <- function(dats, target) {
  disturbance_covars <- dats |>
    select(starts_with("burn_cbi") | 
             starts_with("burn_distance") | 
             #starts_with("biotic") | 
             #starts_with("hd") | 
             ends_with("mean")) |>
    names()
  
  cont_covars <- c(#"aet", "def", 
                   "srtm", "tpi", "chili", "avg_rap_tree_6_yr_pre_fire", "uniqueID",
                   #"lat", "long",
                   "mock_burn_year",  disturbance_covars, grep("^tt_change", names(dats), value = TRUE), grep("^tt_normal", names(dats), value = TRUE))
  cat_covars <- c("frg_reclass_desc", "ecoregion", "nfg_factor")
  trts <- c("fire", "burn_sev")
  
  all_cols <- c(cont_covars, cat_covars, target, trts)

  
  # Scale all continuous covariates
  # filter out anything with NAs
  # ensure categorical covariates are factors
  # encode categorical as one-hot dummies or sufficient representations
  dats_ml <- dats |>
    select(all_of(all_cols)) |>
    filter(dplyr::if_all(everything(), ~ !is.na(.x))) |>
    filter(!is.infinite(!!sym(target))) |>
    mutate(
      across(
        all_of(cont_covars),
        \(x) as.numeric(x)
      )
    ) |>
    mutate(
      across(
        all_of(cat_covars),
        \(x) as.factor(x)
      )
    ) |>
    # fastDummies::dummy_cols(select_columns = c("frg_reclass"),
    #                         remove_first_dummy = FALSE, #Keep full one-hot encoding for tree-based methods
    #                         remove_selected_columns = TRUE) |> 
    sufrep_encode_append(method = 'low_rank', # Use low-rank sufficient representation encoding to reduce to 2 dimensions
                         X_cols = cont_covars,
                         G_col = "nfg_factor",
                         num_components = 2) |> 
    sufrep_encode_append(method = 'low_rank', # Use low-rank sufficient representation encoding to reduce to 2 dimensions
                         X_cols = cont_covars,
                         G_col = "frg_reclass_desc",
                         num_components = 2)
  
  enc_xvars <- dats_ml |>
    select(starts_with("ENC")) |>
    names()
  
  all_covars <- c(cont_covars, enc_xvars)
  
  return(list("dats_ml" = dats_ml,
              "target" = target,
              "covars" = all_covars))
}



#Create operational datasets - we need a separate one for each because the sufrep encodings will be different for each
perc10d <- dats |>
  prep_ml_dats(target = "post_fire_rap_tree_perc_recovery_n6_3_10_yr")

perc15d <- dats |>
  prep_ml_dats(target = "post_fire_rap_tree_perc_recovery_n6_3_15_yr")

perc20d <- dats |>
  prep_ml_dats(target = "post_fire_rap_tree_perc_recovery_n6_3_20_yr")

slope15d <- dats |>
  prep_ml_dats(target = "post_fire_rap_tree_slope_3_15_yrs")


ml_ready_data <- slope15d

target <- ml_ready_data$target

covars <- ml_ready_data$covars

df_use <- ml_ready_data$dats_ml |>
  left_join(geo_dats) |>
  filter(ecoregion == "Idaho Batholith") |>
  select(-uniqueID)#>
  # filter(!!sym(target) > -100 & !!sym(target) < 100) |>
  # filter(burn_sev == "High")

covars <- covars[covars != "uniqueID"]


ggplot(df_use) +
  geom_histogram(aes(x = !!sym(target))) +
  xlim(c(-3, 3))

nrow(df_use)

```


# Tested model tuning time bumps

Tuning all parameters as opposed to a select few increases training time by ~50%.

GRF package tuning parameters:

tune.parameters	
A vector of parameter names to tune. If "all": all tunable parameters are tuned by cross-validation. The following parameters are tunable: ("sample.fraction", "mtry", "min.node.size", "honesty.fraction", "honesty.prune.leaves", "alpha", "imbalance.penalty"). If honesty is FALSE the honesty.* parameters are not tuned. Default is "none" (no parameters are tuned).

tune.num.trees	
The number of trees in each 'mini forest' used to fit the tuning model. Default is 200.

tune.num.reps	
The number of forests used to fit the tuning model. Default is 50.

tune.num.draws	
The number of random parameter values considered when using the model to select the optimal parameters. Default is 1000.



# Assess correlations

```{r}


# Correlation matrix
correlations <- df_use |>
  select(-starts_with("ENC"), -fire, -ecoregion, -burn_sev, -geom) |>
  st_drop_geometry() |>
  cor(use = "pairwise.complete.obs")
png(here(dir_figs, glue('corr_plot.png')), width = 2000, height = 2000)
corrplot(correlations, method = 'circle', type = "lower", diag = FALSE, use = 'pairwise.complete.obs') 
dev.off()



plot_feature_densities_combined <- function(df) {
  
  df_long <- df |>
    select(where(is.numeric)) |>
    mutate(across(everything(), scale)) |>
    pivot_longer(
      cols = everything(),
      names_to = "feature",
      values_to = "value"
    )
  
  ggplot(df_long, aes(x = value, color = feature)) +
    geom_density(linewidth = 1, na.rm = TRUE) +
    labs(
      x = "Z-scaled value",
      y = "Density",
      color = "Feature"
    ) +
    theme_minimal()
}

plot_feature_densities_combined(df_use |>
  select(-starts_with("ENC"), -fire, -ecoregion, -burn_sev, -geom)) +
  xlim(c(-3, 3))

ggsave(here(dir_figs, "scaled_densities.png"))


```

# Simple LM out of curiousity

```{r}

form <- reformulate(termlabels = covars, response = target)
form

lm_fit <- lm(form,
             data = df_use)
               # data = df_use |> filter(post_fire_rap_tree_perc_recovery_n6_3_15_yr < 100 & post_fire_rap_tree_perc_recovery_n6_3_15_yr > -100))
summary(lm_fit)


```


# Full spatial cross-validation and training on test set

```{r}

form <- reformulate(termlabels = covars, response = target)

df_mini <- df_use |> slice_sample(n = 2000)


test_spatial_split <- function(dats, type, cellsize) {
  
  if(type == "SPATIAL") {
    # Spatial train/test split
    data_split <- tidysdm::spatial_initial_split(
      data = dats |> sf::st_as_sf(),
      prop = 0.2, #prop here is for amount in TESTING set
      strategy = spatial_block_cv,
      cellsize = cellsize,
      square = TRUE,
    )
    train_data <- training(spatial_split)
    test_data  <- testing(spatial_split)
  } else {
    # Naive data split
    data_split <- rsample::initial_split(dats, prop = 0.8#, strata = ecoregion
                                         ) #stratify by ecoregion
    train_data <- rsample::training(data_split)
    test_data  <- rsample::testing(data_split)
    
  }
  
  
  rf_recipe <- recipes::recipe(form, data = train_data) |>
    step_zv(all_predictors()) #remove any zero-variance predictors
  
  #Build our random forest specification
  rf_spec <- parsnip::rand_forest(
    mtry  = tune(),
    trees = 1000,
    min_n = tune()) |>
    parsnip::set_engine(
      "ranger",
      importance = "permutation") |>
    parsnip::set_mode("regression")
  
  rf_wf <- workflows::workflow() |>
    workflows::add_model(rf_spec) |>
    workflows::add_recipe(rf_recipe)
  

  
  if(type == "SPATIAL") {
    #SPATIAL FOLDS
    folds <- spatial_block_cv(
      data = train_data |> sf::st_as_sf(),
      v = 5,
      cellsize = cellsize,
      square = TRUE
    )
  } else {
    # NAIVE FOLDS
    folds <- rsample::vfold_cv(train_data, v = 5)
  }
  
  
  mtry_default <- floor(sqrt(length(covars)))
  rf_grid <- dials::grid_regular(
    dials::mtry(c(mtry_default - 2, mtry_default + 2)),
    dials::min_n(),
    levels = 5)
  
  rf_metrics <- yardstick::metric_set(yardstick::rmse, yardstick::rsq)
  
  rf_tuned <- rf_wf |>
    tune::tune_grid(
      resamples = folds,
      grid = rf_grid,
      metrics = rf_metrics)
  
  rf_tuned |>
    tune::collect_metrics()
  
  best_params <- rf_tuned |>
    select_best(metric = "rmse")
  
  rf_final_wf <- rf_wf |>
    finalize_workflow(best_params)
  
  rf_final_fit <- rf_final_wf |>
    last_fit(split = data_split)
  
  final_metrics <- rf_final_fit |>
    collect_metrics() |>
    mutate(cell_size = cellsize,
           type = type,
           name = paste0(type, "_", cellsize))
  
  return(final_metrics)
}


cell_sizes <- c(NA, 10000, 25000, 50000)
types <- c("NON-SPATIAL", "SPATIAL", "SPATIAL", "SPATIAL")


spatial_testing <- map2(types, cell_sizes, \(a,b) test_spatial_split(dats = df_mini, type = a, cellsize = b))

all_sp_test <- bind_rows(spatial_testing)

write_csv(all_sp_test, here(dir_figs, "sp_test.csv"))

ggplot(data = all_sp_test) +
  geom_col(aes(x = name, fill = .metric, y = .estimate), position = 'dodge') +
  labs(title = "Assessing the impact of spatial blocking on RMSE and R^2",
       y = "Estimate", x = "Test",
       caption = "Using minimal test set of 2000 samples") +
  theme_minimal()

ggsave(here(dir_figs, "sp_test.png"))

```



# Build and tune a ranger model
- non-spatial tuning
- spatial assessment


```{r}

form <- reformulate(termlabels = covars, response = target)

tic("Random forest train test")

# Spatial train/test split
spatial_split <- spatial_initial_split(
  data = df_use |> sf::st_as_sf(),
  prop = 0.2,
  strategy = spatial_block_cv,
  cellsize = 25000,
  square = TRUE,
  method = "random"
)
train_data <- training(spatial_split)
test_data  <- testing(spatial_split)

sp_testtrain_p <- autoplot(spatial_split) +
  labs(title = "Spatially-blocked test-train split")

ggsave(sp_testtrain_p, filename = here(dir_figs, "sp_testtrain_p.png"))



rf_recipe <- recipes::recipe(form, data = train_data) |>
  step_zv(all_predictors()) #remove any zero-variance predictors

#Build our random forest specification
rf_spec <- parsnip::rand_forest(
  mtry  = tune(),
  trees = 1000,
  min_n = tune()) |>
  parsnip::set_engine(
    "ranger",
    importance = "permutation") |>
  parsnip::set_mode("regression")

rf_wf <- workflows::workflow() |>
  workflows::add_model(rf_spec) |>
  workflows::add_recipe(rf_recipe)

# Actual folds to use to avoid rapidly growing computation costs of spatial cross-validation
folds <- rsample::vfold_cv(train_data, v = 5)

#EXAMPLE OF SPATIAL FOLDS TO SEE WHAT SPATIAL CROSS-VALIDATION WOULD LOOK LIKE
sample_spatial_folds <- spatial_block_cv(
  data = train_data |> sf::st_as_sf(),
  v = 5,
  cellsize = 25000,
  square = TRUE
)

sp_fold_p <- autoplot(sample_spatial_folds) +
  labs(title = "Example of spatially-blocked cross-validation folds")

ggsave(sp_fold_p, filename = here(dir_figs, "sp_fold_p.png"))




mtry_default <- floor(sqrt(length(covars)))
rf_grid <- dials::grid_regular(
  dials::mtry(c(mtry_default - 2, mtry_default + 2)),
  dials::min_n(),
  levels = 5)

rf_metrics <- yardstick::metric_set(yardstick::rmse, yardstick::rsq)

rf_tuned <- rf_wf |>
  tune::tune_grid(
    resamples = folds,
    grid = rf_grid,
    metrics = rf_metrics)

rf_tuned |>
  tune::collect_metrics()

autoplot(rf_tuned)

show_best(rf_tuned)

best_params <- rf_tuned |>
  select_best(metric = "rmse")

best_params


rf_final_wf <- rf_wf |>
  finalize_workflow(best_params)

rf_final_fit <- rf_final_wf |>
  last_fit(split = spatial_split)

rf_final_fit |>
  collect_metrics()

rf_test_preds <- rf_final_fit |>
  collect_predictions()

rf_fit <- rf_final_fit$.workflow[[1]] |>
  workflows::extract_fit_parsnip()

saveRDS(rf_fit, here(dir_models, "rf_idb_500m_25kmtesttrain.rds"))

toc()

# show permutation-based variable importance (included in model specification)
vip_rf <- vip::vip(rf_fit,
         num_features = 40)

vip_rf




# # Examine interaction importance using IML package
# # Tell future to ignore big R6 reference objects
# options(future.globals.onReference = "ignore")
# 
# # Sequential plan (still uses future, but no workers)
# old_plan <- future::plan()
# future::plan("sequential")
# 
# # Lean ranger engine from parsnip model
# rf_engine <- rf_fit$fit
# 
# # Small, self-contained prediction wrapper using predict.ranger
# pred_fun <- local({
#   predict_ranger <- getFromNamespace("predict.ranger", "ranger")
#   
#   function(object, newdata) {
#     predict_ranger(object, data = newdata)$predictions
#   }
# })
# 
# # Build iml Predictor
# predictor <- iml::Predictor$new(
#   model       = rf_engine,
#   data        = X_train,
#   predict.fun = pred_fun
# )
# 
# # Global interaction strength for all features
# interaction_strength <- iml::Interaction$new(predictor)
# 
# interaction_results <- interaction_strength$results |>
#   arrange(desc(.interaction))
# 
# interaction_results
# plot(interaction_strength)
# 
# # Restore previous future plan
# future::plan(old_plan)




# 
# # NOT WORKING YET
# tic()
# # Compute SHAP values using fastshap
# pfun <- function(object, newdata) {
#   predict(object, data = newdata)$predictions
# }
# 
# shap_vals <- fastshap::explain(
#   object = rf_fit$fit,
#   X = X_train,
#   pred_wrapper = pfun,
#   nsim = 10
# )
# 
# # Wrap for visualization
# sv <- shapviz::shapviz(
#   S = shap_vals,
#   X = X_train
# )
# 
# 
# # Global: beeswarm summary (like SHAP summary plot)
# sv_importance <- shapviz::sv_importance(sv)   # optional
# plot(sv)                             # beeswarm by default
# 
# # Dependence plot for one variable
# shapviz::sv_dependence(sv, "aet")
# 
# # Grouped by feature groups (e.g. "climate", "topography", etc.)
# groups <- c(
#   aet = "climate",
#   def = "climate",
#   srtm = "topography",
#   tpi  = "topography",
#   chili = "topography",
#   avg_rap_TRE_6_yr_pre_fire = "veg"
# )
# sv_groups <- add_groups(sv, groups)
# shapviz::sv_importance(sv_groups, kind = "group")
# 
# toc()







```

# Build and train a boosted gradient model

```{r}


tic("Gradient boost train test")

# USE THE SAME FORMULA AND TEST/TRAIN SPLIT AS BEFORE

# ------------------------------------------------------------
# 2. Recipe (xgboost wants numeric predictors)
# ------------------------------------------------------------

xgb_recipe <- recipes::recipe(form, data = train_data) |>
  recipes::step_zv(all_predictors()) #|>
  #recipes::step_dummy(all_nominal_predictors(), one_hot = TRUE)

# ------------------------------------------------------------
# 3. XGBoost model specification
# ------------------------------------------------------------

xgb_spec <- parsnip::boost_tree(
  trees          = 1000,
  tree_depth     = tune(),
  learn_rate     = tune(),
  loss_reduction = tune(),
  mtry           = tune(),
  min_n          = tune()
) |>
  parsnip::set_engine("xgboost") |>
  parsnip::set_mode("regression")

xgb_wf <- workflows::workflow() |>
  workflows::add_model(xgb_spec) |>
  workflows::add_recipe(xgb_recipe)

# ------------------------------------------------------------
# 4. Resampling
# ------------------------------------------------------------

folds <- rsample::vfold_cv(train_data, v = 5)

# ------------------------------------------------------------
# 5. Hyperparameter ranges and grid
# ------------------------------------------------------------

# Prep recipe to see predictor space after dummy encoding
xgb_prep <- xgb_recipe |>
  recipes::prep(training = train_data)

xgb_train_baked <- recipes::bake(xgb_prep, new_data = NULL)

# Drop outcome column to count predictors
n_pred <- ncol(xgb_train_baked) - 1L
mtry_default <- floor(sqrt(n_pred))

xgb_params <- dials::parameters(
  dials::mtry(range = c(max(1L, mtry_default - 2L), mtry_default + 2L)),
  dials::tree_depth(),
  dials::learn_rate(),
  dials::loss_reduction(),
  dials::min_n()
)

set.seed(123)
xgb_grid <- dials::grid_latin_hypercube(
  xgb_params,
  size = 30
)

# ------------------------------------------------------------
# 6. Tuning
# ------------------------------------------------------------

xgb_metrics <- yardstick::metric_set(yardstick::rmse, yardstick::rsq)

xgb_tuned <- xgb_wf |>
  tune::tune_grid(
    resamples = folds,
    grid = xgb_grid,
    metrics = xgb_metrics
  )

xgb_tuned |>
  tune::collect_metrics()

autoplot(xgb_tuned)

tune::show_best(xgb_tuned, metric = "rmse")

best_params_xgb <- xgb_tuned |>
  tune::select_best(metric = "rmse")

best_params_xgb

# ------------------------------------------------------------
# 7. Finalize workflow and fit on full training data
# ------------------------------------------------------------

xgb_final_wf <- xgb_wf |>
  tune::finalize_workflow(best_params_xgb)

xgb_final_fit <- xgb_final_wf |>
  tune::last_fit(split = spatial_split)

xgb_final_fit |>
  tune::collect_metrics()

xgb_test_preds <- xgb_final_fit |>
  tune::collect_predictions()

xgb_test_preds

# ------------------------------------------------------------
# 8. Extract fitted xgboost model and run VIP
# ------------------------------------------------------------

xgb_fit <- xgb_final_fit$.workflow[[1]] |>
  workflows::extract_fit_parsnip()

saveRDS(xgb_fit, here(dir_models, "xgb_small.rds"))

toc()


vip_xgb <- vip::vip(
  xgb_fit,
  num_features = 40
)

vip_xgb

# 
# xgb_aet <- pdp::partial(
#   object   = xgb_fit,
#   pred.var = "aet",
#   train    = X_train,
#   grid.resolution = 30
# )
# 
# autoplot(rf_aet)



```

# Compare rf and xgboost
```{r}

xgb_m <- xgb_final_fit |>
  tune::collect_metrics() |>
  mutate(model = "xgb")

rf_m <- rf_final_fit |>
  tune::collect_metrics() |>
  mutate(model = "rf")

all_m <- bind_rows(xgb_m, rf_m)

write_csv(all_m, here(dir_figs, "all_metrics.csv"))

ggplot(data = all_m) +
  geom_col(aes(x = model, fill = .metric, y = .estimate), position = 'dodge') +
  labs(title = "XGB and RF performance on spatially-blocked test: RMSE and R^2",
       y = "Estimate", x = "Test",
       caption = "Full dataset") +
  theme_minimal()

ggsave(here(dir_figs, "full_model_performance.png"))

# tuned hyperparameters

best_params_xgb
best_params


# model predictions
all_preds <- bind_rows(xgb_test_preds |> mutate(model = "xgb"),
                       rf_test_preds |> mutate(model = "rf")) |>
  mutate(residual = post_fire_rap_tree_slope_3_15_yrs - .pred)

ggplot(data = all_preds, aes(x = post_fire_rap_tree_slope_3_15_yrs, y = .pred)) +
  geom_point(aes(color = model), alpha = 0.2) +
  geom_smooth(aes(color = model)) +
  geom_abline(intercept = 0, slope = 1) +
  theme_minimal()

ggsave(here(dir_figs, "preds_vs_obs.png"))

ggplot(data = all_preds, aes(x = post_fire_rap_tree_slope_3_15_yrs, y = residual)) +
  geom_point(aes(color = model), alpha = 0.2) +
  geom_smooth(aes(color = model)) +
  geom_abline(intercept = 0, slope = 1) +
  theme_minimal()

ggplot(data = all_preds, aes(x = residual)) +
  geom_density(aes(color = model)) +
  theme_minimal()


# Diagnose what we are seeing in the residuals
diag_summary <- all_preds |>
  group_by(model) |>
  summarize(
    n          = n(),
    mean_obs   = mean(post_fire_rap_tree_slope_3_15_yrs, na.rm = TRUE),
    sd_obs     = sd(post_fire_rap_tree_slope_3_15_yrs, na.rm = TRUE),
    mean_pred  = mean(.pred, na.rm = TRUE),
    sd_pred    = sd(.pred, na.rm = TRUE),
    mean_resid = mean(residual, na.rm = TRUE),
    sd_resid   = sd(residual, na.rm = TRUE),
    cor_y_pred = cor(post_fire_rap_tree_slope_3_15_yrs, .pred,
                     use = "complete.obs"),
    cor_y_resid = cor(post_fire_rap_tree_slope_3_15_yrs, residual,
                      use = "complete.obs"),
    cor_pred_resid = cor(.pred, residual,
                         use = "complete.obs"),
    .groups = "drop"
  )

diag_summary


# compare variable importance
vip_rf +
  labs(title = "RF") +
vip_xgb +
  labs(title = "XGB")


# plot using rank-vs-rank scatterplot

# Extract importance data from vip objects
rf_vi <- vip_rf$data |>
  as_tibble() |>
  transmute(
    variable   = Variable,
    importance = Importance,
    model      = "rf"
  )

xgb_vi <- vip_xgb$data |>
  as_tibble() |>
  transmute(
    variable   = Variable,
    importance = Importance,
    model      = "xgb"
  )

# Combine and compute ranks within each model (1 = most important)
vi_long <- bind_rows(rf_vi, xgb_vi) |>
  group_by(model) |>
  mutate(rank = min_rank(desc(importance))) |>
  ungroup()

# Wide format: one row per variable, ranks from both models
vi_wide <- vi_long |>
  select(variable, model, rank) |>
  pivot_wider(names_from = model, values_from = rank)


# Rank-vs-rank comparison plot
ggplot(vi_wide_plot, aes(x = rf, y = xgb, label = variable)) +
  geom_point() +
  ggrepel::geom_text_repel(size = 3) +
  scale_x_reverse() +
  scale_y_reverse() +
  labs(
    x = "RF importance rank (1 = most important)",
    y = "XGB importance rank (1 = most important)",
    title = "Comparison of variable importance ranks: RF vs XGBoost"
  ) +
  theme_minimal() +
  geom_abline(intersect = 0, slope = 1, color = "gray")

ggsave(here(dir_figs, "pfi_rank_rank.png"))

```



# Set up for robust post-hoc analysis

```{r}

plot_pdp_ice_ale <- function(fit, train, pred_var, fig_dir,
                             grid_resolution = 30,
                             ice_alpha = 0.1,
                             width = 6, height = 9, dpi = 300) {
  # ---------------------------------------------------------------------------
  # 0. Basic checks & setup
  # ---------------------------------------------------------------------------
  if (!inherits(fit, "model_fit")) {
    stop("`fit` should be a parsnip model_fit (e.g., from extract_fit_parsnip()).")
  }
  
  if (!pred_var %in% names(train)) {
    stop("pred_var '", pred_var, "' not found in 'train'.")
  }
  
  if (!dir.exists(fig_dir)) {
    dir.create(fig_dir, recursive = TRUE)
  }
  
  train_df <- as.data.frame(train)
  
  # ---------------------------------------------------------------------------
  # 1. PDP and ICE via pdp using model_fit directly
  #    (this is exactly what you did manually)
  # ---------------------------------------------------------------------------
  pd_pdp <- pdp::partial(
    object   = fit,
    pred.var = pred_var,
    train    = train_df,
    grid.resolution = grid_resolution,
  )
  
  pd_ice <- pdp::partial(
    object   = fit,
    pred.var = pred_var,
    train    = train_df,
    grid.resolution = grid_resolution,
    ice      = TRUE,
  )
  
  # Figure out PDP column names
  # Typically: pred_var + "yhat"
  if (!all(c(pred_var, "yhat") %in% names(pd_pdp))) {
    stop("PDP object does not contain expected columns '", pred_var, "' and 'yhat'.")
  }
  
  # ICE: need pred_var, yhat, and an ID column
  if (!all(c(pred_var, "yhat") %in% names(pd_ice))) {
    stop("ICE object does not contain expected columns '", pred_var, "' and 'yhat'.")
  }
  
  id_col <- NULL
  if (".id" %in% names(pd_ice)) {
    id_col <- ".id"
  } else if ("yhat.id" %in% names(pd_ice)) {
    id_col <- "yhat.id"
  } else {
    pd_ice$.ice_id <- seq_len(nrow(pd_ice))
    id_col <- ".ice_id"
  }
  
  # ---------------------------------------------------------------------------
  # 2. ALE via iml with a parsnip-friendly predict.fun
  # ---------------------------------------------------------------------------
  pred_fun_iml <- function(object, newdata) {
    preds <- predict(object, new_data = newdata)
    if (".pred" %in% names(preds)) {
      preds$.pred
    } else {
      # fallback: first numeric column
      num_cols <- vapply(preds, is.numeric, logical(1))
      if (!any(num_cols)) stop("predict() output has no numeric columns.")
      preds[[which(num_cols)[1]]]
    }
  }
  
  predictor_iml <- iml::Predictor$new(
    model = fit,
    data  = train_df,
    predict.fun = pred_fun_iml
  )
  
  eff_ale <- iml::FeatureEffect$new(
    predictor = predictor_iml,
    feature   = pred_var,
    method    = "ale"
  )
  
  ale_df <- eff_ale$results
  
  if (!pred_var %in% names(ale_df)) {
    x_col <- names(ale_df)[1]
  } else {
    x_col <- pred_var
  }
  y_cols <- setdiff(names(ale_df), x_col)
  if (length(y_cols) < 1) {
    stop("ALE results do not contain an effect column.")
  }
  y_col <- y_cols[1]
  
  names(ale_df)[names(ale_df) == x_col] <- "x"
  names(ale_df)[names(ale_df) == y_col] <- "ale"
  
  # ---------------------------------------------------------------------------
  # 3. Build the three ggplots: PDP, ICE, ALE
  # ---------------------------------------------------------------------------
  
  p_pdp <- ggplot(pd_pdp, aes_string(x = pred_var, y = "yhat")) +
    geom_line() +
    labs(
      title = paste0("Partial Dependence (PDP): ", pred_var),
      x = pred_var,
      y = "Predicted outcome (average)"
    ) +
    theme_minimal()
  
  p_ice <- ggplot(
    pd_ice,
    aes_string(x = pred_var, y = "yhat", group = id_col)
  ) +
    geom_line(alpha = ice_alpha) +
    labs(
      title = paste0("Individual Conditional Expectation (ICE): ", pred_var),
      x = pred_var,
      y = "Predicted outcome"
    ) +
    theme_minimal()
  
  vals <- round(range(ale_df$.value), 2)
  p_ale <- ggplot(ale_df, aes(x = x, y = .value)) +
    geom_line() +
    labs(
      title = paste0("Accumulated Local Effects (ALE): ", pred_var),
      x = pred_var,
      y = "ALE",
      caption = glue("ALE Range: {vals[1]}, {vals[2]}")
    ) +
    theme_minimal()
  
  combined_plot <- p_pdp / p_ice / p_ale
  
  # ---------------------------------------------------------------------------
  # 4. Save PNG
  # ---------------------------------------------------------------------------
  fit_label <- class(fit)[1]
  fit_label <- gsub("[^A-Za-z0-9_]+", "_", fit_label)
  pred_label <- gsub("[^A-Za-z0-9_]+", "_", pred_var)
  
  file_name <- paste0("pdp_ice_ale_", fit_label, "_", pred_label, ".png")
  out_path  <- file.path(fig_dir, file_name)
  
  ggsave(
    filename = out_path,
    plot     = combined_plot,
    width    = width,
    height   = height,
    dpi      = dpi
  )
  
  message("Saved: ", out_path)
  return(list("ale_df" = ale_df))
}


# Top 3 VI

plot_pdp_ice_ale(
  fit      = rf_fit, 
  train    = train_data |> st_drop_geometry(),
  pred_var = "srtm",
  fig_dir  = dir_figs
)

plot_pdp_ice_ale(
  fit      = rf_fit, 
  train    = train_data |> st_drop_geometry(),
  pred_var = "pdsi_annual_5_yrs_prior_mean",
  fig_dir  = dir_figs
)

plot_pdp_ice_ale(
  fit      = rf_fit, 
  train    = train_data |> st_drop_geometry(),
  pred_var = "burn_cbi",
  fig_dir  = dir_figs
)

# 2D PDPs
train_df <- as.data.frame(train_data |> st_drop_geometry())

pdp_2d <- pdp::partial(
  object          = rf_fit,
  pred.var        = c("srtm", "burn_cbi"),
  train           = train_df,
  grid.resolution = 20
)


autoplot(pdp_2d) +
  labs(title = "2D PDP for burn severity and elevation")
ggsave(here(dir_figs, "pdp2d_cbi_srtm.png"))

```

<!-- # Global interactions (NOT INCLUDED IN PROJECT) -->

<!-- ```{r} -->

<!-- set.seed(1) -->
<!-- train_iml <- train_data |> -->
<!--   as.data.frame() |> -->
<!--   sample_n(2000) |> -->
<!--   st_drop_geometry() -->

<!-- # pull out the bare ranger model -->
<!-- rf_engine <- rf_fit$fit -->

<!-- # prediction function for ranger -->
<!-- pred_fun_rf <- function(object, newdata) { -->
<!--   predict(object, data = newdata)$predictions -->
<!-- } -->

<!-- # iml Predictor -->
<!-- predictor_rf <- Predictor$new( -->
<!--   model       = rf_engine, -->
<!--   data        = train_iml |> select(-post_fire_rap_tree_slope_3_15_yrs), -->
<!--   y           = train_iml$post_fire_rap_tree_slope_3_15_yrs, -->
<!--   predict.fun = pred_fun_rf -->
<!-- ) -->

<!-- # global interaction strengths (Friedman's H for each feature) -->
<!-- rf_interactions <- Interaction$new(predictor_rf) -->

<!-- rf_i_r <- rf_interactions$results -->
<!-- write_csv(rf_i_r, here(dir_figs, "rf_global_interactions.csv")) -->

<!-- ``` -->


# Variable interpretation reminder
- TPI: low = valleys, high = ridges
- CHILI: low = cold, high = hot
- SRTM: low = low elevation, high = high elevation
- AET: low = wet, high = dry
- DEF: low = wet, high = dry
- PDSI: low = dry, high = wet



# MIR
```{r}
compute_mir_groups <- function(
  model_spec,
  data_split,
  outcome,
  predictors,
  groups,
  metrics    = yardstick::metric_set(yardstick::rmse),
  recipe_fun = NULL
) {
  # Split data
  train_data <- rsample::training(data_split)
  test_data  <- rsample::testing(data_split)
  
  # Basic checks
  if (!outcome %in% names(train_data)) {
    stop("Outcome '", outcome, "' not found in training data.")
  }
  if (!all(predictors %in% names(train_data))) {
    missing <- setdiff(predictors, names(train_data))
    stop(
      "Some predictors not found in training data: ",
      paste(missing, collapse = ", ")
    )
  }
  
  # Helper: build a recipe for a given predictor set
  build_recipe <- function(preds, data) {
    form <- reformulate(termlabels = preds, response = outcome)
    
    if (is.null(recipe_fun)) {
      recipes::recipe(form, data = data) |>
        recipes::step_zv(recipes::all_predictors())
    } else {
      recipe_fun(form, data)
    }
  }
  
  # ------------------------
  # Baseline model (all predictors)
  # ------------------------
  base_recipe <- build_recipe(predictors, train_data)
  
  base_wf <- workflows::workflow() |>
    workflows::add_model(model_spec) |>
    workflows::add_recipe(base_recipe)
  
  base_fit <- base_wf |>
    fit(data = train_data)
  
  base_pred <- predict(base_fit, new_data = test_data) |>
    dplyr::bind_cols(
      truth = test_data[[outcome]]
    )
  
  base_metrics <- metrics(
    data    = base_pred,
    truth   = truth,
    estimate = .pred
  ) |>
    dplyr::mutate(
      group              = "baseline",
      removed_predictors = NA_character_
    )
  
  # ------------------------
  # Function to evaluate one group
  # ------------------------
  # IMPORTANT: first argument = group vars (.x), second = group name (.y)
  eval_group <- function(g_vars, g_name) {
    preds_reduced <- setdiff(predictors, g_vars)
    
    rec_reduced <- build_recipe(preds_reduced, train_data)
    
    wf_reduced <- workflows::workflow() |>
      workflows::add_model(model_spec) |>
      workflows::add_recipe(rec_reduced)
    
    fit_reduced <- wf_reduced |>
      fit(data = train_data)
    
    pred_reduced <- predict(fit_reduced, new_data = test_data) |>
      dplyr::bind_cols(
        truth = test_data[[outcome]]
      )
    
    met_reduced <- metrics(
      data    = pred_reduced,
      truth   = truth,
      estimate = .pred
    ) |>
      dplyr::mutate(
        group              = g_name,
        removed_predictors = paste(g_vars, collapse = ", ")
      )
    
    met_reduced
  }
  
  # Apply over groups (named list)
  group_results <- purrr::imap_dfr(groups, eval_group)
  
  # ------------------------
  # Combine + compute MIR
  # ------------------------
  combined <- dplyr::bind_rows(base_metrics, group_results)
  
  # Infer metric direction from name (no yardstick::direction)
  combined <- combined |>
    dplyr::mutate(
      direction = dplyr::case_when(
        .metric %in% c("rmse", "mae", "mape", "mse") ~ "minimize",
        .metric %in% c("rsq", "rsq_trad", "rsq_adj") ~ "maximize",
        TRUE ~ "minimize"   # default; adjust if you add others
      )
    )
  
  combined |>
    dplyr::left_join(
      combined |>
        dplyr::filter(group == "baseline") |>
        dplyr::select(.metric, base = .estimate),
      by = ".metric"
    ) |>
    dplyr::mutate(
      delta = dplyr::case_when(
        direction == "minimize" ~ .estimate - base,  # worse = larger value
        direction == "maximize" ~ base - .estimate,  # worse = smaller value
        TRUE ~ NA_real_
      ),
      mir = dplyr::case_when(
        base != 0 ~ delta / abs(base),
        TRUE ~ NA_real_
      )
    ) |>
    dplyr::select(
      group,
      removed_predictors,
      .metric,
      direction,
      base,
      dropped = .estimate,
      delta,
      mir
    ) |>
    dplyr::arrange(.metric, dplyr::desc(mir))
}




rf_spec_final <- rf_final_wf |>
  workflows::extract_spec_parsnip()

# Outcome and predictors based on your existing variables
outcome    <- target       # e.g. "post_fire_rap_TRE_difference_n2_3_yr"
predictors <- covars       # character vector of predictor column names

mir_groups <- list(
  climate_before = c(
    "pdsi_annual_5_yrs_prior_mean",
    "pdsi_summer_5_yrs_prior_mean"
  ),
  
  climate_after = c(
    "pdsi_annual_5_yrs_after_mean",
    "pdsi_summer_5_yrs_after_mean"
  ),
  
  topo = c(
    "srtm",
    "tpi",
    "chili"
  ),
  
  prefire_tree = c(
    "avg_rap_tree_6_yr_pre_fire"
  ),
  
  burn_severity = c(
    "burn_cbi",
    "burn_cbi_500mmwmean",
    "burn_cbi_500mmwstd",
    "burn_distance_forest_after"
  ),
  
  climate_normals = c(
    "tt_change_aet",
    "tt_change_def",
    "tt_normal_aet",
    "tt_normal_def"
  ),
  
  forest_group = c(
    "ENC1_nfg_factor",
    "ENC2_nfg_factor"
  ),
  
  regime_group = c(
    "ENC1_frg_reclass_desc",
    "ENC2_frg_reclass_desc"
  ),
  
  year = c(
    "mock_burn_year"
  )
)

mir_results <- compute_mir_groups(
  model_spec  = rf_spec_final,
  data_split  = spatial_split,
  outcome     = outcome,
  predictors  = predictors,
  groups      = mir_groups,
  metrics     = yardstick::metric_set(yardstick::rmse, yardstick::rsq)
)

mir_results
write_csv(mir_results, here(dir_figs, "mir_results.csv"))

ggplot(mir_results) +
  geom_col(aes(x = mir, y = reorder(group, mir))) +
  facet_grid(~.metric) +
  labs(title = "Leave-one-feature (group)-out-importance")

ggsave(here(dir_figs, "mir_results.png"))


```




#################
# From run with all PDSI data, the cumulative drought sums appear to be more important than the means, which are more important than the thresholds

```{r}

narrower_covars <- all_covars[!str_ends(all_covars, "mean") &
                                !str_ends(all_covars, "threshold") &
                                !str_starts(all_covars, "pdsi_summer")]


form_str <- paste(
  "post_fire_rap_TRE_slope_2_10_yrs ~ burn_sev + ",
  paste(narrower_covars, collapse = " + ")
)

form <- as.formula(form_str)

tic()
test_spmlmRF <- splmRF(form,
                  data = df_test,
                  spcov_type = "exponential",
                  importance = 'permutation',
                  scale.permutation.importance = TRUE,
                  num.threads = 0,
                  mtry = floor(sqrt(ncol(dats) - 1)),
                  num.trees = 1000,
                  seed = seed)
toc()
summary(test_spmlmRF)
print(test_spmlmRF$ranger)
print(test_spmlmRF$call)

#500 trees mtry 3, MSE 213, R2 0.49
#1000 trees mtry 8, basically same
#1000t, mtry8, include disturbance MSE 2.17, R^2 0.14


vip <- vip(test_spmlmRF$ranger,
    num_features = 25)
vip_tbl <- vip(test_spmlmRF$ranger, num_features = 25)$data |>
  as_tibble() |>
  arrange(desc(Importance)) |>   # highest first
  mutate(variable = factor(Variable, levels = Variable))

ggsave(plot = vip, filename = here(dir_rf_figs, "vip_idbath_v1.png"), units = "px", width = 3000, height = 4000, )


pdp_cont <- c(narrower_covars) |>
  set_names() |>
  map_dfr(\(v) {
    # run PDP
    pd <- partial(
      object   = test_spmlmRF$ranger,
      pred.var = v,
      train    = df_test |> st_drop_geometry()
    )

    # rename predictor column to a common name 'x'
    pd |>
      rename(x = !!sym(v)) |>
      mutate(variable = v, .before = x)
  })

pdp_sorted <- pdp_cont |>
  inner_join(
    vip_tbl |> select(variable = Variable),
    by = "variable"
  ) |>
  mutate(variable = factor(variable, levels = vip_tbl$Variable))

plot <- pdp_sorted |>
  ggplot(aes(x = x, y = yhat)) +
  geom_line() +
  facet_wrap(~ variable, scales = "free_x") +
  labs(
    x = "Predictor value",
    y = "Partial dependence (yhat)",
    title = "Partial dependence for continuous covariates"
  )

ggsave(plot = plot, filename = here(dir_rf_figs, "pdp_idbath_v1.png"), units = "px", width = 3000, height = 4000, )


```




<button class="accordion-button" type="button" data-bs-toggle="collapse" data-bs-target="#collapseOne" >Click to see Session Information</button><div id="collapseOne" class="accordion-collapse collapse"><div>

```{r sessionInfo}
sessionInfo()
```