---
title: "06_dydid"
author: Tyler L. McIntosh
date: last-modified
format:
  html:
    embed-resources: true
    html-math-method: katex
    css: styles.css
    toc: true
    toc-depth: 1
    toc-title: Contents
    toc-location: right
    code-fold: true
    code-summary: "Show the code"
---


# Setup

```{r setupchunk}
#| echo: true
#| message: false
#| warning: false
#| results: 'hide'

rm(list = ls())

if (!requireNamespace("here", quietly = TRUE)) install.packages("here")
library(here)

source(here::here("utils", "functions.R"))

install_load_packages(c(
  "tidyverse",
  "glue",
  "arrow",
  "here"
))

# ── Directories ───────────────────────────────────────────────────────────────
# Point these at the same dir_results used in run_analysis.R.

run_name    <- "GEE_resilience_v6_operational_ss500_ts50000"
version     <- "v6"
dir_results <- here::here("results", version)   # shared dir_out from run_analysis.R
dir_figs    <- here::here("figs", run_name, version)

dir_ensure(c(dir_results, dir_figs))

seed <- 1234
set.seed(seed)
```


# Load results tables

```{r load-tables}
dir_all <- file.path(dir_results, "tables", "all")

coef_tbl <- arrow::read_parquet(
  file.path(dir_all, "coef_expanded_vcov.parquet")
)

subgroup_tbl <- arrow::read_parquet(
  file.path(dir_all, "subgroup_run_summary.parquet")
)

support_tbl <- arrow::read_parquet(
  file.path(dir_all, "event_time_support.parquet")
)

registry_tbl <- arrow::read_parquet(
  file.path(dir_all, "run_registry.parquet")
)

spec_index_tbl <- arrow::read_parquet(
  file.path(dir_all, "run_spec_index.parquet")
)

# ── Quick sanity checks ───────────────────────────────────────────────────────

message("coef_tbl columns:     ", paste(names(coef_tbl), collapse = ", "))
message("subgroup_tbl columns: ", paste(names(subgroup_tbl), collapse = ", "))
message("support_tbl columns:  ", paste(names(support_tbl), collapse = ", "))

coef_tbl |> dplyr::count(vcov_id)
coef_tbl |> dplyr::count(subset_id, group_id, subgroup)
support_tbl |> dplyr::arrange(group_id, subgroup, event_time)
```


# Plotting functions

```{r plotting-helpers}

# ── Null coalescing (also exported by portable_sunab.R; safe to redefine) ─────
`%||%` <- function(x, y) if (!is.null(x)) x else y


# ── apply_saved_filters ───────────────────────────────────────────────────────
#
# Filter a data frame by a named list of allowed values.
# Each name must be a column in df; each value is a vector of allowed values,
# NULL (skip), or NA (keep only NA rows).
#
apply_saved_filters <- function(df, filters = list()) {
  if (length(filters) == 0) return(df)

  for (nm in names(filters)) {
    if (!nm %in% names(df)) {
      stop("Filter column not found in data: ", nm)
    }

    vals <- filters[[nm]]
    if (is.null(vals)) next

    if (length(vals) == 1 && is.na(vals)) {
      df <- df |> dplyr::filter(is.na(.data[[nm]]))
    } else {
      df <- df |> dplyr::filter(.data[[nm]] %in% vals)
    }
  }

  df
}


# ── format_n_label ────────────────────────────────────────────────────────────
#
# Format a numeric vector as "prefix=N" or "prefix=min–max" for legend labels.
#
format_n_label <- function(x, prefix = "n") {
  x <- unique(stats::na.omit(x))
  if (length(x) == 0) return(paste0(prefix, "=NA"))
  if (length(x) == 1) return(paste0(prefix, "=", format(x, big.mark = ",")))
  paste0(
    prefix, "=",
    format(min(x), big.mark = ","), "\u2013",
    format(max(x), big.mark = ",")
  )
}


# ── make_legend_labels_with_n ─────────────────────────────────────────────────
#
# Build a named character vector of legend labels that include sample-size
# annotations, keyed by the unique values of `var`.
#
# Column names updated to match new pipeline output:
#   n_pt_col    -> "n_treated_units"   (was "n_treated_units" — unchanged)
#   n_event_col -> "n_treated_events"  (was "n_treated_events" — unchanged)
#
make_legend_labels_with_n <- function(df,
                                       var,
                                       n_pt_col    = "n_treated_units",
                                       n_event_col = "n_treated_events") {
  if (is.null(var)) return(NULL)

  for (col in c(var, n_pt_col, n_event_col)) {
    if (!col %in% names(df)) stop("Column not found in data: ", col)
  }

  df |>
    dplyr::distinct(
      .data[[var]],
      subgroup_run_id,
      .data[[n_pt_col]],
      .data[[n_event_col]]
    ) |>
    dplyr::group_by(.data[[var]]) |>
    dplyr::summarise(
      n_pt_lab    = format_n_label(.data[[n_pt_col]],    prefix = "pts"),
      n_event_lab = format_n_label(.data[[n_event_col]], prefix = "events"),
      .groups = "drop"
    ) |>
    dplyr::mutate(
      label = paste0(.data[[var]], " (", n_pt_lab, ", ", n_event_lab, ")")
    ) |>
    (\(tbl) stats::setNames(tbl$label, tbl[[var]]))()
}


# ── build_event_study_caption ─────────────────────────────────────────────────
#
# Build a multi-line plot caption summarising the key estimation choices.
# Renamed from build_sunab_caption; now also shows estimator_type when known.
#
build_event_study_caption <- function(filters           = list(),
                                       vcov_label        = NULL,
                                       estimator_type    = NULL,
                                       show_ci           = TRUE,
                                       ci_level          = 0.95,
                                       n_min_ptids_event      = 0,
                                       n_min_fireids_event    = 0,
                                       n_min_ptids_subgroup   = 0,
                                       n_min_fireids_subgroup = 0) {
  ci_pct <- round(ci_level * 100)

  estimator_line <- if (!is.null(estimator_type)) {
    paste0("Estimator: ", estimator_type)
  } else {
    "Estimator: Sun-Abraham"
  }

  base_lines <- c(
    estimator_line,
    if (show_ci) paste0(ci_pct, "% confidence intervals shown") else
                 paste0(ci_pct, "% confidence intervals not shown"),
    paste0(
      "N limits (event-time): ",
      format(n_min_ptids_event,    big.mark = ","), " pts, ",
      format(n_min_fireids_event,  big.mark = ","), " events"
    ),
    paste0(
      "N limits (subgroup): ",
      format(n_min_ptids_subgroup,   big.mark = ","), " pts, ",
      format(n_min_fireids_subgroup, big.mark = ","), " events"
    )
  )

  if (!is.null(vcov_label)) {
    base_lines <- c(base_lines, paste0("Standard errors: ", vcov_label))
  }

  if (length(filters) > 0) {
    filter_lines <- purrr::imap_chr(filters, \(v, nm) {
      if (is.null(v))                    return(paste0(nm, ": NULL"))
      if (length(v) == 1 && is.na(v))   return(paste0(nm, ": NA"))
      paste0(nm, ": ", paste(as.character(v), collapse = ", "))
    })
    base_lines <- c(base_lines, filter_lines)
  }

  paste(base_lines, collapse = "\n")
}


# ── plot_event_study ──────────────────────────────────────────────────────────
#
# General event-study plot from a coef_tbl produced by portable_sunab.R.
#
# Key column name changes vs. the old pipeline:
#   analysis_id  -> subset_id   (the analysis subset identifier)
#
# All other column names used as aesthetics or filters are unchanged:
#   outcome, group_id, subgroup, model_id, vcov_id, vcov_label,
#   term_value, term_matches_pattern, estimate, ci_lower, ci_upper,
#   n_ptids, n_fireids, n_treated_units, n_treated_events,
#   subgroup_run_id, coef_run_id, estimator_type.
#
plot_event_study <- function(coef_tbl,
                              filters            = list(),
                              color_var,
                              linetype_var       = NULL,
                              facet_var          = NULL,
                              facet_ncol         = NULL,
                              facet_scales       = "fixed",
                              show_ci            = TRUE,
                              ci_style           = c("ribbon", "lines"),
                              title              = NULL,
                              subtitle           = NULL,
                              caption            = NULL,
                              baseline           = -6,
                              xlim               = NULL,
                              ylim               = NULL,
                              color_values       = NULL,
                              linetype_values    = NULL,
                              legend_title_color    = NULL,
                              legend_title_linetype = NULL,
                              show_n             = TRUE,
                              n_legend_var       = NULL,
                              n_min_ptids_event       = 0,
                              n_min_fireids_event     = 0,
                              n_min_ptids_subgroup    = 0,
                              n_min_fireids_subgroup  = 0,
                              drop_na_estimate   = TRUE,
                              drop_baseline_term = FALSE,
                              line_size          = 0.9,
                              point_size         = 1.8,
                              ribbon_alpha       = 0.18,
                              ci_line_alpha      = 0.7,
                              ci_level           = 0.95) {

  ci_style <- match.arg(ci_style)

  # ── Column checks ────────────────────────────────────────────────────────────

  needed_cols <- c(
    "term_value", "estimate", "ci_lower", "ci_upper",
    "subgroup_run_id", "coef_run_id",
    "n_ptids", "n_fireids",
    "n_treated_units", "n_treated_events"
  )
  missing_needed <- setdiff(needed_cols, names(coef_tbl))
  if (length(missing_needed) > 0) {
    stop("coef_tbl is missing required columns: ",
         paste(missing_needed, collapse = ", "))
  }

  aes_vars <- stats::na.omit(c(color_var, linetype_var, facet_var))
  missing_aes <- setdiff(aes_vars, names(coef_tbl))
  if (length(missing_aes) > 0) {
    stop("Aesthetic columns missing from coef_tbl: ",
         paste(missing_aes, collapse = ", "))
  }

  # ── Filter and trim ──────────────────────────────────────────────────────────

  d <- coef_tbl |> apply_saved_filters(filters = filters)

  if ("term_matches_pattern" %in% names(d)) {
    d <- d |> dplyr::filter(.data$term_matches_pattern)
  }

  d <- d |> dplyr::filter(!is.na(.data$term_value))

  if (drop_na_estimate)   d <- d |> dplyr::filter(!is.na(.data$estimate))
  if (drop_baseline_term) d <- d |> dplyr::filter(.data$term_value != baseline)
  if (!is.null(xlim)) {
    d <- d |> dplyr::filter(
      .data$term_value >= xlim[1], .data$term_value <= xlim[2]
    )
  }

  # ── Sample-size filters ──────────────────────────────────────────────────────

  subgroup_keep <- d |>
    dplyr::distinct(subgroup_run_id, n_treated_units, n_treated_events) |>
    dplyr::filter(
      .data$n_treated_units >= n_min_ptids_subgroup,
      is.na(.data$n_treated_events) |
        .data$n_treated_events >= n_min_fireids_subgroup
    ) |>
    dplyr::pull(subgroup_run_id)

  d <- d |>
    dplyr::filter(.data$subgroup_run_id %in% subgroup_keep) |>
    dplyr::filter(
      .data$n_ptids >= n_min_ptids_event,
      is.na(.data$n_fireids) | .data$n_fireids >= n_min_fireids_event
    )

  # ── Build caption ────────────────────────────────────────────────────────────

  estimator_types <- if ("estimator_type" %in% names(d)) {
    unique(stats::na.omit(as.character(d$estimator_type)))
  } else {
    NULL
  }
  estimator_caption <- if (length(estimator_types) == 1) estimator_types else NULL

  vcov_labels <- unique(stats::na.omit(as.character(
    if (nrow(d) > 0) d$vcov_label else coef_tbl$vcov_label
  )))
  vcov_caption <- if (length(vcov_labels) == 1) {
    vcov_labels
  } else if (length(vcov_labels) > 1) {
    paste(vcov_labels, collapse = ", ")
  } else {
    NULL
  }

  caption_auto <- build_event_study_caption(
    filters                = filters,
    vcov_label             = vcov_caption,
    estimator_type         = estimator_caption,
    show_ci                = show_ci,
    ci_level               = ci_level,
    n_min_ptids_event      = n_min_ptids_event,
    n_min_fireids_event    = n_min_fireids_event,
    n_min_ptids_subgroup   = n_min_ptids_subgroup,
    n_min_fireids_subgroup = n_min_fireids_subgroup
  )
  if (!is.null(caption)) caption_auto <- paste(caption_auto, caption, sep = "\n")

  # ── Empty-data fallback ──────────────────────────────────────────────────────

  if (nrow(d) == 0) {
    p <- ggplot2::ggplot() +
      ggplot2::geom_blank() +
      ggplot2::geom_vline(xintercept = 0,        linewidth = 0.6,
                          linetype = "dotted", color = "red") +
      ggplot2::geom_hline(yintercept = 0,        linewidth = 0.4,
                          linetype = "solid",  color = "grey40") +
      ggplot2::labs(title = title, subtitle = subtitle,
                    caption = caption_auto, x = "Event time", y = "Effect") +
      ggplot2::theme_minimal(base_size = 12) +
      ggplot2::theme(panel.grid.minor = ggplot2::element_blank(),
                     legend.position = "none")
    if (!is.null(xlim) || !is.null(ylim)) {
      p <- p + ggplot2::coord_cartesian(xlim = xlim, ylim = ylim)
    }
    return(p)
  }

  # ── Y-axis label ─────────────────────────────────────────────────────────────

  outcomes <- unique(stats::na.omit(as.character(d$outcome)))
  y_label  <- if (length(outcomes) == 1) paste0("Effect on ", outcomes) else "Effect"

  # ── Legend labels (with optional N annotation) ───────────────────────────────

  if (is.null(n_legend_var)) n_legend_var <- color_var %||% linetype_var

  color_labels         <- NULL
  linetype_labels      <- NULL
  color_values_plot    <- NULL
  linetype_values_plot <- NULL
  color_breaks         <- NULL
  linetype_breaks      <- NULL

  if (!is.null(color_var)) {
    color_labels <- if (show_n && identical(n_legend_var, color_var)) {
      make_legend_labels_with_n(d, color_var)
    } else {
      vals <- d |>
        dplyr::distinct(.data[[color_var]]) |>
        dplyr::pull(.data[[color_var]]) |>
        as.character()
      stats::setNames(vals, vals)
    }

    d <- d |>
      dplyr::mutate(
        color_plot = dplyr::recode(as.character(.data[[color_var]]), !!!color_labels)
      )
    color_breaks <- unname(color_labels)

    if (!is.null(color_values)) {
      if (is.null(names(color_values))) stop("color_values must be a named vector.")
      missing_nms <- setdiff(names(color_labels), names(color_values))
      if (length(missing_nms) > 0) {
        stop("color_values is missing names for: ", paste(missing_nms, collapse = ", "))
      }
      color_values_plot <- stats::setNames(
        color_values[names(color_labels)],
        unname(color_labels)
      )
    }
  }

  if (!is.null(linetype_var)) {
    linetype_labels <- if (show_n && identical(n_legend_var, linetype_var)) {
      make_legend_labels_with_n(d, linetype_var)
    } else {
      vals <- d |>
        dplyr::distinct(.data[[linetype_var]]) |>
        dplyr::pull(.data[[linetype_var]]) |>
        as.character()
      stats::setNames(vals, vals)
    }

    d <- d |>
      dplyr::mutate(
        linetype_plot = dplyr::recode(as.character(.data[[linetype_var]]), !!!linetype_labels)
      )
    linetype_breaks <- unname(linetype_labels)

    if (!is.null(linetype_values)) {
      if (is.null(names(linetype_values))) stop("linetype_values must be a named vector.")
      missing_nms <- setdiff(names(linetype_labels), names(linetype_values))
      if (length(missing_nms) > 0) {
        stop("linetype_values is missing names for: ", paste(missing_nms, collapse = ", "))
      }
      linetype_values_plot <- stats::setNames(
        linetype_values[names(linetype_labels)],
        unname(linetype_labels)
      )
    }
  }

  # ── group aesthetic ───────────────────────────────────────────────────────────
  # Build a plain character group ID rather than using interaction(), which
  # produces a factor whose levels may not align across rows after filtering,
  # causing geom_line to silently draw nothing.

  d <- d |>
    dplyr::arrange(.data$term_value) |>
    dplyr::mutate(
      plot_group = paste(.data$subgroup_run_id, .data$vcov_id, sep = "__")
    )

  # ── Base plot ─────────────────────────────────────────────────────────────────

  p <- ggplot2::ggplot(
    d,
    ggplot2::aes(x = .data$term_value, y = .data$estimate, group = .data$plot_group)
  )

  # ── CI ribbon ────────────────────────────────────────────────────────────────
  # geom_ribbon fill is added as a separate aes() call per branch so that NULL
  # is never passed inside aes() — that silently breaks rendering.

  if (show_ci && ci_style == "ribbon") {
    if (!is.null(color_var)) {
      p <- p + ggplot2::geom_ribbon(
        ggplot2::aes(ymin = .data$ci_lower, ymax = .data$ci_upper,
                     fill = .data$color_plot),
        alpha = ribbon_alpha, linewidth = 0, show.legend = FALSE
      )
    } else {
      p <- p + ggplot2::geom_ribbon(
        ggplot2::aes(ymin = .data$ci_lower, ymax = .data$ci_upper),
        alpha = ribbon_alpha, linewidth = 0, show.legend = FALSE
      )
    }
  }

  # ── CI lines ─────────────────────────────────────────────────────────────────

  if (show_ci && ci_style == "lines") {
    if (!is.null(color_var) && !is.null(linetype_var)) {
      p <- p +
        ggplot2::geom_line(
          ggplot2::aes(y = .data$ci_lower,
                       color = .data$color_plot, linetype = .data$linetype_plot),
          alpha = ci_line_alpha, linewidth = 0.45
        ) +
        ggplot2::geom_line(
          ggplot2::aes(y = .data$ci_upper,
                       color = .data$color_plot, linetype = .data$linetype_plot),
          alpha = ci_line_alpha, linewidth = 0.45
        )
    } else if (!is.null(color_var)) {
      p <- p +
        ggplot2::geom_line(
          ggplot2::aes(y = .data$ci_lower, color = .data$color_plot),
          alpha = ci_line_alpha, linewidth = 0.45
        ) +
        ggplot2::geom_line(
          ggplot2::aes(y = .data$ci_upper, color = .data$color_plot),
          alpha = ci_line_alpha, linewidth = 0.45
        )
    } else if (!is.null(linetype_var)) {
      p <- p +
        ggplot2::geom_line(
          ggplot2::aes(y = .data$ci_lower, linetype = .data$linetype_plot),
          alpha = ci_line_alpha, linewidth = 0.45
        ) +
        ggplot2::geom_line(
          ggplot2::aes(y = .data$ci_upper, linetype = .data$linetype_plot),
          alpha = ci_line_alpha, linewidth = 0.45
        )
    } else {
      p <- p +
        ggplot2::geom_line(
          ggplot2::aes(y = .data$ci_lower),
          alpha = ci_line_alpha, linewidth = 0.45
        ) +
        ggplot2::geom_line(
          ggplot2::aes(y = .data$ci_upper),
          alpha = ci_line_alpha, linewidth = 0.45
        )
    }
  }

  # ── Estimate lines + points ───────────────────────────────────────────────────

  if (!is.null(color_var) && !is.null(linetype_var)) {
    p <- p +
      ggplot2::geom_line(
        ggplot2::aes(color = .data$color_plot, linetype = .data$linetype_plot),
        linewidth = line_size
      ) +
      ggplot2::geom_point(
        ggplot2::aes(color = .data$color_plot),
        size = point_size
      )
  } else if (!is.null(color_var)) {
    p <- p +
      ggplot2::geom_line(
        ggplot2::aes(color = .data$color_plot),
        linewidth = line_size
      ) +
      ggplot2::geom_point(
        ggplot2::aes(color = .data$color_plot),
        size = point_size
      )
  } else if (!is.null(linetype_var)) {
    p <- p +
      ggplot2::geom_line(
        ggplot2::aes(linetype = .data$linetype_plot),
        linewidth = line_size
      ) +
      ggplot2::geom_point(size = point_size)
  } else {
    p <- p +
      ggplot2::geom_line(linewidth = line_size) +
      ggplot2::geom_point(size = point_size)
  }

  # ── Reference lines ───────────────────────────────────────────────────────────

  p <- p +
    ggplot2::geom_hline(yintercept = 0,        linewidth = 0.4,
                        linetype = "solid",  color = "grey40") +
    ggplot2::geom_vline(xintercept = baseline, linewidth = 0.4,
                        linetype = "dashed", color = "grey40") +
    ggplot2::geom_vline(xintercept = 0,        linewidth = 0.6,
                        linetype = "dotted", color = "red")

  # ── Facets ────────────────────────────────────────────────────────────────────

  if (!is.null(facet_var)) {
    p <- p + ggplot2::facet_wrap(
      stats::as.formula(paste("~", facet_var)),
      ncol   = facet_ncol,
      scales = facet_scales
    )
  }

  # ── Scales ────────────────────────────────────────────────────────────────────

  if (!is.null(color_var)) {
    if (!is.null(color_values_plot)) {
      p <- p +
        ggplot2::scale_color_manual(
          values = color_values_plot,
          breaks = color_breaks,
          name   = legend_title_color %||% color_var
        )
      if (show_ci && ci_style == "ribbon") {
        p <- p +
          ggplot2::scale_fill_manual(
            values = color_values_plot,
            breaks = color_breaks,
            guide  = "none"
          )
      }
    } else {
      p <- p + ggplot2::labs(color = legend_title_color %||% color_var)
    }
  }

  if (!is.null(linetype_var)) {
    if (!is.null(linetype_values_plot)) {
      p <- p +
        ggplot2::scale_linetype_manual(
          values = linetype_values_plot,
          breaks = linetype_breaks,
          name   = legend_title_linetype %||% linetype_var
        )
    } else {
      p <- p + ggplot2::labs(linetype = legend_title_linetype %||% linetype_var)
    }
  }

  # ── Labels and theme ──────────────────────────────────────────────────────────

  p <- p +
    ggplot2::labs(
      title    = title,
      subtitle = subtitle,
      caption  = caption_auto,
      x        = "Event time",
      y        = y_label
    ) +
    ggplot2::theme_minimal(base_size = 12) +
    ggplot2::theme(
      panel.grid.minor = ggplot2::element_blank(),
      legend.position  = "right",
      strip.text       = ggplot2::element_text(face = "bold")
    )

  # ── Coord cartesian (only added when needed) ──────────────────────────────────
  # Never add coord_cartesian conditionally with + if(...) — when the condition
  # is FALSE that evaluates to p + FALSE which corrupts the plot object.

  if (!is.null(xlim) || !is.null(ylim)) {
    p <- p + ggplot2::coord_cartesian(xlim = xlim, ylim = ylim)
  }

  p
}
```


# Palette helpers

```{r palette-helpers}

# ── make_named_palette ────────────────────────────────────────────────────────
#
# Build a named colour vector from a set of level values.
#
make_named_palette <- function(levels, palette = "Okabe-Ito") {
  levels <- unique(as.character(levels))
  vals   <- grDevices::palette.colors(length(levels), palette = palette)
  stats::setNames(vals, levels)
}


# ── make_named_palette_from_tbl ───────────────────────────────────────────────
#
# Build a named colour vector by pulling unique values from a table column.
# Updated to use `subset_id` as the default column (was `analysis_id`).
#
make_named_palette_from_tbl <- function(tbl,
                                         var,
                                         palette     = "Okabe-Ito",
                                         levels      = NULL,
                                         sort_levels = TRUE) {
  if (!var %in% names(tbl)) stop("Column not found in table: ", var)

  if (is.null(levels)) {
    levels <- tbl |>
      dplyr::distinct(.data[[var]]) |>
      dplyr::pull(.data[[var]]) |>
      as.character()
    if (sort_levels) levels <- sort(levels)
  } else {
    levels <- as.character(levels)
  }

  if (length(levels) == 0) stop("No levels found for palette.")

  vals <- grDevices::palette.colors(n = length(levels), palette = palette)
  stats::setNames(vals, levels)
}
```


# Plot parameters

```{r plot-params}

grDevices::palette.pals()   # list available palettes

# ── Subgroup palette ──────────────────────────────────────────────────────────
subgroup_pal <- c(
  f   = "red3",
  bf  = "forestgreen",
  df  = "goldenrod2",
  bdf = "purple"
)

# ── Subset (ecoregion / analysis partition) palette ───────────────────────────
# Column is now `subset_id` instead of `analysis_id`.
subset_pal <- make_named_palette_from_tbl(
  coef_tbl,
  var     = "subset_id",
  palette = "Polychrome 36"
)

# ── Global N thresholds ───────────────────────────────────────────────────────
n_min_ptids_event      <- 20
n_min_fireids_event    <- 2
n_min_ptids_subgroup   <- 50
n_min_fireids_subgroup <- 5
```


# Outcome comparison plots

Compare RAP tree vs VCF tree for each analysis subset × grouping scheme,
one panel per outcome.

```{r outcome-compare}
#| message: false
#| warning: false
#| fig-width: 10
#| fig-height: 5

dir_outcome_compare <- file.path(dir_figs, "outcome_compare")
dir_ensure(dir_outcome_compare)

# Iterate over subset_id (replaces analysis_id in old script)
for (sid in sort(unique(coef_tbl$subset_id))) {

  p <- plot_event_study(
    coef_tbl  = coef_tbl,
    filters   = list(
      subset_id = sid,
      group_id  = "b10_pdsisumn10",
      vcov_id   = "cluster_eco4",
      model_id  = "b_pdsi_sunab_twfe"
    ),
    color_var             = "subgroup",
    color_values          = subgroup_pal,
    linetype_var          = NULL,
    facet_var             = "outcome",
    facet_ncol            = 2,
    facet_scales          = "fixed",
    legend_title_color    = "Subgroup",
    legend_title_linetype = NULL,
    n_min_ptids_event     = n_min_ptids_event,
    n_min_fireids_event   = n_min_fireids_event,
    n_min_ptids_subgroup  = n_min_ptids_subgroup,
    n_min_fireids_subgroup = n_min_fireids_subgroup,
    show_ci               = TRUE,
    ci_style              = "ribbon",
    title                 = sid,
    subtitle              = NULL,
    baseline              = -6,
    xlim                  = c(-10, 20),
    ylim                  = c(-50, 10),
    show_n                = TRUE,
    n_legend_var          = NULL,
    line_size             = 1,
    point_size            = 1.8,
    ribbon_alpha          = 0.2,
    ci_line_alpha         = 0.7,
    ci_level              = 0.95
  )

  print(p)

  ggplot2::ggsave(
    file.path(dir_outcome_compare, glue::glue("outcome_compare_{sid}.png")),
    plot   = p,
    units  = "px",
    width  = 4000,
    height = 2000
  )
}
```
compare weighted to non-weighted

```{r outcome-compare}
#| message: false
#| warning: false
#| fig-width: 10
#| fig-height: 5

dir_outcome_compare <- file.path(dir_figs, "outcome_compare")
dir_ensure(dir_outcome_compare)

# Iterate over subset_id (replaces analysis_id in old script)
for (sid in sort(unique(coef_tbl$subset_id))) {

  p <- plot_event_study(
    coef_tbl  = coef_tbl,
    filters   = list(
      subset_id = sid,
      group_id  = "b10_pdsin3t1",
      vcov_id   = "cluster_eco4",
      model_id  = "b_pdsi_sunab_twfe"
    ),
    color_var             = "subgroup",
    color_values          = subgroup_pal,
    linetype_var          = NULL,
    facet_var             = "outcome",
    facet_ncol            = 2,
    facet_scales          = "fixed",
    legend_title_color    = "Subgroup",
    legend_title_linetype = NULL,
    n_min_ptids_event     = n_min_ptids_event,
    n_min_fireids_event   = n_min_fireids_event,
    n_min_ptids_subgroup  = n_min_ptids_subgroup,
    n_min_fireids_subgroup = n_min_fireids_subgroup,
    show_ci               = TRUE,
    ci_style              = "ribbon",
    title                 = sid,
    subtitle              = NULL,
    baseline              = -6,
    xlim                  = c(-10, 20),
    ylim                  = c(-50, 10),
    show_n                = TRUE,
    n_legend_var          = NULL,
    line_size             = 1,
    point_size            = 1.8,
    ribbon_alpha          = 0.2,
    ci_line_alpha         = 0.7,
    ci_level              = 0.95
  )

  print(p)

  ggplot2::ggsave(
    file.path(dir_outcome_compare, glue::glue("outcome_compare_{sid}.png")),
    plot   = p,
    units  = "px",
    width  = 4000,
    height = 2000
  )
}
```

# Grouping scheme comparison plots

Compare the four treatment grouping schemes for a fixed outcome and subset.

```{r grouping-compare}
#| message: false
#| warning: false
#| fig-width: 10
#| fig-height: 5

dir_group_compare <- file.path(dir_figs, "group_compare")
dir_ensure(dir_group_compare)

for (sid in sort(unique(coef_tbl$subset_id))) {
  for (outcome_val in unique(coef_tbl$outcome)) {

    p <- plot_event_study(
      coef_tbl  = coef_tbl,
      filters   = list(
        subset_id = sid,
        outcome   = outcome_val,
        vcov_id   = "cluster_eco4",
        model_id  = "b_pdsi_sunab_twfe"
      ),
      color_var             = "subgroup",
      color_values          = subgroup_pal,
      linetype_var          = NULL,
      facet_var             = "group_id",
      facet_ncol            = 2,
      facet_scales          = "fixed",
      legend_title_color    = "Subgroup",
      n_min_ptids_event     = n_min_ptids_event,
      n_min_fireids_event   = n_min_fireids_event,
      n_min_ptids_subgroup  = n_min_ptids_subgroup,
      n_min_fireids_subgroup = n_min_fireids_subgroup,
      show_ci               = TRUE,
      ci_style              = "ribbon",
      title                 = glue::glue("{sid} — {outcome_val}"),
      baseline              = -6,
      xlim                  = c(-10, 20),
      ylim                  = c(-50, 10),
      show_n                = TRUE,
      line_size             = 1,
      point_size            = 1.8,
      ribbon_alpha          = 0.2,
      ci_level              = 0.95
    )

    print(p)

    ggplot2::ggsave(
      file.path(
        dir_group_compare,
        glue::glue("group_compare_{sid}_{outcome_val}.png")
      ),
      plot   = p,
      units  = "px",
      width  = 4000,
      height = 2000
    )
  }
}
```


# Vcov sensitivity plots

Compare standard error estimators for a fixed subset × outcome × grouping.

```{r vcov-sensitivity}
#| message: false
#| warning: false
#| fig-width: 12
#| fig-height: 5

dir_vcov_compare <- file.path(dir_figs, "vcov_compare")
dir_ensure(dir_vcov_compare)

for (sid in sort(unique(coef_tbl$subset_id))) {
  for (outcome_val in unique(coef_tbl$outcome)) {

    p <- plot_event_study(
      coef_tbl  = coef_tbl,
      filters   = list(
        subset_id = sid,
        outcome   = outcome_val,
        group_id  = "b10_pdsin3t1",
        model_id  = "b_pdsi_sunab_twfe",
        subgroup  = "f"
      ),
      color_var             = "vcov_label",
      linetype_var          = NULL,
      facet_var             = NULL,
      legend_title_color    = "SE estimator",
      n_min_ptids_event     = n_min_ptids_event,
      n_min_fireids_event   = n_min_fireids_event,
      n_min_ptids_subgroup  = n_min_ptids_subgroup,
      n_min_fireids_subgroup = n_min_fireids_subgroup,
      show_ci               = TRUE,
      ci_style              = "lines",
      title                 = glue::glue("{sid} — {outcome_val} — vcov sensitivity"),
      baseline              = -6,
      xlim                  = c(-10, 20),
      ylim                  = c(-50, 10),
      show_n                = FALSE,
      line_size             = 0.9,
      point_size            = 1.5,
      ci_line_alpha         = 0.5,
      ci_level              = 0.95
    )

    print(p)

    ggplot2::ggsave(
      file.path(
        dir_vcov_compare,
        glue::glue("vcov_compare_{sid}_{outcome_val}.png")
      ),
      plot   = p,
      units  = "px",
      width  = 4000,
      height = 2000
    )
  }
}
```


# Cross-subset comparison (e.g. all ecoregions, one subgroup)

Compare a single subgroup × outcome across all analysis subsets (ecoregions),
using `subset_id` on the colour or facet aesthetic.

```{r cross-subset}
#| message: false
#| warning: false
#| fig-width: 14
#| fig-height: 6

dir_subset_compare <- file.path(dir_figs, "subset_compare")
dir_ensure(dir_subset_compare)

for (outcome_val in unique(coef_tbl$outcome)) {

  p <- plot_event_study(
    coef_tbl  = coef_tbl,
    filters   = list(
      outcome  = outcome_val,
      group_id = "b10_pdsin3t1",
      vcov_id  = "cluster_eco4",
      model_id = "b_pdsi_sunab_twfe",
      subgroup = "f"
    ),
    color_var             = "subset_id",
    color_values          = subset_pal,
    linetype_var          = NULL,
    facet_var             = NULL,
    legend_title_color    = "Subset",
    n_min_ptids_event     = n_min_ptids_event,
    n_min_fireids_event   = n_min_fireids_event,
    n_min_ptids_subgroup  = n_min_ptids_subgroup,
    n_min_fireids_subgroup = n_min_fireids_subgroup,
    show_ci               = FALSE,    # too cluttered with many lines
    title                 = glue::glue("All subsets — {outcome_val} — subgroup f"),
    baseline              = -6,
    xlim                  = c(-10, 20),
    ylim                  = c(-50, 10),
    show_n                = TRUE,
    line_size             = 0.8,
    point_size            = 1.4,
    ci_level              = 0.95
  )

  print(p)

  ggplot2::ggsave(
    file.path(
      dir_subset_compare,
      glue::glue("subset_compare_{outcome_val}.png")
    ),
    plot   = p,
    units  = "px",
    width  = 5000,
    height = 2500
  )
}
```
