
---
title: "06_dydid_v6_testestimators"
author: Tyler L. McIntosh
date: last-modified
format:
  html:
    embed-resources: true
    html-math-method: katex
    css: styles.css
    toc: true
    toc-depth: 1
    toc-title: Contents
    toc-location: right
    code-fold: true
    code-summary: "Show the code"
---

Compare different DyDiD estimators - Sierra Nevada & North cascades as test case
- basic TWFE dydid
- Sun-Abraham
- Calloway Sant'anna

From prior Sunab models, we see two different cases in SN vs NC:
- SN: notable pre-fire biotic drop
- NC: poor pre-cd period parallelality

# Setup for analysis

``` {r setupchunk}
#| echo: true
#| message: false
#| warning: false
#| results: 'hide'

rm(list = ls())

cyverse = FALSE

if(!requireNamespace("here", quietly = TRUE)) {
  install.packages("here")
}
library(here)

source(here::here("utils", "functions.R"))


install_load_packages(c(
  "tidyverse",
  "tictoc",
  "glue",
  "fixest",
  "arrow",
  "did",
  "didimputation"
))

source(here::here("code", "portable_sunab2.R"))

run <- "GEE_resilience_v6_operational_ss500_ts50000"


if(cyverse) {
  dir_figs <- file.path("~/data-store/data/iplant/home/shared/earthlab/macrosystems/tlm", run, "figs")
  dir_derived <- file.path("~/data-store/data/iplant/home/shared/earthlab/macrosystems/tlm", run, "data/derived")
  dir_parquet <- here::here(dir_derived, 'parquet')
  dir_manual <- file.path("~/data-store/data/iplant/home/shared/earthlab/macrosystems/tlm", run, "data/manual")
  dir_raw <- file.path("~/data-store/data/iplant/home/shared/earthlab/macrosystems/tlm", run, "data/raw")
  dir_results <- file.path("~/data-store/data/iplant/home/shared/earthlab/macrosystems/tlm", run, "results")
} else {
  dir_figs <- here::here('figs', run)
  dir_derived <- here::here('data', 'derived', run)
  dir_parquet <- here::here(dir_derived, 'parquet')
  dir_raw <- here::here('data', 'raw')
  dir_manual <- here::here('data', 'manual')
  dir_results <- here::here('results')
}

dir_ensure(c(dir_figs,
             dir_derived,
             dir_parquet,
             dir_manual,
             dir_raw,
             dir_results))


seed = 1234
set.seed(seed)

```

# CD group function
```{r}
set_cd_groups <- function(df,
                          b_nm,
                          d_nm,
                          cd_nm,
                          b_threshold,
                          d_threshold) {
  
  if(d_threshold < 0) {
    df_new <- df |>
      dplyr::mutate(
        "{cd_nm}" := dplyr::case_when(
          .data[[b_nm]] <  b_threshold & fire == 1 & .data[[d_nm]] >  d_threshold ~ "f",
          .data[[b_nm]] >= b_threshold & fire == 1 & .data[[d_nm]] >  d_threshold ~ "bf",
          .data[[b_nm]] <  b_threshold & fire == 1 & .data[[d_nm]] <= d_threshold ~ "df",
          .data[[b_nm]] >= b_threshold & fire == 1 & .data[[d_nm]] <= d_threshold ~ "bdf",
          TRUE ~ NA_character_
        )
      )
  } else {
    df_new <- df |>
      dplyr::mutate(
        "{cd_nm}" := dplyr::case_when(
          .data[[b_nm]] <  b_threshold & fire == 1 & .data[[d_nm]] <  d_threshold ~ "f",
          .data[[b_nm]] >= b_threshold & fire == 1 & .data[[d_nm]] <  d_threshold ~ "bf",
          .data[[b_nm]] <  b_threshold & fire == 1 & .data[[d_nm]] >= d_threshold ~ "df",
          .data[[b_nm]] >= b_threshold & fire == 1 & .data[[d_nm]] >= d_threshold ~ "bdf",
          TRUE ~ NA_character_
        )
      )
    
  }
  
  df_new <- df_new |>
    dplyr::mutate(
      "{cd_nm}" := forcats::fct_relevel(.data[[cd_nm]], c("f", "bf", "df", "bdf"))
    )
  
  return(df_new)
}


set_cd_groups_with_control <- function(df,
                          b_nm,
                          d_nm,
                          cd_nm,
                          b_threshold,
                          d_threshold) {
  
  if(d_threshold < 0) {
    df_new <- df |>
      dplyr::mutate(
        "{cd_nm}" := dplyr::case_when(
          .data[[b_nm]] <  b_threshold & fire == 1 & .data[[d_nm]] >  d_threshold ~ "f",
          .data[[b_nm]] >= b_threshold & fire == 1 & .data[[d_nm]] >  d_threshold ~ "bf",
          .data[[b_nm]] <  b_threshold & fire == 1 & .data[[d_nm]] <= d_threshold ~ "df",
          .data[[b_nm]] >= b_threshold & fire == 1 & .data[[d_nm]] <= d_threshold ~ "bdf",
          fire == 0 ~ "control",
          TRUE ~ NA_character_
        )
      )
  } else {
    df_new <- df |>
      dplyr::mutate(
        "{cd_nm}" := dplyr::case_when(
          .data[[b_nm]] <  b_threshold & fire == 1 & .data[[d_nm]] <  d_threshold ~ "f",
          .data[[b_nm]] >= b_threshold & fire == 1 & .data[[d_nm]] <  d_threshold ~ "bf",
          .data[[b_nm]] <  b_threshold & fire == 1 & .data[[d_nm]] >= d_threshold ~ "df",
          .data[[b_nm]] >= b_threshold & fire == 1 & .data[[d_nm]] >= d_threshold ~ "bdf",
          fire == 0 ~ "control",
          TRUE ~ NA_character_
        )
      )
    
  }
  
  df_new <- df_new |>
    dplyr::mutate(
      "{cd_nm}" := forcats::fct_relevel(.data[[cd_nm]], c("control", "f", "bf", "df", "bdf"))
    )
  
  return(df_new)
}



```

# Sunab functions
```{r}
# library(dplyr)
# library(ggplot2)
# library(purrr)
# library(rlang)
# library(stringr)
# library(glue)
# library(fixest)
# 
# sunab_estimator_subg <- function(subgrp_col,
#                                  grp,
#                                  df,
#                                  fml,
#                                  trt_col = "fire",
#                                  id_col = "uniqueID",
#                                  model_fun = fixest::feols,
#                                  model_args = list(
#                                    mem.clean = TRUE,
#                                    lean = FALSE
#                                  ),
#                                  vcov = NULL) {
# 
#   col_sym <- sym(subgrp_col)
#   trt_col_sym <- sym(trt_col)
# 
#   df_never_trt <- df |>
#     filter(!!trt_col_sym == 0)
# 
#   df_subset <- df |>
#     filter(!!trt_col_sym == 1, !!col_sym == grp)
# 
#   df_use <- bind_rows(df_never_trt, df_subset)
# 
#   call_args <- c(
#     list(
#       fml = fml,
#       data = df_use
#     ),
#     model_args
#   )
# 
#   if (!is.null(vcov)) {
#     call_args$vcov <- vcov
#   }
# 
#   model <- do.call(model_fun, call_args)
# 
#   n_treated_subgroup_units <- dplyr::n_distinct(df_subset[[id_col]])
#   n_never_treated_units <- dplyr::n_distinct(df_never_trt[[id_col]])
# 
#   attr(model, "n_treated_subgroup_units") <- n_treated_subgroup_units
#   attr(model, "n_never_treated_units") <- n_never_treated_units
#   attr(model, "subgroup_value") <- grp
#   attr(model, "subgroup_column") <- subgrp_col
# 
#   model
# }
# 
# 
# sunab_plot_models <- function(model_list,
#                               model_labels = NULL,
#                               title = "Dynamic-DiD Event-Study Estimates from Multiple Models",
#                               ribbon_type = "ribbon",
#                               caption = NULL) {
# 
#   if (is.null(model_labels)) {
#     model_labels <- paste0("Model ", seq_along(model_list))
#   }
# 
#   has_attr <- function(x, attr_name) {
#     !is.null(attr(x, attr_name))
#   }
# 
#   plot_data <- map2_dfr(model_list, model_labels, function(model, label) {
#     df_coef <- as.data.frame(as.matrix(model$coeftable))
#     df_coef$term <- rownames(model$coeftable)
# 
#     # if (has_attr(model, "n_treated_subgroup_units")) {
#     #   label <- glue("{label} (N = {attr(model, 'n_treated_subgroup_units')})")
#     # }
# 
#     df_coef |>
#       filter(str_detect(term, "^year::")) |>
#       mutate(
#         Time = as.numeric(str_extract(term, "-?\\d+")),
#         CI_upper = Estimate + 1.96 * `Std. Error`,
#         CI_lower = Estimate - 1.96 * `Std. Error`,
#         Model = label
#       )
#   })
# 
#   if (ribbon_type == "ribbon") {
#     ggplot(plot_data, aes(x = Time, y = Estimate, color = Model, fill = Model)) +
#       geom_line(linewidth = 1) +
#       geom_point(size = 2) +
#       geom_ribbon(aes(ymin = CI_lower, ymax = CI_upper), alpha = 0.2, color = NA) +
#       geom_vline(xintercept = 0, linetype = "dashed", color = "red") +
#       geom_vline(xintercept = -6, linetype = "dashed", color = "gray40") +
#       geom_hline(yintercept = 0, color = "black") +
#       labs(
#         x = "Event Time",
#         y = "Treatment effect of fire on forest cover",
#         title = title,
#         caption = caption
#       ) +
#       theme_minimal(base_size = 13) +
#       theme(legend.position = "bottom")
#   } else {
#     ggplot(plot_data, aes(x = Time, y = Estimate, color = Model)) +
#       geom_line(linewidth = 1) +
#       geom_point(size = 2) +
#       geom_line(aes(y = CI_upper), linetype = "dotted", linewidth = 0.8) +
#       geom_line(aes(y = CI_lower), linetype = "dotted", linewidth = 0.8) +
#       geom_vline(xintercept = -1, linetype = "dashed", color = "red") +
#       geom_vline(xintercept = -6, linetype = "dashed", color = "gray40") +   
#       geom_hline(yintercept = 0, color = "black") +
#       labs(
#         x = "Event Time",
#         y = "Treatment effect of fire on forest cover",
#         title = title,
#         caption = caption
#       ) +
#       theme_minimal(base_size = 13) +
#       theme(legend.position = "bottom") +
#       xlim(c(-10, 20))
#   }
# }
# 
# 
# sunab_plot_all_together_did <- function(subg_col,
#                                         df,
#                                         fml,
#                                         title = NULL,
#                                         trt_col = "fire",
#                                         id_col = "uniqueID",
#                                         model_fun = fixest::feols,
#                                         model_args = list(
#                                           mem.clean = TRUE,
#                                           lean = FALSE
#                                         ),
#                                         vcov = NULL,
#                                         ribbon_type = "ribbon",
#                                         caption = NULL) {
# 
#   if (is.null(title)) {
#     title <- glue("Effect estimates by subgroup in column {subg_col}")
#   }
# 
#   subgs <- df |>
#     pull(all_of(subg_col)) |>
#     unique()
# 
#   subgs <- subgs[!is.na(subgs)]
# 
#   iw_results <- subgs |>
#     set_names() |>
#     map(\(x) {
#       sunab_estimator_subg(
#         subgrp_col = subg_col,
#         grp = x,
#         df = df,
#         fml = fml,
#         trt_col = trt_col,
#         id_col = id_col,
#         model_fun = model_fun,
#         model_args = model_args,
#         vcov = vcov
#       )
#     })
# 
#   sunab_plot_models(
#     model_list = iw_results,
#     model_labels = subgs,
#     title = title,
#     ribbon_type = ribbon_type,
#     caption = caption
#   )
# }

```


# prep dats

```{r}

test_dats_sn <- arrow::read_parquet(here(dir_derived, "parquet", "dats_long_sierranevada.parquet"))
test_dats_sn_short <- arrow::read_parquet(here(dir_derived, "parquet", "dats_short_sierranevada.parquet"))
test_dats_nc <- arrow::read_parquet(here(dir_derived, "parquet", "dats_long_northcascades.parquet"))

#test_dats_sr <- arrow::read_parquet(here(dir_derived, "parquet", "dats_long_southernrockies.parquet"))



test_dats_sn <- test_dats_sn |>
  set_cd_groups(
    b_nm = "biotic_relaxedforestnorm_5_yrs_prior_sum_yot",
    d_nm = "pdsi_annual_5_yrs_prior_threshold_n3_yot",
    cd_nm = "b10_pdsin3t1",
    b_threshold = 10,
    d_threshold = 1
  ) |>
  filter(year >= 1997) #|>
  #mutate(use_tv_covars = ifelse(Time_to_Treatment <= -5 & Time_to_Treatment <= 0, 0, 1))


test_dats_nc <- test_dats_nc |>
  set_cd_groups(
    b_nm = "biotic_relaxedforestnorm_5_yrs_prior_sum_yot",
    d_nm = "pdsi_annual_5_yrs_prior_threshold_n3_yot",
    cd_nm = "b10_pdsin3t1",
    b_threshold = 10,
    d_threshold = 1
  ) |>
  filter(year >= 1997) #|>
  #mutate(use_tv_covars = ifelse(Time_to_Treatment <= -5 & Time_to_Treatment <= 0, 0, 1))


# WHY ARE SOME UNITS NA FOR THE BIOTIC/PDSI SUMS??
test_dats_sn |> group_by(b10_pdsin3t1) |> summarize(n = n())
why_na <- test_dats_sn |>
  filter(is.na(b10_pdsin3t1)) |>
  select(pdsi_annual_5_yrs_prior_threshold_n3_yot, biotic_relaxedforestnorm_5_yrs_prior_sum_yot, fire, year, pt_id, lat, long)
length(unique(why_na$pt_id))
length(unique(test_dats_sn$pt_id))
why_na_sf <- why_na |>
  distinct(pt_id, .keep_all = TRUE) |>
  sf::st_as_sf(coords = c("long", "lat"), crs = 4326)

mapview::mapview(why_na_sf)
write_csv(why_na_sf, file = here::here(dir_derived, "sierra_nevada_biotic_or_pdsi_nas_unique_pt_id.csv"))

# 
# test_dats_sn_gwf <- test_dats_sn |>
#   mutate(gwf_cat = case_when((frg_reclass == 1 | frg_reclass == 3) & burn_cbi <= 2.25 ~ "LowSevGWF",
#                              (frg_reclass == 2 | frg_reclass == 4) & burn_cbi <= 2.25 ~ "LowSevInReplacement",
#                              (frg_reclass == 2 | frg_reclass == 4) & burn_cbi > 2.25 ~ "HighSevGWF",
#                              (frg_reclass == 1 | frg_reclass == 3) & burn_cbi > 2.25 ~ "HighSevInLow",
#          ))
# 
# test_dats_nc_gwf <- test_dats_nc |>
#   mutate(gwf_cat = case_when((frg_reclass == 1 | frg_reclass == 3) & burn_cbi <= 2.25 ~ "LowSevGWF",
#                              (frg_reclass == 2 | frg_reclass == 4) & burn_cbi <= 2.25 ~ "LowSevInReplacement",
#                              (frg_reclass == 2 | frg_reclass == 4) & burn_cbi > 2.25 ~ "HighSevGWF",
#                              (frg_reclass == 1 | frg_reclass == 3) & burn_cbi > 2.25 ~ "HighSevInLow",
#          ))
# 
# test_dats_sr_gwf <- test_dats_sr |>
#   mutate(gwf_cat = case_when((frg_reclass == 1 | frg_reclass == 3) & burn_cbi <= 2.25 ~ "LowSevGWF",
#                              (frg_reclass == 2 | frg_reclass == 4) & burn_cbi <= 2.25 ~ "LowSevInReplacement",
#                              (frg_reclass == 2 | frg_reclass == 4) & burn_cbi > 2.25 ~ "HighSevGWF",
#                              (frg_reclass == 1 | frg_reclass == 3) & burn_cbi > 2.25 ~ "HighSevInLow",
#          ))

```


# Quick GWF test
```{r}

# 
# sunab_gwf_fml <- rap_tree ~ sunab(FirstTreat, year, ref.p = -1) + pdsi_annual | pt_id + year
# 
# 
# sunab_gwf_args <- list(
#   fml = sunab_gwf_fml,
#   vcov = ~ us_l4code,
#   model_fun = fixest::feols,
#   model_args = list(
#     mem.clean = TRUE,
#     lean = FALSE
#   ),
#   term_pattern = "^year::",
#   time_extract_regex = "-?\\d+",
#   treatment_time = 0,
#   ref_line = -1,
#   caption = deparse(sunab_gwf_fml)
# )
# 
# sunab_gwf_p1 <- plot_all_together_did(
#   subg_col = "gwf_cat",
#   df = test_dats_sn_gwf,
#   backend = "sunab",
#   backend_args = sunab_gwf_args,
#   title = "Sierra Nevada test: Sun-Ab, GWF"
# )
# 
# 
# sunab_gwf_p2 <- plot_all_together_did(
#   subg_col = "gwf_cat",
#   df = test_dats_nc_gwf,
#   backend = "sunab",
#   backend_args = sunab_gwf_args,
#   title = "North Cascades test: Sun-Ab, GWF"
# )
# 
# 
# sunab_gwf_p3 <- plot_all_together_did(
#   subg_col = "gwf_cat",
#   df = test_dats_sr_gwf,
#   backend = "sunab",
#   backend_args = sunab_gwf_args,
#   title = "Southern Rockies test: Sun-Ab, GWF"
# )
# 
# 
# sunab_gwf_p1 + xlim(-5, 25)
# sunab_gwf_p2 + xlim(-5, 25)
# sunab_gwf_p3 + xlim(-5, 25)

```


# Test one group only - F
```{r}

td_sn_1 <- test_dats_sn |>
  filter(b10_pdsin3t1 == "f" | is.na(b10_pdsin3t1)) |>
  mutate(use_tv_covars = ifelse(Time_to_Treatment <= -5 & Time_to_Treatment <= 0, 0, 1))



est_sunab <- feols(
  rap_tree ~ sunab(FirstTreat, year, ref.p = -6) + biotic_relaxedforestnorm + pdsi_annual | pt_id + year,
  data = td_sn_1,
  mem.clean = TRUE,
  lean = FALSE,
  cluster = ~ us_l4code
)

est_sunab2 <- feols(
  rap_tree ~ sunab(FirstTreat, year, ref.p = -6) + biotic_relaxedforestnorm:use_tv_covars + pdsi_annual:use_tv_covars | pt_id + year,
  data = td_sn_1,
  mem.clean = TRUE,
  lean = FALSE,
  cluster = ~ us_l4code
)


est_twfe <- feols(
  rap_tree ~ i(factor_var = Time_to_Treatment, var = treated, ref = -6) + biotic_relaxedforestnorm + pdsi_annual | pt_id + year,
  data = td_sn_1,
  mem.clean = TRUE,
  lean = FALSE,
  cluster = ~ us_l4code
)



# est_sunab_together_f <- feols(
#   rap_tree ~ sunab(FirstTreat, year, ref.p = -6):b10_pdsin3t1 | pt_id + year,
#   data = td_sn_1 |> mutate(b10_pdsin3t1 = ifelse(is.na(b10_pdsin3t1), "none", b10_pdsin3t1)),
#   mem.clean = TRUE,
#   lean = FALSE,
#   cluster = ~ us_l4code
# )

# 
# test_dats_sn_simple <- test_dats_sn |> select(b10_pdsin3t1, FirstTreat, year, pt_id, fire) |>
#     mutate(
#       b10_pdsin3t1 = forcats::fct_na_value_to_level(b10_pdsin3t1, level = "control")
#     ) 
# test_sample <- test_dats_sn_simple |>
#     distinct(pt_id) |>
#     slice_sample(n = 10000) |>
#     inner_join(test_dats_sn, by = "pt_id")
  



# 
# # Try to run sunab interaction models; CODE DOESN"T WORK
# est_sunab_together <- feols(
#   rap_tree ~ sunab(FirstTreat, year, ref.p = -6):b10_pdsin3t1 | pt_id + year,
#   data = test_sample,
#   mem.clean = TRUE,
#   lean = FALSE,
#   cluster = NULL #~ us_l4code
# )
# 
# est_sunab_together$coeftable
# 
# iplot(est_sunab_together, sep = TRUE)
# fixest::iplot(
#   est_sunab_together,
#   group = list(
#     f   = "b10_pdsin3t1f$",
#     bf  = "b10_pdsin3t1bf$",
#     df  = "b10_pdsin3t1df$",
#     bdf = "b10_pdsin3t1bdf$"
#   )
# )
# 
# dyn_custom <- aggregate(
#   est_sunab_together,
#   agg = "(year::-?[0-9]+):cohort::[0-9]+:(b10_pdsin3t1[a-z]+)"
# )
# 
# dyn_df <- as.data.frame(dyn_custom) |>
#   rownames_to_column("term")
# 
# names(dyn_df) <- c("term", "estimate", "std_error", "stat", "p_value")
# 
# dyn_df <- dyn_df |>
#   mutate(
#     event_time = as.numeric(str_match(term, "year::(-?\\d+)")[, 2]),
#     subgroup = str_match(term, "(b10_pdsin3t1[a-z]+)$|(b10_pdsin3t1[a-z]+)$")[, 1],
#     conf_low = estimate - 1.96 * std_error,
#     conf_high = estimate + 1.96 * std_error
#   )
# 
# dyn_df <- dyn_df |>
#   mutate(
#     subgroup = sub("^b10_pdsin3t1", "", subgroup)
#   )
# 
# ggplot(dyn_df, aes(x = event_time, y = estimate, color = subgroup, fill = subgroup)) +
#   geom_hline(yintercept = 0, linetype = "dashed") +
#   geom_vline(xintercept = -6, linetype = "dotted") +
#   geom_ribbon(aes(ymin = conf_low, ymax = conf_high), alpha = 0.15, color = NA) +
#   geom_line() +
#   geom_point() +
#   theme_minimal() +
#   ylim(-40, 50)





#CS Model
est_cs <- att_gt(
  yname = "rap_tree",
  tname = "year",
  idname = "pt_id_num",
  gname = "FirstTreat",
  xformla = ~ biotic_relaxedforestnorm + pdsi_annual,
  data = td_sn_1 |>
    mutate(
      pt_id_num = as.numeric(gsub("\\D", "", pt_id)),
      FirstTreat = FirstTreat - 5,
      FirstTreat = ifelse(FirstTreat == 995, 0, FirstTreat)
    ),
  panel = TRUE,
  control_group = "nevertreated",
  base_period = "universal",
  anticipation = 0,
  allow_unbalanced_panel = FALSE,
  est_method = "dr"
)

es_cs <- aggte(
  est_cs,
  type = "dynamic"
)

summary(es_cs)





#BJS model
est_bjs <- did_imputation(
  data = td_sn_1 |>
    mutate(FirstTreat = FirstTreat - 5,
          FirstTreat = ifelse(FirstTreat == 995, NA, FirstTreat)),
  yname = "rap_tree",
  gname = "FirstTreat",
  tname = "year",
  idname = "pt_id",
  first_stage = rap_tree ~ biotic_relaxedforestnorm + pdsi_annual | pt_id + year,
  horizon = -10:20,
  pretrends = -10:-1,
  cluster_var = "pt_id"
)


# PLOTS
sn_f_cs_p <- did::ggdid(es_cs) + 
  xlim(-10, 20) +
  labs(title = "Callaway Sant'Anna F, Sierra Nevada")

sn_f_twfe_p <- ggfixest::ggiplot(list('Sunab' = est_sunab, 'TWFE' = est_twfe, 'Sunab2' = est_sunab2),
        ci_level = 0.95,
        #geom_style = "ribbon",
        col = c("turquoise", "pink", "yellow")) +
  xlim(-10, 20) +
  ylim(-30, 20) +
  labs(title = "Sun-Ab & TWFE, F, Sierra Nevada")


sn_f_twfe_p


est_bjs |>
  mutate(term_num = as.numeric(term)) |>
  ggplot(aes(x = term_num, y = estimate)) +
  geom_hline(yintercept = 0, linetype = 2) +
  geom_vline(xintercept = -0.5, linetype = 2) +
  geom_point() +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0.15) +
  labs(
    x = "Event time",
    y = "Estimated effect",
    title = "Borusyak imputation event-study"
  )

```


# Test one group only - BF
```{r}

td_sn_1 <- test_dats_sn |>
  filter(b10_pdsin3t1 == "bf" | is.na(b10_pdsin3t1))



est_sunab <- feols(
  rap_tree ~ sunab(FirstTreat, year, ref.p = -6) + biotic_relaxedforestnorm + pdsi_annual | pt_id + year,
  data = td_sn_1,
  mem.clean = TRUE,
  lean = FALSE,
  cluster = ~ us_l4code
)


est_twfe <- feols(
  rap_tree ~ i(factor_var = Time_to_Treatment, var = treated, ref = -6) + biotic_relaxedforestnorm + pdsi_annual | pt_id + year,
  data = td_sn_1,
  mem.clean = TRUE,
  lean = FALSE,
  cluster = ~ us_l4code
)



est_cs <- att_gt(
  yname = "rap_tree",
  tname = "year",
  idname = "pt_id_num",
  gname = "FirstTreat",
  xformla = ~ biotic_relaxedforestnorm + pdsi_annual,
  data = td_sn_1 |>
    mutate(
      pt_id_num = as.numeric(gsub("\\D", "", pt_id)),
      FirstTreat = FirstTreat - 5,
      FirstTreat = ifelse(FirstTreat == 995, 0, FirstTreat)
    ),
  panel = TRUE,
  control_group = "nevertreated",
  base_period = "universal",
  anticipation = 0,
  allow_unbalanced_panel = FALSE,
  est_method = "dr"
)

es_cs <- aggte(
  est_cs,
  type = "dynamic"
)

summary(es_cs)


sn_bf_cs_p <- did::ggdid(es_cs) + 
  xlim(-10, 20) +
  labs(title = "Callaway Sant'Anna F, Sierra Nevada")

sn_bf_twfe_p <- ggfixest::ggiplot(list('Sunab' = est_sunab, 'TWFE' = est_twfe),
        ci_level = 0.95,
        #geom_style = "ribbon",
        col = c("turquoise", "pink")) +
  xlim(-10, 20) +
  ylim(-30, 20) +
  labs(title = "Sun-Ab & TWFE, F, Sierra Nevada")


  

```


# Sun-abraham

```{r}
# 
# # Formulas
# sunab_fml1 <- rap_tree ~ sunab(FirstTreat, year, ref.p = -6)
# sunab_fml2 <- rap_tree ~ sunab(FirstTreat, year, ref.p = -6) | pt_id + year
# sunab_fml3 <- rap_tree ~ sunab(FirstTreat, year, ref.p = -6) + biotic_relaxedforestnorm + pdsi_annual | pt_id + year
# sunab_fml4 <- rap_tree ~ sunab(FirstTreat, year, ref.p = -6) + biotic_relaxedforestnorm:use_tv_covars  + pdsi_annual:use_tv_covars | pt_id + year
# 
# 
# 
# sn_sunab_p1 <- sunab_plot_all_together_did(
#   subg_col = "b10_pdsin3t1",
#   df = test_dats_sn,
#   fml = sunab_fml1,
#   vcov = ~ us_l4code,
#   title = "Sierra Nevada test: Sun-Ab, naive",
#   caption = as.character(sunab_fml1)[3]
# )
# 
# sn_sunab_p2 <- sunab_plot_all_together_did(
#   subg_col = "b10_pdsin3t1",
#   df = test_dats_sn,
#   fml = sunab_fml2,
#   vcov = ~ us_l4code,
#   title = "Sierra Nevada test: Sun-Ab, twfe",
#   caption = as.character(sunab_fml2)[3]
# )
# 
# sn_sunab_p3 <- sunab_plot_all_together_did(
#   subg_col = "b10_pdsin3t1",
#   df = test_dats_sn,
#   fml = sunab_fml3,
#   vcov = ~ us_l4code,
#   title = "Sierra Nevada test: Sun-Ab, twfe + covars",
#   caption = as.character(sunab_fml3)[3]
# )
# 
# sn_sunab_p4 <- sunab_plot_all_together_did(
#   subg_col = "b10_pdsin3t1",
#   df = test_dats_sn,
#   fml = sunab_fml4,
#   vcov = ~ us_l4code,
#   title = "Sierra Nevada test: Sun-Ab, twfe + covars:usetv",
#   caption = as.character(sunab_fml4)[3]
# )
# 
# 
# nc_sunab_p1 <- sunab_plot_all_together_did(
#   subg_col = "b10_pdsin3t1",
#   df = test_dats_nc,
#   fml = sunab_fml1,
#   vcov = ~ us_l4code,
#   title = "North Cascades test: Sun-Ab, naive",
#   caption = as.character(sunab_fml1)[3]
# )
# 
# nc_sunab_p2 <- sunab_plot_all_together_did(
#   subg_col = "b10_pdsin3t1",
#   df = test_dats_nc,
#   fml = sunab_fml2,
#   vcov = ~ us_l4code,
#   title = "North Cascades test: Sun-Ab, twfe",
#   caption = as.character(sunab_fml2)[3]
# )
# 
# nc_sunab_p3 <- sunab_plot_all_together_did(
#   subg_col = "b10_pdsin3t1",
#   df = test_dats_nc,
#   fml = sunab_fml3,
#   vcov = ~ us_l4code,
#   title = "North Cascades test: Sun-Ab, twfe + covars",
#   caption = as.character(sunab_fml3)[3]
# )


```
# Sierra Nevada results

```{r}
# sn_sunab_p1
# sn_sunab_p2
# sn_sunab_p3
# sn_sunab_p4

```

# North Cascades results

```{r}
nc_sunab_p1
nc_sunab_p2
nc_sunab_p3

```

# Code that works with all 3 estimators

```{r}

library(dplyr)
library(ggplot2)
library(purrr)
library(rlang)
library(stringr)
library(glue)
library(fixest)
library(did)

`%||%` <- function(x, y) {
  if (is.null(x)) y else x
}


filter_subgroup_data <- function(df,
                                 subgrp_col,
                                 grp,
                                 trt_col = "fire",
                                 treated_value = 1,
                                 untreated_value = 0) {

  col_sym <- sym(subgrp_col)
  trt_col_sym <- sym(trt_col)

  df_never_trt <- df |>
    filter(!!trt_col_sym == untreated_value)

  df_subset <- df |>
    filter(!!trt_col_sym == treated_value, !!col_sym == grp)

  df_use <- bind_rows(df_never_trt, df_subset)

  list(
    df_use = df_use,
    df_subset = df_subset,
    df_never_trt = df_never_trt
  )
}


fit_event_study_model <- function(df,
                                  backend = c("sunab", "twfe", "cs"),
                                  backend_args) {

  backend <- match.arg(backend)

  if (backend %in% c("sunab", "twfe")) {
    model_fun <- backend_args$model_fun %||% fixest::feols
    model_args <- backend_args$model_args %||% list(mem.clean = TRUE, lean = FALSE)

    call_args <- c(
      list(
        fml = backend_args$fml,
        data = df
      ),
      model_args
    )

    if (!is.null(backend_args$vcov)) {
      call_args$vcov <- backend_args$vcov
    }

    out <- do.call(model_fun, call_args)

    attr(out, "estimator_backend") <- backend
    attr(out, "term_pattern") <- backend_args$term_pattern
    attr(out, "time_extract_regex") <- backend_args$time_extract_regex
    attr(out, "treatment_time") <- backend_args$treatment_time %||% 0
    attr(out, "ref_line") <- backend_args$ref_line
    attr(out, "caption") <- backend_args$caption

    return(out)
  }

  if (backend == "cs") {
    att_gt_args <- backend_args$att_gt_args %||% list()
    aggte_args <- backend_args$aggte_args %||% list(type = "dynamic")

    att_gt_call <- c(
      list(data = df),
      att_gt_args
    )

    att_gt_obj <- do.call(did::att_gt, att_gt_call)

    aggte_call <- c(
      list(MP = att_gt_obj),
      aggte_args
    )

    aggte_obj <- do.call(did::aggte, aggte_call)

    out <- list(
      att_gt = att_gt_obj,
      aggte = aggte_obj
    )

    class(out) <- c("subgroup_did_es", class(out))

    attr(out, "estimator_backend") <- backend
    attr(out, "treatment_time") <- backend_args$treatment_time %||% 0
    attr(out, "ref_line") <- backend_args$ref_line
    attr(out, "caption") <- backend_args$caption

    return(out)
  }
}


estimate_subgroup_model <- function(subgrp_col,
                                    grp,
                                    df,
                                    backend = c("sunab", "twfe", "cs"),
                                    backend_args,
                                    trt_col = "fire",
                                    id_col = "uniqueID",
                                    treated_value = 1,
                                    untreated_value = 0) {

  backend <- match.arg(backend)

  d_sub <- filter_subgroup_data(
    df = df,
    subgrp_col = subgrp_col,
    grp = grp,
    trt_col = trt_col,
    treated_value = treated_value,
    untreated_value = untreated_value
  )

  model <- fit_event_study_model(
    df = d_sub$df_use,
    backend = backend,
    backend_args = backend_args
  )

  n_treated_subgroup_units <- dplyr::n_distinct(d_sub$df_subset[[id_col]])
  n_never_treated_units <- dplyr::n_distinct(d_sub$df_never_trt[[id_col]])

  attr(model, "n_treated_subgroup_units") <- n_treated_subgroup_units
  attr(model, "n_never_treated_units") <- n_never_treated_units
  attr(model, "subgroup_value") <- grp
  attr(model, "subgroup_column") <- subgrp_col

  model
}


extract_fixest_event_study <- function(model,
                                       label = NULL,
                                       term_pattern = NULL,
                                       time_extract_regex = NULL) {

  backend <- attr(model, "estimator_backend")

  term_pattern <- term_pattern %||%
    attr(model, "term_pattern") %||%
    switch(
      backend,
      sunab = "^year::",
      twfe = NULL,
      NULL
    )

  time_extract_regex <- time_extract_regex %||%
    attr(model, "time_extract_regex") %||%
    "-?\\d+"

  if (is.null(term_pattern)) {
    stop("For fixest TWFE models, provide `term_pattern` in `backend_args` so event-time coefficients can be identified.")
  }

  coef_df <- as.data.frame(as.matrix(model$coeftable))
  coef_df$term <- rownames(model$coeftable)

  if (is.null(label)) {
    label <- attr(model, "subgroup_value") %||% "model"
  }

  # if (!is.null(attr(model, "n_treated_subgroup_units"))) {
  #   label <- glue("{label} (N = {attr(model, 'n_treated_subgroup_units')})")
  # }

  coef_df |>
    filter(str_detect(term, term_pattern)) |>
    mutate(
      Time = as.numeric(str_extract(term, time_extract_regex)),
      Estimate = Estimate,
      Std_Error = `Std. Error`,
      CI_lower = Estimate - 1.96 * Std_Error,
      CI_upper = Estimate + 1.96 * Std_Error,
      Model = label,
      backend = backend
    ) |>
    arrange(Time)
}


extract_did_event_study <- function(model, label = NULL) {

  dyn <- model$aggte

  egt <- dyn$egt

  att_vals <- dyn$att.egt %||% dyn$att_egt
  se_vals <- dyn$se.egt %||% dyn$se_egt
  crit_val <- dyn$crit.val %||% 1.96

  if (is.null(att_vals) || is.null(se_vals)) {
    stop("Could not find dynamic ATT or SE vectors in the `aggte` object.")
  }

  if (is.null(label)) {
    label <- attr(model, "subgroup_value") %||% "model"
  }

  # if (!is.null(attr(model, "n_treated_subgroup_units"))) {
  #   label <- glue("{label} (N = {attr(model, 'n_treated_subgroup_units')})")
  # }

  tibble(
    Time = egt,
    Estimate = att_vals,
    Std_Error = se_vals,
    CI_lower = Estimate - crit_val * Std_Error,
    CI_upper = Estimate + crit_val * Std_Error,
    Model = label,
    backend = "cs"
  ) |>
    arrange(Time)
}


extract_event_study_data <- function(model, label = NULL) {

  if (inherits(model, "fixest")) {
    return(extract_fixest_event_study(model = model, label = label))
  }

  if (inherits(model, "subgroup_did_es")) {
    return(extract_did_event_study(model = model, label = label))
  }

  stop("Unsupported model type for event-study extraction.")
}


plot_event_study_models <- function(model_list,
                                    model_labels = NULL,
                                    title = "Dynamic-DiD event-study estimates",
                                    ylab = "Treatment effect",
                                    ribbon_type = "ribbon",
                                    caption = NULL,
                                    treatment_time = NULL,
                                    ref_line = NULL,
                                    xlim = NULL) {

  if (is.null(model_labels)) {
    model_labels <- names(model_list)

    if (is.null(model_labels)) {
      model_labels <- paste0("Model ", seq_along(model_list))
    }
  }

  plot_data <- map2_dfr(model_list, model_labels, extract_event_study_data)
  
  ##
  all_models <- unique(model_labels)

  plot_data <- plot_data |>
    mutate(Model = factor(Model, levels = all_models))
  
  palette_vals <- scales::hue_pal()(length(all_models))
  names(palette_vals) <- all_models
  ##

  treatment_time <- treatment_time %||%
    attr(model_list[[1]], "treatment_time")

  ref_line <- ref_line %||%
    attr(model_list[[1]], "ref_line")

  caption <- caption %||%
    attr(model_list[[1]], "caption")

  if (ribbon_type == "ribbon") {
    p <- ggplot(plot_data, aes(x = Time, y = Estimate, color = Model, fill = Model)) +
      geom_line(linewidth = 1) +
      geom_point(size = 2) +
      geom_ribbon(aes(ymin = CI_lower, ymax = CI_upper), alpha = 0.2, color = NA)
  } else {
    p <- ggplot(plot_data, aes(x = Time, y = Estimate, color = Model)) +
      geom_line(linewidth = 1) +
      geom_point(size = 2) +
      geom_line(aes(y = CI_upper), linetype = "dotted", linewidth = 0.8) +
      geom_line(aes(y = CI_lower), linetype = "dotted", linewidth = 0.8)
  }

  p <- p +
    geom_hline(yintercept = 0, color = "black")

  if (!is.null(treatment_time)) {
    p <- p +
      geom_vline(xintercept = treatment_time, linetype = "dashed", color = "red")
  }

  if (!is.null(ref_line)) {
    p <- p +
      geom_vline(xintercept = ref_line, linetype = "dashed", color = "gray40")
  }

  p <- p +
    labs(
      x = "Event Time",
      y = ylab,
      title = title,
      caption = caption
    ) +
    theme_minimal(base_size = 13) +
    theme(legend.position = "bottom")

  if (!is.null(xlim)) {
    p <- p + coord_cartesian(xlim = xlim)
  }
  
  ##
  p <- p +
    scale_color_manual(values = palette_vals) +
    scale_fill_manual(values = palette_vals)
  ##
  
  p
}


plot_all_together_did <- function(subg_col,
                                  df,
                                  backend = c("sunab", "twfe", "cs"),
                                  backend_args,
                                  title = NULL,
                                  trt_col = "fire",
                                  id_col = "uniqueID",
                                  treated_value = 1,
                                  untreated_value = 0,
                                  ribbon_type = "ribbon",
                                  ylab = "Treatment effect of fire on forest cover",
                                  caption = NULL,
                                  treatment_time = NULL,
                                  ref_line = NULL,
                                  xlim = NULL) {

  backend <- match.arg(backend)

  if (is.null(title)) {
    title <- glue("Effect estimates by subgroup in column {subg_col}")
  }

  subgs <- df |>
    pull(all_of(subg_col)) |>
    unique()

  subgs <- subgs[!is.na(subgs)]

  model_list <- subgs |>
    set_names() |>
    map(\(x) {
      estimate_subgroup_model(
        subgrp_col = subg_col,
        grp = x,
        df = df,
        backend = backend,
        backend_args = backend_args,
        trt_col = trt_col,
        id_col = id_col,
        treated_value = treated_value,
        untreated_value = untreated_value
      )
    })

  plot_event_study_models(
    model_list = model_list,
    model_labels = subgs,
    title = title,
    ylab = ylab,
    ribbon_type = ribbon_type,
    caption = caption,
    treatment_time = treatment_time,
    ref_line = ref_line,
    xlim = xlim
  )
}


# SUNAB
sunab_fml1 <- rap_tree ~ sunab(FirstTreat, year, ref.p = -6)
sunab_fml2 <- rap_tree ~ sunab(FirstTreat, year, ref.p = -6) | pt_id + year
sunab_fml3 <- rap_tree ~ sunab(FirstTreat, year, ref.p = -6) +
  biotic_relaxedforestnorm + pdsi_annual | pt_id + year

sunab_args1 <- list(
  fml = sunab_fml1,
  vcov = ~ us_l4code,
  model_fun = fixest::feols,
  model_args = list(
    mem.clean = TRUE,
    lean = FALSE
  ),
  term_pattern = "^year::",
  time_extract_regex = "-?\\d+",
  treatment_time = 0,
  ref_line = -6,
  caption = deparse(sunab_fml1)
)

sunab_args2 <- list(
  fml = sunab_fml2,
  vcov = ~ us_l4code,
  model_fun = fixest::feols,
  model_args = list(
    mem.clean = TRUE,
    lean = FALSE
  ),
  term_pattern = "^year::",
  time_extract_regex = "-?\\d+",
  treatment_time = 0,
  ref_line = -6,
  caption = deparse(sunab_fml2)
)

sunab_args3 <- list(
  fml = sunab_fml3,
  vcov = ~ us_l4code,
  model_fun = fixest::feols,
  model_args = list(
    mem.clean = TRUE,
    lean = FALSE
  ),
  term_pattern = "^year::",
  time_extract_regex = "-?\\d+",
  treatment_time = 0,
  ref_line = -6,
  caption = deparse(sunab_fml3)
)




# TWFE

twfe_fml1 <- rap_tree ~ i(Time_to_Treatment, fire, ref = -6) | pt_id + year

twfe_args1 <- list(
  fml = twfe_fml1,
  vcov = ~ us_l4code,
  model_fun = fixest::feols,
  model_args = list(
    mem.clean = TRUE,
    lean = FALSE
  ),
  term_pattern = "^Time_to_Treatment::",
  time_extract_regex = "-?\\d+",
  treatment_time = -5,
  ref_line = -6,
  caption = deparse(twfe_fml1)
)


twfe_fml3 <- rap_tree ~ i(Time_to_Treatment, fire, ref = -6) +
  biotic_relaxedforestnorm + pdsi_annual | pt_id + year

twfe_args3 <- list(
  fml = twfe_fml3,
  vcov = ~ us_l4code,
  model_fun = fixest::feols,
  model_args = list(
    mem.clean = TRUE,
    lean = FALSE
  ),
  term_pattern = "^Time_to_Treatment::",
  time_extract_regex = "-?\\d+",
  treatment_time = -5,
  ref_line = -6,
  caption = deparse(twfe_fml3)
)



# CA-SA
# These specifications include bstrap and cband set to FALSE to estimate pointwise confidence intervals rather than uniform confidence bands; pointwise confidence intervals is what is most common in application settings (noted by CA-SA here: https://bcallaway11.github.io/did/articles/pre-testing.html)
cs_args1 <- list(
  att_gt_args = list(
    yname = "rap_tree",
    tname = "year",
    idname = "pt_id_num",
    gname = "FirstTreat",
    xformla = ~ 1,
    panel = TRUE,
    control_group = "nevertreated",
    est_method = "dr",
    clustervars = "us_l4code",
    faster_mode = TRUE,
    bstrap = FALSE,
    cband = FALSE
  ),
  aggte_args = list(
    type = "dynamic"
  ),
  treatment_time = 0,
  ref_line = -1,
  caption = "did::att_gt(yname = 'rap_tree', tname = 'year', idname = 'pt_id', gname = 'FirstTreat', xformla = ~ 1)"
)

cs_args2 <- list(
  att_gt_args = list(
    yname = "rap_tree",
    tname = "year",
    idname = "pt_id_num",
    gname = "FirstTreat",
    xformla = ~ biotic_relaxedforestnorm + pdsi_annual,
    panel = TRUE,
    control_group = "nevertreated",
    est_method = "dr",
    clustervars = "us_l4code",
    faster_mode = TRUE,
    bstrap = FALSE,
    cband = FALSE
  ),
  aggte_args = list(
    type = "dynamic"
  ),
  treatment_time = 0,
  ref_line = -1,
  caption = "did::att_gt(..., xformla = ~ biotic_relaxedforestnorm + pdsi_annual)"
)

cs_args2_nyt <- list(
  att_gt_args = list(
    yname = "rap_tree",
    tname = "year",
    idname = "pt_id_num",
    gname = "FirstTreat",
    xformla = ~ biotic_relaxedforestnorm + pdsi_annual,
    panel = TRUE,
    control_group = "notyettreated",
    est_method = "dr",
    clustervars = "us_l4code",
    faster_mode = TRUE,
    bstrap = FALSE,
    cband = FALSE
  ),
  aggte_args = list(
    type = "dynamic"
  ),
  treatment_time = 0,
  ref_line = -1,
  caption = "did::att_gt(..., xformla = ~ biotic_relaxedforestnorm + pdsi_annual), notyettreated"
)


```

# use code


```{r}

#Sunab
sn_sunab_p1 <- plot_all_together_did(
  subg_col = "b10_pdsin3t1",
  df = test_dats_sn,
  backend = "sunab",
  backend_args = sunab_args1,
  title = "Sierra Nevada test: Sun-Ab, naive"
)

sn_sunab_p2 <- plot_all_together_did(
  subg_col = "b10_pdsin3t1",
  df = test_dats_sn,
  backend = "sunab",
  backend_args = sunab_args2,
  title = "Sierra Nevada test: Sun-Ab, twfe"
)

sn_sunab_p3 <- plot_all_together_did(
  subg_col = "b10_pdsin3t1",
  df = test_dats_sn,
  backend = "sunab",
  backend_args = sunab_args3,
  title = "Sierra Nevada test: Sun-Ab, twfe + covars"
)



#TWFE
sn_twfe_p1 <- plot_all_together_did(
  subg_col = "b10_pdsin3t1",
  df = test_dats_sn,
  backend = "twfe",
  backend_args = twfe_args1,
  title = "Sierra Nevada test: TWFE event study"
)

sn_twfe_p3 <- plot_all_together_did(
  subg_col = "b10_pdsin3t1",
  df = test_dats_sn,
  backend = "twfe",
  backend_args = twfe_args3,
  title = "Sierra Nevada test: TWFE event study + covars"
)




#Callaway Sant'Anna
sn_cs_p1 <- plot_all_together_did(
  subg_col = "b10_pdsin3t1",
  df = test_dats_sn |>
    mutate(
      pt_id_num = as.numeric(gsub("\\D", "", pt_id)),
      FirstTreat = FirstTreat - 5,
      FirstTreat = ifelse(FirstTreat == 995, 0, FirstTreat)
    ),
  backend = "cs",
  backend_args = cs_args1,
  title = "Sierra Nevada test: Callaway-Sant'Anna"
)

sn_cs_p2 <- plot_all_together_did(
  subg_col = "b10_pdsin3t1",
  df = test_dats_sn |>
    mutate(
      pt_id_num = as.numeric(gsub("\\D", "", pt_id)),
      FirstTreat = FirstTreat - 5,
      FirstTreat = ifelse(FirstTreat == 995, 0, FirstTreat)
    ),
  backend = "cs",
  backend_args = cs_args2,
  title = "Sierra Nevada test: Callaway-Sant'Anna + covars"
)

sn_cs_p2_nyt <- plot_all_together_did(
  subg_col = "b10_pdsin3t1",
  df = test_dats_sn |>
    mutate(
      pt_id_num = as.numeric(gsub("\\D", "", pt_id)),
      FirstTreat = FirstTreat - 5,
      FirstTreat = ifelse(FirstTreat == 995, 0, FirstTreat)
    ),
  backend = "cs",
  backend_args = cs_args2_nyt,
  title = "Sierra Nevada test: Callaway-Sant'Anna + covars, notyettreated"
)


```





```{r}



nc_sunab_p1 <- plot_all_together_did(
  subg_col = "b10_pdsin3t1",
  df = test_dats_nc,
  backend = "sunab",
  backend_args = sunab_args1,
  title = "NCascades test: Sun-Ab, naive"
)

nc_sunab_p2 <- plot_all_together_did(
  subg_col = "b10_pdsin3t1",
  df = test_dats_nc,
  backend = "sunab",
  backend_args = sunab_args2,
  title = "NCascades test: Sun-Ab, twfe"
)

nc_sunab_p3 <- plot_all_together_did(
  subg_col = "b10_pdsin3t1",
  df = test_dats_nc,
  backend = "sunab",
  backend_args = sunab_args3,
  title = "NCascades test: Sun-Ab, twfe + covars"
)


nc_twfe_p3 <- plot_all_together_did(
  subg_col = "b10_pdsin3t1",
  df = test_dats_nc,
  backend = "twfe",
  backend_args = twfe_args3,
  title = "NCascades test: TWFE event study"
)



nc_cs_p1 <- plot_all_together_did(
  subg_col = "b10_pdsin3t1",
  df = test_dats_nc |>
    mutate(
      pt_id_num = as.numeric(gsub("\\D", "", pt_id)),
      FirstTreat = FirstTreat - 5,
      FirstTreat = ifelse(FirstTreat == 995, 0, FirstTreat)
    ),
  backend = "cs",
  backend_args = cs_args1,
  title = "NCascades test: Callaway-Sant'Anna"
)

nc_cs_p2 <- plot_all_together_did(
  subg_col = "b10_pdsin3t1",
  df = test_dats_nc |>
    mutate(
      pt_id_num = as.numeric(gsub("\\D", "", pt_id)),
      FirstTreat = FirstTreat - 5,
      FirstTreat = ifelse(FirstTreat == 995, 0, FirstTreat)
    ),
  backend = "cs",
  backend_args = cs_args2,
  title = "NCascades test: Callaway-Sant'Anna + covars"
)

nc_cs_p3 <- plot_all_together_did(
  subg_col = "b10_pdsin3t1",
  df = test_dats_nc |>
    mutate(
      pt_id_num = as.numeric(gsub("\\D", "", pt_id)),
      FirstTreat = FirstTreat - 5,
      FirstTreat = ifelse(FirstTreat == 995, 0, FirstTreat)
    ),
  backend = "cs",
  backend_args = cs_args3,
  title = "NCascades test: Callaway-Sant'Anna + covars, notyettreated"
)



nc_sunab_p1
nc_sunab_p2
nc_sunab_p3
nc_twfe_p3
nc_cs_p1 + ylim(-40, 20)
nc_cs_p2 + ylim(-40, 20)
nc_cs_p3 + ylim(-40, 20)

```


# Test weighting

```{r}

dir_weight_test <- here::here(dir_figs, "weighting_test")
dir_ensure(dir_weight_test)

library(WeightIt)
library(cobalt)

pre_cd_cover <- test_dats_sn_short |> select(pt_id, gam_rap_tree_pre6_fit)


weight_test_sn <- test_dats_sn |>
  set_cd_groups_with_control(
    b_nm = "biotic_relaxedforestnorm_5_yrs_prior_sum_yot",
    d_nm = "pdsi_annual_5_yrs_prior_threshold_n3_yot",
    cd_nm = "b10_pdsin3t1_control",
    b_threshold = 10,
    d_threshold = 1
  ) |>
  distinct(pt_id, .keep_all = TRUE) |>
  filter(!is.na(b10_pdsin3t1_control)) |>
  left_join(pre_cd_cover)




# generalized overlap weighting via multinomial propensity scores
w_out_glm <- weightit(
  b10_pdsin3t1_control ~ chili + def + aet + srtm + tpi + nfg_factor,
  data = weight_test_sn,
  method = "glm",
  estimand = "ATO"
)

summary(w_out_glm)


w_out_glm_ate <- weightit(
  b10_pdsin3t1_control ~ chili + def + aet + srtm + tpi + nfg_factor,
  data = weight_test_sn,
  method = "glm",
  estimand = "ATE"
)

summary(w_out_glm_ate)


w_out_glm_precov <- weightit(
  b10_pdsin3t1_control ~ chili + def + aet + srtm + tpi + nfg_factor + gam_rap_tree_pre6_fit,
  data = weight_test_sn,
  method = "glm",
  estimand = "ATO"
)

summary(w_out_glm_precov)




# not well-balanced yet, try GBM

w_out_gbm <- weightit(
  b10_pdsin3t1_control ~ chili + def + aet + srtm + tpi + nfg_factor,
  data = weight_test_sn,
  method = "gbm",
  estimand = "ATO"
)

summary(w_out_gbm)


w_out_gbm_precov <- weightit(
  b10_pdsin3t1_control ~ chili + def + aet + srtm + tpi + nfg_factor + gam_rap_tree_pre6_fit,
  data = weight_test_sn,
  method = "gbm",
  estimand = "ATO"
)

summary(w_out_gbm_precov)


weight_test_sn <- weight_test_sn |>
  mutate(glm_weights = w_out_glm$weights,
         gbm_weights = w_out_gbm$weights,
         glm_weights_precov = w_out_glm_precov$weights,
         gbm_weights_precov = w_out_gbm_precov$weights,
         glm_weights_ate = w_out_glm_ate$weights)


test_dats_sn_withweights <- test_dats_sn |>
  left_join(weight_test_sn |>
              select(pt_id, glm_weights, gbm_weights, glm_weights_precov, gbm_weights_precov, glm_weights_ate), by = c("pt_id"))




# balance tables

bal_glm <- bal.tab(
  w_out_glm,
  un = TRUE,
  pairwise = TRUE,
  abs = TRUE,
  stats = c("mean.diffs", "variance.ratios"),
  thresholds = c(m = .1, v = 2)
)

print(bal_glm)

# For all pairwise data
bal_glm$Pair.Balance


bal_gbm <- bal.tab(
  w_out_gbm,
  un = TRUE,
  pairwise = TRUE,
  abs = TRUE,
  stats = c("mean.diffs", "variance.ratios"),
  thresholds = c(m = .1, v = 2)
)

print(bal_gbm)

bal_glm_precov <- bal.tab(
  w_out_glm_precov,
  un = TRUE,
  pairwise = TRUE,
  abs = TRUE,
  stats = c("mean.diffs", "variance.ratios"),
  thresholds = c(m = .1, v = 2)
)

print(bal_glm_precov)


bal_gbm_precov <- bal.tab(
  w_out_gbm_precov,
  un = TRUE,
  pairwise = TRUE,
  abs = TRUE,
  stats = c("mean.diffs", "variance.ratios"),
  thresholds = c(m = .1, v = 2)
)

print(bal_gbm_precov)


bal_sn_test_obs <- bal_glm$Observations |> 
  mutate(weights = "glm") |> 
  rbind(bal_gbm$Observations |> 
          mutate(weights = "gbm")) |>
  rbind(bal_glm_precov$Observations |>
          mutate(weights = "glm_precov")) |>
  rbind(bal_gbm_precov$Observations |>
          mutate(weights = "gbm_precov")) %>%
  mutate(adjustment = rownames(.))

write_csv(bal_sn_test_obs, file = here(dir_weight_test, "obs_counts_sn_test.csv"))


plot_df <- bal_sn_test_obs |>
  pivot_longer(
    cols = c(control, f, bf, df, bdf),
    names_to = "group",
    values_to = "n"
  ) |>
  mutate(
    model = case_when(
      weights == "glm" & adjustment == "Unadjusted" ~ "Unadjusted",
      weights == "glm" & adjustment == "Adjusted"   ~ "GLM adjusted",
      #weights == "glm_precov" & grepl("Adjusted", adjustment)  ~ "GLM_precov adjusted",
      weights == "gbm" & grepl("Adjusted", adjustment) ~ "GBM adjusted",
      #weights == "gbm_precov" & grepl("Adjusted", adjustment) ~ "GBM_precov adjusted",
      TRUE ~ NA_character_
    )
  ) |>
  filter(!is.na(model))

ggplot(plot_df, aes(x = model, y = n, fill = model)) +
  geom_col(width = 0.7) +
  facet_wrap(~ group, scales = "free_y") +
  labs(x = NULL, y = "Count") +
  theme_bw() +
  theme(
    legend.position = "none",
    axis.text.x = element_text(angle = 45, hjust = 1)
  ) +
  labs(title = "Sierra Nevada effective sample size with weighting")

ggsave(filename = here(dir_weight_test, "sn_sample_size_weighted.png"))

# Love plots

love.plot(
  w_out_glm,
  stats = "mean.diffs",
  abs = TRUE,
  pairwise = TRUE,
  thresholds = c(m = .1),
  var.order = "unadjusted",
  which.treat = .all
) +
  labs(title = "GLM-weighted Sierra Nevada, pairwise\n Weighted for Average Treatment Effect in the Overlap (ATO)")

ggsave(filename = here(dir_weight_test, "love_plot_glm_pairwise.png"),
       units = "px",
       width = 5000,
       height = 2000,
       bg = "white")

love.plot(
  w_out_glm,
  stats = "mean.diffs",
  abs = TRUE,
  pairwise = TRUE,
  thresholds = c(m = .1),
  var.order = "unadjusted"
) +
  labs(title = "GLM-weighted Sierra Nevada, max\n Weighted for Average Treatment Effect in the Overlap (ATO)")

ggsave(filename = here(dir_weight_test, "love_plot_glm_max.png"),
       units = "px",
       width = 2000,
       height = 2000,
       bg = "white")


love.plot(
  w_out_glm_ate,
  stats = "mean.diffs",
  abs = TRUE,
  pairwise = TRUE,
  thresholds = c(m = .1),
  var.order = "unadjusted"
) +
  labs(title = "GLM-weighted Sierra Nevada, max\n Weighted for Average Treatment Effect (ATE)")

ggsave(filename = here(dir_weight_test, "love_plot_glm_ate_max.png"),
       units = "px",
       width = 2000,
       height = 2000,
       bg = "white")

love.plot(
  w_out_gbm,
  stats = "mean.diffs",
  abs = TRUE,
  pairwise = TRUE,
  thresholds = c(m = .1),
  var.order = "unadjusted"
)+
  labs(title = "GBM-weighted Sierra Nevada, max\n Weighted for Average Treatment Effect in the Overlap (ATO)")

ggsave(filename = here(dir_weight_test, "love_plot_gbm_max.png"),
       units = "px",
       width = 2000,
       height = 2000,
       bg = "white")



love.plot(
  w_out_glm_precov,
  stats = "mean.diffs",
  abs = TRUE,
  pairwise = TRUE,
  thresholds = c(m = .1),
  var.order = "unadjusted"
)+
  labs(title = "GLM-weighted Sierra Nevada - with pre-cover, max\n Weighted for Average Treatment Effect in the Overlap (ATO)")

ggsave(filename = here(dir_weight_test, "love_plot_glm_precov_max.png"),
       units = "px",
       width = 2000,
       height = 2000,
       bg = "white")

love.plot(
  w_out_gbm_precov,
  stats = "mean.diffs",
  abs = TRUE,
  pairwise = TRUE,
  thresholds = c(m = .1),
  var.order = "unadjusted"
)+
  labs(title = "GBM-weighted Sierra Nevada - with pre-cover, max\n Weighted for Average Treatment Effect in the Overlap (ATO)")

ggsave(filename = here(dir_weight_test, "love_plot_gbm_precov_max.png"),
       units = "px",
       width = 2000,
       height = 2000,
       bg = "white")



# Weighted distributions


# ATE

ggplot(weight_test_sn |> filter(!is.na(b10_pdsin3t1)), aes(x = aet,
                          color = b10_pdsin3t1)) +
  geom_density() +
  labs(title = "Raw distribution of AET")

ggplot(weight_test_sn |> filter(!is.na(b10_pdsin3t1)), aes(x = aet,
                          color = b10_pdsin3t1,
                          weight = glm_weights )) +
  geom_density() +
  labs(title = "Weighted distribution of AET (overlap region using GLM) - ATO")


ggplot(weight_test_sn |> filter(!is.na(b10_pdsin3t1)), aes(x = aet,
                          color = b10_pdsin3t1,
                          weight = gbm_weights)) +
  geom_density() +
  labs(title = "Weighted distribution of AET (overlap region using GBM) - ATO")


ggplot(weight_test_sn |> filter(!is.na(b10_pdsin3t1)), aes(x = aet,
                          color = b10_pdsin3t1,
                          weight = glm_weights_ate)) +
  geom_density() +
  labs(title = "Weighted distribution of ATE (overlap region using GLM) - ATE")




# SRTM
ggplot(weight_test_sn |> filter(!is.na(b10_pdsin3t1)), aes(x = srtm,
                          color = b10_pdsin3t1)) +
  geom_density() +
  labs(title = "Raw distribution of SRTM")

ggplot(weight_test_sn |> filter(!is.na(b10_pdsin3t1)), aes(x = srtm,
                          color = b10_pdsin3t1,
                          weight = glm_weights)) +
  geom_density() +
  labs(title = "Weighted distribution of SRTM (overlap region using GLM) - ATO")


ggplot(weight_test_sn |> filter(!is.na(b10_pdsin3t1)), aes(x = srtm,
                          color = b10_pdsin3t1,
                          weight = gbm_weights)) +
  geom_density() +
  labs(title = "Weighted distribution of SRTM (overlap region using GBM) - ATO")

ggplot(weight_test_sn |> filter(!is.na(b10_pdsin3t1)), aes(x = srtm,
                          color = b10_pdsin3t1,
                          weight = glm_weights_ate)) +
  geom_density() +
  labs(title = "Weighted distribution of SRTM (overlap region using GLM) - ATE")



plot_df <- bind_rows(
  weight_test_sn |> 
    filter(!is.na(b10_pdsin3t1)) |>
    mutate(type = "Unweighted",
           w = 1),
  
  weight_test_sn |> 
    filter(!is.na(b10_pdsin3t1)) |>
    mutate(type = "GLM Weighted",
           w = glm_weights),
  
  weight_test_sn |> 
    filter(!is.na(b10_pdsin3t1)) |>
    mutate(type = "GBM Weighted",
           w = gbm_weights)
)

ggplot(plot_df, aes(x = srtm,
                    color = b10_pdsin3t1,
                    linetype = type,
                    weight = w)) +
  geom_density() +
  labs(title = "Distribution of SRTM: Weighted vs Unweighted",
       linetype = "Type",
       color = "Group")

```


# Weighted CS-A estimator

```{r}
# without pre-cover

cs_args_weighted_glm <- list(
  att_gt_args = list(
    yname = "rap_tree",
    tname = "year",
    idname = "pt_id_num",
    gname = "FirstTreat",
    xformla = ~ biotic_relaxedforestnorm + pdsi_annual,
    panel = TRUE,
    control_group = "notyettreated",
    est_method = "dr",
    clustervars = "us_l4code",
    faster_mode = TRUE,
    bstrap = FALSE,
    cband = FALSE,
    weightsname = "glm_weights"
  ),
  aggte_args = list(
    type = "dynamic"
  ),
  treatment_time = 0,
  ref_line = -1,
  caption = "did::att_gt(..., xformla = ~ biotic_relaxedforestnorm + pdsi_annual), notyettreated"
)


cs_args_weighted_gbm <- list(
  att_gt_args = list(
    yname = "rap_tree",
    tname = "year",
    idname = "pt_id_num",
    gname = "FirstTreat",
    xformla = ~ biotic_relaxedforestnorm + pdsi_annual,
    panel = TRUE,
    control_group = "notyettreated",
    est_method = "dr",
    clustervars = "us_l4code",
    faster_mode = TRUE,
    bstrap = FALSE,
    cband = FALSE,
    weightsname = "gbm_weights"
  ),
  aggte_args = list(
    type = "dynamic"
  ),
  treatment_time = 0,
  ref_line = -1,
  caption = "did::att_gt(..., xformla = ~ biotic_relaxedforestnorm + pdsi_annual), notyettreated"
)



sn_cs_weighted_covars_glm <- plot_all_together_did(
  subg_col = "b10_pdsin3t1",
  df = test_dats_sn_withweights |>
    mutate(
      pt_id_num = as.numeric(gsub("\\D", "", pt_id)),
      FirstTreat = FirstTreat - 5,
      FirstTreat = ifelse(FirstTreat == 995, 0, FirstTreat)
    ),
  backend = "cs",
  backend_args = cs_args_weighted_glm,
  title = "Sierra Nevada test: Callaway-Sant'Anna, GLM-weighted"
)

sn_cs_weighted_covars_gbm <- plot_all_together_did(
  subg_col = "b10_pdsin3t1",
  df = test_dats_sn_withweights |>
    mutate(
      pt_id_num = as.numeric(gsub("\\D", "", pt_id)),
      FirstTreat = FirstTreat - 5,
      FirstTreat = ifelse(FirstTreat == 995, 0, FirstTreat)
    ),
  backend = "cs",
  backend_args = cs_args_weighted_gbm,
  title = "Sierra Nevada test: Callaway-Sant'Anna, GBM-weighted"
)


# with pre-cover

cs_args_weighted_glm_precov <- list(
  att_gt_args = list(
    yname = "rap_tree",
    tname = "year",
    idname = "pt_id_num",
    gname = "FirstTreat",
    xformla = ~ biotic_relaxedforestnorm + pdsi_annual,
    panel = TRUE,
    control_group = "notyettreated",
    est_method = "dr",
    clustervars = "us_l4code",
    faster_mode = TRUE,
    bstrap = FALSE,
    cband = FALSE,
    weightsname = "glm_weights_precov"
  ),
  aggte_args = list(
    type = "dynamic"
  ),
  treatment_time = 0,
  ref_line = -1,
  caption = "did::att_gt(..., xformla = ~ biotic_relaxedforestnorm + pdsi_annual), notyettreated"
)


cs_args_weighted_gbm_precov <- list(
  att_gt_args = list(
    yname = "rap_tree",
    tname = "year",
    idname = "pt_id_num",
    gname = "FirstTreat",
    xformla = ~ biotic_relaxedforestnorm + pdsi_annual,
    panel = TRUE,
    control_group = "notyettreated",
    est_method = "dr",
    clustervars = "us_l4code",
    faster_mode = TRUE,
    bstrap = FALSE,
    cband = FALSE,
    weightsname = "gbm_weights_precov"
  ),
  aggte_args = list(
    type = "dynamic"
  ),
  treatment_time = 0,
  ref_line = -1,
  caption = "did::att_gt(..., xformla = ~ biotic_relaxedforestnorm + pdsi_annual), notyettreated"
)



sn_cs_weighted_covars_glm_precov <- plot_all_together_did(
  subg_col = "b10_pdsin3t1",
  df = test_dats_sn_withweights |>
    mutate(
      pt_id_num = as.numeric(gsub("\\D", "", pt_id)),
      FirstTreat = FirstTreat - 5,
      FirstTreat = ifelse(FirstTreat == 995, 0, FirstTreat)
    ),
  backend = "cs",
  backend_args = cs_args_weighted_glm_precov,
  title = "Sierra Nevada test: Callaway-Sant'Anna, GLM-weighted"
)

sn_cs_weighted_covars_gbm_precov <- plot_all_together_did(
  subg_col = "b10_pdsin3t1",
  df = test_dats_sn_withweights |>
    mutate(
      pt_id_num = as.numeric(gsub("\\D", "", pt_id)),
      FirstTreat = FirstTreat - 5,
      FirstTreat = ifelse(FirstTreat == 995, 0, FirstTreat)
    ),
  backend = "cs",
  backend_args = cs_args_weighted_gbm_precov,
  title = "Sierra Nevada test: Callaway-Sant'Anna, GBM-weighted"
)


# ATE


cs_args_weighted_glm_ate <- list(
  att_gt_args = list(
    yname = "rap_tree",
    tname = "year",
    idname = "pt_id_num",
    gname = "FirstTreat",
    xformla = ~ biotic_relaxedforestnorm + pdsi_annual,
    panel = TRUE,
    control_group = "notyettreated",
    est_method = "dr",
    clustervars = "us_l4code",
    faster_mode = TRUE,
    bstrap = FALSE,
    cband = FALSE,
    weightsname = "glm_weights_ate"
  ),
  aggte_args = list(
    type = "dynamic"
  ),
  treatment_time = 0,
  ref_line = -1,
  caption = "did::att_gt(..., xformla = ~ biotic_relaxedforestnorm + pdsi_annual), notyettreated"
)


sn_cs_weighted_covars_glm_precov <- plot_all_together_did(
  subg_col = "b10_pdsin3t1",
  df = test_dats_sn_withweights |>
    mutate(
      pt_id_num = as.numeric(gsub("\\D", "", pt_id)),
      FirstTreat = FirstTreat - 5,
      FirstTreat = ifelse(FirstTreat == 995, 0, FirstTreat)
    ),
  backend = "cs",
  backend_args = cs_args_weighted_glm_precov,
  title = "Sierra Nevada test: Callaway-Sant'Anna, GLM-weighted"
)

```

# Weighted Sun-Abraham estimator???

```{r}

sunab_args3_glmweights <- list(
  fml = sunab_fml3,
  vcov = ~ us_l4code,
  model_fun = fixest::feols,
  model_args = list(
    mem.clean = TRUE,
    lean = FALSE,
    weights = ~ glm_weights
  ),
  term_pattern = "^year::",
  time_extract_regex = "-?\\d+",
  treatment_time = 0,
  ref_line = -6,
  caption = deparse(sunab_fml3)
)


sunab_args3_gbmweights <- list(
  fml = sunab_fml3,
  vcov = ~ us_l4code,
  model_fun = fixest::feols,
  model_args = list(
    mem.clean = TRUE,
    lean = FALSE,
    weights = ~ gbm_weights
  ),
  term_pattern = "^year::",
  time_extract_regex = "-?\\d+",
  treatment_time = 0,
  ref_line = -6,
  caption = deparse(sunab_fml3)
)


sn_sunab_covars_glmweights <- plot_all_together_did(
  subg_col = "b10_pdsin3t1",
  df = test_dats_sn_withweights,
  backend = "sunab",
  backend_args = sunab_args3_glmweights,
  title = "Sierra Nevada test: Sun-Ab, twfe + covars, GLM weights"
)


sn_sunab_covars_gbmweights <- plot_all_together_did(
  subg_col = "b10_pdsin3t1",
  df = test_dats_sn_withweights,
  backend = "sunab",
  backend_args = sunab_args3_gbmweights,
  title = "Sierra Nevada test: Sun-Ab, twfe + covars, GBM weights"
)




# No covars
sunab_args2_glmweights <- list(
  fml = sunab_fml2,
  vcov = ~ us_l4code,
  model_fun = fixest::feols,
  model_args = list(
    mem.clean = TRUE,
    lean = FALSE,
    weights = ~ glm_weights
  ),
  term_pattern = "^year::",
  time_extract_regex = "-?\\d+",
  treatment_time = 0,
  ref_line = -6,
  caption = deparse(sunab_fml2)
)


sn_sunab_nocovars_glmweights <- plot_all_together_did(
  subg_col = "b10_pdsin3t1",
  df = test_dats_sn_withweights,
  backend = "sunab",
  backend_args = sunab_args2_glmweights,
  title = "Sierra Nevada test: Sun-Ab, twfe, GLM weights"
)



sunab_args2_gbmweights <- list(
  fml = sunab_fml2,
  vcov = ~ us_l4code,
  model_fun = fixest::feols,
  model_args = list(
    mem.clean = TRUE,
    lean = FALSE,
    weights = ~ gbm_weights
  ),
  term_pattern = "^year::",
  time_extract_regex = "-?\\d+",
  treatment_time = 0,
  ref_line = -6,
  caption = deparse(sunab_fml2)
)


sn_sunab_nocovars_gbmweights <- plot_all_together_did(
  subg_col = "b10_pdsin3t1",
  df = test_dats_sn_withweights,
  backend = "sunab",
  backend_args = sunab_args2_gbmweights,
  title = "Sierra Nevada test: Sun-Ab, twfe, GBM weights"
)


```



# Compare plots

```{r}

sn_sunab_p1 + ylim(-50, 20)
sn_sunab_p2 + ylim(-50, 20)
sn_sunab_p3 + ylim(-50, 20)

sn_twfe_p1 + ylim(-50, 20)
sn_twfe_p3 + ylim(-50, 20)

sn_cs_p1 + ylim(-50, 20)
sn_cs_p2 + ylim(-50, 20)
sn_cs_p2_nyt + ylim(-50, 20)

sn_cs_weighted_covars_glm + ylim(-50, 20)
sn_cs_weighted_covars_gbm + ylim(-50, 20)

sn_sunab_nocovars_glmweights + ylim(-50, 20)
sn_sunab_nocovars_gbmweights + ylim(-50, 20)

sn_sunab_covars_glmweights + ylim(-50, 20)
sn_sunab_covars_gbmweights + ylim(-50, 20)

sn_f_cs_p + ylim(-50, 20)
sn_f_twfe_p + ylim(-50, 20)
sn_bf_cs_p + ylim(-50, 20)
sn_bf_twfe_p + ylim(-50, 20)

ggsave(here(dir_weight_test, "sn_cs_weighted_covars_glm.png"), sn_cs_weighted_covars_glm, width = 8, height = 5, dpi = 300)
ggsave(here(dir_weight_test, "sn_cs_weighted_covars_gbm.png"), sn_cs_weighted_covars_gbm, width = 8, height = 5, dpi = 300)


ggsave(here(dir_weight_test, "sn_twfe_nocovars.png"), sn_twfe_p1, width = 8, height = 5, dpi = 300)
ggsave(here(dir_weight_test, "sn_twfe_covars.png"), sn_twfe_p3, width = 8, height = 5, dpi = 300)

ggsave(here(dir_weight_test, "sn_cs_UNweighted_nocovars.png"), sn_cs_p1, width = 8, height = 5, dpi = 300)
ggsave(here(dir_weight_test, "sn_cs_UNweighted_covars.png"), sn_cs_p2, width = 8, height = 5, dpi = 300)

ggsave(plot = sn_sunab_p2, here(dir_weight_test, "sn_sunab_nocovars.png"), width = 8, height = 5, dpi = 300)
ggsave(plot = sn_sunab_p3, here(dir_weight_test, "sn_sunab_covars.png"), width = 8, height = 5, dpi = 300)

ggsave(plot = sn_sunab_covars_glmweights, here(dir_weight_test, "sn_sunab_covars_weighted_glm.png"), width = 8, height = 5, dpi = 300)
ggsave(plot = sn_sunab_covars_gbmweights, here(dir_weight_test, "sn_sunab_covars_weighted_gbm.png"), width = 8, height = 5, dpi = 300)

```

<button class="accordion-button" type="button" data-bs-toggle="collapse" data-bs-target="#collapseOne" >Click to see Session Information</button><div id="collapseOne" class="accordion-collapse collapse"><div>

```{r sessionInfo}
sessionInfo()
```