---
title: "02_prep_data"
author: Tyler L. McIntosh
date: last-modified
format:
  html:
    embed-resources: true
    html-math-method: katex
    css: styles.css
    toc: true
    toc-depth: 1
    toc-title: Contents
    toc-location: right
    code-fold: true
    code-summary: "Show the code"
---

# Setup and read in data

``` {r setupchunk}
#| echo: true
#| message: false
#| warning: false
#| results: 'hide'

rm(list = ls())

if(!requireNamespace("here", quietly = TRUE)) {
  install.packages("here")
}
library(here)

source(here::here("code", "functions.R"))

install_and_load_packages(c(
  "tidyverse",
  "terra",
  "mapview",
  "jsonlite",
  "sf",
  "janitor",
  "data.table",
  "grf",
  "tictoc"))

dir_figs <- here::here('figs')
dir_derived <- here::here('data', 'derived')
dir_raw <- here::here('data', 'raw')
dir_manual <- here::here('data', 'manual')
dir_ensure(c(dir_figs,
             dir_derived,
             dir_manual,
             dir_raw))



# Load data
dats <- readr::read_csv(here::here(dir_manual, "forest_fire_points_SRockies_250.csv"))
biotic_sev <- terra::rast(here::here(dir_derived, "biotic_gridded_1km_all_years_severity.tif"))

# Crop biotic to s_rockies for now
eco_l3 <- access_data_epa_l3_ecoregions_vsi()
s_rockies <- eco_l3 |>
  filter(str_detect(US_L3NAME, "Southern Rockies"))
biotic_sev <- biotic_sev |>
  terra::crop(sf::st_bbox(s_rockies |> sf::st_transform(terra::crs(biotic_sev))))

epsg <- terra::crs(biotic_sev)


```


# THINGS TO ADD
- Fire event ID -- need fixed effect
- Day of ignition

# Format data from GEE so it's more friendly to use

```{r data-format}

#glimpse(dats)


# Remove all instances with re-burns
dats_no_reburn <- dats %>%
  filter(
    rowSums(!is.na(select(., starts_with("cbi_year")))) <= 1
  )

# Get burn year & burn year bcCBI
  # Identify CBI columns and corresponding years
  cbi_cols <- grep("^cbi_year_\\d{4}$", names(dats_no_reburn), value = TRUE)
  years <- as.integer(sub(".*_(\\d{4})$", "\\1", cbi_cols))
  
  # Extract the relevant subset
  cbi_matrix <- as.data.frame(dats_no_reburn[cbi_cols], stringsAsFactors = FALSE)
  
  # Count non-NA values per row
  non_na_counts <- rowSums(!is.na(cbi_matrix))
  
  # Rows with exactly one non-NA value
  valid_idx <- which(non_na_counts == 1)
  
  # Get the column index (position) of the non-NA value
  non_na_col_index <- max.col(!is.na(cbi_matrix[valid_idx, , drop = FALSE]), ties.method = "first")
  
  # Initialize result columns
  dats_no_reburn$burn_year <- NA_integer_
  dats_no_reburn$burn_year_cbi <- NA_real_
  
  # Assign values for valid rows only
  dats_no_reburn$burn_year[valid_idx] <- years[non_na_col_index]
  dats_no_reburn$burn_year_cbi[valid_idx] <- cbi_matrix[cbind(valid_idx, non_na_col_index)]
  
  rm(cbi_matrix)

# Remove CBI data now that we don't need it, filter out a few that have NAs
  ### WHY ARE THERE A FEW NAs?
dats_no_reburn <- dats_no_reburn |>
  select(-starts_with("cbi_year")) |>
  filter(!is.na(burn_year))

# Make spatial object
# Extract coordinates from JSON strings
dats_no_reburn <- dats_no_reburn |>
  mutate(
    coords = map(.geo, ~ parse_json(.x)$coordinates),
    x = map_dbl(coords, 1),
    y = map_dbl(coords, 2)
  )

# convert to sf
dats_no_reburn_sf <- sf::st_as_sf(dats_no_reburn, coords = c("x", "y"), crs = 4326) |>
  sf::st_transform(epsg)

```

# Extract biotic data at samples

```{r extract-biotic}

dats_no_reburn_vect <- terra::vect(dats_no_reburn_sf)

dats_with_biotic <- terra::extract(x = biotic_sev,
                                        y = dats_no_reburn_vect,
                                        bind = TRUE) |>
  sf::st_as_sf() |>
  sf::st_drop_geometry()



```

# Continue working with data now that we have biotic data

```{r multi-disturbance-data-format}


# Only keep biotic severity values >= 25 and remove multi-biotic
dats_with_biotic_new <- dats_with_biotic %>%
  mutate(across(starts_with("biotic_"), ~ ifelse(. < 25, NA, .))) %>%
  filter(
    rowSums(!is.na(select(., starts_with("biotic_")))) <= 1
  )


# Get biotic year & biotic severity
  # Identify CBI columns and corresponding years
  biotic_cols <- grep("^biotic_\\d{4}$", names(dats_with_biotic_new), value = TRUE)
  years <- as.integer(sub(".*_(\\d{4})$", "\\1", biotic_cols))
  
  # Extract the relevant subset
  biotic_matrix <- as.data.frame(dats_with_biotic_new[biotic_cols], stringsAsFactors = FALSE)
  
  # Count non-NA values per row
  non_na_counts <- rowSums(!is.na(biotic_matrix))
  
  # Rows with exactly one non-NA value
  valid_idx <- which(non_na_counts == 1)
  
  # Get the column index (position) of the non-NA value
  non_na_col_index <- max.col(!is.na(biotic_matrix[valid_idx, , drop = FALSE]), ties.method = "first")
  
  # Initialize result columns
  dats_with_biotic_new$bio_year <- NA_integer_
  dats_with_biotic_new$bio_year_sev <- NA_real_
  
  # Assign values for valid rows only
  dats_with_biotic_new$bio_year[valid_idx] <- years[non_na_col_index]
  dats_with_biotic_new$bio_year_sev[valid_idx] <- biotic_matrix[cbind(valid_idx, non_na_col_index)]
  
  rm(biotic_matrix)

# Remove CBI data now that we don't need it, filter out a few that have NAs
  ### WHY ARE THERE A FEW NAs?
dats_with_biotic_new <- dats_with_biotic_new |>
  select(-starts_with("biotic_")) |>
  rename(biotic_year = bio_year,
         biotic_year_sev = bio_year_sev) |>
  mutate(biotic_fire_gap = biotic_year - burn_year)

#Filter out those with FRG = 3 - not useful

dats_with_biotic_new <- dats_with_biotic_new |>
  filter(frg_reclass != 3)

hist(dats_with_biotic_new$biotic_year_sev)
hist(dats_with_biotic_new$biotic_year)

hist(dats_with_biotic_new$biotic_fire_gap)

```

# Continue working with data to manage response variable

```{r response-format}


# A function to calculate the average rap tree cover percent for a time period pre or post fire
# Uses data.table for efficiency
# If all columns aren't available, provide NA (i.e. avoid averaging only one or two values if 3 are requested)

# compute_rap_offset_avg_dt <- function(dt, offsets, new_col_name, prefix = "rap_TRE_") {
#   # Ensure dt is a data.table
#   dt <- as.data.table(copy(dt))
#   dt[, id__temp := .I]  # Create unique ID for merge
# 
#   # Melt relevant columns (e.g., "rap_TRE_1985" to "rap_TRE_2022")
#   melt_cols <- grep(paste0("^", prefix, "\\d{4}$"), names(dt), value = TRUE)
#   
#   long_dt <- melt(
#     dt,
#     id.vars = c("id__temp", "burn_year"),
#     measure.vars = melt_cols,
#     variable.name = "year_col",
#     value.name = "value"
#   )
# 
#   # Extract year from column name
#   long_dt[, year := as.integer(sub(prefix, "", year_col))]
# 
#   # Calculate offset from burn year
#   long_dt[, offset := year - burn_year]
# 
#   # Keep only rows with offset in desired range
#   long_dt <- long_dt[offset %in% offsets]
# 
#   # Summarise by id only if there are ≥ 2 non-NA values
#   summary_dt <- long_dt[!is.na(value), .(
#     n = .N,
#     avg = mean(value, na.rm = TRUE)
#   ), by = id__temp][n >= 2, .(id__temp, avg)]
# 
#   # Join back and assign new column
#   dt <- merge(dt, summary_dt, by = "id__temp", all.x = TRUE)
#   setnames(dt, "avg", new_col_name)
#   dt[, id__temp := NULL]
# 
#   return(as_tibble(dt))
# }
# 
compute_rap_offset_avg_dt <- function(dt, offsets, new_col_name, prefix = "rap_TRE_") {
  # Ensure dt is a data.table
  dt <- as.data.table(copy(dt))
  dt[, id__temp := .I]  # Create unique ID for merge

  # Melt relevant columns
  melt_cols <- grep(paste0("^", prefix, "\\d{4}$"), names(dt), value = TRUE)

  long_dt <- melt(
    dt,
    id.vars = c("id__temp", "burn_year"),
    measure.vars = melt_cols,
    variable.name = "year_col",
    value.name = "value"
  )

  # Extract year from column name
  long_dt[, year := as.integer(sub(prefix, "", year_col))]

  # Calculate offset from burn year
  long_dt[, offset := year - burn_year]

  # Only keep rows with desired offsets
  long_dt <- long_dt[offset %in% offsets]

  # Now, summarize carefully
  summary_dt <- long_dt[, .(
    n_available = sum(!is.na(value)),
    n_expected = length(offsets),
    avg = if (sum(!is.na(value)) == length(offsets)) mean(value, na.rm = TRUE) else NA_real_
  ), by = id__temp]

  # Keep only id and avg
  summary_dt <- summary_dt[, .(id__temp, avg)]

  # Merge back
  dt <- merge(dt, summary_dt, by = "id__temp", all.x = TRUE)
  setnames(dt, "avg", new_col_name)
  dt[, id__temp := NULL]

  return(as_tibble(dt))
}



tic()
dats_response_clean <- dats_with_biotic_new |>
  compute_rap_offset_avg_dt(4:6, "avg_rap_TRE_5_yr_post_fire") |>
  compute_rap_offset_avg_dt(9:11, "avg_rap_TRE_10_yr_post_fire") |>
  compute_rap_offset_avg_dt(14:16, "avg_rap_TRE_15_yr_post_fire") |>
  compute_rap_offset_avg_dt(-7:-5, "avg_rap_TRE_6_yr_pre_fire")
toc()

hist(dats_response_clean$avg_rap_TRE_5_yr_post_fire)
hist(dats_response_clean$avg_rap_TRE_10_yr_post_fire)
hist(dats_response_clean$avg_rap_TRE_15_yr_post_fire)
hist(dats_response_clean$avg_rap_TRE_6_yr_pre_fire)

dats_response_clean <- dats_response_clean |>
  dplyr::mutate(post_fire_TRE_difference_5_yr = avg_rap_TRE_5_yr_post_fire - avg_rap_TRE_6_yr_pre_fire,
                post_fire_TRE_difference_10_yr = avg_rap_TRE_10_yr_post_fire - avg_rap_TRE_6_yr_pre_fire,
                post_fire_TRE_difference_15_yr = avg_rap_TRE_15_yr_post_fire - avg_rap_TRE_6_yr_pre_fire)


hist(dats_response_clean$post_fire_TRE_difference_5_yr)
hist(dats_response_clean$post_fire_TRE_difference_10_yr)
hist(dats_response_clean$post_fire_TRE_difference_15_yr)


dats_response_clean <- dats_response_clean %>%
  mutate(
    biotic_fire_cat = case_when(
      biotic_fire_gap >= -10 & biotic_fire_gap <= -6 ~ "biotic_6_10_prior",
      biotic_fire_gap >= -5  & biotic_fire_gap <= 0  ~ "biotic_0_5_prior",
      biotic_fire_gap >= 1   & biotic_fire_gap <= 5  ~ "biotic_1_5_after",
      biotic_fire_gap >= 6   & biotic_fire_gap <= 10  ~ "biotic_6_10_after",
      TRUE ~ "no_biotic_or_outside_timeframe"
    ),
    biotic_fire_cat = factor(
      biotic_fire_cat,
      levels = c(
        "no_biotic_or_outside_timeframe",
        "biotic_6_10_prior",
        "biotic_0_5_prior",
        "biotic_1_5_after",
        "biotic_6_10_after"
      )
    ),
    burn_year_sev = case_when(
      burn_year_cbi >= 2.25 ~ "high-sev",
      burn_year_cbi < 0.1 ~ "unburned",
      burn_year_cbi >= 0.1 & burn_year_cbi < 2.25 ~ "low-moderate-sev"
    ),
    beneficial_fire = case_when(
      burn_year_sev == "high-sev" & frg_reclass == 2 ~ "replacement bf",
      burn_year_sev == "high-sev" & frg_reclass == 1 ~ "too severe",
      burn_year_sev == "low-moderate-sev" & frg_reclass == 2 ~ "not severe enough",
      burn_year_sev == "low-moderate-sev" & frg_reclass == 1 ~ "low sev bf",
    )
  )


compute_post_fire_rap_slope_dt <- function(
  dt,
  offsets = 2:10,
  req_yrs = 5,
  new_col_name = "post_fire_rap_slope",
  r2_col_name = "post_fire_rap_r2",
  prefix = "rap_TRE_"
) {
  dt <- as.data.table(copy(dt))
  dt[, id__temp := .I]  # Unique ID

  # Melt RAP TRE columns
  melt_cols <- grep(paste0("^", prefix, "\\d{4}$"), names(dt), value = TRUE)
  
  long_dt <- melt(
    dt,
    id.vars = c("id__temp", "burn_year"),
    measure.vars = melt_cols,
    variable.name = "year_col",
    value.name = "value"
  )

  # Extract year
  long_dt[, year := as.integer(sub(prefix, "", year_col))]

  # Calculate offset from burn year
  long_dt[, offset := year - burn_year]

  # Keep only desired offsets
  long_dt <- long_dt[offset %in% offsets]

  # Group and compute slope + R2 if enough data
  slope_dt <- long_dt[
    !is.na(value),
    {
      if (.N >= req_yrs) {
        model <- lm(value ~ offset)
        list(
          slope = coef(model)[2],
          r2 = summary(model)$r.squared
        )
      } else {
        list(
          slope = NA_real_,
          r2 = NA_real_
        )
      }
    },
    by = id__temp
  ]

  # Merge results back
  dt <- merge(dt, slope_dt, by = "id__temp", all.x = TRUE)
  setnames(dt, c("slope", "r2"), c(new_col_name, r2_col_name))
  dt[, id__temp := NULL]

  return(as_tibble(dt))
}



dats_response_clean <- dats_response_clean |>
  compute_post_fire_rap_slope_dt(
    offsets = 2:10,
    req_yrs = 9,
    new_col_name = "post_fire_rap_slope_2_10_yrs",
    r2_col_name = "post_fire_rap_2_10_yrs_r2",
    prefix = "rap_TRE_") |>
  compute_post_fire_rap_slope_dt(
    offsets = 2:15,
    req_yrs = 9,
    new_col_name = "post_fire_rap_slope_2_15_yrs",
    r2_col_name = "post_fire_rap_2_15_yrs_r2",
    prefix = "rap_TRE_")

hist(dats_response_clean$post_fire_rap_slope_2_10_yrs)
hist(dats_response_clean$post_fire_rap_2_10_yrs_r2)
hist(dats_response_clean$post_fire_rap_slope_2_15_yrs)
hist(dats_response_clean$post_fire_rap_2_15_yrs_r2)


# Create long version of data
dats_response_clean_long <- dats_response_clean |>
  pivot_longer(
    cols = starts_with("rap_TRE_"),
    names_to = "year",
    names_prefix = "rap_TRE_",
    names_transform = list(year = as.integer),
    values_to = "rap_TRE"
  ) |>
  mutate(year_from_fire_index = year - burn_year)

dats_response_clean_short <- dats_response_clean |>
  select(-starts_with("rap_TRE"))

```


# Create long-form cross-tabulation

```{r n-tables}


# Basic counts
create_cross_tab <- function(dats) {
    
  summary_table <- dats %>%
    count(frg_reclass, biotic_fire_cat, burn_year_sev) %>%
    mutate(across(c(frg_reclass, biotic_fire_cat, burn_year_sev), as.character))
  
  # Subtotals
  subtotal_frg_reclass <- dats %>%
    count(frg_reclass) %>%
    mutate(
      frg_reclass = as.character(frg_reclass),
      biotic_fire_cat = "ALL",
      burn_year_sev = "ALL"
    )
  
  subtotal_biotic_fire_cat <- dats %>%
    count(biotic_fire_cat) %>%
    mutate(
      frg_reclass = "ALL",
      biotic_fire_cat = as.character(biotic_fire_cat),
      burn_year_sev = "ALL"
    )
  
  subtotal_burn_year_sev <- dats %>%
    count(burn_year_sev) %>%
    mutate(
      frg_reclass = "ALL",
      biotic_fire_cat = "ALL",
      burn_year_sev = as.character(burn_year_sev)
    )
  
  # Grand total
  grand_total <- dats %>%
    summarise(n = n()) %>%
    mutate(
      frg_reclass = "ALL",
      biotic_fire_cat = "ALL",
      burn_year_sev = "ALL"
    )
  
  # Combine all together
  full_summary_table <- bind_rows(
    summary_table,
    subtotal_frg_reclass,
    subtotal_biotic_fire_cat,
    subtotal_burn_year_sev,
    grand_total
  ) %>%
    arrange(frg_reclass, biotic_fire_cat, burn_year_sev)
    
  return(full_summary_table)
}

full_summary_table <- create_cross_tab(dats_response_clean_short)
high_sev_biotic_summary_table <- create_cross_tab(dats_response_clean_short |> filter(biotic_year_sev >= 50))

print(full_summary_table, n = Inf)
print(high_sev_biotic_summary_table, n = Inf)

readr::write_csv(full_summary_table,
                 append = FALSE,
                 file = here::here(dir_derived, "cross_tabs.csv"))

```


# Write data for analysis

```{r write-out}

dats_response_clean_short_sf <- dats_response_clean_short |>
  left_join(dats_no_reburn_sf |>
              select(`system:index`),
            by = join_by(`system:index`)) |>
  sf::st_as_sf()


readr::write_csv(dats_response_clean_short,
                 file = here::here(dir_derived, "ard_short_v1.csv"))
sf::st_write(dats_response_clean_short_sf,
             here::here(dir_derived, "ard_short_sf_v1.gpkg"),
             append = FALSE)
readr::write_csv(dats_response_clean_long,
                 file = here::here(dir_derived, "ard_long_v1.csv"))
```


<button class="accordion-button" type="button" data-bs-toggle="collapse" data-bs-target="#collapseOne" >Click to see Session Information</button><div id="collapseOne" class="accordion-collapse collapse"><div>

```{r sessionInfo}
sessionInfo()
```

</div></div>