---
title: "03_prep_data"
author: Tyler L. McIntosh
date: last-modified
format:
  html:
    embed-resources: true
    html-math-method: katex
    css: styles.css
    toc: true
    toc-depth: 1
    toc-title: Contents
    toc-location: right
    code-fold: true
    code-summary: "Show the code"
---

# Variable interpretation reminders
- TPI: low = valleys, high = ridges
- CHILI: low = cold/wet, high = hot/dry
- SRTM: low = low elevation, high = high elevation
- AET: low = wet, high = dry
- DEF: low = wet, high = dry
- PDSI: low = dry, high = wet **************


# Setup and read in data

``` {r setupchunk}
#| echo: true
#| message: false
#| warning: false
#| results: 'hide'

rm(list = ls())

if(!requireNamespace("here", quietly = TRUE)) {
  install.packages("here")
}
library(here)

source(here::here("utils", "functions.R"))

install_load_packages(c(
  "tidyverse",
  "terra",
  "mapview",
  "jsonlite",
  "sf",
  "janitor",
  "data.table",
  "googledrive",
  "mblm",
  "arrow",
  "forcats",
  "tictoc",
  "glue",
  "rlang",
  "tibble",
  "future",
  "furrr",
  "readr",
  "vroom",
  "fs",
  "data.table"))

set.seed(1234)
gdrive_folder <- "GEE_resilience_v6_operational_ss500_ts50000"
service_account_file <- here('config', 'secrets', 'tymc5571-utils-project-692f27c034dd.json')

dir_figs <- here::here('figs')
dir_derived <- here::here('data', 'derived', gdrive_folder)
dir_raw <- here::here('data', 'raw')
dir_manual <- here::here('data', 'manual')
dir_ensure(c(dir_figs,
             dir_derived,
             dir_manual,
             dir_raw))

```

# Regions

```{r}

forested_ecoregions <- tibble(
  na_l3name = c(
    "Blue Mountains",
    "Cascades",
    "Coast Range",
    "Eastern Cascades Slopes and Foothills",
    "Klamath Mountains",
    "North Cascades",
    "Straight of Georgia/Puget Lowland",
    "Willamette Valley",
    "California Coastal Sage, Chaparral, and Oak Woodlands",
    "Sierra Nevada",
    "Southern and Baja California Pine-Oak Mountains",
    "Canadian Rockies",
    "Idaho Batholith",
    "Middle Rockies",
    "Columbia Mountains/Northern Rockies",
    "Southern Rockies",
    "Wasatch and Uinta Mountains",
    "Arizona/New Mexico Mountains",
    "Colorado Plateaus"
  ),
  na_l3code = c(
    "6.2.9", "6.2.7", "7.1.8", "6.2.8", "6.2.11", "6.2.5", "7.1.7",
    "7.1.9", "11.1.1", "6.2.12", "11.1.3", "6.2.4", "6.2.15",
    "6.2.10", "6.2.3", "6.2.14", "6.2.13", "13.1.1", "10.1.6"
  ),
  short_name = c(
    "Blue Mtns", "Cascades", "Coast Range", "Eastern Cascades",
    "Klamath Mtns", "North Cascades", "Puget Lowland",
    "Willamette Valley", "Central California Mtns", "Sierra Nevada",
    "Southern California Mtns", "Canadian Rockies", "Idaho Batholith",
    "Middle Rockies", "Northern Rockies", "Southern Rockies",
    "Wasatch Uinta Mtns", "AZ/NM Mtns", "Colorado Plateaus"
  ),
  region = c(
    "Pacific Northwest", "Pacific Northwest", "Pacific Northwest",
    "Pacific Northwest", "Pacific Northwest", "Pacific Northwest",
    "Pacific Northwest", "Pacific Northwest",
    "California", "California", "California",
    "Upper Rockies", "Upper Rockies", "Upper Rockies", "Upper Rockies",
    "Lower Rockies", "Lower Rockies",
    "Southwest", "Southwest"
  ),
  code_name = c(
    "bluemtns", "cascades", "coastrange", "eastcascades",
    "klamathmtns", "northcascades", "pugetlowland",
    "willamettevalley", "centralcaliforniamtns", "sierranevada",
    "southerncaliforniamtns", "canadianrockies", "idahobatholith",
    "middlerockies", "northernrockies", "southernrockies",
    "wasatchuintamtns", "aznmmtns", "coloradoplateaus"
  ),
  us_l3code = c(
    "11", "4", "1", "9", "78", "77", "2", "3", "6", "5", "8",
    "41", "16", "17", "15", "21", "19", "23", "20"
  )
)



```


# Load data from GDrive

```{r}
# #### Run to authenticate gdrive access - if receive scope permissions issues, de-authenticate (drive_deauth()) and try again
# drive_auth(
#   scopes = "https://www.googleapis.com/auth/drive.readonly"
# )
# ####
# 
# 
# for(er in forested_ecoregions$code_name) {
#   print(er)
#   er_dats <- read_matching_csvs_from_gdrive(drive_folder = gdrive_folder,
#                                             pattern = paste0("fc_tile_", er),
#                                             force_cols = c("us_l4code"),
#                                             force_types = c("character"))
#   write_csv(er_dats, file = here(dir_derived, paste0("all_", er, ".csv")))
# }
# 
# all_er_csvs <- list.files(path = dir_derived, full.names = TRUE)
# 
# 
# #There is a problem where some instances of '1's with NAs in the CBI columns can end up being treated as logical; ensure doesn't happen
# read_csv_ensure_cbi_dbl <- function(path) {
#   # Preview column names without reading all data
#   col_names <- names(read_csv(path, n_max = 0))
#   
#   # Find columns that start with "CBI"
#   cbi_cols <- grep("^cbi", col_names, value = TRUE)
#   
#   # Create a list of column specs: all CBI columns as double
#   col_spec <- do.call(cols, c(
#     setNames(rep(list(col_double()), length(cbi_cols)), cbi_cols),
#     .default = col_guess()
#   ))
#   
#   # Read the CSV with specified col_types
#   dats <- read_csv(path, col_types = col_spec)
#   
#   return(dats)
# }
# 
# 
# #Create full raw dats file
# all_dats_fl <- here(dir_derived, "all_dats.parquet")
# files <- list.files(path = dir_derived, full.names = TRUE)
# 
# dats <- list()
# for(erf in files) {
#   erf_dats <- read_csv_ensure_cbi_dbl(erf)
#   dats$erf <- erf_dats
# }
# dats <- bind_rows(dats)
# write_parquet(dats, sink = all_dats_fl)
# 
# 
# 
# 
# 
# Use local download
all_files <- list.files(dir_derived, full.names = TRUE, pattern = ".csv")
files_with_data <- all_files[fs::file_size(all_files) > 2]
all_dats <- lapply(files_with_data, fread) |>
  bind_rows()

arrow::write_parquet(x = all_dats, sink = here(dir_derived, "merged_dats_raw.parquet"))
write_csv(x = all_dats, file = here(dir_derived, "merged_dats_raw.csv"))
```

# Examine GEE data & clean NAs

```{r}

dats <- arrow::read_parquet(here(dir_derived, "merged_dats.parquet"))
glimpse(dats)

# check NA counts
#colSums(is.na(dats))

# Almost all rows being dropped are those with NFG == NA; could fix most of these by backfilling NFG
dats <- dats |>
  #select(-rand) |>
  tidyr::drop_na()

# Check strange CBI negative values
check_cbi <- dats |>
  select(starts_with("cbi_")) |>
  pivot_longer(
    cols = everything(),
    names_to = "variable",
    values_to = "value"
  ) |>
  filter(value != -1)

# Count of negative CBI values that aren't -1: #3678 - WHERE IS THIS COMING FROM?
check_cbi |>
  summarise(n_negative = sum(value < 0, na.rm = TRUE))

# check_cbi |>
#   ggplot(aes(x = value)) +
#   geom_density() +
#   labs(
#     x = "CBI value",
#     y = "Density",
#     title = "Density of all CBI-year values (excluding -1)"
#   )

# for now, remove all rows with values that are - but not -1 in CBI
dats <- dats |>
  filter(
    !if_any(
      starts_with("cbi_year"),
      ~ .x < 0 & .x != -1
    )
  )

```

# Deal with fire and fire heterogeneity data

```{r}

# replace -1 signifier in CBI bands with NA
dats <- dats %>%
  mutate(across(starts_with("cbi_"), ~ na_if(.x, -1)))

# Remove all instances with re-burns
dats_no_reburn <- dats %>%
  filter(
    rowSums(!is.na(select(., starts_with("cbi_")))) <= 1
  )

# Get burn year & burn year bcCBI
# Identify CBI columns and corresponding years
cbi_cols <- grep("^cbi_\\d{4}$", names(dats_no_reburn), value = TRUE)
years <- as.integer(sub(".*_(\\d{4})$", "\\1", cbi_cols))

# Extract the relevant subset (data.table-safe)
cbi_matrix <- as.data.frame(dats_no_reburn[, ..cbi_cols], stringsAsFactors = FALSE)

# Count non-NA values per row
non_na_counts <- rowSums(!is.na(cbi_matrix))

# Rows with exactly one non-NA value
valid_idx <- which(non_na_counts == 1)

# Get the column index (position) of the non-NA value
non_na_col_index <- max.col(!is.na(cbi_matrix[valid_idx, , drop = FALSE]), ties.method = "first")

# Initialize result columns
dats_no_reburn$burn_year <- NA_integer_
dats_no_reburn$burn_year_cbi <- NA_real_

# Assign values for valid rows only
dats_no_reburn$burn_year[valid_idx] <- years[non_na_col_index]
dats_no_reburn$burn_year_cbi[valid_idx] <- cbi_matrix[cbind(valid_idx, non_na_col_index)]

rm(cbi_matrix)



# REMOVE 'YR' and "YEAR" from names
names(dats_no_reburn) <- gsub("(_yr_|_year_)", "_", names(dats_no_reburn))
# 
# # GET THE OTHER CBI CONTEXT DATA
# 
# # Function to extract burn-year-aligned values from year-tagged columns
# extract_burn_year_value <- function(df, burn_year, prefix) {
#   # Get matching columns like 'cbi_1kmmwmedian_2008'
#   pattern <- paste0("^", prefix, "_\\d{4}$")
#   cols <- grep(pattern, names(df), value = TRUE)
#   
#   # Extract years from column names
#   years <- as.integer(sub(".*_(\\d{4})$", "\\1", cols))
#   
#   # Create value matrix (rows = obs, cols = years)
#   value_mat <- as.data.frame(df[cols], stringsAsFactors = FALSE)
#   
#   # Create lookup index: for each row, match burn_year to column year
#   col_index <- match(burn_year, years)
#   
#   # Use cbind(row, col) indexing to extract values
#   result <- mapply(function(row, col) {
#     if (!is.na(col)) value_mat[row, col] else NA
#   }, row = seq_along(burn_year), col = col_index)
#   
#   return(result)
# }
# 
# # Apply the function to create the new columns
# dats_no_reburn$burn_cbi_500mmwmean <- extract_burn_year_value(
#   dats_no_reburn,
#   burn_year = dats_no_reburn$burn_year,
#   prefix = "cbi_500mmwmean"
# )
# 
# dats_no_reburn$burn_cbi_500mmwstd <- extract_burn_year_value(
#   dats_no_reburn,
#   burn_year = dats_no_reburn$burn_year,
#   prefix = "cbi_500mmwstd"
# )
# 
# dats_no_reburn$burn_distance_forest_after <- extract_burn_year_value(
#   dats_no_reburn,
#   burn_year = dats_no_reburn$burn_year,
#   prefix = "distance_forest_after"
# )

# Generate mock_burn_year
dats_no_reburn$mock_burn_year <- dats_no_reburn$burn_year

na_idx <- which(is.na(dats_no_reburn$mock_burn_year))
dats_no_reburn$mock_burn_year[na_idx] <- sample(1984:2020, length(na_idx), replace = TRUE)


# RESOLVED THESE
# # Replace NAs with 0 in specified columns
# cols_to_fill <- c("burn_cbi", "burn_cbi_1kmmwmedian", "burn_cbi_1kmmwstd")
# 
# for (col in cols_to_fill) {
#   dats_no_reburn[[col]][is.na(dats_no_reburn[[col]])] <- 0
# }

# # Remove the few rows that still have the strange fire but no burn year phenomenon
# x <- dats_no_reburn |>
#   filter(!(fire == 1 & is.na(burn_year)))

```



# Rescale data and deal with factors etc

```{r}

dats_no_reburn <- dats_no_reburn |>
  mutate(
    across(
      c(starts_with("vcf_tree_SD"), starts_with("pdsi_")),
      ~ .x * 0.01
    )
  )

# NFG metadata: https://data.fs.usda.gov/geodata/rastergateway/forest_type/conus_forest_type_group_metadata.php
nfg_lookup <- c(
  "100" = "White/Red/Jack Pine Group",
  "120" = "Spruce/Fir Group",
  "140" = "Longleaf/Slash Pine Group",
  "160" = "Loblolly/Shortleaf Pine Group",
  "180" = "Pinyon/Juniper Group",
  "200" = "Douglas-fir Group",
  "220" = "Ponderosa Pine Group",
  "240" = "Western White Pine Group",
  "260" = "Fir/Spruce/Mountain Hemlock Group",
  "280" = "Lodgepole Pine Group",
  "300" = "Hemlock/Sitka Spruce Group",
  "320" = "Western Larch Group",
  "340" = "Redwood Group",
  "360" = "Other Western Softwood Group",
  "370" = "California Mixed Conifer Group",
  "380" = "Exotic Softwoods Group",
  "400" = "Oak/Pine Group",
  "500" = "Oak/Hickory Group",
  "600" = "Oak/Gum/Cypress Group",
  "700" = "Elm/Ash/Cottonwood Group",
  "800" = "Maple/Beech/Birch Group",
  "900" = "Aspen/Birch Group",
  "910" = "Alder/Maple Group",
  "920" = "Western Oak Group",
  "940" = "Tanoak/Laurel Group",
  "950" = "Other Western Hardwoods Group",
  "980" = "Tropical Hardwoods Group",
  "990" = "Exotic Hardwoods Group"
)

dats_no_reburn <- dats_no_reburn |>
  mutate(
    nfg_factor = factor(nfg_lookup[as.character(nfg)])
  )

dats_no_reburn <- dats_no_reburn %>%
  mutate(
    burn_sev = case_when(
      fire == 1 & burn_cbi >= 2.25 ~ "High",
      fire == 1 & burn_cbi < 0.1 ~ "Negligible",
      fire == 1 & burn_cbi >= 0.1 & burn_cbi < 2.25 ~ "Low/Moderate",
      #fire == 1 & burn_cbi >= 0.1 & burn_cbi <  ~ "Low",
      #fire == 1 * burn_cbi >= & burn_cbi < 2.25 ~ "Moderate"
      TRUE ~ "Never"
    )
  )

dats_no_reburn <- dats_no_reburn


#Clean up factors & years as integers 
dats_no_reburn <- dats_no_reburn |>
  mutate(frg_reclass_desc = case_when(frg_reclass == 1 ~ "<35 return, low/mixed",
                                 frg_reclass == 2 ~ "<35 return, replacement",
                                 frg_reclass == 3 ~ "35-200 return, low/mixed",
                                 frg_reclass == 4 ~ "35-200 return replacement",
                                 frg_reclass == 5 ~ ">200 return any"),
         frg_reclass_desc = as.factor(frg_reclass_desc),
         burn_sev = fct_relevel(as.factor(burn_sev), c("Never", "Negligible", "Low/Moderate", "High")),
         burn_year = as.integer(burn_year),
         mock_burn_year = as.integer(mock_burn_year))

```


# Get biotic & drought sums and means in time before/after burn
```{r}

transform_annual_to_priorafter <- function(df, column_pattern, summary_fn_name, nyears) {
  # Validate the summary function
  if (!exists(summary_fn_name, mode = "function")) {
    stop("Summary function '", summary_fn_name, "' is not a valid function.")
  }
  summary_fn <- match.fun(summary_fn_name)
  
  # Identify columns matching pattern
  matched_cols <- grep(column_pattern, names(df), value = TRUE)
  if (length(matched_cols) == 0) return(df)
  
  # Extract base name
  base_name <- sub("^(.*?)_\\d{4}$", "\\1", matched_cols[1])
  
  # Map year to column names
  col_years <- as.integer(sub(".*_(\\d{4})$", "\\1", matched_cols))
  col_year_map <- setNames(matched_cols, col_years)
  
  # Initialize result vectors
  n <- nrow(df)
  prior_vals <- rep(NA_real_, n)
  after_vals <- rep(NA_real_, n)
  
  for (i in seq_len(n)) {
    byr <- df$mock_burn_year[i]
    if (!is.na(byr)) {
      # Prior 5 years
      yrs_prior <- (byr - nyears):(byr - 1)
      if (all(yrs_prior %in% col_years)) {
        cols <- col_year_map[as.character(yrs_prior)]
        vals <- unlist(df[i, ..cols])
        if (!any(is.na(vals))) prior_vals[i] <- summary_fn(vals, na.rm = TRUE)
      }
      
      # After 5 years
      yrs_after <- (byr + 1):(byr + nyears)
      if (all(yrs_after %in% col_years)) {
        cols <- col_year_map[as.character(yrs_after)]
        vals <- unlist(df[i, ..cols])
        if (!any(is.na(vals))) after_vals[i] <- summary_fn(vals, na.rm = TRUE)
      }
    }
  }
  
  # Construct column names with function name
  fn_suffix <- tolower(summary_fn_name)
  prior_col <- paste0(base_name, "_", nyears, "_yrs_prior_", fn_suffix)
  after_col <- paste0(base_name, "_", nyears, "_yrs_after_", fn_suffix)
  
  df[[prior_col]] <- prior_vals
  df[[after_col]] <- after_vals
  
  # DO NOT DO THIS ANYMORE TO ALLOW FOR NON-CD ANALYSES
  # # Filter out rows where either summary column is NA
  # df <- df[!is.na(df[[prior_col]]) & !is.na(df[[after_col]]), ]
  
  return(df)
}


n_beyond_threshold_x_yrs <- function(df, column_pattern, nyears, threshold) {
  # Validate and parse the threshold expression
  threshold_fn <- tryCatch(
    {
      eval(parse(text = paste0("function(x) x", threshold)))
    },
    error = function(e) stop("Invalid threshold expression: ", threshold)
  )
  
  # Identify columns matching the pattern
  matched_cols <- grep(column_pattern, names(df), value = TRUE)
  if (length(matched_cols) == 0) return(df)
  
  # Extract base name and year info
  base_name <- sub("^(.*?)_\\d{4}$", "\\1", matched_cols[1])
  col_years <- as.integer(sub(".*_(\\d{4})$", "\\1", matched_cols))
  col_year_map <- setNames(matched_cols, col_years)
  
  # Initialize result vectors
  n <- nrow(df)
  prior_counts <- integer(n)
  after_counts <- integer(n)
  
  for (i in seq_len(n)) {
    byr <- df$mock_burn_year[i]
    if (!is.na(byr)) {
      # Prior years
      yrs_prior <- (byr - nyears):(byr - 1)
      if (all(yrs_prior %in% col_years)) {
        cols <- col_year_map[as.character(yrs_prior)]
        vals <- unlist(df[i, ..cols])
        prior_counts[i] <- sum(threshold_fn(vals), na.rm = TRUE)
      } else {
        prior_counts[i] <- NA_integer_
      }
      
      # After years
      yrs_after <- (byr + 1):(byr + nyears)
      if (all(yrs_after %in% col_years)) {
        cols <- col_year_map[as.character(yrs_after)]
        vals <- unlist(df[i, ..cols])
        after_counts[i] <- sum(threshold_fn(vals), na.rm = TRUE)
      } else {
        after_counts[i] <- NA_integer_
      }
    } else {
      prior_counts[i] <- NA_integer_
      after_counts[i] <- NA_integer_
    }
  }
  
  # Create new column names
  prior_col <- paste0(base_name, "_", nyears, "_yrs_prior_threshold")
  after_col <- paste0(base_name, "_", nyears, "_yrs_after_threshold")
  
  # Append to df
  df[[prior_col]] <- prior_counts
  df[[after_col]] <- after_counts
  
  return(df)
}


dats_simple_added <- transform_annual_to_priorafter(
  df = dats_no_reburn,
  column_pattern = "^biotic_\\d{4}$",
  summary_fn_name = "sum",
  nyears = 5
)

dats_simple_added <- transform_annual_to_priorafter(
  df = dats_simple_added,
  column_pattern = "^hd_fingerprint_\\d{4}$",
  summary_fn_name = "sum",
  nyears = 5
)

dats_simple_added <- transform_annual_to_priorafter(
  df = dats_simple_added,
  column_pattern = "^hd_fingerprint_\\d{4}$",
  summary_fn_name = "mean",
  nyears = 5
)

dats_simple_added <- transform_annual_to_priorafter(
  df = dats_simple_added,
  column_pattern = "^pdsi_annual_\\d{4}$",
  summary_fn_name = "sum",
  nyears = 5
)

dats_simple_added <- transform_annual_to_priorafter(
  df = dats_simple_added,
  column_pattern = "^pdsi_annual_\\d{4}$",
  summary_fn_name = "mean",
  nyears = 5
)

# dats_simple_added <- transform_annual_to_priorafter(
#   df = dats_simple_added,
#   column_pattern = "^pdsi_summer_\\d{4}$",
#   summary_fn_name = "sum",
#   nyears = 5
# )
# 
# dats_simple_added <- transform_annual_to_priorafter(
#   df = dats_simple_added,
#   column_pattern = "^pdsi_summer_\\d{4}$",
#   summary_fn_name = "mean",
#   nyears = 5
# )



# Add thresholded data
dats_simple_added <- n_beyond_threshold_x_yrs(
  df = dats_simple_added,
  column_pattern = "^pdsi_annual_\\d{4}$",
  nyears = 5,
  threshold = "<=-4"
)

# dats_simple_added <- n_beyond_threshold_x_yrs(
#   df = dats_simple_added,
#   column_pattern = "^pdsi_summer_\\d{4}$",
#   nyears = 5,
#   threshold = "<=-4"
# )


dats_simple_added <- n_beyond_threshold_x_yrs(
  df = dats_simple_added,
  column_pattern = "^hd_fingerprint_\\d{4}$",
  nyears = 5,
  threshold = ">=6"
)


# hist(dats_simple_added$hd_fingerprint_5_yrs_prior_threshold)
# hist(dats_simple_added$hd_fingerprint_5_yrs_after_threshold)
# hist(dats_simple_added$pdsi_annual_5_yrs_prior_threshold)
# hist(dats_simple_added$pdsi_annual_5_yrs_after_threshold)

```


# Deal with response data (VCF & RAP) - GAMs

```{r}


compute_gam_stats <- function(
  df,
  column_prefix,
  reference_time_col,
  k = 10,
  min_n = 8,
  post_min_search_max_year = Inf,
  gam_method = "REML",
  parallel = FALSE,
  n_cores = max(1L, parallel::detectCores() - 1L),
  debug = FALSE
) {
  stopifnot(is.data.frame(df))
  if (!requireNamespace("data.table", quietly = TRUE)) stop("Please install 'data.table'.")
  if (!requireNamespace("mgcv", quietly = TRUE)) stop("Please install 'mgcv'.")
  if (!requireNamespace("tibble", quietly = TRUE)) stop("Please install 'tibble'.")

  dt <- data.table::as.data.table(df)

  if (!reference_time_col %in% names(dt)) {
    stop("reference_time_col not found in df: ", reference_time_col)
  }

  ts_cols <- grep(paste0("^", column_prefix, "\\d{4}$"), names(dt), value = TRUE)
  if (length(ts_cols) == 0) {
    stop("No time-series columns found matching prefix+YYYY for prefix: ", column_prefix)
  }

  prefix_tag <- sub("_$", "", column_prefix)

  nm_pre6_fit <- paste0("gam_", prefix_tag, "_pre6_fit")
  nm_post_min <- paste0("gam_", prefix_tag, "_post_min")
  nm_post_min_year <- paste0("gam_", prefix_tag, "_post_min_year")
  nm_yrs_ref_to_post_min <- paste0("gam_", prefix_tag, "_yrs_ref_to_post_min")
  nm_diff_pre6_to_post_min <- paste0("gam_", prefix_tag, "_diff_pre6_to_post_min")

  nm_fit_ref_p10 <- paste0("gam_", prefix_tag, "_fit_ref_p10")
  nm_fit_ref_p15 <- paste0("gam_", prefix_tag, "_fit_ref_p15")
  nm_fit_ref_p20 <- paste0("gam_", prefix_tag, "_fit_ref_p20")

  nm_diff_post_min_to_p10 <- paste0("gam_", prefix_tag, "_diff_post_min_to_p10")
  nm_diff_post_min_to_p15 <- paste0("gam_", prefix_tag, "_diff_post_min_to_p15")
  nm_diff_post_min_to_p20 <- paste0("gam_", prefix_tag, "_diff_post_min_to_p20")

  nm_slope_post_min_to_p10 <- paste0("gam_", prefix_tag, "_slope_post_min_to_p10")
  nm_slope_post_min_to_p15 <- paste0("gam_", prefix_tag, "_slope_post_min_to_p15")
  nm_slope_post_min_to_p20 <- paste0("gam_", prefix_tag, "_slope_post_min_to_p20")

  # --- NEW metric names ---
  nm_post_min_after_p10 <- paste0("gam_", prefix_tag, "_post_min_after_p10")
  nm_post_min_after_p15 <- paste0("gam_", prefix_tag, "_post_min_after_p15")
  nm_post_min_after_p20 <- paste0("gam_", prefix_tag, "_post_min_after_p20")

  nm_min_at_end <- paste0("gam_", prefix_tag, "_min_at_end")

  nm_perc_recov_p10 <- paste0("gam_", prefix_tag, "_perc_recov_p10")
  nm_perc_recov_p15 <- paste0("gam_", prefix_tag, "_perc_recov_p15")
  nm_perc_recov_p20 <- paste0("gam_", prefix_tag, "_perc_recov_p20")

  dt[, id__temp := .I]

  long_dt <- data.table::melt(
    dt,
    id.vars = c("id__temp", reference_time_col),
    measure.vars = ts_cols,
    variable.name = "year_col",
    value.name = "value"
  )

  long_dt[, year := as.integer(sub(column_prefix, "", year_col))]
  long_dt[, ref_year := as.integer(get(reference_time_col))]

  long_dt <- long_dt[!is.na(ref_year)]

  dbg <- new.env(parent = emptyenv())
  dbg$fail_bad_ref <- 0L
  dbg$fail_min_n <- 0L
  dbg$fail_years_lt3 <- 0L
  dbg$fail_gam <- 0L
  dbg$ok <- 0L
  dbg$first_gam_error <- NULL

  na_row <- function(id) {
    out <- list(
      id__temp = id,

      NA_real_, NA_real_, NA_real_, NA_real_, NA_real_,
      NA_real_, NA_real_, NA_real_,
      NA_real_, NA_real_, NA_real_,
      NA_real_, NA_real_, NA_real_,

      NA, NA, NA,
      NA,
      NA_real_, NA_real_, NA_real_
    )
    names(out) <- c(
      "id__temp",

      nm_pre6_fit,
      nm_post_min,
      nm_post_min_year,
      nm_yrs_ref_to_post_min,
      nm_diff_pre6_to_post_min,
      nm_fit_ref_p10,
      nm_fit_ref_p15,
      nm_fit_ref_p20,
      nm_diff_post_min_to_p10,
      nm_diff_post_min_to_p15,
      nm_diff_post_min_to_p20,
      nm_slope_post_min_to_p10,
      nm_slope_post_min_to_p15,
      nm_slope_post_min_to_p20,

      nm_post_min_after_p10,
      nm_post_min_after_p15,
      nm_post_min_after_p20,

      nm_min_at_end,

      nm_perc_recov_p10,
      nm_perc_recov_p15,
      nm_perc_recov_p20
    )
    data.table::as.data.table(out)
  }

  fit_one <- function(d) {
    id <- unique(d$id__temp)[1]

    ref <- unique(d$ref_year)
    if (length(ref) != 1L || is.na(ref)) {
      if (isTRUE(debug)) dbg$fail_bad_ref <- dbg$fail_bad_ref + 1L
      return(na_row(id))
    }

    d_fit <- d[!is.na(value) & !is.na(year)]
    if (nrow(d_fit) < min_n) {
      if (isTRUE(debug)) dbg$fail_min_n <- dbg$fail_min_n + 1L
      return(na_row(id))
    }

    years_obs <- sort(unique(d_fit$year))
    if (length(years_obs) < 3) {
      if (isTRUE(debug)) dbg$fail_years_lt3 <- dbg$fail_years_lt3 + 1L
      return(na_row(id))
    }

    d_fit[, t := year - ref]
    t_obs <- sort(unique(d_fit$t))

    k_eff <- min(as.integer(k), max(3L, length(t_obs) - 1L))

    g <- tryCatch(
      mgcv::gam(value ~ s(t, k = k_eff), data = d_fit, method = gam_method),
      error = function(e) {
        if (isTRUE(debug) && is.null(dbg$first_gam_error)) dbg$first_gam_error <- conditionMessage(e)
        NULL
      }
    )
    if (is.null(g)) {
      if (isTRUE(debug)) dbg$fail_gam <- dbg$fail_gam + 1L
      return(na_row(id))
    }

    t_min <- min(t_obs)
    t_max <- max(t_obs)

    t_pre6 <- -6L
    pre6_fit <- if (t_pre6 >= t_min && t_pre6 <= t_max) {
      as.numeric(stats::predict(g, newdata = data.frame(t = t_pre6), type = "response"))
    } else {
      NA_real_
    }

    t_cap <- if (is.finite(post_min_search_max_year)) post_min_search_max_year - ref else Inf
    post_ts <- t_obs[t_obs > 0 & t_obs <= t_cap]

    post_min_val <- NA_real_
    post_min_t <- NA_real_

    if (length(post_ts) > 0) {
      post_fits <- as.numeric(stats::predict(g, newdata = data.frame(t = post_ts), type = "response"))
      i_min <- which.min(post_fits)
      post_min_val <- post_fits[i_min]
      post_min_t <- post_ts[i_min]
    }

    post_min_year <- if (!is.na(post_min_t)) ref + post_min_t else NA_real_
    yrs_ref_to_min <- if (!is.na(post_min_t)) post_min_t else NA_real_

    # Your convention: negative if decreases from before to after
    diff_pre6_to_min <- if (!is.na(pre6_fit) && !is.na(post_min_val)) post_min_val - pre6_fit else NA_real_

    t_p10 <- 10L
    t_p15 <- 15L
    t_p20 <- 20L

    fit_p10 <- if (t_p10 >= t_min && t_p10 <= t_max) {
      as.numeric(stats::predict(g, newdata = data.frame(t = t_p10), type = "response"))
    } else {
      NA_real_
    }

    fit_p15 <- if (t_p15 >= t_min && t_p15 <= t_max) {
      as.numeric(stats::predict(g, newdata = data.frame(t = t_p15), type = "response"))
    } else {
      NA_real_
    }

    fit_p20 <- if (t_p20 >= t_min && t_p20 <= t_max) {
      as.numeric(stats::predict(g, newdata = data.frame(t = t_p20), type = "response"))
    } else {
      NA_real_
    }

    diff_min_to_p10 <- if (!is.na(post_min_val) && !is.na(fit_p10)) fit_p10 - post_min_val else NA_real_
    diff_min_to_p15 <- if (!is.na(post_min_val) && !is.na(fit_p15)) fit_p15 - post_min_val else NA_real_
    diff_min_to_p20 <- if (!is.na(post_min_val) && !is.na(fit_p20)) fit_p20 - post_min_val else NA_real_

    slope_to_p10 <- if (!is.na(post_min_t) && !is.na(post_min_val) && !is.na(fit_p10) && (t_p10 != post_min_t)) {
      (fit_p10 - post_min_val) / (t_p10 - post_min_t)
    } else {
      NA_real_
    }

    slope_to_p15 <- if (!is.na(post_min_t) && !is.na(post_min_val) && !is.na(fit_p15) && (t_p15 != post_min_t)) {
      (fit_p15 - post_min_val) / (t_p15 - post_min_t)
    } else {
      NA_real_
    }

    slope_to_p20 <- if (!is.na(post_min_t) && !is.na(post_min_val) && !is.na(fit_p20) && (t_p20 != post_min_t)) {
      (fit_p20 - post_min_val) / (t_p20 - post_min_t)
    } else {
      NA_real_
    }


    # min_at_end: post-min is the last point in the series (within the post-min search window)
    # note: post_ts is already the post-reference grid that was searched for minima
    min_at_end <- if (!is.na(post_min_t) && length(post_ts) > 0) {
      isTRUE(post_min_t == max(post_ts))
    } else {
      NA
    }

    # percent recovery: (diff min->px / abs(diff pre6->min)) * 100
    denom <- abs(diff_pre6_to_min)
    perc_recov_p10 <- if (!is.na(diff_min_to_p10) && !is.na(denom) && denom > 0) (diff_min_to_p10 / denom) * 100 else NA_real_
    perc_recov_p15 <- if (!is.na(diff_min_to_p15) && !is.na(denom) && denom > 0) (diff_min_to_p15 / denom) * 100 else NA_real_
    perc_recov_p20 <- if (!is.na(diff_min_to_p20) && !is.na(denom) && denom > 0) (diff_min_to_p20 / denom) * 100 else NA_real_

    if (isTRUE(debug)) dbg$ok <- dbg$ok + 1L

    out <- list(
      id__temp = id,

      pre6_fit,
      post_min_val,
      post_min_year,
      yrs_ref_to_min,
      diff_pre6_to_min,

      fit_p10,
      fit_p15,
      fit_p20,

      diff_min_to_p10,
      diff_min_to_p15,
      diff_min_to_p20,

      slope_to_p10,
      slope_to_p15,
      slope_to_p20,

      min_at_end,

      perc_recov_p10,
      perc_recov_p15,
      perc_recov_p20
    )

    names(out) <- c(
      "id__temp",

      nm_pre6_fit,
      nm_post_min,
      nm_post_min_year,
      nm_yrs_ref_to_post_min,
      nm_diff_pre6_to_post_min,

      nm_fit_ref_p10,
      nm_fit_ref_p15,
      nm_fit_ref_p20,

      nm_diff_post_min_to_p10,
      nm_diff_post_min_to_p15,
      nm_diff_post_min_to_p20,

      nm_slope_post_min_to_p10,
      nm_slope_post_min_to_p15,
      nm_slope_post_min_to_p20,

      nm_min_at_end,

      nm_perc_recov_p10,
      nm_perc_recov_p15,
      nm_perc_recov_p20
    )

    data.table::as.data.table(out)
  }

  split_list <- split(long_dt, long_dt$id__temp)

  if (isTRUE(parallel)) {
    if (!requireNamespace("future.apply", quietly = TRUE)) stop("Please install 'future.apply' for parallel=TRUE.")
    if (!requireNamespace("future", quietly = TRUE)) stop("Please install 'future' for parallel=TRUE.")
    old_plan <- future::plan()
    on.exit(future::plan(old_plan), add = TRUE)
    future::plan(future::multisession, workers = n_cores)

    res_list <- future.apply::future_lapply(split_list, fit_one)
  } else {
    res_list <- lapply(split_list, fit_one)
  }

  stats_dt <- data.table::rbindlist(res_list, use.names = TRUE, fill = TRUE)

  out <- merge(dt, stats_dt, by = "id__temp", all.x = TRUE)
  out[, id__temp := NULL]

  if (isTRUE(debug)) {
    message(
      "compute_gam_stats debug counts:\n",
      "ok: ", dbg$ok, "\n",
      "fail_bad_ref: ", dbg$fail_bad_ref, "\n",
      "fail_min_n: ", dbg$fail_min_n, "\n",
      "fail_years_lt3: ", dbg$fail_years_lt3, "\n",
      "fail_gam: ", dbg$fail_gam
    )
    if (!is.null(dbg$first_gam_error)) message("First GAM error: ", dbg$first_gam_error)
  }

  tibble::as_tibble(out)
}

dats_simple_added <- compute_gam_stats(
  df = dats_simple_added,
  column_prefix = "rap_tree_",
  reference_time_col = "burn_year",
  k = 8,
  min_n = 15,
  gam_method = "REML",
  parallel = FALSE,
  debug = TRUE
)

dats_simple_added <- compute_gam_stats(
  df = dats_simple_added,
  column_prefix = "vcf_tree_",
  reference_time_col = "burn_year",
  k = 8,
  min_n = 15,
  gam_method = "REML",
  parallel = FALSE,
  debug = TRUE
)

```

# Check GAMs

```{r}

test <- dats_simple_added |> filter(burn_year == 2005) |> slice_head(n = 100)
test <- dats_no_reburn |> slice_head(n = 10000)

tic()
test_gam <- compute_gam_stats(
  df = test,
  column_prefix = "rap_tree_",
  reference_time_col = "burn_year",
  k = 8,
  min_n = 15,
  gam_method = "REML",
  parallel = FALSE,
  debug = TRUE
)
toc()

test_gam <- compute_gam_stats(
  df = test_gam,
  column_prefix = "vcf_tree_",
  reference_time_col = "burn_year",
  k = 8,
  min_n = 15,
  gam_method = "REML",
  parallel = FALSE,
  debug = TRUE
)

# Visual QC for compute_gam_stats(): pick 5 rows, refit GAM, plot TS + GAM,
# and annotate the computed metrics on each panel.

# ---- helpers ----

extract_year <- function(x, prefix) {
  as.integer(sub(paste0("^", prefix), "", x))
}

# Detect "gam_<prefix>_*" columns in a row and map to a standard set of metric keys.
# Returns a list: prefix_tag, cols (named vector of columns found)
detect_gam_metric_cols <- function(row_df) {
  nm <- names(row_df)

  # Candidate metric endings we know how to use
  endings <- c(
    "pre6_fit",
    "post_min",
    "post_min_year",
    "yrs_ref_to_post_min",
    "diff_pre6_to_post_min",
    "fit_ref_p10",
    "fit_ref_p15",
    "fit_ref_p20"
  )

  # Find any columns matching gam_<something>_<ending>
  # Capture <something> as prefix_tag
  rx <- paste0("^gam_([^_].*?)_(", paste(endings, collapse = "|"), ")$")
  hits <- grep(rx, nm, value = TRUE)
  if (length(hits) == 0) {
    return(list(prefix_tag = NA_character_, cols = setNames(character(0), character(0))))
  }

  # Pick the prefix_tag with the most matching columns
  tag <- sub(rx, "\\1", hits)
  tbl <- sort(table(tag), decreasing = TRUE)
  best_tag <- names(tbl)[1]

  cols_found <- setNames(character(0), character(0))
  for (end in endings) {
    colname <- paste0("gam_", best_tag, "_", end)
    if (colname %in% nm) cols_found[end] <- colname
  }

  list(prefix_tag = best_tag, cols = cols_found)
}

# Refit GAM and recompute metrics for ONE row (unchanged logic)
refit_gam_and_metrics_one <- function(row_df,
                                     column_prefix = "rap_tree_",
                                     reference_time_col = "burn_year",
                                     k = 8,
                                     gam_method = "REML") {
  stopifnot(nrow(row_df) == 1)

  ts_cols <- grep(paste0("^", column_prefix, "\\d{4}$"), names(row_df), value = TRUE)
  if (length(ts_cols) == 0) stop("No time-series columns found for prefix: ", column_prefix)

  ref <- as.integer(row_df[[reference_time_col]][[1]])
  if (is.na(ref)) stop("Reference year is NA for this row.")

  long <- tibble::tibble(
    year = extract_year(ts_cols, column_prefix),
    value = as.numeric(row_df[1, ts_cols, drop = TRUE])
  ) |>
    arrange(year) |>
    filter(!is.na(value)) |>
    mutate(t = year - ref)

  if (nrow(long) < 3) stop("Too few non-NA observations to fit GAM for this row.")

  t_obs <- sort(unique(long$t))
  k_eff <- min(as.integer(k), max(3L, length(t_obs) - 1L))

  g <- mgcv::gam(value ~ s(t, k = k_eff), data = long, method = gam_method)

  pred_grid <- tibble::tibble(year = sort(unique(long$year))) |>
    mutate(
      t = year - ref,
      gam_fit = as.numeric(predict(g, newdata = data.frame(t = t), type = "response"))
    )

  pre6_fit <- if (-6L %in% pred_grid$t) pred_grid$gam_fit[pred_grid$t == -6L][1] else NA_real_

  post_pred <- pred_grid |> filter(t > 0)
  if (nrow(post_pred) > 0) {
    i_min <- which.min(post_pred$gam_fit)
    post_min_val <- post_pred$gam_fit[i_min]
    post_min_year <- post_pred$year[i_min]
    post_min_t <- post_pred$t[i_min]
  } else {
    post_min_val <- NA_real_
    post_min_year <- NA_real_
    post_min_t <- NA_real_
  }

  yrs_ref_to_min <- if (!is.na(post_min_t)) post_min_t else NA_real_
  diff_pre6_to_min <- if (!is.na(pre6_fit) && !is.na(post_min_val)) pre6_fit - post_min_val else NA_real_

  fit_p10 <- if (10L %in% pred_grid$t) pred_grid$gam_fit[pred_grid$t == 10L][1] else NA_real_
  fit_p15 <- if (15L %in% pred_grid$t) pred_grid$gam_fit[pred_grid$t == 15L][1] else NA_real_
  fit_p20 <- if (20L %in% pred_grid$t) pred_grid$gam_fit[pred_grid$t == 20L][1] else NA_real_

  diff_min_to_p10 <- if (!is.na(post_min_val) && !is.na(fit_p10)) fit_p10 - post_min_val else NA_real_
  diff_min_to_p15 <- if (!is.na(post_min_val) && !is.na(fit_p15)) fit_p15 - post_min_val else NA_real_
  diff_min_to_p20 <- if (!is.na(post_min_val) && !is.na(fit_p20)) fit_p20 - post_min_val else NA_real_

  slope_to_p10 <- if (!is.na(post_min_t) && !is.na(post_min_val) && !is.na(fit_p10) && (10L != post_min_t)) {
    (fit_p10 - post_min_val) / (10L - post_min_t)
  } else NA_real_

  slope_to_p15 <- if (!is.na(post_min_t) && !is.na(post_min_val) && !is.na(fit_p15) && (15L != post_min_t)) {
    (fit_p15 - post_min_val) / (15L - post_min_t)
  } else NA_real_

  slope_to_p20 <- if (!is.na(post_min_t) && !is.na(post_min_val) && !is.na(fit_p20) && (20L != post_min_t)) {
    (fit_p20 - post_min_val) / (20L - post_min_t)
  } else NA_real_

  metrics <- tibble::tibble(
    ref_year = ref,
    gam_pre6_fit = pre6_fit,
    gam_post_min = post_min_val,
    gam_post_min_year = post_min_year,
    gam_yrs_ref_to_post_min = yrs_ref_to_min,
    gam_diff_pre6_to_post_min = diff_pre6_to_min,
    gam_fit_ref_p10 = fit_p10,
    gam_fit_ref_p15 = fit_p15,
    gam_fit_ref_p20 = fit_p20,
    gam_diff_post_min_to_p10 = diff_min_to_p10,
    gam_diff_post_min_to_p15 = diff_min_to_p15,
    gam_diff_post_min_to_p20 = diff_min_to_p20,
    gam_slope_post_min_to_p10 = slope_to_p10,
    gam_slope_post_min_to_p15 = slope_to_p15,
    gam_slope_post_min_to_p20 = slope_to_p20
  )

  list(long_obs = long, pred = pred_grid, metrics = metrics)
}

# Plot one row with annotations + overlay prior metric points in red
plot_one_gam_qc <- function(row_df,
                            row_id,
                            column_prefix = "rap_tree_",
                            reference_time_col = "burn_year",
                            k = 8,
                            gam_method = "REML") {
  out <- refit_gam_and_metrics_one(
    row_df = row_df,
    column_prefix = column_prefix,
    reference_time_col = reference_time_col,
    k = k,
    gam_method = gam_method
  )

  obs <- out$long_obs
  pred <- out$pred
  m <- out$metrics
  ref <- m$ref_year[1]

  # Pull previously generated metric points from row_df (any gam_<prefix>_* set)
  det <- detect_gam_metric_cols(row_df)
  prev_prefix <- det$prefix_tag
  prev_cols <- det$cols

  # Build points from ref +/- offsets using the *previous* computed fits (red points)
  prev_points <- tibble::tibble(
    label = character(0),
    year = numeric(0),
    value = numeric(0)
  )

  if (!is.na(prev_prefix) && length(prev_cols) > 0) {
    # Helper to safely pull a scalar numeric from row_df
    pull_num <- function(col) {
      if (is.null(col) || !nzchar(col) || !(col %in% names(row_df))) return(NA_real_)
      as.numeric(row_df[[col]][[1]])
    }

    prev_pre6 <- pull_num(prev_cols[["pre6_fit"]])
    prev_post_min <- pull_num(prev_cols[["post_min"]])
    prev_post_min_year <- pull_num(prev_cols[["post_min_year"]])
    prev_fit_p10 <- pull_num(prev_cols[["fit_ref_p10"]])
    prev_fit_p15 <- pull_num(prev_cols[["fit_ref_p15"]])
    prev_fit_p20 <- pull_num(prev_cols[["fit_ref_p20"]])

    prev_points <- tibble::tibble(
      label = c("prev_pre6", "prev_post_min", "prev_ref+10", "prev_ref+15", "prev_ref+20"),
      year = c(ref - 6L, prev_post_min_year, ref + 10L, ref + 15L, ref + 20L),
      value = c(prev_pre6, prev_post_min, prev_fit_p10, prev_fit_p15, prev_fit_p20)
    ) |>
      filter(!is.na(year) & !is.na(value))
  }

  # Newly computed metric points (default styling; distinct from red previous points)
  mark_df <- tibble::tibble(
    label = c("pre6", "post_min", "ref+10", "ref+15", "ref+20"),
    year = c(ref - 6L, m$gam_post_min_year[1], ref + 10L, ref + 15L, ref + 20L),
    value = c(m$gam_pre6_fit[1], m$gam_post_min[1], m$gam_fit_ref_p10[1], m$gam_fit_ref_p15[1], m$gam_fit_ref_p20[1])
  ) |>
    filter(!is.na(year) & !is.na(value))

  # Annotation text: show both "new" (recomputed) and "prev" prefix used for overlay
  ann <- paste0(
    "ref=", ref,
    "\n[new] pre6=", signif(m$gam_pre6_fit[1], 4),
    "\n[new] post_min=", signif(m$gam_post_min[1], 4), " @", m$gam_post_min_year[1],
    "\n[new] Δt(ref→min)=", signif(m$gam_yrs_ref_to_post_min[1], 4),
    "\n[new] pre6 - min=", signif(m$gam_diff_pre6_to_post_min[1], 4),
    "\n[new] fit+10=", signif(m$gam_fit_ref_p10[1], 4),
    "\n[new] fit+15=", signif(m$gam_fit_ref_p15[1], 4),
    "\n[new] fit+20=", signif(m$gam_fit_ref_p20[1], 4),
    "\n[prev set] ", ifelse(is.na(prev_prefix), "none", prev_prefix),
    "\nfire_sev=", row_df$burn_sev
  )

  ggplot() +
    # raw data
    geom_point(data = obs, aes(x = year, y = value), alpha = 0.7) +
    # newly refit GAM curve
    geom_line(data = pred, aes(x = year, y = gam_fit), linewidth = 0.8) +
    geom_vline(xintercept = ref, linetype = "dashed") +
    # newly computed metric points (not red)
    geom_point(data = mark_df, aes(x = year, y = value), size = 2.2) +
    geom_text(
      data = mark_df,
      aes(x = year, y = value, label = label),
      nudge_y = 1.0,
      size = 3,
      check_overlap = TRUE
    ) +
    # previously generated metric points (red)
    geom_point(
      data = prev_points,
      aes(x = year, y = value),
      size = 2.6,
      color = "red"
    ) +
    geom_text(
      data = prev_points,
      aes(x = year, y = value, label = label),
      nudge_y = -1.2,
      size = 3,
      color = "red",
      check_overlap = TRUE
    ) +
    annotate(
      "text",
      x = min(pred$year, na.rm = TRUE),
      y = max(c(obs$value, pred$gam_fit), na.rm = TRUE),
      hjust = 0, vjust = 1,
      label = ann,
      size = 3.0
    ) +
    labs(
      title = paste0("GAM QC row ", row_id),
      subtitle = paste0("k=", k, " (effective k may be reduced per row), dashed = reference year"),
      x = "Year",
      y = column_prefix
    ) +
    theme_minimal(base_size = 11)
}

# ---- run: sample rows from test_gam and plot ----
set.seed(123)

# Pick a stable row identifier to label panels.
# Prefer pt_id if present; else use row number.
id_col <- if ("pt_id" %in% names(test_gam)) "pt_id" else NULL

rows <- test_gam |>
  slice_sample(n = 10)

plots <- purrr::map(seq_len(nrow(rows)), function(i) {
  row_df <- rows[i, , drop = FALSE]
  row_id <- if (!is.null(id_col)) as.character(row_df[[id_col]][[1]]) else as.character(i)

  plot_one_gam_qc(
    row_df = row_df,
    row_id = row_id,
    column_prefix = "rap_tree_",
    reference_time_col = "burn_year",
    k = 8,
    gam_method = "REML"
  )
})

# Print plots (in RStudio this will show them one by one)
for (p in plots) print(p)

# Optional: arrange in a grid if you have patchwork installed
# install.packages("patchwork")
# library(patchwork)
# wrap_plots(plots, ncol = 1)

test_gam$gam_pre6_fit


# ---- run: sample rows from test_gam and plot ----
set.seed(123)

# Pick a stable row identifier to label panels.
# Prefer pt_id if present; else use row number.
id_col <- if ("pt_id" %in% names(test_gam)) "pt_id" else NULL

rows <- test_gam |>
  slice_sample(n = 10)

plots <- purrr::map(seq_len(nrow(rows)), function(i) {
  row_df <- rows[i, , drop = FALSE]
  row_id <- if (!is.null(id_col)) as.character(row_df[[id_col]][[1]]) else as.character(i)

  plot_one_gam_qc(
    row_df = row_df,
    row_id = row_id,
    column_prefix = "vcf_tree_",
    reference_time_col = "burn_year",
    k = 8,
    gam_method = "REML"
  )
})

# Print plots (in RStudio this will show them one by one)
for (p in plots) print(p)

# Optional: arrange in a grid if you have patchwork installed
# install.packages("patchwork")
# library(patchwork)
# wrap_plots(plots, ncol = 1)

test_gam$gam_pre6_fit


plot_one_gam_qc(
  row_df = test_gam |> filter(pt_id == "r10632_c6397"),
  row_id = "r10632_c6397",
  column_prefix = "rap_tree_",
  reference_time_col = "burn_year",
  k = 8,
  gam_method = "REML"
)

```
# New QAQC
```{r}


extract_year <- function(x, prefix) {
  as.integer(sub(paste0("^", prefix), "", x))
}

pick_gam_tag_for_base <- function(nm, base_prefix_tag) {
  endings <- c(
    "pre6_fit", "post_min", "post_min_year", "yrs_ref_to_post_min",
    "diff_pre6_to_post_min", "fit_ref_p10", "fit_ref_p15", "fit_ref_p20"
  )
  rx <- paste0("^gam_(", base_prefix_tag, ".*)_(", paste(endings, collapse = "|"), ")$")
  hits <- grep(rx, nm, value = TRUE)
  if (length(hits) == 0) return(NA_character_)
  tags <- sub(rx, "\\1", hits)
  names(sort(table(tags), decreasing = TRUE))[1]
}

metric_col_map <- function(nm, gam_tag) {
  endings <- c(
    "pre6_fit", "post_min", "post_min_year", "yrs_ref_to_post_min",
    "diff_pre6_to_post_min", "fit_ref_p10", "fit_ref_p15", "fit_ref_p20"
  )
  out <- setNames(rep(NA_character_, length(endings)), endings)
  for (e in endings) {
    col <- paste0("gam_", gam_tag, "_", e)
    if (col %in% nm) out[e] <- col
  }
  out
}

refit_gam_one_prefix <- function(row_df,
                                 column_prefix,
                                 reference_time_col,
                                 k = 8,
                                 gam_method = "REML") {
  stopifnot(nrow(row_df) == 1)

  ts_cols <- grep(paste0("^", column_prefix, "\\d{4}$"), names(row_df), value = TRUE)
  if (length(ts_cols) == 0) stop("No time-series columns found for prefix: ", column_prefix)

  ref <- as.integer(row_df[[reference_time_col]][[1]])
  if (is.na(ref)) stop("Reference year is NA for this row.")

  obs <- tibble::tibble(
    year = extract_year(ts_cols, column_prefix),
    value = as.numeric(row_df[1, ts_cols, drop = TRUE])
  ) |>
    arrange(year) |>
    filter(!is.na(value)) |>
    mutate(t = year - ref)

  if (nrow(obs) < 3) stop("Too few non-NA observations to fit GAM for this row/prefix: ", column_prefix)

  t_obs <- sort(unique(obs$t))
  k_eff <- min(as.integer(k), max(3L, length(t_obs) - 1L))

  g <- mgcv::gam(value ~ s(t, k = k_eff), data = obs, method = gam_method)

  pred <- tibble::tibble(year = sort(unique(obs$year))) |>
    mutate(
      t = year - ref,
      gam_fit = as.numeric(predict(g, newdata = data.frame(t = t), type = "response"))
    )

  list(ref_year = ref, obs = obs, pred = pred)
}

pull_prev_metric_points <- function(row_df, ref_year, gam_tag) {
  nm <- names(row_df)
  if (is.na(gam_tag)) {
    return(tibble::tibble(label = character(0), year = numeric(0), value = numeric(0)))
  }

  cols <- metric_col_map(nm, gam_tag)

  pull_num <- function(col) {
    if (is.na(col) || !(col %in% nm)) return(NA_real_)
    as.numeric(row_df[[col]][[1]])
  }

  pre6 <- pull_num(cols["pre6_fit"])
  post_min <- pull_num(cols["post_min"])
  post_min_year <- pull_num(cols["post_min_year"])
  fit_p10 <- pull_num(cols["fit_ref_p10"])
  fit_p15 <- pull_num(cols["fit_ref_p15"])
  fit_p20 <- pull_num(cols["fit_ref_p20"])

  tibble::tibble(
    label = c("pre6", "post_min", "ref+10", "ref+15", "ref+20"),
    year  = c(ref_year - 6L, post_min_year, ref_year + 10L, ref_year + 15L, ref_year + 20L),
    value = c(pre6, post_min, fit_p10, fit_p15, fit_p20)
  ) |>
    filter(!is.na(year) & !is.na(value))
}

# ---- MAIN: overlay or facet multiple prefixes ----
plot_multi_gam_qc <- function(df,
                             sample_id,
                             id_col = "pt_id",
                             column_prefixes,
                             colors,
                             reference_time_col,
                             k = 8,
                             gam_method = "REML",
                             gam_tags = NULL,
                             facet = FALSE) {
  stopifnot(length(column_prefixes) >= 1)
  stopifnot(length(colors) == length(column_prefixes))

  if (!id_col %in% names(df)) stop("id_col not found in df: ", id_col)
  if (!reference_time_col %in% names(df)) stop("reference_time_col not found in df: ", reference_time_col)

  row_df <- df[df[[id_col]] == sample_id, , drop = FALSE]
  if (nrow(row_df) != 1) stop("Expected exactly 1 row for sample_id; got: ", nrow(row_df))

  base_tags <- sub("_$", "", column_prefixes)
  series_names <- base_tags
  nm <- names(row_df)

  if (!is.null(gam_tags)) {
    stopifnot(length(gam_tags) == length(column_prefixes))
    used_gam_tags <- gam_tags
  } else {
    used_gam_tags <- purrr::map_chr(base_tags, ~ pick_gam_tag_for_base(nm, .x))
  }

  pieces <- purrr::pmap(
    list(column_prefixes, series_names, colors, used_gam_tags),
    function(col_pref, series, col, gam_tag) {
      fit <- refit_gam_one_prefix(
        row_df = row_df,
        column_prefix = col_pref,
        reference_time_col = reference_time_col,
        k = k,
        gam_method = gam_method
      )

      obs_df <- fit$obs |>
        mutate(series = series)

      pred_df <- fit$pred |>
        mutate(series = series)

      dia_df <- pull_prev_metric_points(
        row_df = row_df,
        ref_year = fit$ref_year,
        gam_tag = gam_tag
      ) |>
        mutate(series = series)

      list(
        ref_year = fit$ref_year,
        obs = obs_df,
        pred = pred_df,
        diamonds = dia_df,
        series = series,
        color = col,
        gam_tag = gam_tag
      )
    }
  )

  ref_lines <- purrr::map_dfr(pieces, ~ tibble::tibble(series = .x$series, ref_year = .x$ref_year))
  obs_all <- purrr::map_dfr(pieces, "obs")
  pred_all <- purrr::map_dfr(pieces, "pred")
  dia_all <- purrr::map_dfr(pieces, "diamonds")

  color_map <- setNames(colors, series_names)

  tag_txt <- purrr::map_chr(pieces, ~ paste0(.x$series, " diamonds=", ifelse(is.na(.x$gam_tag), "none", .x$gam_tag)))
  subtitle_txt <- paste(tag_txt, collapse = " | ")

  p <- ggplot() +
    # raw TS (colored by series)
    geom_point(
      data = obs_all,
      aes(x = year, y = value, color = series),
      alpha = 0.7
    ) +
    # newly refit GAM curve (colored by series)
    geom_line(
      data = pred_all,
      aes(x = year, y = gam_fit, color = series),
      linewidth = 0.8
    ) +
    # reference-year dashed line per series (colored)
    geom_vline(
      data = ref_lines,
      aes(xintercept = ref_year, color = series),
      linetype = "dashed"
    ) +
    # previously-generated metric points as diamonds (shape 18), colored by series
    geom_point(
      data = dia_all,
      aes(x = year, y = value, color = series),
      shape = 18,
      size = 3.0
    ) +
    geom_text(
      data = dia_all,
      aes(x = year, y = value, label = label, color = series),
      nudge_y = -1.0,
      size = 3,
      check_overlap = TRUE
    ) +
    scale_color_manual(values = color_map) +
    labs(
      title = paste0("GAM QC overlay: ", id_col, " = ", sample_id),
      subtitle = subtitle_txt,
      x = "Year",
      y = "Value"
    ) +
    theme_minimal(base_size = 11)

  if (isTRUE(facet)) {
    p <- p + facet_wrap(~ series, scales = "free_y")
  }

  p
}

# ---- Examples ----
# 1) Overlay different variables (2 raw series + 2 GAM fits + 2 diamond sets)
plot_multi_gam_qc(
  df = test_gam,
  sample_id = test_gam$pt_id[1],
  id_col = "pt_id",
  column_prefixes = c("rap_tree_", "vcf_tree_"),
  colors = c("steelblue", "darkorange"),
  reference_time_col = "burn_year",
  k = 8,
  gam_method = "REML",
  facet = FALSE
)
#
#2) Same, but faceted (cleanest when units differ)
plot_multi_gam_qc(
  df = test_gam,
  sample_id = test_gam$pt_id[1],
  id_col = "pt_id",
  column_prefixes = c("rap_tree_", "vcf_tree_"),
  colors = c("steelblue", "darkorange"),
  reference_time_col = "burn_year",
  k = 8,
  gam_method = "REML",
  facet = TRUE
)





for(i in 1:10) {
  p <- plot_multi_gam_qc(
    df = test_gam,
    sample_id = test_gam$pt_id[i],
    id_col = "pt_id",
    column_prefixes = c("rap_tree_", "vcf_tree_"),
    colors = c("steelblue", "darkorange"),
    reference_time_col = "burn_year",
    k = 8,
    gam_method = "REML",
    facet = FALSE
  )
  print(p)
}


t <- test_gam |> slice_head(n = 3) |> select(pt_id, dplyr::starts_with("rap_"), dplyr::starts_with("gam_"))


```


# Deal with response data (VCF & RAP) - OLD (avgs, etc)


```{r}

compute_forest_response_offset <- function(dt, offsets, new_col_name, prefix = "rap_TRE_") {
  # Ensure dt is a data.table
  dt <- as.data.table(copy(dt))
  dt[, id__temp := .I]  # Create unique ID for merge

  # Melt relevant columns
  melt_cols <- grep(paste0("^", prefix, "\\d{4}$"), names(dt), value = TRUE)

  long_dt <- melt(
    dt,
    id.vars = c("id__temp", "mock_burn_year"),
    measure.vars = melt_cols,
    variable.name = "year_col",
    value.name = "value"
  )

  # Extract year from column name
  long_dt[, year := as.integer(sub(prefix, "", year_col))]

  # Calculate offset from burn year
  long_dt[, offset := year - mock_burn_year]

  # Only keep rows with desired offsets
  long_dt <- long_dt[offset %in% offsets]

  # Now, summarize carefully
  summary_dt <- long_dt[, .(
    n_available = sum(!is.na(value)),
    n_expected = length(offsets),
    avg = if (sum(!is.na(value)) == length(offsets)) mean(value, na.rm = TRUE) else NA_real_
  ), by = id__temp]

  # Keep only id and avg
  summary_dt <- summary_dt[, .(id__temp, avg)]

  # Merge back
  dt <- merge(dt, summary_dt, by = "id__temp", all.x = TRUE)
  setnames(dt, "avg", new_col_name)
  dt[, id__temp := NULL]

  return(as_tibble(dt))
}


#' Compute Post-Fire Forest Trend Slope from forest cover Time Series
#'
#' This function calculates the slope and R-squared value of a thiel-sen regression
#' fitted to post-fire remote sensing time series data (e.g., RAP TRE) over
#' specified year offsets. The output includes the computed slope and R² for each
#' observation where sufficient data is available.
#'
#' @param dt A `data.table` or `data.frame` containing RAP data and a `mock_burn_year` column.
#' @param offsets Integer vector indicating year offsets from `mock_burn_year` to include in the slope calculation. Defaults to `2:10`.
#' @param req_yrs Integer indicating the minimum number of non-NA observations required to fit a regression. Defaults to `5`.
#' @param new_col_name Name of the column to store computed slope values. Defaults to `"post_fire_rap_slope"`.
#' @param r2_col_name Name of the column to store R-squared values. Defaults to `"post_fire_rap_r2"`.
#' @param prefix Prefix of column names containing yearly response values. These should be named like `prefixYYYY` (e.g., `"rap_TRE_2012"`).
#'
#' @return A `tibble` with the original data and two additional columns containing slope and R-squared values.
#'
#' @details
#' The function internally melts the  time series into long format, filters values
#' by offset from the `mock_burn_year`, and fits a thiel-sen model (`value ~ offset`) for each
#' row. If the number of non-NA years is below `req_yrs`, NA values are returned.
#'
#' @importFrom data.table as.data.table copy melt setnames
#' @importFrom stats lm coef summary
#' @importFrom tibble as_tibble
#'
#' @examples
#' \dontrun{
#' dt <- data.table(
#'   burn_year = c(2010, 2012),
#'   rap_TRE_2011 = c(0.5, 0.6),
#'   rap_TRE_2012 = c(0.6, 0.65),
#'   rap_TRE_2013 = c(0.7, NA),
#'   rap_TRE_2014 = c(0.8, 0.75)
#' )
#' compute_post_fire_forest_slope_dt(dt, offsets = 1:4, req_yrs = 3)
#' }
#'
#' @export
compute_post_fire_forest_slope_dt <- function(
  dt,
  offsets = 2:10,
  req_yrs = 5,
  new_col_name = "post_fire_rap_slope",
  r2_col_name = "post_fire_rap_r2",
  prefix = "rap_TRE_"
) {
  dt <- as.data.table(copy(dt))
  dt[, id__temp := .I]  # Unique ID

  # Melt RAP TRE columns
  melt_cols <- grep(paste0("^", prefix, "\\d{4}$"), names(dt), value = TRUE)

  long_dt <- melt(
    dt,
    id.vars = c("id__temp", "mock_burn_year"),
    measure.vars = melt_cols,
    variable.name = "year_col",
    value.name = "value"
  )

  # Extract year
  long_dt[, year := as.integer(sub(prefix, "", year_col))]

  # Calculate offset from burn year
  long_dt[, offset := year - mock_burn_year]

  # Keep only desired offsets
  long_dt <- long_dt[offset %in% offsets]

  # Group and compute slope + R2 if enough data
  slope_dt <- long_dt[
    !is.na(value),
    {
      if (.N >= req_yrs) {
        # model <- lm(value ~ offset, data = .SD) ####################SWITCH TO USE THEIL-SEN REGRESSION
        # list(
        #   slope = coef(model)[["offset"]],
        #   r2 = summary(model)$r.squared
        # )
        model <- mblm::mblm(value ~ offset, data = .SD, repeated = FALSE)
        preds <- predict(model)
        r2 <- 1 - sum((.SD$value - preds)^2, na.rm = TRUE) / sum((.SD$value - mean(.SD$value, na.rm = TRUE))^2, na.rm = TRUE)

        list(
          slope = coef(model)[["offset"]],
          r2 = r2
        )
      } else {
        list(
          slope = NA_real_,
          r2 = NA_real_
        )
      }
    },
    by = id__temp
  ]

  # Merge results back
  dt <- merge(dt, slope_dt, by = "id__temp", all.x = TRUE)
  setnames(dt, c("slope", "r2"), c(new_col_name, r2_col_name))
  dt[, id__temp := NULL]

  return(as_tibble(dt))
}

compute_avgs <- function(dats, prefix) {
  dats <- dats |>
  compute_forest_response_offset(offsets = 9:11,
                                 new_col_name = glue("avg_{prefix}10_yr_post_fire"),
                                 prefix = prefix) |>
  compute_forest_response_offset(offsets = 14:16,
                                 new_col_name = glue("avg_{prefix}15_yr_post_fire"),
                                 prefix = prefix) |>
  compute_forest_response_offset(offsets = 19:21,
                                 new_col_name = glue("avg_{prefix}20_yr_post_fire"),
                                 prefix = prefix) |>
  compute_forest_response_offset(offsets = -7:-5,
                                 new_col_name = glue("avg_{prefix}6_yr_pre_fire"),
                                 prefix = prefix) |>
  compute_forest_response_offset(offsets = 2:4, 
                                 new_col_name = glue("avg_{prefix}3_yr_post_fire"),  # USE 3 YRS POST FIRE SO THAT WINDOW CAPTURES 2-4, missing most delayed mortality
                                 prefix = prefix)
  return(dats)
}

compute_response_differences_recovery <- function(dats, prefix) {
  dats <- dats |>
    mutate(
      # n6 -> 3 year difference
      !!sym(glue("post_fire_{prefix}difference_n6_3_yr")) :=
        !!sym(glue("avg_{prefix}3_yr_post_fire")) -
        !!sym(glue("avg_{prefix}6_yr_pre_fire")),

      # increase from 3 -> 10 years post-fire
      !!sym(glue("post_fire_{prefix}increase_3_10_yr")) :=
        !!sym(glue("avg_{prefix}10_yr_post_fire")) -
        !!sym(glue("avg_{prefix}3_yr_post_fire")),

      # increase from 3 -> 15 years post-fire
      !!sym(glue("post_fire_{prefix}increase_3_15_yr")) :=
        !!sym(glue("avg_{prefix}15_yr_post_fire")) -
        !!sym(glue("avg_{prefix}3_yr_post_fire")),
      
      # increase from 3 -> 20 years post-fire
      !!sym(glue("post_fire_{prefix}increase_3_20_yr")) :=
        !!sym(glue("avg_{prefix}20_yr_post_fire")) -
        !!sym(glue("avg_{prefix}3_yr_post_fire")),

      # percent recovery (3–10yr)
      !!sym(glue("post_fire_{prefix}perc_recovery_n6_3_10_yr")) :=
        (!!sym(glue("post_fire_{prefix}increase_3_10_yr")) /
        (-1 * !!sym(glue("post_fire_{prefix}difference_n6_3_yr")))) * 100,

      # percent recovery (3–15yr)
      !!sym(glue("post_fire_{prefix}perc_recovery_n6_3_15_yr")) :=
        (!!sym(glue("post_fire_{prefix}increase_3_15_yr")) /
        (-1 * !!sym(glue("post_fire_{prefix}difference_n6_3_yr")))) * 100,
      
      # percent recovery (3–20yr)
      !!sym(glue("post_fire_{prefix}perc_recovery_n6_3_20_yr")) :=
        (!!sym(glue("post_fire_{prefix}increase_3_20_yr")) /
        (-1 * !!sym(glue("post_fire_{prefix}difference_n6_3_yr")))) * 100
    )
  return(dats)
}  

# Allow for up to 1 missing year for each slope computation
compute_response_slopes <- function(dats, prefix) {
  dats <- dats |>
    compute_post_fire_forest_slope_dt(
      offsets = 3:10,
      req_yrs = 7,
      new_col_name = glue("post_fire_{prefix}slope_3_10_yrs"),
      r2_col_name = glue("post_fire_{prefix}3_10_yrs_r2"),
      prefix = prefix) |>
    compute_post_fire_forest_slope_dt(
      offsets = 3:15,
      req_yrs = 12,
      new_col_name = glue("post_fire_{prefix}slope_3_15_yrs"),
      r2_col_name = glue("post_fire_{prefix}3_15_yrs_r2"),
      prefix = prefix) |>
    compute_post_fire_forest_slope_dt(
      offsets = 3:20,
      req_yrs = 17,
      new_col_name = glue("post_fire_{prefix}slope_3_20_yrs"),
      r2_col_name = glue("post_fire_{prefix}3_20_yrs_r2"),
      prefix = prefix)
  return(dats)
}


# Use functions
tic() #161 seconds for 36k row dataset
dats_response <- dats_simple_added |>
  compute_avgs(prefix = "rap_tree_") |>
  compute_avgs(prefix = "vcf_tree_") |>
  compute_response_differences_recovery(prefix = "rap_tree_")  |>
  compute_response_differences_recovery(prefix = "vcf_tree_") |>
  compute_response_slopes(prefix = "rap_tree_") |>
  compute_response_slopes(prefix = "vcf_tree_")
toc()



dats_response <- dats_simple_added




```



# Clean and write short data

```{r}
dats_clean <- dats_response |>
  mutate(uniqueID = row_number()) |> # apply a unique ID to be sure it has one
  remove_columns_with_prefix(prefixes = c("cbi", "distance_forest_after"))


arrow::write_parquet(x = dats_clean, sink = here(dir_derived, "all_dats_cleaned.parquet"))
#dats_clean <- read_parquet(here(dir_derived, "test_dats_cleaned.parquet"))



# Clean up unnecessary columns
dats_short <- dats_clean |>
  remove_columns_with_prefix(prefixes = c("^pdsi_summer_\\d{4}$",
                                          "^pdsi_annual_\\d{4}$",
                                          "^hd_fingerprint_\\d{4}$",
                                          "^biotic_\\d{4}$",
                                          "rap_tree",
                                          "vcf_tree"))


arrow::write_parquet(x = dats_short, sink = here(dir_derived, "test_dats_short.parquet"))



# Create gpkg for points

gpkg <- dats_short |>
  select(uniqueID, lat, long, ecoregion) |>
  st_as_sf(
    coords = c("long", "lat"),
    crs = 4326,
    remove = FALSE
  )

sf::st_write(gpkg, here(dir_derived, "test_dats_sf.gpkg"), append = FALSE)

```


# Long data
```{r}
dats_long <- dats_clean |>
  pivot_longer(
    cols = matches("_(19|20)\\d{2}$"),        # ends with a 4-digit year
    names_to = c(".value", "year"),           # .value = "pdsi_summer", "vpd_fall", ...
    names_pattern = "(.*)_(\\d{4})"           # capture "pdsi_summer" and the year
  ) |>
  mutate(year = as.integer(year)) |>
  mutate(year_from_fire_index = year - mock_burn_year,
         fire_timeframe = ifelse(year_from_fire_index < 0, "before_burn", ifelse(year_from_fire_index > 0, "after_burn", "year_of_burn")),
         treated = ifelse(year_from_fire_index >= 0 & fire == 1, 1, 0))

tic()
arrow::write_parquet(x = dats_long, sink = here(dir_derived, "all_dats_long.parquet"))
toc()

```





<button class="accordion-button" type="button" data-bs-toggle="collapse" data-bs-target="#collapseOne" >Click to see Session Information</button><div id="collapseOne" class="accordion-collapse collapse"><div>

```{r sessionInfo}
sessionInfo()
```

</div></div>