---
title: "03_test_analysis"
author: Tyler L. McIntosh
date: last-modified
format:
  html:
    embed-resources: true
    html-math-method: katex
    css: styles.css
    toc: true
    toc-depth: 1
    toc-title: Contents
    toc-location: right
    code-fold: true
    code-summary: "Show the code"
---



# Setup and read in data

``` {r setupchunk}
#| echo: true
#| message: false
#| warning: false
#| results: 'hide'

rm(list = ls())

if(!requireNamespace("here", quietly = TRUE)) {
  install.packages("here")
}
library(here)

source(here::here("code", "functions.R"))

install_and_load_packages(c(
  "tidyverse",
  "emmeans", # For Dunnett's test
  "faraway", # For vif
  "car", # For partial residual plots
  "effects", # For marginal effect prediction plots
  "data.table",
  "gstat",
  "spdep",
  "randomForest",
  "spmoran",
  "forcats",
  "tictoc"))

dir_figs <- here::here('figs')
dir_derived <- here::here('data', 'derived')
dir_raw <- here::here('data', 'raw')
dir_manual <- here::here('data', 'manual')
dir_ensure(c(dir_figs,
             dir_derived,
             dir_manual,
             dir_raw))


set_factors <- function(dats) {
  newdats <- dats |>
  mutate(
      biotic_fire_cat = factor(
      biotic_fire_cat,
      levels = c(
        "no_biotic_or_outside_timeframe",
        "biotic_6_10_prior",
        "biotic_0_5_prior",
        "biotic_1_5_after",
        "biotic_6_10_after"
      )
    )
  ) |>
  mutate(frg_reclass = factor(frg_reclass,
                         levels = c(
                           1, 2)), #1: low/mixed, 2: replacement, 3: indeterminate / barren / etc (filtered out in script 02)
         lf_bps = factor(lf_bps),
         burn_year_sev = factor(burn_year_sev),
         beneficial_fire = factor(beneficial_fire),
         biotic_sev_binned = cut(
            biotic_year_sev,
            breaks = seq(20, 100, by = 20),
            right = FALSE,
            labels = paste0(seq(20, 80, by = 20), "-", seq(40, 100, by = 20)),
            include.lowest = TRUE
          ),
          biotic_sev_binned = fct_explicit_na(biotic_sev_binned, na_level = "Missing"),
          biotic_sev_binned = fct_relevel(biotic_sev_binned, "Missing", after = 0),
         biotic_fire_gap2 = ifelse(biotic_fire_gap > 10 & biotic_fire_gap < -10, NA, biotic_fire_gap),
         biotic_fire_gap_binned = cut(
            biotic_fire_gap2,
            breaks = seq(-10, 10, by = 5),
            right = FALSE,
            labels = paste0(seq(-10, 5, by = 5), "-", seq(-5, 10, by = 5)),
            include.lowest = TRUE
          ),
          biotic_fire_gap_binned = fct_explicit_na(biotic_fire_gap_binned, na_level = "Beyond 10 years from fire"),
          biotic_fire_gap_binned = fct_relevel(biotic_fire_gap_binned, "Beyond 10 years from fire", after = 0)) 
  return(newdats)
}


filter_out_edge_cases <- function(dats) {
  newdats <- dats |>
    filter(burn_year_sev != "unburned") |>
    filter(biotic_fire_cat != "biotic_6_10_after")
  return(newdats)
}


# Load data
dats_short <- readr::read_csv(here::here(dir_derived, "ard_short_v1.csv")) |>
  set_factors() |>
  filter_out_edge_cases()

dats_long <- readr::read_csv(here::here(dir_derived, "ard_long_v1.csv")) |>
  set_factors() |>
  filter_out_edge_cases()

dats_short_sf <- sf::st_read(here::here(dir_derived, "ard_short_sf_v1.gpkg")) |>
  set_factors() |>
  filter_out_edge_cases()


```

# Plotting
```{r plot}

ggplot() +
  geom_sf(data = dats_short_sf, 
          aes(color = post_fire_TRE_difference_10_yr), 
          size = 0.3) +
  scale_color_viridis_c(option = "D") +
  theme_minimal() +
  labs(title = "Tree cover difference from 5 years pre-fire to 10 years post-fire")

ggplot() +
  geom_sf(data = dats_short_sf, 
          aes(color = biotic_fire_cat), 
          size = 0.3) +
  theme_minimal() +
  labs(title = "Biotic-fire-category")


```



# ANOVA - using short data frame

While our data appears to be non-normal, shapiro-wilk tests suggest that it some of our samples are close enough to normal.

We do seem to have differences between our 5-year post-fire differences, although visually these differences seem to collapse at 10 an 15 years post-fire. To do a quick test we would use a parametric anova. We can follow this with a post-hoc test; here we are interested in comparing multiple treatment groups against a control group, so we will use a Dunnett's test.

From our analysis we see that there is a statistically significant difference between our treatment groups and the control group for 5 and 10 year post-fire tree cover difference, but not at 15 years.


```{r anova}

# QQ plot helper with tidy evaluation
run_qq <- function(biotic_fire_cat_value, response_col) {
  dats_filtered <- dats_short %>%
    filter(biotic_fire_cat == biotic_fire_cat_value)

  qq_data <- pull(dats_filtered, {{ response_col }})

  qqnorm(qq_data, main = paste("QQ Plot:", biotic_fire_cat_value))
  qqline(qq_data)

  if (length(qq_data) < 5000) {
    print(shapiro.test(qq_data))
  } else {
    message("Too many samples to run Shapiro-Wilk test")
  }
}

# Main test function
basic_tests <- function(response_col) {
  
  response_name <- deparse(substitute(response_col))
  message("-------------------------------------------------\nRunning tests on: ", response_name)
  
  # Boxplot
  bp <- ggplot(dats_short, aes(x = biotic_fire_cat, y = {{ response_col }})) +
    geom_boxplot() +
    labs(title = "Boxplot by Biotic Fire Category")
  print(bp)
  
  # QQ plots + normality test
  run_qq("biotic_1_5_after", {{ response_col }})
  run_qq("no_biotic_or_outside_timeframe", {{ response_col }})

  # Fit ANOVA using tidy evaluation
  formula <- as.formula(paste0(deparse(substitute(response_col)), " ~ biotic_fire_cat"))
  aov_model <- aov(formula, data = dats_short)

  print(summary(aov_model))

  # Dunnett's test with emmeans
  emm <- emmeans::emmeans(aov_model, "biotic_fire_cat")
  print(emmeans::contrast(emm, method = "dunnett", ref = "no_biotic_or_outside_timeframe"))
}


basic_tests(post_fire_TRE_difference_5_yr)
basic_tests(post_fire_TRE_difference_10_yr)
basic_tests(post_fire_TRE_difference_15_yr)


```

# Simple regressions

First, let's not worry about the spatial components and just quickly look at some simple regressions.

Let's take a look at the relationship of CBI with post-fire tree cover difference; it looks like a 2nd degree polynomial fits a bit better than linear, but 3rd doesn't add much value.

Let's also look at the relationship of time between fire and biotic agent with post-fire tree cover difference; it looks like a 2nd degree polynomial fits A LOT better than linear, but 3rd doesn't add much value.

All of these relationships are significant! Let's see whether or not this remains the case when we start to take into account more variables.



```{r simple-regressions}

basic_regressions <- function(response_col) {
  response_name <- deparse(substitute(response_col))
  message("-------------------------------------------------\nRunning tests on: ", response_name)

  for (i in 1:3) {
    # Build formula for lm
    form <- as.formula(paste(response_name, "~ poly(burn_year_cbi,", i, ")"))
    
    cbi_lm <- lm(form, data = dats_short)
    print(summary(cbi_lm))
    print(paste("AIC", AIC(cbi_lm)))

    # Plot
    p <- ggplot(data = dats_short, aes(x = burn_year_cbi, y = !!ensym(response_col))) +
      geom_point() +
      stat_smooth(method = 'lm', formula = y ~ poly(x, i))
    print(p)
  }

  for (i in 1:3) {
    gap_dats <- dats_short |>
      filter(!is.na(biotic_fire_gap))

    form_gap <- as.formula(paste(response_name, "~ poly(biotic_fire_gap,", i, ")"))
    
    gap_lm <- lm(form_gap, data = gap_dats)
    print(summary(gap_lm))
    print(paste("AIC", AIC(gap_lm)))

    p <- ggplot(data = gap_dats, aes(x = biotic_fire_gap, y = !!ensym(response_col))) +
      geom_point() +
      stat_smooth(method = 'lm', formula = y ~ poly(x, i))
    print(p)
  }
}

  

basic_regressions(post_fire_TRE_difference_5_yr)
basic_regressions(post_fire_TRE_difference_10_yr)
basic_regressions(post_fire_TRE_difference_15_yr)
basic_regressions(post_fire_rap_slope_2_10_yrs)





```
# Before moving into doing any MLR, let's take a look at what the data looks like from a time series perspective

```{r long-time-series}


# A function to create a time series plot from the data, along with the number of samples used
time_series_plot <- function(dat_subset, title, col_split) {
  p <- ggplot(data = dat_subset,
              aes(x = year_from_fire_index, y = rap_TRE, col = {{ col_split }})) +
              #geom_point() +
              stat_smooth(method = 'lm', formula = y ~ poly(x, 5)) +
              ylim(0, 75) +
              xlim(-20, 20) +
              xlab("Normalized year from fire") +
              ylab("Percent tree cover")
              labs(title = title, caption = paste0("N: ", length(unique(dat_subset$`system:index`))))
  print(p)
}

time_series_plot2 <- function(dat_subset, title) {
  p <- ggplot(data = dat_subset,
              aes(x = year_from_fire_index, y = rap_TRE)) +
              #geom_point() +
              stat_smooth(method = 'lm', formula = y ~ poly(x, 4)) +
              ylim(0, 75) +
              xlim(-20, 20) +
              xlab("Normalized year from fire") +
              ylab("Percent tree cover")
              labs(title = title, caption = paste0("N: ", length(unique(dat_subset$`system:index`))))
  print(p)
}






time_series_plot2(
  dat_subset = dats_long,
  title = "Generalized trajectory"
)


# All
time_series_plot(
  dat_subset = dats_long,
  title = "All dats",
  col_split = biotic_fire_cat
)

bf <- time_series_plot(
  dat_subset = dats_long |>
    filter(beneficial_fire != "not severe enough" &
             beneficial_fire != "replacement bf"),
  title = "All dats",
  col_split = beneficial_fire
)


# By severity

time_series_plot(
  dat_subset = dats_long |>
         filter(burn_year_cbi <1),
  title = "Low sev",
  col_split = biotic_fire_cat
)

time_series_plot(
  dat_subset = dats_long |>
         filter(burn_year_cbi >2),
  title = "High sev",
  col_split = biotic_fire_cat
)


# By FRG

time_series_plot(
  dat_subset = dats_long |>
         filter(frg_reclass == 1),
  title = "Low/mixed regime",
  col_split = biotic_fire_cat
)

time_series_plot(
  dat_subset = dats_long |>
         filter(frg_reclass == 2),
  title = "Replacement regime",
  col_split = biotic_fire_cat
)


# By FRG & Sev

time_series_plot(
  dat_subset = dats_long |>
         filter(frg_reclass == 1 & burn_year_cbi <1),
  title = "Low/mixed regime, low sev",
  col_split = biotic_fire_cat
)

time_series_plot(
  dat_subset = dats_long |>
         filter(frg_reclass == 2 & burn_year_cbi <1),
  title = "Replacement regime, low sev",
  col_split = biotic_fire_cat
)

time_series_plot(
  dat_subset = dats_long |>
         filter(frg_reclass == 1 & burn_year_cbi >2),
  title = "Low/mixed regime, high sev",
  col_split = biotic_fire_cat
)

time_series_plot(
  dat_subset = dats_long |>
         filter(frg_reclass == 2 & burn_year_cbi >2),
  title = "Replacement regime, high sev",
  col_split = biotic_fire_cat
)


```




```{r beneficial fire}

time_series_plot_recovery_only <- function(dat_subset, title, col_split) {
  p <- ggplot(data = dat_subset |>
                filter(year_from_fire_index >= 1),
              aes(x = year_from_fire_index, y = rap_TRE, col = {{ col_split }})) +
              #geom_point() +
              stat_smooth(method = 'lm', formula = y ~ poly(x, 4)) +
              ylim(0, 75) +
              xlim(0, 40) +
              xlab("Time since fire (years)") +
              ylab("Percent tree cover")
              labs(title = title, caption = paste0("N: ", length(unique(dat_subset$`system:index`))))
  print(p)
}

time_series_plot_recovery_only <- function(dat_subset, title, col_split) {
  library(ggplot2)
  library(mgcv)
  
  p <- ggplot(
    data = dat_subset |>
      dplyr::filter(year_from_fire_index >= 0),
    aes(x = year_from_fire_index, y = rap_TRE, col = {{ col_split }})
  ) +
    stat_smooth(method = "gam", formula = y ~ s(x, bs = "cs", k = 5), method.args = list(family = gaussian())) +
    ylim(0, 75) +
    xlim(0, 40) +
    xlab("Time since fire (years)") +
    ylab("Percent tree cover") +
    labs(
      title = title,
      caption = paste0("N: ", length(unique(dat_subset$`system:index`)))
    ) +
    theme_minimal()

  #print(p)
}


bf <- time_series_plot_recovery_only(
  dat_subset = dats_long |>
    filter(beneficial_fire != "not severe enough" &
             beneficial_fire != "replacement bf"),
  title = "All dats",
  col_split = beneficial_fire
)
bf <- bf +
  scale_color_manual(
    name = "Beneficial wildfire category",
    values = c("#dc0ab4ff", "navy"),
    labels = c("Low Severity Good Fire", "High Severity Detrimental Fire")
)
bf


```


```{r}
# Load required packages
library(splines)
library(mgcv)
library(ggplot2)

# Step 1: Generate sample data
set.seed(123)
n <- 200
x <- sort(runif(n, 0, 10))
y <- sin(x) + rnorm(n, sd = 0.3)
data <- data.frame(x = x, y = y)

# Step 2: Fit natural spline model and compute CIs
fit_spline <- lm(y ~ ns(x, df = 5), data = data)
pred_spline <- predict(fit_spline, newdata = data, se.fit = TRUE)
data$fit_spline <- pred_spline$fit
data$lower_spline <- pred_spline$fit - 1.96 * pred_spline$se.fit
data$upper_spline <- pred_spline$fit + 1.96 * pred_spline$se.fit

# Step 3: Fit GAM and compute CIs
fit_gam <- gam(y ~ s(x), data = data)
pred_gam <- predict(fit_gam, newdata = data, se.fit = TRUE)
data$fit_gam <- pred_gam$fit
data$lower_gam <- pred_gam$fit - 1.96 * pred_gam$se.fit
data$upper_gam <- pred_gam$fit + 1.96 * pred_gam$se.fit

# Step 4: Plot both fits with confidence intervals
ggplot(data, aes(x, y)) +
  geom_point(color = "gray70", alpha = 0.7) +
  
  # Spline fit and CI
  geom_ribbon(aes(ymin = lower_spline, ymax = upper_spline), fill = "blue", alpha = 0.2) +
  geom_line(aes(y = fit_spline), color = "blue", linetype = "dashed", size = 1) +
  
  # GAM fit and CI
  geom_ribbon(aes(ymin = lower_gam, ymax = upper_gam), fill = "red", alpha = 0.2) +
  geom_line(aes(y = fit_gam), color = "red", size = 1) +
  
  labs(title = "Natural Spline (blue dashed) vs GAM (red) with Confidence Intervals",
       x = "x", y = "y") +
  theme_minimal()




```




# Multiple linear regressions

The full model that we first try appears to be more or less all significant, with a stepwise reduction removing no features.

However, SRTM (elevation) appears to be quite collinear with our other variables. Let's remove it. Our adjusted R^2 doesn't change much, suggesting that the more parsimonious model should be used.

We can also try including the 2nd degree polynomial biotic_fire_gap that we identified above. When we do this we see that while the 2nd degree is significant, it doesn't improve our adjusted R^2 very much.


```{r multiple-linear-regressions}

# Check for NAs
col_vec <- c("post_fire_TRE_difference_5_yr", "post_fire_TRE_difference_10_yr", "post_fire_TRE_difference_15_yr", "aet", "def", "chili", "frg_reclass", "srtm", "burn_year_cbi", "biotic_sev_binned", "biotic_fire_cat", "avg_rap_TRE_6_yr_pre_fire")
na_counts <- dats_short %>%
  select(all_of(col_vec)) %>%
  summarise(across(everything(), ~sum(is.na(.)), .names = "na_{.col}"))
t(na_counts)


# When we add avg_rap_TRE_6_yr_pre_fire frg_reclass and the frg*cbi becomes highly colinear
# We get essentially the same results using the binned biotic data and our biotic category factor... not a good sign, there is very little additional useful information encoded in the biotic severity
lm_full <- lm(post_fire_TRE_difference_5_yr ~ aet + def + chili + frg_reclass + srtm + burn_year_cbi + biotic_sev_binned + biotic_fire_gap_binned + burn_year_cbi * frg_reclass + avg_rap_TRE_6_yr_pre_fire,
              data = dats_short)
summary(lm_full) #0.733 r2

lm_full <- lm(post_fire_TRE_difference_5_yr ~ aet + def + chili + frg_reclass + srtm + burn_year_cbi + biotic_fire_cat + avg_rap_TRE_6_yr_pre_fire,
              data = dats_short)
summary(lm_full) #0.733 r2

reduced_lm_full <- step(lm_full)

faraway::vif(reduced_lm_full)

reduced_lm_full <- lm(post_fire_TRE_difference_5_yr ~ chili +
                        frg_reclass + 
                        def + 
                        srtm +
                        burn_year_cbi + 
                          biotic_sev_binned + 
                          biotic_fire_gap_binned + 
                        avg_rap_TRE_6_yr_pre_fire,
    data = dats_short) #remove aet from step and interaction for high vif

summary(reduced_lm_full) # a-r2: 0.73


# Test mlr with biotic info but without CBI - adjusted R^2 drops from 0.73 to 0.48.
lm_no_cbi <- lm(post_fire_TRE_difference_5_yr ~ chili +
                        frg_reclass + def + srtm +
    biotic_sev_binned + 
      biotic_fire_gap_binned + 
    avg_rap_TRE_6_yr_pre_fire,
    data = dats_short)
summary(lm_no_cbi)

# With CBI but not biotic - 0.73
lm_no_bio <- lm(post_fire_TRE_difference_5_yr ~ chili +
                        frg_reclass + def + srtm +
    burn_year_cbi + 
    avg_rap_TRE_6_yr_pre_fire,
    data = dats_short)
summary(lm_no_bio)

# Test mlr without cbi - without CBI OR  biotic info adjusted R^2 drops to 0.01; with all biotic info added back in it goes back up to 0.44. But if we only have biotic_fire_cat it only goes back up to 0.02. So there's important information coded in biotic_year_sev and biotic_fire_gap. When we try adding these individually we get an r^2 of 0.17 with only biotic_fire_gap, and an r^2 of 0.1 with only biotic_year_sev. Biotic fire gap + biotic_year_sev brings us up to 0.18.
lm_no_cbi_or_biotic <- lm(post_fire_TRE_difference_5_yr ~ aet + def + chili + frg_reclass + srtm,
              data = dats_short)
summary(lm_no_cbi_or_biotic)











# Test on other response variables
lm_full <- lm(post_fire_TRE_difference_10_yr ~ aet + def + chili + frg_reclass + srtm + burn_year_cbi + biotic_sev_binned + biotic_fire_gap_binned + burn_year_cbi * frg_reclass + avg_rap_TRE_6_yr_pre_fire + tpi,
              data = dats_short)
summary(lm_full) 

reduced_lm_full <- step(lm_full)

faraway::vif(reduced_lm_full)

reduced_lm_full <- lm(post_fire_TRE_difference_10_yr ~ chili +
                        #frg_reclass + 
                        #srtm +
                        #def + 
                        aet +
                        tpi +
                        burn_year_cbi + 
                          biotic_sev_binned + 
                          biotic_fire_gap_binned + 
                        avg_rap_TRE_6_yr_pre_fire,
    data = dats_short) #remove aet from step and interaction for high vif
summary(reduced_lm_full)
faraway::vif(reduced_lm_full)


reduced_lm_full <- lm(post_fire_TRE_difference_10_yr ~ chili +
                        aet +
                        tpi +
                          biotic_sev_binned + 
                          biotic_fire_gap_binned + 
                        avg_rap_TRE_6_yr_pre_fire,
    data = dats_short) 
summary(reduced_lm_full)















lm_full <- lm(post_fire_TRE_difference_15_yr ~ aet + def + chili + frg_reclass + srtm + burn_year_cbi + biotic_sev_binned + biotic_fire_gap_binned + burn_year_cbi * frg_reclass + avg_rap_TRE_6_yr_pre_fire,
              data = dats_short)
summary(lm_full) 


#The models are REALLY bad at predicting the recovery slope...
# A) probably need to aggregate the RAP data etc to appropriate resolution for insect interaction
# B) There's likely some drought impacts?...
lm_full <- lm(post_fire_rap_slope_2_10_yrs ~ aet + def + chili + frg_reclass + srtm + burn_year_cbi + biotic_sev_binned + biotic_fire_gap_binned + burn_year_cbi * frg_reclass + avg_rap_TRE_6_yr_pre_fire,
              data = dats_short)
summary(lm_full) 

lm_full <- lm(post_fire_rap_slope_2_15_yrs ~ aet + def + chili + frg_reclass + srtm + burn_year_cbi + biotic_sev_binned + biotic_fire_gap_binned + burn_year_cbi * frg_reclass + avg_rap_TRE_6_yr_pre_fire,
              data = dats_short)
summary(lm_full) 



# With 2nd degree polynomial fire gap
lm_reduced_poly <- lm(post_fire_TRE_difference_5_yr ~ aet + def + chili + frg_reclass + burn_year_cbi + biotic_year_sev + poly(biotic_fire_gap, 2) + biotic_fire_cat,
              data = dats_short  |>
      filter(!is.na(biotic_fire_gap)))
summary(lm_reduced_poly)


# Partial and marginal effects

plot(effects::allEffects(reduced_lm_full))
car::crPlots(reduced_lm_full, terms = ~ burn_year_cbi)
car::crPlots(reduced_lm_full, terms = ~ biotic_year_sev)
car::crPlots(reduced_lm_full, terms = ~ biotic_fire_cat)


lm_reduced_poly_raw <- lm(
  post_fire_TRE_difference_5_yr ~ 
    aet + def + chili + frg_reclass + burn_year_cbi + 
    biotic_year_sev + poly(biotic_fire_gap, 2, raw = TRUE) + 
    biotic_fire_cat,
  data = dats_short |> filter(!is.na(biotic_fire_gap))
)

# CR plot for both biotic_fire_gap and biotic_fire_gap^2
crPlots(lm_reduced_poly_raw, terms = ~ poly(biotic_fire_gap, 2, raw = TRUE))









lm_full <- lm(post_fire_TRE_difference_5_yr ~ aet + def + chili + frg_reclass + burn_year_cbi + biotic_fire_cat + frg_reclass * burn_year_cbi,
              data = dats_short)
summary(lm_full)
crPlots(lm_full, terms = ~ frg_reclass2:burn_year_cbi)



lm_full <- lm(post_fire_TRE_difference_10_yr ~ aet + def + chili + frg_reclass + srtm + burn_year_cbi + biotic_year_sev + biotic_fire_gap + biotic_fire_cat,
              data = dats_short)
summary(lm_full)

lm_full <- lm(post_fire_TRE_difference_15_yr ~ aet + def + chili + frg_reclass + srtm + burn_year_cbi + biotic_year_sev + biotic_fire_gap + biotic_fire_cat,
              data = dats_short)
summary(lm_full)



#model CBI
lm_full <- lm(burn_year_cbi ~ avg_rap_TRE_6_yr_pre_fire + aet + def + chili + frg_reclass + srtm + biotic_sev_binned + biotic_fire_gap_binned,
              data = dats_short |>
                filter(biotic_fire_cat != 'biotic_1_5_after'))
summary(lm_full)
```






# VALIDATE AGAINST VCF
# ADD FIRE HETEROGENEITY
# FILTER TO ONLY INCLUDE HIGH SEV BIOTIC



# Thinking spatially

- Test variogram
- Try spatial lag and spatial error models

```{r spatial}

# sample variogram based on Cressie method, it is often used to account for outliers
png(filename = here::here(dir_figs, "p_vgm_plot.png"), width = 800, height = 600, res = 150)
p.vgm <- variogram(post_fire_TRE_difference_10_yr ~ 1, data = dats_short_sf |>
                     filter(!is.na(post_fire_TRE_difference_10_yr)), cressie=T)
plot(p.vgm, type='l')
dev.off()


# Spatial modeling
# Select columns, keeping outcome_var by name
dats_short_sp_ready <- dats_short_sf |>
  select(all_of(c("post_fire_TRE_difference_10_yr", "chili", "frg_reclass", "def", "aet", "srtm", 
                  "burn_year_cbi", "biotic_sev_binned", "tpi",
                  "biotic_fire_gap_binned", "avg_rap_TRE_6_yr_pre_fire", "biotic_fire_cat"))) |>
  drop_na() |>
  filter(biotic_fire_cat != "no_biotic_or_outside_timeframe")

#Create clean LM to use

reduced_lm_full <- lm(post_fire_TRE_difference_10_yr ~ 
                        chili +
                        aet +
                        tpi +
                        burn_year_cbi + 
                          biotic_sev_binned + 
                          biotic_fire_gap_binned + 
                        avg_rap_TRE_6_yr_pre_fire,
    data = dats_short_sp_ready) #remove aet from step and interaction for high vif
summary(reduced_lm_full)

reduced_lm_full_no_cbi <- lm(post_fire_TRE_difference_10_yr ~ 
                        chili +
                        aet +
                        tpi +
                        #burn_year_cbi + 
                          biotic_sev_binned + 
                          biotic_fire_gap_binned + 
                        avg_rap_TRE_6_yr_pre_fire,
    data = dats_short_sp_ready) #remove aet from step and interaction for high vif
summary(reduced_lm_full_no_cbi)
  
reduced_lm_cbi <- lm(burn_year_cbi ~ 
                       chili +
                        aet +
                        tpi +
                          biotic_sev_binned + 
                          biotic_fire_gap_binned + 
                        avg_rap_TRE_6_yr_pre_fire,
    data = dats_short_sp_ready) #remove aet from step and interaction for high vif
summary(reduced_lm_cbi)
  
  
# Create weights matrix
tic()
coords <- sf::st_coordinates(dats_short_sp_ready)

# 8 nearest neighbors
knn_8 <- knearneigh(coords, k = 8)
nb_8 <- knn2nb(knn_8)

# Create a spatial weights list object (row-standardized by default)
lw_8 <- nb2listw(nb_8, style = "W", zero.policy = TRUE)
toc()


# Run Moran's I
#The result shows the value Moran’s I is about 0.49 with a highly significant p-value, indicating a significant positive spatial autocorrelation
moran_result <- moran.test(dats_short_sp_ready$post_fire_TRE_difference_10_yr, lw_8, zero.policy = TRUE)
print(moran_result)



# Run LM test to see whether we need an error or lag model
lm_test <- spdep::lm.LMtests(model = reduced_lm_full,
                             listw = lw_8,
                             test = "all")
summary(lm_test) #We should use a spatial error model

lm_test <- spdep::lm.LMtests(model = reduced_lm_full_no_cbi,
                             listw = lw_8,
                             test = "all")
summary(lm_test) #We should use a spatial error model


lm_test <- spdep::lm.LMtests(model = reduced_lm_cbi,
                             listw = lw_8,
                             test = "all")
summary(lm_test) #We should use a spatial error model


# Tries to allocate large amounts of RAM; not a scalable solution
tic()
errorsar <- spatialreg::errorsarlm(post_fire_TRE_difference_10_yr ~
                        chili +
                        #frg_reclass +
                        #def +
                        aet +
                        tpi +
                        burn_year_cbi +
                          biotic_sev_binned +
                          biotic_fire_gap_binned +
                        avg_rap_TRE_6_yr_pre_fire,
    data = dats_short_sp_ready,
                               listw = lw_8)
summary(errorsar)
toc()

tic()
errorsar <- spatialreg::errorsarlm(post_fire_TRE_difference_10_yr ~
                        chili +
                        #frg_reclass +
                        #def +
                        aet +
                        tpi +
                        #burn_year_cbi +
                          biotic_sev_binned +
                          biotic_fire_gap_binned +
                        avg_rap_TRE_6_yr_pre_fire,
    data = dats_short_sp_ready,
                               listw = lw_8)
summary(errorsar)
toc()

tic()
errorsar <- spatialreg::errorsarlm(burn_year_cbi ~ chili +
                        frg_reclass +
                        def +
                        aet +
                        tpi +
                          biotic_sev_binned +
                          biotic_fire_gap_binned +
                        avg_rap_TRE_6_yr_pre_fire,
    data = dats_short_sp_ready,
                               listw = lw_8)
summary(errorsar)
toc()







```




# Tree-based methods

- Try RF, sRF, and boosted regression



```{r}


dats_short_rf <- dats_short_sf |>
  select(all_of(c("post_fire_TRE_difference_10_yr", "chili", "frg_reclass", "def", "aet", "srtm", 
                  "burn_year_cbi", "biotic_sev_binned", "tpi", "biotic_fire_cat",
                  "biotic_fire_gap_binned", "avg_rap_TRE_6_yr_pre_fire"))) |>
  drop_na()


run_rf <- function(dats, outcome_var) {
  set.seed(1)
  
  train <- sample(1:nrow(dats), nrow(dats)/2)
  
  d_test <- dats |> 
    slice(-train) |> 
    pull(outcome_var)  
  
  # Construct formula from outcome_var
  rf_formula <- as.formula(
    paste(outcome_var, "~ chili + frg_reclass + def + srtm + burn_year_cbi +
          biotic_sev_binned + biotic_fire_gap_binned + avg_rap_TRE_6_yr_pre_fire")
    # paste(outcome_var, "~ .")
  )
  
  rf <- randomForest(
    formula = rf_formula,
    data = dats,
    subset = train,
    mtry = 12 / 3,
    importance = TRUE
  )
  
  plot(rf)
  yhat_rf <- predict(rf, newdata = dats[-train, ])
  mse_rf <- mean((yhat_rf - d_test)^2)
  print(mse_rf)
  
  print(importance(rf))
  varImpPlot(rf)
  
  return(rf)
}


run_rf(dats_short_rf, "post_fire_TRE_difference_10_yr")

tic()
run_rf(dats_short_rf, "post_fire_rap_slope_2_10_yrs")
toc()



tic()
run_rf(dats_short_rf |>
         filter(biotic_fire_cat != "no_biotic_or_outside_timeframe"),
       "post_fire_TRE_difference_10_yr")
toc()

tic()
run_rf(dats_short_rf |>
         filter(biotic_fire_cat != "no_biotic_or_outside_timeframe"),
       "post_fire_TRE_difference_10_yr")
toc()


```



# Additional things

- Fixed effect for fire ID!
- Host-agent relationship
- Forest type
- Ecoregion level fixed effects (L3 and L4)







