---
title: "05_dydid"
author: Tyler L. McIntosh
date: last-modified
format:
  html:
    embed-resources: true
    html-math-method: katex
    css: styles.css
    toc: true
    toc-depth: 1
    toc-title: Contents
    toc-location: right
    code-fold: true
    code-summary: "Show the code"
---

# Setup for analysis

``` {r setupchunk}
#| echo: true
#| message: false
#| warning: false
#| results: 'hide'

rm(list = ls())

if(!requireNamespace("here", quietly = TRUE)) {
  install.packages("here")
}
library(here)

source(here::here("utils", "functions.R"))

install_load_packages(c(
  "tidyverse",
  "tictoc",
  "hstats",
  "glue",
  "shapviz",
  "patchwork",
  "cobalt",
  "purrr",
  "corrplot",
  "gridExtra",
  "tictoc",
  "fect",
  "fixest",
  "did",
  "panelView",
  "PanelMatch",
  "bacondecomp",
  "didimputation",
  "doParallel",
  "HonestDid",
  "HonestDidFEct",
  "DIDmultiplegtDYN",
  "sf",
  "spdep",
  "purrr",
  "gstat",
  "scales",
  "broom",
  "ggfixest"))

# # install packages from CRAN
# packages <- c("fixest", "did", "didimputation", 
#          "panelView", "bacondecomp", "HonestDiD",
#               "DIDmultiplegtDYN", "PanelMatch")
# install.packages(setdiff(packages, rownames(installed.packages())))  
# 
# # install most up-to-date "fect" from Github
# if ("fect" %in% rownames(installed.packages()) == FALSE) {
#   devtools:: install_github("xuyiqing/fect")
# }
# 
# # install forked "HonestDiD" package compatible with "fect"
# if ("HonestDiDFEct" %in% rownames(installed.packages()) == FALSE) {
#   devtools:: install_github("lzy318/HonestDiDFEct")
# }


dir_figs <- here::here('figs')
dir_derived <- here::here('data', 'derived')
dir_raw <- here::here('data', 'raw')
dir_manual <- here::here('data', 'manual')
dir_ensure(c(dir_figs,
             dir_derived,
             dir_manual,
             dir_raw))

seed = 1234
set.seed(seed)


```

#

```{r}
  
dats_long <- read_csv(here(dir_derived, "test_dats_long.csv"))
dats_short <- read_csv(here(dir_derived, "test_dats_short.csv"))


```



# Test key assumptions

- Parallel trends
- No anticipatory behavior
- Treatment effect heterogeneity - “Since treatment effects heterogeneity violates Assumption 3 and can alter how we interpret μg , it is important to think through when different cohorts likely experience different paths of treatment effect. Such heterogeneity could arise for many reasons. For example, cohorts may differ in their covariates, which affect how they respond to treatment. We will explore a concrete example in our application: if treatment effects differ with age, and there is variation in age across units first treated at different times, we will have heterogeneous effects (see Section 5 for details)” ([Sun and Abraham, 2021, p. 182](zotero://select/library/items/86R62KQ4)) ([pdf](zotero://open-pdf/library/items/3WRC5CP7?page=8&annotation=QUPBLF69))

IT APPEARS THAT WE HAVE A PROBLEM OF HETEROGENEOUS TREATMENT EFFECTS - We need to use an interaction-weighted estimator (i.e. dynamic DiD from Sun & Abraham)



```{r}


# Parallel trends - see results from 04_causal_forest


# Heterogeneous treatment effects?

ggplot() +
  geom_boxplot(data = dats_short |> filter(fire == 1), aes(x = as.factor(burn_year), y = biotic_5_yrs_prior_sum))

ggplot() +
  geom_boxplot(data = dats_short |> filter(fire == 1), aes(x = as.factor(burn_year), y = biotic_5_yrs_after_sum))

ggplot() +
  geom_boxplot(data = dats_short |> filter(fire == 1), aes(x = as.factor(burn_year), y = pdsi_annual_5_yrs_prior_sum))

ggplot() +
  geom_boxplot(data = dats_short |> filter(fire == 1), aes(x = as.factor(burn_year), y = pdsi_annual_5_yrs_after_sum))

ggplot() +
  geom_boxplot(data = dats_short |> filter(fire == 1), aes(x = as.factor(burn_year), y = chili))

ggplot() +
  geom_boxplot(data = dats_short |> filter(fire == 1), aes(x = as.factor(burn_year), y = def))


# Number of points for disturbance interactions
n_bio_greater_10 <- dats_short |>
  filter(fire == 1 & biotic_5_yrs_prior_sum >= 10)
n_hd_greater_17 <- dats_short |>
  filter(fire == 1 & hd_fingerprint_5_yrs_after_sum >= 17)

n <- dats_short |>
  filter(fire == 1 & pdsi_annual_5_yrs_after_threshold >= 1)


only_burned <- dats_short |>
  filter(fire == 1)

ggplot(only_burned, aes(x = biotic_5_yrs_prior_sum, y = burn_cbi)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", se = TRUE, color = "blue") +
  labs(
    x = "Biotic Sum (5 Years Prior)",
    y = "Burn Severity",
    title = "Linear Relationship: Burn Severity vs Biotic Sum"
  ) +
  theme_minimal()



burn_sev_model <- lm(burn_cbi ~ biotic_5_yrs_prior_sum + def + aet + avg_forest_2_yr_pre_fire + pdsi_annual_5_yrs_prior_sum + tpi + ecoregion + chili + frg_reclass, data = only_burned)
summary(burn_sev_model)

```

```{r}



# #Visualize treatment status over time
# panelview(rap_TRE ~ treated, data = testset, index = c("uniqueID", "year"),
#   xlab = "Year", ylab = "Unit", display.all = T,
#   gridOff = TRUE, by.timing = TRUE)
# 
# 
# hist(testset |> filter(fire_timeframe == "year_of_burn") |> pull(year))

```

# Run Dynamic DiD with Sun-Abraham estimators and two-way fixed effects

PROBLEM: Our error bars pre-event are often NOT covering zero; this suggests that we are violating the parallel trends assumption. We are missing important covariates (?) and/or need to perform pre-fire matching/weighting.



```{r}

# model.twfe.0 <- feols(rap_TRE~treated|uniqueID+year,
#                       data=dats_long, cluster = "uniqueID") 
# print(model.twfe.0)


# # We don't have any always-treated units, so we don't need to worry about the Goodman-Bacon decomposition


dats_long_did <- get.cohort(dats_long, D = "treated",
                                 index = c("uniqueID", "year"),
                                 start0 = TRUE)
dats_long_did[which(is.na(dats_long_did$FirstTreat)),"FirstTreat"] <- 1000 
df.use <- dats_long_did |>
  mutate(biotic_before_10t = case_when(biotic_5_yrs_prior_sum >= 10 ~ "biotic_prior_10t",
                                  biotic_5_yrs_prior_sum < 10 ~ "no_biotic_prior_10t"),
         biotic_before_20t = case_when(biotic_5_yrs_prior_sum >= 20 ~ "biotic_prior_20t",
                                  biotic_5_yrs_prior_sum < 20 ~ "no_biotic_prior_20t"),
         hd_after_17t = case_when(hd_fingerprint_5_yrs_after_sum >= 17 ~ "hd_after_17t",
                              hd_fingerprint_5_yrs_after_sum < 17 ~ "no_hd_after_17t"),
         pdsi_before_n500t = case_when(pdsi_annual_5_yrs_prior_sum <= -500 ~ "pdsi_before_n500t",
                                 pdsi_annual_5_yrs_prior_sum > -500 ~ "no_pdsi_before_n500t"),
         pdsi_before_n1000t = case_when(pdsi_annual_5_yrs_prior_sum <= -1000 ~ "pdsi_before_n1000t",
                                 pdsi_annual_5_yrs_prior_sum > -1000 ~ "no_pdsi_before_n1000t"))



# 
# n_bio_greater_10 <- dats_short |>
#     filter(ecoregion == "IDBatholith") |>
#   filter(fire == 1 & biotic_5_yrs_prior_sum >= 10)
# n_bio_greater_10_control <- dats_short |>
#     filter(ecoregion == "IDBatholith") |>
#   filter(fire == 0 & biotic_5_yrs_prior_sum >= 10)

# # Dynamic TWFE
# df.twfe <- df.use
# 
# # drop always treated units
# #df.twfe$treat <- as.numeric(df.twfe$treatment_mean>0) 
# df.twfe[which(is.na(df.twfe$Time_to_Treatment)),'Time_to_Treatment'] <- -999 # can be an arbitrary value
# twfe.est <- feols(rap_TRE ~ i(Time_to_Treatment, fire, ref = c(-1, -999))| uniqueID + year, 
#                   data = df.twfe,
#                   cluster = "uniqueID")
# 
# twfe.output <- as.matrix(twfe.est$coeftable)
# print(round(twfe.output, 3))
# 
# twfe.est$collin.var
# 
# table(df.twfe$treated, useNA = "always")
# table(df.twfe$treated, is.na(df.twfe$Time_to_Treatment))
# 
# 
# twfe.output <- as.data.frame(twfe.output)
# twfe.output$Time <- c(c(-29:-2),c(0:21))+1 
# p.twfe <- esplot(twfe.output,Period = 'Time',Estimate = 'Estimate',
#                                SE = 'Std. Error', xlim = c(-29,21))
# p.twfe




# 
# model <- feols(
#   rap_TRE ~ sunab(FirstTreat, year) | year + uniqueID,
#   data = testset |>
#     filter(biotic_before == 0),
#   cluster = "uniqueID"
# )
# summary(model)
# plot_sunab_models(model_list = list(model),
#                   model_labels = "test2 - no biotic before")


```

# Functions
```{r}

# Multi-model functions ----


# This function runs a sun-abraham dynamic difference-in-difference analysis on dataframe df using ref_period.
# It will compute the model using all 0 values in trt_col (treatment column) as the control and all 1 values in trt_col with grp in subgrp_col as the estimating group 
iw_estimator_subg <- function(subgrp_col,
                              grp,
                              trt_col = "fire",
                              ref_period = -1,
                              df) {
  # Convert string to symbol
  col_sym <- sym(subgrp_col)
  trt_col_sym <- sym(trt_col)
  
  df_never_trt <- df |>
    filter(!!trt_col_sym == 0)
  
  df_subset <- df |>
    filter(!!trt_col_sym == 1 & !!col_sym == grp)
  
  df_use <- bind_rows(df_never_trt, df_subset)

  model <- feols(
    rap_TRE ~ sunab(FirstTreat, year, ref.p = ref_period) | uniqueID + year,
    data = df_use,
    mem.clean = TRUE,
    lean = FALSE,
    vcov = fixest::vcov_conley(lat = "lat",
                               lon = "long",
                               cutoff = 125,
                               pixel = 20,
                               distance = "triangular"
                               )
  )
  
  n_treated_subgroup_units <- length(unique(df_subset$uniqueID))
  n_never_treated_units <- length(unique(df_never_trt$uniqueID))
  attr(model, "n_treated_subgroup_units") <- n_treated_subgroup_units
  attr(model, "n_never_treated_units") <- n_never_treated_units
  
  return(model)
}



# Model list and model_labels should be the same length!
# If item in model_list contains attribute "n_treated_subgroup_units", the attribute will be appended to the model name in the legend

plot_sunab_models <- function(model_list,
                              model_labels = NULL,
                              title = "Dynamic-DiD Event-Study Estimates from Multiple Models",
                              ribbon_type = "ribbon", #either "ribbon" or "dashed",
                              caption = "95% confidence intervals shown;\nInteraction-weighted (Sun-Abraham) estimates used;\nConley standard errors with 125km cutoff, 20km pixel"
                              ) {
  
  # If no labels provided, auto-label as Model 1, Model 2, etc.
  if (is.null(model_labels)) {
    model_labels <- paste0("Model ", seq_along(model_list))
  }
  
  # Extract coefficients for each model and attach model label
  plot_data <- map2_dfr(model_list, model_labels, function(model, label) {
    df <- as.data.frame(as.matrix(model$coeftable))
    df$term <- rownames(model$coeftable)
    
    has_attr <- function(x, attr_name) {
      !is.null(attr(x, attr_name))
    }
    
    if (has_attr(model, "n_treated_subgroup_units")) {
      label <- glue("{label} (N = {attr(model, 'n_treated_subgroup_units')})")
    }
    
    df %>%
      filter(str_detect(term, "^year::")) %>%
      mutate(
        Time = as.numeric(str_extract(term, "-?\\d+")),
        CI_upper = Estimate + 1.96 * `Std. Error`,
        CI_lower = Estimate - 1.96 * `Std. Error`,
        Model = label
      )
  })

  # Plot
  if(ribbon_type == "ribbon") {
    ggplot(plot_data, aes(x = Time, y = Estimate, color = Model, fill = Model)) +
      geom_line(linewidth = 1) +
      geom_point(size = 2) +
      geom_ribbon(aes(ymin = CI_lower, ymax = CI_upper), alpha = 0.2, color = NA) +
      geom_vline(xintercept = -1, linetype = "dashed", color = "gray40") +
      geom_hline(yintercept = 0, color = "black") +
      labs(
        x = "Event Time",
        y = "Treatment Effect of fire on forest cover",
        title = title,
        caption = caption
      ) +
      theme_minimal(base_size = 13) +
      theme(legend.position = "bottom")    
  } else {
    ggplot(plot_data, aes(x = Time, y = Estimate, color = Model)) +
      geom_line(linewidth = 1) +
      geom_point(size = 2) +
      
      # CI as dotted lines instead of ribbon
      geom_line(aes(y = CI_upper), linetype = "dotted", linewidth = 0.8) +
      geom_line(aes(y = CI_lower), linetype = "dotted", linewidth = 0.8) +
      
      geom_vline(xintercept = -1, linetype = "dashed", color = "gray40") +
      geom_hline(yintercept = 0, color = "black") +
      labs(
        x = "Event Time",
        y = "Treatment Effect of fire on forest cover",
        title = title,
        caption = caption
      ) +
      theme_minimal(base_size = 13) +
      theme(legend.position = "bottom") +
      xlim(c(-10, 20))
  }

}


plot_all_together_did <- function(subg_col, df, title = NULL) {
  
  if(is.null(title)) {
    title = glue("Effect estimates by subgroup in column {subg_col}")
  }
  
  subgs <- unique(df |>
                    pull(subg_col))
  
  subgs <- subgs[!is.na(subgs)]

  
    # Fit models for each subgroup level
  iw_results <- subgs %>%
    set_names() %>%
    map(~ iw_estimator_subg(grp = .x,
                            subgrp_col = subg_col,
                            df = df))
  
  # Generate the overlay plot
  p <- plot_sunab_models(
    model_list = iw_results,
    model_labels = subgs,
    title = title
  )
  return(p)
}


plot_ecoregion_did <- function(ecoregion_name, subg_col, df) {
  df_ecoregion <- df %>% filter(ecoregion == ecoregion_name)
  
  if(subg_col == "burn_sev_tagged") {
    subgs <- c("low-moderate-sev", "unburned", "high-sev")
  } else {
    subgs <- unique(df_ecoregion |> pull(subg_col))
  }
  
  # Fit models for each subgroup level
  iw_results <- subgs %>%
    set_names() %>%
    map(~ iw_estimator_subg(grp = .x,
                            subgrp_col = subg_col,
                            df = df_ecoregion))
  
  # Generate the overlay plot
  p <- plot_sunab_models(
    model_list = iw_results,
    model_labels = subgs,
    title = ecoregion_name
  )
  return(p)
}

plot_all_ecoregions_did <- function(subg_col, df, title = "Interaction-Weighted Estimates by Ecoregion") {
  # Get unique ecoregions
  ecoregions <- unique(df$ecoregion)
  
  # Generate a plot per ecoregion
  ecoregion_plots <- map(ecoregions, ~ plot_ecoregion_did(ecoregion_name = .x,
                                                          subg_col = subg_col,
                                                          df = df))
  
  # Combine into grid
  final_plot <- wrap_plots(ecoregion_plots, ncol = 2) +
    plot_annotation(title = title )
  
  return(final_plot)
  
}
```



# Test for spatial autocorrelation in residuals
- We find that there is significant autocorrelation in the residuals of all test ecoregions, in all years
- Semivariograms for each year/ecoregion finds different patterns
- Conley error sensitivity analyses show different peaks for each ecoregion; might be worth using a different cutoff for each. If we were to choose a single universal cutoff, ~125km looks good (?)
- We find that the Conley standard errors don't change very much by altering pixel (i.e. grid aggregation of residuals), so we may as well use 20km (more computationally efficient)

```{r}

dir_sptests <- here(dir_figs, "spatial_dependence_tests")
dir_ensure(dir_sptests)

ecoregions <- unique(df.use$ecoregion)

#Function to apply moran's I tests to each year and compute semivariograms
test_spdep_ecoregion <- function(ecor, dats, dir_results) {
  testset <- dats |>
    filter(ecoregion == ecor)
  model <- feols(
    rap_TRE ~ sunab(FirstTreat, year) | uniqueID + year,
    data = testset,
    cluster = "uniqueID")
  
  # Test for spatial autocorrelation in residuals - attach residuals to data
  res <- resid(model)
  testset$residuals <- res
  
  # Make sf data
  sf_dat <- st_as_sf(testset, coords = c("long", "lat"), crs = 4326) |>
    st_transform(5070)
  
  # Compute moran results
  moran_results <- sf_dat |>
    group_by(year) |>
    group_modify(~ {
      nb <- knearneigh(st_coordinates(.x), k = 5) |>
        knn2nb()
      lw <- nb2listw(nb, style = "W")
  
      m <- moran.test(.x$residuals, lw)
      tibble(statistic = m$statistic, p.value = m$p.value)
    })
  
  write_csv(moran_results, here(dir_results, glue("moran_results_{ecor}.csv")))
  
  # 2) Compute empirical variogram per year
  #    We pick a year-specific cutoff = half the spatial diagonal,
  #    and ~15 bins
  vg_emp <- sf_dat |>
    group_by(year) |>
    group_modify(~{
      d <- .x |> filter(!is.na(residuals))
      n <- nrow(d)
      if (n < 10) return(tibble(dist = numeric(), gamma = numeric(), np = integer()))
      
      bb <- st_bbox(d)
      diag_len <- sqrt((bb["xmax"] - bb["xmin"])^2 + (bb["ymax"] - bb["ymin"])^2)
      cutoff <- 0.5 * diag_len
      width  <- cutoff / 15
  
      v <- gstat::variogram(residuals ~ 1, data = d, cutoff = cutoff, width = width)
      as_tibble(v)  # columns: dist, gamma, dir.hor, dir.ver, np, ...
    }) |>
    ungroup()
  
  ggplot(vg_emp, aes(dist, gamma)) +
    geom_point(aes(size = np), alpha = 0.8) +
    geom_line() +
    facet_wrap(~ year, scales = "free") +
    scale_x_continuous(labels = label_comma()) +   # ← HERE
    labs(
      x = "Distance (m)",
      y = "Semivariance γ(h)",
      size = "Pairs (np)"
    ) +
    theme_bw()
  
  ggsave(filename = here(dir_results, glue("variograms_{ecor}.png")), units = "px", width = 6000, height = 6000)
    
}




test_conley <- function(dats,
                        ecor,
                        cutoffs,
                        pixel,
                        distance,
                        dir_results,
                        flnm = "conley_test",
                        event_years = c(5, 10, 15),
                        vcov_fix = TRUE) {
  # deps guard (optional)
  if (!requireNamespace("fixest", quietly = TRUE)) stop("fixest required.")
  if (!dir.exists(dir_results)) dir.create(dir_results, recursive = TRUE)

  # 0) data subset
  testset <- dats |>
    dplyr::filter(ecoregion == ecor)

  # 1) fit DiD ONCE (ID-cluster baseline), keep object light
  base_model <- fixest::feols(
    rap_TRE ~ sunab(FirstTreat, year) | uniqueID + year,
    data = testset,
    cluster = "uniqueID",
    lean = FALSE
  )

  # Helper to extract Sun–Abraham rows from a coeftable -> slim tibble
  extract_sa_rows <- function(ct, model_label, term_pattern = "^year::", ci_z = 1.96) {
    rn <- rownames(ct)
    keep <- grepl(term_pattern, rn, perl = TRUE)
    if (!any(keep)) return(tibble::tibble())

    se_col <- if ("Std. Error" %in% colnames(ct)) "Std. Error" else colnames(ct)[2]

    tibble::tibble(
      term   = rn[keep],
      Time   = as.numeric(sub(".*year::(-?\\d+).*", "\\1", rn[keep])),
      Estimate = ct[keep, "Estimate", drop = TRUE],
      SE       = ct[keep, se_col,    drop = TRUE],
      CI_lower = ct[keep, "Estimate", drop = TRUE] - ci_z * ct[keep, se_col, drop = TRUE],
      CI_upper = ct[keep, "Estimate", drop = TRUE] + ci_z * ct[keep, se_col, drop = TRUE],
      Model    = model_label
    )
  }

  # 2a) Baseline (clustered-by-ID) plot-data (no Conley)
  base_ct <- fixest::coeftable(base_model)  # uses clustered SEs from fit
  plot_rows <- list(
    extract_sa_rows(base_ct, model_label = "ID_cluster")
  )
  rm(base_ct); gc(verbose = FALSE)

  # 2b) Stream through Conley cutoffs:
  #     for each cutoff: build vcov -> get coeftable -> append slim rows -> drop vcov immediately
  se_rows <- vector("list", length(cutoffs))   # for SE-vs-cutoff of event_years
  for (i in seq_along(cutoffs)) {
    c0 <- cutoffs[i]

    # Heavy step
    vc <- fixest::vcov_conley(
      base_model,
      lat = "lat", lon = "long",
      cutoff = c0, pixel = pixel, distance = distance, vcov_fix = vcov_fix
    )

    # Lightweight table from that vcov
    ct <- fixest::coeftable(base_model, vcov = vc)

    # Immediately free the big vcov
    rm(vc); gc(verbose = FALSE)

    # Append slim rows for the comparison plot
    plot_rows[[length(plot_rows) + 1L]] <- extract_sa_rows(ct, model_label = paste0("Conley_", c0))

    # Extract SEs for requested event_years and stash tiny tibble
    rn   <- rownames(ct)
    pat  <- paste0("^year::(", paste(event_years, collapse = "|"), ")\\b")
    keep <- grepl(pat, rn, perl = TRUE)
    if (any(keep)) {
      se_col <- if ("Std. Error" %in% colnames(ct)) "Std. Error" else colnames(ct)[2]
      se_rows[[i]] <- tibble::tibble(
        coef   = rn[keep],
        event  = as.integer(sub("^year::(-?\\d+).*", "\\1", rn[keep])),
        se     = ct[keep, se_col, drop = TRUE],
        cutoff = c0
      )
    } else {
      se_rows[[i]] <- tibble::tibble(coef = character(), event = integer(), se = numeric(), cutoff = numeric())
    }

    # Drop the coeftable too (small, but keep peak low)
    rm(ct); gc(verbose = FALSE)
  }

  # Bind the small tibbles
  plot_data <- dplyr::bind_rows(plot_rows)
  se_df     <- dplyr::bind_rows(se_rows)

  # 3) Comparison plot (baseline + all Conley), built from slim plot_data (no model objects kept)
  p_models <- ggplot2::ggplot(
    plot_data,
    ggplot2::aes(x = Time, y = Estimate, color = Model)
  ) +
    ggplot2::geom_line(linewidth = 1) +
    ggplot2::geom_point(size = 2) +
    # CI as dotted bounds (lighter than ribbons)
    ggplot2::geom_line(ggplot2::aes(y = CI_upper), linetype = "dotted", linewidth = 0.7) +
    ggplot2::geom_line(ggplot2::aes(y = CI_lower), linetype = "dotted", linewidth = 0.7) +
    ggplot2::geom_vline(xintercept = -1, linetype = "dashed", color = "gray40") +
    ggplot2::geom_hline(yintercept = 0, color = "black") +
    ggplot2::labs(
      x = "Event Time",
      y = "Treatment Effect of fire on forest cover",
      title = "Dynamic-DiD Event-Study Estimates (Baseline vs Conley cutoffs)",
      caption = "95% CIs; baseline uses clustered-by-unit SEs; Conley varies by cutoff"
    ) +
    ggplot2::theme_minimal(base_size = 13) +
    ggplot2::theme(legend.position = "bottom")

  ggplot2::ggsave(
    plot     = p_models,
    filename = here::here(dir_results, glue::glue("{flnm}_{ecor}.png")),
    width = 8, height = 5, dpi = 300
  )

  # 4) SE vs cutoff (event_years)
  p_se <- NULL
  if (nrow(se_df) > 0) {
    p_se <- ggplot2::ggplot(
      se_df,
      ggplot2::aes(x = cutoff, y = se, group = factor(event), linetype = factor(event), color = factor(event))
    ) +
      ggplot2::geom_line() +
      ggplot2::geom_point() +
      ggplot2::labs(
        x = "Conley cutoff (km)",
        y = "Standard error",
        linetype = "Event time (years)",
        color = "Event time (years)",
        title = glue::glue("Conley SE vs cutoff — {ecor}"),
        subtitle = paste0("pixel = ", pixel, " km; distance = ", distance)
      ) +
      ggplot2::scale_x_continuous(breaks = cutoffs) +
      ggplot2::theme_minimal(base_size = 12)

    ggplot2::ggsave(
      plot     = p_se,
      filename = here::here(dir_results, glue::glue("{flnm}_{ecor}_se_by_cutoff.png")),
      width = 7, height = 5, dpi = 300
    )
  }

  # 5) Return only small things
  list(
    plot_data    = plot_data,         # slim data behind comparison plot
    se_by_cutoff = se_df,             # tiny table for requested event years
    plots        = list(comparison = p_models, se_cutoff = p_se)
  )
}



# Use functions on all ecoregions
#Moran & variograms
ecoregions |>
  walk( ~ test_spdep_ecoregion(.x, df.use, dir_sptests))

#Conley tests
# 765 seconds for 4 test ecoregions
tic()
tcs <- ecoregions |>
  map( ~ test_conley(dats = df.use,
                      ecor = .x,
                      cutoffs = c(seq.int(25, 250, 25)),
                      pixel = 20, #pool observations into Xkm grids to speed up calculations
                      distance = "triangular", #use triangular rather than spherical distance to speed up calculations
                      dir_results = dir_sptests)
  )
toc()  

# 790 seconds for 4 test ecoregions
tic()
tcs <- ecoregions |>
  map( ~ test_conley(dats = df.use,
                      ecor = .x,
                      cutoffs = c(seq.int(25, 250, 25)),
                      pixel = 10, #pool observations into Xkm grids to speed up calculations
                      distance = "triangular", #use triangular rather than spherical distance to speed up calculations
                     flnm = "conley_test_pixel10", 
                     dir_results = dir_sptests)
  )
toc()  


```




# Operational across ecoregions

```{r}

#run_nm <- paste0("did_conley_", timestamp('for_file'))
run_nm <- paste0("did_conley_125km_20km")
dir_run <- here(dir_figs, run_nm)
dir_ensure(dir_run)

# Operate ----


x <- df.use |> filter(ecoregion == "SierraNevada")
p <- plot_all_together_did(subg_col = "burn_sev", df = x, title = NULL)
p


model <- feols(
  rap_TRE ~ sunab(FirstTreat, year) | uniqueID + year,
  data = x,
  cluster = "uniqueID")

model <- feols(
  rap_TRE ~ sunab(FirstTreat, year) | uniqueID + year,
  data = x,
  cluster = "uniqueID")

vc <- fixest::vcov_conley(
  model,
  lat = "lat", lon = "long",
  cutoff = 125, pixel = 20, distance = "triangular", vcov_fix = TRUE
)

ggcoefplot(model)



ggiplot(
  list('TWFE' = est_twfe, 'Sun & Abraham (2020)' = est_sa20),
  main = 'Staggered treatment', ref.line = -1, pt.join = TRUE
)

# USE FUNCTIONS TO VIEW SEVERITY BY ECOREGION

all_by_sev_p <- plot_all_ecoregions_did(subg_col = "burn_sev",
                                        df = df.use,
                                        title = "Interaction-Weighted Estimates \nby Ecoregion & Burn Severity, conley100km")


ggsave(plot = all_by_sev_p,
       filename = here(dir_run, "all_did_by_eco_and_sev_plot.png"),
       bg = 'white',
       units = "px",
       width = 4000,
       height = 4000)




all_by_biotic_p <- plot_all_ecoregions_did(subg_col = "biotic_before_10t",
                                        df = df.use,
                                        title = "Interaction-Weighted Estimates by \nEcoregion & Pre-Fire Biotic (10%, 5-yr), conley100km")

ggsave(plot = all_by_biotic_p,
       filename = here(dir_run, "all_did_by_eco_and_biotic10t_plot.png"),
       bg = 'white',
       units = "px",
       width = 4000,
       height = 4000)



all_by_hd_p <- plot_all_ecoregions_did(subg_col = "hd_after",
                                        df = df.use,
                                        title = "Interaction-Weighted Estimates by Ecoregion & Post-fire HD (17, sum, 5-yr)")

ggsave(plot = all_by_hd_p,
       filename = here(dir_run, "all_did_by_eco_and_hd_plot.png"),
       bg = 'white',
       units = "px",
       width = 4000,
       height = 4000)



all_by_pdsi_p <- plot_all_ecoregions_did(subg_col = "pdsi_before",
                                        df = df.use,
                                        title = "Interaction-Weighted Estimates by Ecoregion & Pre-Fire PDSI (-500, sum, 5-yr)")

ggsave(plot = all_by_pdsi_p,
       filename = here(dir_run, "all_did_by_eco_and_pdsi_plot.png"),
       bg = 'white',
       units = "px",
       width = 4000,
       height = 4000)




# RESULTS FOR ALL ECOREGIONS MERGED
together_by_sev_p <- plot_all_together_did(subg_col = "burn_sev_tagged",
                                           df = df.use,
                                           title = "Interaction-Weighted Estimates for all Ecoregions by severity")

ggsave(plot = together_by_sev_p,
       filename = here(dir_run, "together_did_plot_sev.png"),
       bg = 'white')




together_by_biotic_p <- plot_all_together_did(subg_col = "biotic_before",
                                           df = df.use,
                                           title = "Interaction-Weighted Estimates for all Ecoregions by pre-fire biotic (10%, 5-yr)")

ggsave(plot = together_by_biotic_p,
       filename = here(dir_run, "together_did_plot_biotic.png"),
       bg = 'white')




together_by_hd_p <- plot_all_together_did(subg_col = "hd_after",
                                           df = df.use,
                                           title = "Interaction-Weighted Estimates for all Ecoregions by post-fire hd (17, sum, 5yr)")

ggsave(plot = together_by_hd_p,
       filename = here(dir_run, "together_did_plot_hd.png"),
       bg = 'white')





together_by_pdsi_p <- plot_all_together_did(subg_col = "pdsi_before",
                                           df = df.use,
                                           title = "Interaction-Weighted Estimates for all Ecoregions by pre-fire PDSI (-500, sum, 5yr)")

ggsave(plot = together_by_pdsi_p,
       filename = here(dir_run, "together_did_plot_pdsi.png"),
       bg = 'white')






# MULTI-SPLIT

for(burn_l in c("low-moderate-sev", "unburned", "high-sev")) {
  df <- df.use |>
    filter(burn_sev_tagged == burn_l | burn_sev_tagged == "unburned-mock")
  
  by_biotic_sev_p <- plot_all_ecoregions_did(subg_col = "biotic_before_20t",
                                          df = df,
                                          title = paste0(burn_l, ": Interaction-Weighted Estimates by Ecoregion & Pre-Fire Biotic (10%, 5-yr)"))
  
  ggsave(plot = by_biotic_sev_p,
         filename = here(dir_run, paste0("did_by_eco_and_biotic_20t_plot_", burn_l, ".png")),
         bg = 'white',
         units = "px",
         width = 4000,
         height = 4000)
  
  gc()

}


# 
# objs <- ls(envir = .GlobalEnv)
# sizes <- sapply(objs, function(x) object.size(get(x)))
# sizes_df <- data.frame(
#   Object = objs,
#   Size_Bytes = as.numeric(sizes),
#   Size_GB = round(as.numeric(sizes) / 1024^3, 3)
# )
# sizes_df[order(sizes_df$Size_Bytes, decreasing = TRUE), ]

```

```{r}


# SIMPLE
testset <- df.use |>
  filter(ecoregion == "IDBatholith")
model <- feols(
  rap_TRE ~ sunab(FirstTreat, year) | uniqueID + year,
  data = testset,
  cluster = "uniqueID")

# INTERACTED
model_interact <- feols(
  rap_TRE ~ sunab(FirstTreat, year):biotic_before_10t | uniqueID + year,
  data = testset,
  cluster = "uniqueID")

# SUBGROUPED
p_test <- plot_all_together_did(subg_col = "biotic_before_10t",
                      df = testset, title = "Test subg")


# RESULTS
summary(model)
plot_sunab_models(model_list = list(model),
                  model_labels = "test")

summary(model_interact)
interacted_coef <- summary(model_interact)$coefficients
interacted_coef


plot_sunab_models(model_list = list(model_interact),
                  model_labels = "test interact")

p_test


# TRYING TO PLOT THE INTERACTED RESULTS

etable(model_interact, keep = "%year::")


agg_main <- aggregate(model_interact, agg = "cohort", fun = "mean", keep = "^(?!.*biotic_before_10t)")
agg_inter <- aggregate(model_interact, agg = "cohort", fun = "mean", keep = "biotic_before_10t")


tidy_main <- broom::tidy(agg_main) %>%
  mutate(type = "Baseline (biotic_before_10t = 0)")

tidy_inter <- broom::tidy(agg_inter) %>%
  mutate(type = "Effect of biotic_before_10t (per unit)")



plot_df <- bind_rows(tidy_main, tidy_inter) %>%
  mutate(rel_year = as.numeric(gsub("year::(-?\\d+).*", "\\1", term)))

ggplot(plot_df, aes(x = rel_year, y = estimate, color = type)) +
  geom_line() +
  geom_point() +
  geom_errorbar(aes(ymin = estimate - 1.96 * std.error,
                    ymax = estimate + 1.96 * std.error),
                width = 0.2) +
  geom_vline(xintercept = -1, linetype = "dashed") +
  theme_minimal() +
  labs(x = "Event time (years)", y = "Estimated effect",
       color = "Series")

```




<button class="accordion-button" type="button" data-bs-toggle="collapse" data-bs-target="#collapseOne" >Click to see Session Information</button><div id="collapseOne" class="accordion-collapse collapse"><div>

```{r sessionInfo}
sessionInfo()
```