---
title: "05_dydid"
author: Tyler L. McIntosh
date: last-modified
format:
  html:
    embed-resources: true
    html-math-method: katex
    css: styles.css
    toc: true
    toc-depth: 1
    toc-title: Contents
    toc-location: right
    code-fold: true
    code-summary: "Show the code"
---


# THIS VERSION OF THE CODE EXPLORES USING THE INTERNAL PLOTTING FUNCTIONS
# THEY ARE USEFUL FOR ECOREGION SPLITS, BUT NOT FOR DIFFERENT TREATMENT TYPES


# Setup for analysis

``` {r setupchunk}
#| echo: true
#| message: false
#| warning: false
#| results: 'hide'

rm(list = ls())

if(!requireNamespace("here", quietly = TRUE)) {
  install.packages("here")
}
library(here)

source(here::here("utils", "functions.R"))

install_load_packages(c(
  "tidyverse",
  "tictoc",
  "hstats",
  "glue",
  "shapviz",
  "patchwork",
  "cobalt",
  "purrr",
  "corrplot",
  "gridExtra",
  "tictoc",
  "fect",
  "fixest",
  "did",
  "arrow",
  "panelView",
  "PanelMatch",
  "bacondecomp",
  "didimputation",
  "doParallel",
  "HonestDid",
  "HonestDidFEct",
  "DIDmultiplegtDYN",
  "sf",
  "spdep",
  "purrr",
  "gstat",
  "scales",
  "broom",
  "ggfixest"))

# # install packages from CRAN
# packages <- c("fixest", "did", "didimputation", 
#          "panelView", "bacondecomp", "HonestDiD",
#               "DIDmultiplegtDYN", "PanelMatch")
# install.packages(setdiff(packages, rownames(installed.packages())))  
# 
# # install most up-to-date "fect" from Github
# if ("fect" %in% rownames(installed.packages()) == FALSE) {
#   devtools:: install_github("xuyiqing/fect")
# }
# 
# # install forked "HonestDiD" package compatible with "fect"
# if ("HonestDiDFEct" %in% rownames(installed.packages()) == FALSE) {
#   devtools:: install_github("lzy318/HonestDiDFEct")
# }


dir_figs <- here::here('figs')
dir_derived <- here::here('data', 'derived')
dir_raw <- here::here('data', 'raw')
dir_manual <- here::here('data', 'manual')
dir_ensure(c(dir_figs,
             dir_derived,
             dir_manual,
             dir_raw))

seed = 1234
set.seed(seed)


```

#

```{r}
  
dats_long <- arrow::read_parquet(here(dir_derived, "test_dats_long.parquet"))


```


```{r}

# model.twfe.0 <- feols(rap_TRE~treated|uniqueID+year,
#                       data=dats_long, cluster = "uniqueID") 
# print(model.twfe.0)


# # We don't have any always-treated units, so we don't need to worry about the Goodman-Bacon decomposition


dats_long_did <- get.cohort(dats_long, D = "treated",
                                 index = c("uniqueID", "year"),
                                 start0 = TRUE)
dats_long_did[which(is.na(dats_long_did$FirstTreat)),"FirstTreat"] <- 1000 
df.use <- dats_long_did |>
  mutate(biotic_before_10t = case_when(biotic_5_yrs_prior_sum >= 10 ~ "biotic_prior_10t",
                                  biotic_5_yrs_prior_sum < 10 ~ "no_biotic_prior_10t"),
         biotic_before_20t = case_when(biotic_5_yrs_prior_sum >= 20 ~ "biotic_prior_20t",
                                  biotic_5_yrs_prior_sum < 20 ~ "no_biotic_prior_20t"),
         hd_after_17t = case_when(hd_fingerprint_5_yrs_after_sum >= 17 ~ "hd_after_17t",
                              hd_fingerprint_5_yrs_after_sum < 17 ~ "no_hd_after_17t"),
         pdsi_before_n500t = case_when(pdsi_annual_5_yrs_prior_sum <= -500 ~ "pdsi_before_n500t",
                                 pdsi_annual_5_yrs_prior_sum > -500 ~ "no_pdsi_before_n500t"),
         pdsi_before_n1000t = case_when(pdsi_annual_5_yrs_prior_sum <= -1000 ~ "pdsi_before_n1000t",
                                 pdsi_annual_5_yrs_prior_sum > -1000 ~ "no_pdsi_before_n1000t"))


test <- get.cohort(dats_long |> slice_sample(n = 10000), D = "treated",
                                 index = c("uniqueID", "year"),
                                 start0 = TRUE,
                   drop.always.treat = FALSE,
                   varname = c("FCohort", "F_TtT", "help", "eek"))


# 
# n_bio_greater_10 <- dats_short |>
#     filter(ecoregion == "IDBatholith") |>
#   filter(fire == 1 & biotic_5_yrs_prior_sum >= 10)
# n_bio_greater_10_control <- dats_short |>
#     filter(ecoregion == "IDBatholith") |>
#   filter(fire == 0 & biotic_5_yrs_prior_sum >= 10)

# # Dynamic TWFE
# df.twfe <- df.use
# 
# # drop always treated units
# #df.twfe$treat <- as.numeric(df.twfe$treatment_mean>0) 
# df.twfe[which(is.na(df.twfe$Time_to_Treatment)),'Time_to_Treatment'] <- -999 # can be an arbitrary value
# twfe.est <- feols(rap_TRE ~ i(Time_to_Treatment, fire, ref = c(-1, -999))| uniqueID + year, 
#                   data = df.twfe,
#                   cluster = "uniqueID")
# 
# twfe.output <- as.matrix(twfe.est$coeftable)
# print(round(twfe.output, 3))
# 
# twfe.est$collin.var
# 
# table(df.twfe$treated, useNA = "always")
# table(df.twfe$treated, is.na(df.twfe$Time_to_Treatment))
# 
# 
# twfe.output <- as.data.frame(twfe.output)
# twfe.output$Time <- c(c(-29:-2),c(0:21))+1 
# p.twfe <- esplot(twfe.output,Period = 'Time',Estimate = 'Estimate',
#                                SE = 'Std. Error', xlim = c(-29,21))
# p.twfe




# 
# model <- feols(
#   rap_TRE ~ sunab(FirstTreat, year) | year + uniqueID,
#   data = testset |>
#     filter(biotic_before == 0),
#   cluster = "uniqueID"
# )
# summary(model)
# plot_sunab_models(model_list = list(model),
#                   model_labels = "test2 - no biotic before")


```

# Operate using internal fixest commands and ggfixest

# Test ATE model on Sierra Nevada and compare to TWFE without sun abraham estimate
We find that there is a meaningful difference between the two estimates
```{r}

x <- df.use |>
  filter(ecoregion == "SierraNevada")

est_sunab <- feols(
  rap_TRE ~ sunab(FirstTreat, year, ref.p = -1) | uniqueID + year,
  data = x,
  mem.clean = TRUE,
  lean = FALSE
)


ggiplot(sunab, ci_level = 0.95, geom_style = "ribbon", col = "turquoise") +
  xlim(-10, 20)

est_twfe <- feols(
  rap_TRE ~ i(factor_var = Time_to_Treatment, var = treated, ref = -1) | uniqueID + year,
  data = x,
  mem.clean = TRUE,
  lean = FALSE
)


ggiplot(list('Sunab' = sunab, 'TWFE' = twfe),
        ci_level = 0.95,
        #geom_style = "ribbon",
        col = c("turquoise", "pink")) +
  xlim(-10, 20)


```

# Use internal functions to test Conley cutoffs
NOTE: Doesn't work, still need to iterate over 


```{r}

# 1) Build Conley "requests" (not matrices!)
V25_req  <- vcov_conley(lat = ~lat, lon = ~long, cutoff = 25,  pixel = 20, distance = "spherical")
V150_req <- vcov_conley(lat = ~lat, lon = ~long, cutoff = 150, pixel = 20, distance = "spherical")

# 2) Replicate the model into a fixest_list with different VCOVs
mods <- rep(
  est_sunab,
  times = 3,
  vcov = list("hc1", "Conley 25km" = V25_req, "Conley 150km" = V150_req)
)

# 3) Plot both with ggiplot, faceting by model label
ggiplot(mods) +
  theme(legend.position = "right") +
  xlim(-10, 20)

```



# Test split samples using internal fixest functions
The internal functions work well for splitting on things like ecoregion, but I will need custom code to split on anything related to the treatment
```{r}

xx <- df.use |>
  filter(ecoregion == "Idaho Batholith" | ecoregion == "Southern Rockies") |>
  slice_sample(prop = 0.25)

tic()
est_sunab_eco_test <- feols(
  rap_tree ~ sunab(FirstTreat, year, ref.p = -1) | uniqueID + year,
  data = xx,
  mem.clean = TRUE,
  lean = FALSE,
  split = ~ ecoregion,
  vcov = fixest::vcov_conley(lat = "lat",
                             lon = "long",
                             cutoff = 125,
                             pixel = 20,
                             distance = "triangular"
                             )
)
toc()

p1 <- ggiplot(
  est_sunab_eco_test,
  ref.line = -1,
  main = 'Staggered treatment: Split multi-sample',
  xlab = 'Time to treatment',
  multi_style = 'facet',
  geom_style = 'ribbon',
  facet_args = list(labeller = labeller(id = function(x) gsub(".*: ", "", x))),
  theme = theme_minimal()) +
  xlim(-10, 20)

p1



xx_ls <- df.use |>
  filter(ecoregion == "SierraNevada" | ecoregion == "IDBatholith") |>
  filter(is.na(burn_sev) | burn_sev == "low-moderate-sev")

xx_hs <- df.use |>
  filter(ecoregion == "SierraNevada" | ecoregion == "IDBatholith") |>
  filter(is.na(burn_sev) | burn_sev == "high-sev")

# 76.83 sec to run
tic()
est_sunab_eco_test_ls <- feols(
  rap_TRE ~ sunab(FirstTreat, year, ref.p = -1) | uniqueID + year,
  data = xx_ls,
  mem.clean = TRUE,
  lean = FALSE,
  split = ~ ecoregion,
  vcov = fixest::vcov_conley(lat = "lat",
                             lon = "long",
                             cutoff = 125,
                             pixel = 20,
                             distance = "triangular"
                             )
)
toc()

# 63.05 sec
tic()
est_sunab_eco_test_hs <- feols(
  rap_TRE ~ sunab(FirstTreat, year, ref.p = -1) | uniqueID + year,
  data = xx_hs,
  mem.clean = TRUE,
  lean = FALSE,
  split = ~ ecoregion,
  vcov = fixest::vcov_conley(lat = "lat",
                             lon = "long",
                             cutoff = 125,
                             pixel = 20,
                             distance = "triangular"
                             )
)
toc()

ggiplot(
  list('LS' = est_sunab_eco_test_ls, 'HS' = est_sunab_eco_test_hs),
  ref.line = -1,
  main = 'Staggered treatment: Split multi-sample',
  xlab = 'Time to treatment',
  #multi_style = 'facet',
  geom_style = 'ribbon',
  facet_args = list(labeller = labeller(id = function(x) gsub(".*: ", "", x))),
  theme = theme_minimal()) +
  xlim(-10, 20)




# 1) Extract data for each split set
df_ls <- iplot_data(est_sunab_eco_test_ls) |> mutate(Severity = "Low/Moderate")
df_hs <- iplot_data(est_sunab_eco_test_hs) |> mutate(Severity = "High")

# 2) Bind and clean facet labels
df_all <- bind_rows(df_ls, df_hs) |>
  mutate(
    # 'sample' contains strings like "ecoregion: SierraNevada"
    ecoregion = sub(".*: ", "", sample)
  )

# 3) Plot: one facet per ecoregion; both severities overlaid
ggplot(df_all, aes(x = estimate_names, y = estimate,
                   #ymin = ci_low, ymax = ci_high,
                   color = Severity, fill = Severity, linetype = Severity)) +
  geom_hline(yintercept = 0, linetype = 2) +
  geom_ribbon(aes(ymin = ci_low, ymax = ci_high, fill = Severity),
              alpha = 0.3, color = NA) +
  geom_line() +
  facet_wrap(~ ecoregion) +
  coord_cartesian(xlim = c(-10, 20)) +
  labs(x = "Time to treatment", y = "Forest cover effect estimate",
       title = "Staggered treatment: LS vs HS overlaid by ecoregion",
      caption = "95% CIs, Conley standard errors with 125km cutoff and 20km pixel") +
  theme_minimal() +
  guides(fill = guide_legend(title = "Severity"),
         color = guide_legend(title = "Severity"),
         linetype = guide_legend(title = "Severity"))
  

```

# Test computation time with drought & biotic time series included

```{r}


tic()
est_sunab_eco_test_ls_controlled <- feols(
  rap_TRE ~ biotic_sum + pdsi_annual + sunab(FirstTreat, year, ref.p = -1) | uniqueID + year,
  data = xx_ls,
  mem.clean = TRUE,
  lean = FALSE,
  split = ~ ecoregion,
  vcov = fixest::vcov_conley(lat = "lat",
                             lon = "long",
                             cutoff = 125,
                             pixel = 20,
                             distance = "triangular"
                             )
)
toc()

tic()
est_sunab_eco_test_hs_controlled <- feols(
  data = xx_hs,
  mem.clean = TRUE,
  lean = FALSE,
  split = ~ ecoregion,
  vcov = fixest::vcov_conley(lat = "lat",
                             lon = "long",
                             cutoff = 125,
                             pixel = 20,
                             distance = "triangular"
                             )
)
toc()





# 1) Extract data for each split set
df_ls_c <- iplot_data(est_sunab_eco_test_ls_controlled) |> mutate(Severity = "Low/Moderate")
df_hs_c <- iplot_data(est_sunab_eco_test_hs_controlled) |> mutate(Severity = "High")

# 2) Bind and clean facet labels
df_all_c <- bind_rows(df_ls_c, df_hs_c) |>
  mutate(
    # 'sample' contains strings like "ecoregion: SierraNevada"
    ecoregion = sub(".*: ", "", sample)
  )

# 3) Plot: one facet per ecoregion; both severities overlaid
ggplot(df_all_c, aes(x = estimate_names, y = estimate,
                   #ymin = ci_low, ymax = ci_high,
                   color = Severity, fill = Severity, linetype = Severity)) +
  geom_hline(yintercept = 0, linetype = 2) +
  geom_ribbon(aes(ymin = ci_low, ymax = ci_high, fill = Severity),
              alpha = 0.3, color = NA) +
  geom_line() +
  facet_wrap(~ ecoregion) +
  coord_cartesian(xlim = c(-10, 20)) +
  labs(x = "Time to treatment", y = "Forest cover effect estimate",
       title = "Staggered treatment: LS vs HS overlaid by ecoregion",
      caption = "95% CIs, Conley standard errors with 125km cutoff and 20km pixel") +
  theme_minimal() +
  guides(fill = guide_legend(title = "Severity"),
         color = guide_legend(title = "Severity"),
         linetype = guide_legend(title = "Severity"))
  

```
# Functions for treatment split

```{r}



```


# Test pre-fire disturbance with -5 as baseline

```{r}


xx_f <- df.use |>
  filter(ecoregion == "Idaho Batholith" | ecoregion == "Southern Rockies") |>
  filter(is.na(burn_sev) | burn_sev == "low-moderate-sev")

xx_bf <- df.use |>
  filter(ecoregion == "SierraNevada" | ecoregion == "IDBatholith") |>
  filter(is.na(burn_sev) | burn_sev == "high-sev")

# 76.83 sec to run
tic()
est_sunab_eco_test_ls <- feols(
  rap_TRE ~ sunab(FirstTreat, year, ref.p = -1) | uniqueID + year,
  data = xx_ls,
  mem.clean = TRUE,
  lean = FALSE,
  split = ~ ecoregion,
  vcov = fixest::vcov_conley(lat = "lat",
                             lon = "long",
                             cutoff = 125,
                             pixel = 20,
                             distance = "triangular"
                             )
)
toc()

# 63.05 sec
tic()
est_sunab_eco_test_hs <- feols(
  rap_TRE ~ sunab(FirstTreat, year, ref.p = -1) | uniqueID + year,
  data = xx_hs,
  mem.clean = TRUE,
  lean = FALSE,
  split = ~ ecoregion,
  vcov = fixest::vcov_conley(lat = "lat",
                             lon = "long",
                             cutoff = 125,
                             pixel = 20,
                             distance = "triangular"
                             )
)
toc()

ggiplot(
  list('LS' = est_sunab_eco_test_ls, 'HS' = est_sunab_eco_test_hs),
  ref.line = -1,
  main = 'Staggered treatment: Split multi-sample',
  xlab = 'Time to treatment',
  #multi_style = 'facet',
  geom_style = 'ribbon',
  facet_args = list(labeller = labeller(id = function(x) gsub(".*: ", "", x))),
  theme = theme_minimal()) +
  xlim(-10, 20)




# 1) Extract data for each split set
df_ls <- iplot_data(est_sunab_eco_test_ls) |> mutate(Severity = "Low/Moderate")
df_hs <- iplot_data(est_sunab_eco_test_hs) |> mutate(Severity = "High")

# 2) Bind and clean facet labels
df_all <- bind_rows(df_ls, df_hs) |>
  mutate(
    # 'sample' contains strings like "ecoregion: SierraNevada"
    ecoregion = sub(".*: ", "", sample)
  )

# 3) Plot: one facet per ecoregion; both severities overlaid
ggplot(df_all, aes(x = estimate_names, y = estimate,
                   #ymin = ci_low, ymax = ci_high,
                   color = Severity, fill = Severity, linetype = Severity)) +
  geom_hline(yintercept = 0, linetype = 2) +
  geom_ribbon(aes(ymin = ci_low, ymax = ci_high, fill = Severity),
              alpha = 0.3, color = NA) +
  geom_line() +
  facet_wrap(~ ecoregion) +
  coord_cartesian(xlim = c(-10, 20)) +
  labs(x = "Time to treatment", y = "Forest cover effect estimate",
       title = "Staggered treatment: LS vs HS overlaid by ecoregion",
      caption = "95% CIs, Conley standard errors with 125km cutoff and 20km pixel") +
  theme_minimal() +
  guides(fill = guide_legend(title = "Severity"),
         color = guide_legend(title = "Severity"),
         linetype = guide_legend(title = "Severity"))
  





```


<button class="accordion-button" type="button" data-bs-toggle="collapse" data-bs-target="#collapseOne" >Click to see Session Information</button><div id="collapseOne" class="accordion-collapse collapse"><div>

```{r sessionInfo}
sessionInfo()
```